### 2021-11-09 main.tex
Initial Overleaf Import

\documentclass[a4paper,11point]{article}
\usepackage[margin=2cm]{geometry}

%% Fonts, etc.

%% Language and font encodings
\usepackage[english]{babel}
\usepackage[utf8x]{inputenc}
\usepackage[T1]{fontenc}
% \usepackage{microtype}      % fine pdf font control

\usepackage[scaled=0.92]{helvet} \renewcommand{\familydefault}{\sfdefault}
\usepackage{xcolor} % -- loaded by tikz
% \usepackage{soul} % \ul \st \hl

%% Maths
% \usepackage{amssymb,amsmath}
% \usepackage{algorithm,algpseudocode}
% \usepackage{mathtools}
% \usepackage{mathalfa}
% \usepackage{bbm}
% \usepackage{mathrsfs}
% \usepackage{gensymb}
% \input m.macros
% \input m.symbols

%% Other typsetting
% \usepackage{booktabs}       % professional-quality tables
% \usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{enumitem}

%% Graphics
% \usepackage{graphicx,tikz}
% \usetikzlibrary{positioning}
% \usetikzlibrary{fit}
% \usetikzlibrary{calc}

%% Floats
% \usepackage{wrapfig}

%% Margin bubbles
% \usepackage{mbubble}
% \mbubblestyle{ms}{font=\tiny,label=ms,color=blue}
% \newcommand{\comment}[1]{}

%% Sections
\setcounter{secnumdepth}{-2}
\usepackage[small]{titlesec} % format section titles
% \titleformat{command}[shape]{format}{label}{sep}{before-code}[after-code]

% \usepackage{caption,subcaption}
% \DeclareCaptionSubType[Alph]{figure}
% \captionsetup[subfigure]{labelformat=simple,labelfont={sf,bf,large},textfont={sf}}
% \renewcommand{\thesubfigure}{{\sffamily\alph{subfigure}}}

%% Spacing
% \usepackage[doublespacing]{setspace}
\parindent 0pt
\parskip 1ex

%% Citations, refs, hyperlinks
\usepackage[colorlinks=true, allcolors=blue]{hyperref}
% \usepackage{natbib}
% \usepackage[capitalize]{cleveref}
% \crefname{equation}{}{}
% \crefname{Equation}{Equation}{Equations}

%% Title

\title{}
\author{}
% \usepackage{filemod} % access to file modification times
% \date{\Filemodtoday{\jobname}}  % replace \jobname if compiling from wrapper

% \usepackage{titling} % modify title(page) appearance
%\pretitle{\begin{center}\LARGE}
%\posttitle{\par\end{center}\vskip 0.5em}
%\preauthor{\begin{center} \large \lineskip 0.5em \begin{tabular}[t]{c}}
%\postauthor{\end{tabular}\par\end{center}}
%\predate{\begin{center}\large}
%\postdate{\par\end{center}}

\def\ii#1\par{{\color{blue!40}\sl #1}\par}
\def\iibf#1\par{{\color{blue!40}\sl\bfseries #1}\par}
\def\iitem#1\par{\ii\begin{itemize}[nosep]\item #1 \end{itemize}\par}

\begin{document}

\section{Notes}

(This page is not for submission)

Required items:
\begin{itemize}  
  \item case for support (up to eight pages)
  \item justification of resources (two pages)
  \item data management plan (DMP) (up to three pages, data management plan template)
        if you opt not to use the template for your plan, the topics listed in the template must be addressed in the DMP document you do provide
  \item diagrammatic workplan (one page)
  \item management structure (one page)
  \item narrative CV (up to two pages per staff member)
  \item community letters of support or demand (up to 10)  
  \item proposal cover letter.
\end{itemize}

\newpage

% \maketitle

\section{Case for support}

\ii Instructions and guidance:
  \url{https://www.ukri.org/wp-content/uploads/2021/09/BBSRC-150921-FundingOpp-BioinformaticsBiologicalResourcesFund-GuidanceDocument.pdf}

\iibf Assessment criteria.

\ii \url{https://www.ukri.org/wp-content/uploads/2021/09/BBSRC-150921-FundingOpp-BioinformaticsBiologicalResourcesFund-AssessmentCriteria.pdf}

\iibf Scientific quality and strategic relevance of the resource.

\ii Including:
\begin{itemize}[nosep]
    \item the extent to which the resource meets the highest international standards in resource provision in its field
    \item how well the resource is demonstrated to be unique or complementary to other similar existing resources
    \item the extent to which the resource addresses the research and policy priority areas of BBSRC.
\end{itemize}

\iibf Cost effectiveness, particularly considerations for long-term
sustainability beyond BBSRC funding.

\ii The extent to which:
\begin{itemize}[nosep]
    \item the resource delivers value for money relative to the anticipated scientific gains it will provide
    \item long term sustainability options have been considered, addressed, and planned where appropriate, particularly for existing resources.
\end{itemize}

\iibf Potential for economic and social impact beyond the academic
community.

\ii Including:
\begin{itemize}
    \item the extent to which the outputs from the resource will contribute to knowledge and potential for economic return or social impact
    \item how well the proposal has outlined methods of engagement and measures of success in developing milestones and timelines of associated activities.
\end{itemize}

\iibf Fit to the scope.

\ii How well the proposal addresses the scope of the opportunity.

\iibf Assessment criteria adapted to new or existing resources

\ii To allow for a more nuanced assessment between new and existing
resources, the use of the following assessment criteria will be
adapted accordingly.

\ii For new resources, these criteria will assess the ‘plans, potential, and promise’ of the resources.

\iibf Quality of the overall arrangements for resource management, advisory functions, as well as user access and engagement

\ii Including:
\begin{itemize}
    \item the extent to which the proposal has evidenced or planned interaction with relevant users and the broader research community to ensure the aims of the resource are realised and there is sufficient uptake and continued development
    \item the extent to which adequate user access arrangements have been discussed and considered
    \item the set-up of project management and advisory structures of the resource to ensure longevity in delivering the resource to a broad user base.
\end{itemize}

\iibf Need or demand, and potential benefit to the UK academic research community

\ii The extent to which:
\begin{itemize}
    \item the community has demonstrated demand for the proposed resource, relative to the total community size (in particular, proposals for new resources should have consulted their prospective community prior to application)
    \item the proposed resource will deliver and benefit the wider BBSRC community indicating how the proposed resource will help to deliver high-quality research.
\end{itemize}

\newpage
\ii Section headings taken from guide.  We can omit some if clearly not necessary, but should do so with caution.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Background to the Resource}
\ii Introduction of the proposed resource, including its academic and wider 
economic and societal context.

\ii Overview of past and current resource(s) in the subject area in both the UK 
and abroad, including any alternative community resources currently 
available. You should indicate the community size of the intended resource 
and how this relates to the field in which it operates.

\subsection{Details of Resource}

\ii The case for support should outline the full details of the resource and 
associated work packages presented in the proposal.

\iitem Indication whether the project proposed to develop a new
resource or is in support of an existing one.

Existing if general Bonsai enhancements.  If specific to analysis we
may be able to argue for "new" if helpful, while still making
reference to Bonsai installed base.

\iitem Objectives for the proposal should be detailed including
individual measurable targets against which the outcome of the work
will be assessed. This should refer to the objectives set out in the
Je-S proposal form.

\iitem{Significant technical details for the development, maintenance or 
enhancement of the resource must be clearly outlined and indicated 
how this is of internationally exceptional quality. }

\iitem{If applicable, outline any proposed research efforts and how they 
  directly facilitate development of the resource. }

\iitem{For proposals looking to focus on maintaining status quo for an 
existing resource instead of suggesting further development, you 
should detail evidence of why significant upgrades are not required at 
this time and detail why the resource needs continued support to 
maintain world-leading functionality.}

\ii{Additional questions that may be considered: }

\iitem Does the facility begin or continue to support a growing field of 
bioscience, what is the anticipated growth and does the proposal 
adequately accommodate this?

\iitem How will the resource accelerate science within its field and
  beyond?

\iitem What would be the impact on the scientific community if the resource 
did not exist? How would this impact other, possibly dependent 
resources?

\subsection{Community Demand}
\ii Evidence of community demand should be primarily driven from UK 
academic researchers working largely within BBSRC remit – see our 
Forward look for UK Bioscience for more detail on research areas covered by 
BBSRC. Demand from other users (such as academic communities outside 
of BBSRC remit or industrial users) may be appropriate to provide additional 
support, especially in highlighting the potential for economic, commercial or 
societal impact, but should not be the focus of the demand demonstration. 
Evidence provided should highlight examples of the high-quality science that 
the resource will underpin or has underpinned. Where possible and relevant, 
examples should be drawn from a wide research community to illustrate the 
broad impact of the resource to support high-quality internationally excellent 
science.

\ii The level of community demand should be benchmarked against other 
relevant resources and/or the size of the community. This will allow the fair 
assessment of resources with different user bases. The types of evidence 
that may be appropriate to provide will be different for new and existing 
resources.

\ii Evidence of wider consultation of the prospective community (e.g community 
surveys) is encouraged.

\iibf New Resources

\ii New resources should estimate the number of researchers who may engage and benefit from the resource. Evidence, where possible, would be of benefit and may include: 
\begin{itemize}
    \item  Datasets (or samples) in public or private repositories 
    \item Citations or acknowledgments 
    \item Gap analysis with existing resources 
    \item Pilot project uptake or feedback from potential users.  
\end{itemize}

\ii In particular, proposals for new resources should have consulted their prospective community prior to application

\iibf Existing Resources

\ii For existing resources this should include usage data of the
current resource. Data types may include:

\iitem Access requests from independent users/ sites

\iitem Citations or acknowledgements

\iitem Other public resources providing links to the resource

\iitem New major acquisitions captured by the resource

\ii{Additionally, existing resources need 
to evidence why this resource needs 
to be maintained/updated by the 
current grant. This could include: }

\iitem Survey data from users on what upgrades are needed

\iitem Evidence of an expanding 
user base, which requires 
additional resource

\iitem Recent developments in the 
field, which require upgrades 
to be integrated into the 
resource.

\subsection{User engagement}
\ii Discussion of user engagement provision should aim to answer the following 
questions:

\iitem Is there awareness of the resource within the user community?

\iitem How do you plan to develop the engagement strategy within the 
proposal timeframe to expand user engagement?

\iitem How have access mechanisms to ensure usability of the resource 
been considered?

\iitem How have user needs been incorporated into this proposal to ensure 
it is fit for purpose and will deliver on expectations?

\iibf New Resources

\ii Evidence should be provided as to how the resource plans to engage
stakeholders and ensure that the resource meets their needs and is
used by the community targeted by the resource.

\iibf Existing Resources

\ii Evidence should be provided as to whether the resource has
achieved the level of engagement it originally anticipated, and
consideration is given how the additional investment would change
this.

\subsection{Long term sustainability planning}
\ii{
In addition, the case for support should outline considerations for the long-
term sustainability of the resource beyond UKRI-BBSRC funding, as well as 
the true cost of running and maintaining the resource in question. 
The proposal should include: }

\iitem Cost recovery plans, where appropriate, or an explanation why
not if not viable. Evidence of clear business planning with a focus on
at least partial cost recovery is required, especially when applying
as an existing resource.

\iitem Details for alternative support plans, aside from UKRI-BBSRC 
funding.

\iitem The level of support the resource is projected to require for expected 
maintenance and/or subsequent maturation/enhancement activities. 
Clear arguments as to why UKRI-BBSRC should support the resource now 
should be provided, if other cost recovery and support plans are deemed 
unsuitable.

\subsection{Potential for economic and societal impact}
\ii{Outline how the outputs of the proposed resource will contribute to 
knowledge and how this may have the potential for economic return or 
societal benefits. Impact activities should be integrated into appropriate 
sections of the case for support, not presented as an independent work 
package. }

\iitem All proposals are expected to demonstrate clear plans with recorded 
milestones and timelines for associated activities to develop 
economic, commercial and societal impacts.  
August 2021 7

\iitem Methods of engagement and measures of success should be outlined 
including how these will be regularly reviewed throughout the project 
in order to deliver the most impact.

\iitem Any planned activities should be fully justified within the Justification 
of Resources attachment

\subsection{Track Record}
\ii{
The majority of the track record relevant to the project should be located 
within the Narrative CV and should not be repeated within the case for 
support. You may, however, want to describe: }

\iitem Track record of the team working together

\iitem Specific expertise available at the host organisation and any 
proposed partner organisations to enable the successful delivery of 
the project.

\iitem The specific role of each applicant and collaborator in the project

\end{document}

### 2021-11-10 main.tex
Updates from Overleaf
Added text from EPSRC proposal

@@
 %% Sections
-\setcounter{secnumdepth}{-2}
-\usepackage[small]{titlesec} % format section titles
+% \setcounter{secnumdepth}{-2}
+% \usepackage[small]{titlesec} % format section titles
 % \titleformat{command}[shape]{format}{label}{sep}{before-code}[after-code]
@@
 % \usepackage[doublespacing]{setspace}
-\parindent 0pt
-\parskip 1ex
+% \parindent 0pt
+% \parskip 1ex
 
@@
 \usepackage[colorlinks=true, allcolors=blue]{hyperref}
+\usepackage{natbib}
+\usepackage{apalike}
+\usepackage{graphicx}
 % \usepackage{natbib}
@@
 
-\section{Notes}
+\noindent\textbf{Notes}
 
@@
 \end{itemize}
+
+\tableofcontents
 
@@
 
+Experimental control and neural data analysis are currently highly
+disassociated.  Most neuroscientists collect their data and only afterward
+analyze it. This split has had dramatic consequences for the progress of
+neuroscience. First, most experimental control methods currently do not use
+advanced data analysis methods to enhance their functionality. Second, the vast
+majority of neural data analysis methods are offline and do not provide key
+functionality needed for online experimental control. Using advanced data
+analysis methods in experimental control will improve the sophistication of
+experiments that are currently possible, and generate new demands for neural
+data analysis methods that will generate advances in the field.
+
+A central goal of the software development proposed here is to address this
+problem, by providing Bonsai~(an experimental control software ecosystem;
+Figure~\ref{fig:bonsai} and Section~\ref{sec:bonsai}) with state-of-the-art online (and offline) data
+analysis methods (Figure~\ref{fig:proposedBonsaiExtensions} and 
+Section~\ref{sec:toBeAddedMachineIntelligenceFunctionality}).
+%
+It is unfortunate that neuroscientists currently do not have software that
+allows them to easily compare the performance of different data analysis
+methods for the characterization of their datasets. The second central goal
+of this proposal is to address this limitation by adding to Bonsai a generic
+mechanism for comparing the performance of different methods for the analysis
+of specific datasets (Section~\ref{sec:comparisonOfMultipleMethods}).
+
+\begin{figure}
+    \begin{center}
+        \includegraphics[width=7in]{figures/bonsai.pdf}
+        \label{fig:bonsai}
+    \end{center}
+\end{figure}
+
+\setcounter{figure}{1}
+
+\begin{figure}
+    \begin{center}
+        \includegraphics[width=4in]{figures/proposed_bonsai_extensions.png}
+        \caption{Proposed extensions to Bonsai}
+        \label{fig:proposedBonsaiExtensions}
+    \end{center}
+\end{figure}
+
+Currently Bonsai has its own package manager, and its community of users have
+already extended Bonsai's functionality with several contributed packages
+(e.g., \ldots \textcolor{red}{Goncalo please reference the best contributed
+packages developed by the Bonsai community}). However, these packages need to
+be written in C\#, while most current neuroscience data analysis methods are
+written in Python, R or Matlab.  We propose to add to Bonsai capabilities to
+communicate with software written in these languages
+(Section~\ref{sec:bonsaiPythonMatlabCommunication}).  These capabilities will
+allow a large number of neuroscience data analysis methods to be easily
+integrated into the Bonsai ecosystem and will provide Bonsai users a large
+repertoire of advance machine learning methods for their behavioral and neural
+data analysis.
+
+Bonsai is an excellent tool for reproducible data acquisition and
+experimental control. An experiment implemented in Bonsai can be replicated in
+a new laboratory by just sharing a Bonsai configuration file. With the addition
+of the proposed machine learning functionality Bonsai will extend this
+reproducibility to the domain of data analysis
+(Section~\ref{sec:reproducibility}).
+
+\subsubsection{Bonsai}
+\label{sec:bonsai}
+
+Bonsai~\citep{lopesEtAl15,lopesAndMonteiro21} is a free and open-source visual
+programming language that emphasizes performance, flexibility, and ease-of-use,
+allowing scientists with no previous programming experience to quickly develop
+their own high-performance data acquisition and experimental control systems
+(Figure 1c). Bonsai combines a high-level event algebra for data streams with
+an integrated development environment (IDE) and a strong library of plugins
+supporting multiple hardware and software packages used by the biomedical
+research community (Figures 1a, 1b).
+
+Bonsai has a large user base in the systems neuroscience community.
+Quantifications of this user base are provided in Figures 1d-1f.  In the last
+year alone, we estimate more than 1,000 new users have started to incorporate
+Bonsai into their experimental protocols across the world. The surprising rate
+of adoption of Bonsai in non-programmer experimental labs highlights the need
+for accessible programming tools that enable state-of-the-art technology but
+also allow researchers to stay in control and change their experimental
+paradigms.  Many open-source software tools are either inaccessible to
+non-programmers, or too constrained to be of general use outside their narrow
+domain of application. Bonsai has been successful because it has straddled this
+gap to some extent.
+
+The language has also helped to potentiate the growing wave of foundational
+open hardware initiatives, such as the OpenEphys \citep{siegleEtAl17} and UCLA
+Miniscope~\citep{caiEtAl16}, making it possible to quickly combine and
+integrate these tools into new experiments~\citep{buccinoEtAl18}.
+%
+Bonsai has been adopted in large neuroscience undertakings like the
+International Brain
+Laboratory\footnote{\href{https://www.internationalbrainlab.com/}{https://www.internationalbrainlab.com/}}
+and the Allen Institute for Brain
+Science\footnote{\href{https://alleninstitute.org/what-we-do/brain-science/}{https://alleninstitute.org/what-we-do/brain-science/}}.
+%
+More recently, Bonsai has started to expand outside the domain of neuroscience
+into biomedical research and biotechnology tool development, and even outside
+academia into public outreach and education programs.
+
 \subsection{Details of Resource}
@@
 resources?
+
+\subsubsection{Video-based behavioral analysis}
+\label{sec:videoBasedBehavioralAnalysis}
+
+Precisely quantifying animal behavior is an essential step toward understanding
+brain function.  Deeplabcut is a Python-based software for tracking animal body
+parts \citep{mathisEtAl18}, which it is currently well integrated with
+Bonsai~\citep{kaneEtAl20}. Here we propose extensions to Bonsai to extract
+other informative features from video recordings.
+
+\begin{description}
+
+    \item[Motion sequencing]\citep[MoSeq;][]{wiltschkoEtAl15}: extracts patterns
+        of behavior that repeat over time (i.e., syllables of behavior) from
+        video data. For instance, it parses (in an unsupervised way) the
+        behavior of a mouse in an arena into segments of time where a mouse
+        was running, rearing and grooming. It uses a hidden Markov model and it
+        is implemented in Python\footnote{Code for MoSeq can be requested from
+        the Datta laboratory, as indicated at
+        \href{http://datta.hms.harvard.edu/research/behavioral-analysis/}{http://datta.hms.harvard.edu/research/behavioral-analysis/}.}.
+        After trained MoSeq can be used to detect behavioral syllables online.
+
+    \item[Decoding behavior from neural
+        activity]\citep[BehaveNet;][]{battyEtAl19} combines hidden Markov
+        models with convolutional autoencoders and discriminative models to
+        decode video data from neural recordings. It is implemented in
+        Python\footnote{\href{https://github.com/themattinthehatt/behavenet}{https://github.com/themattinthehatt/behavenet}}.
+        After trained it can be used online to decode video data and detect
+        behavioral syllables.
+
+    \item[Interpretable latents for behavioral videos]\citep[Partitioned
+        Subspace Variational Autoencoder, PS-VAE;][]{whitewayEtAl21}: produces
+        interpretable low-dimensional representations of behavioral videos by
+        combining the output of pose-estimation algorithms (e.g., DeepLabCut)
+        with unsupervised dimensionality reduction methods. These
+        low-dimensional representations facilitate downstream behavioral and
+        neural analyses. It is based on autoencoders and is implemented in
+        Python\footnote{code for PS-VAE is embedded in the BehaveNet code
+        \href{https://github.com/themattinthehatt/behavenet}{https://github.com/themattinthehatt/behavenet}.
+        Please refer to class \texttt{PSVAE} in 
+        \texttt{behavenet.models.vaes.py}.}. After trained it can be used
+        online to extract low-dimensional features and perform downstream
+        processing.
+
+\end{description}
+
+\subsubsection{Spike sorting}
+\label{sec:spikeSorting}
+
+\noindent\textcolor{red}{TODO}
+
+\subsubsection{Low-dimensional representations of spiking data}
+\label{sec:lowDimensionalRepresentationsOfSpikingData}
+
+Bonsai is well integrated with OpenEphys~\citep{siegleEtAl17}, which allows
+scientists to record neural data from a large number of devices. However, it
+lacks functionality to process these recordings. Here we describe software that
+we propose to integrate into Bonsai to extract interpretable summary statistics
+(i.e., latent variables) from neural spiking activity.
+
+\begin{description}
+
+    \item[Gaussian Linear Dynamical
+        System]\citep[GLDS;][]{andersonAndMoore12}: with sufficiently large
+        bin sizes, spike counts can be modeled as Gaussian random processes.
+        This Gaussianity assumption greatly simplifies the estimation of
+        parameters of linear dynamical system (LDS) models, as well as
+        inferences about their states. After models parameters have been
+        learned, GLDS can be used online. A unique feature of GLDS is that the
+        posterior distribution of states given all observation up to the
+        present can be calculated efficiently. This estimate of the posterior
+        distribution can be used online for experimental control, as we
+        propose in Section~\ref{sec:comparisonOfMultipleMethods}. We will use an R implementation of GLDS
+        which allows the use of external
+        inputs\footnote{\href{https://github.com/joacorapela/kalmanFilter}{https://github.com/joacorapela/kalmanFilter}}.
+
+    \item[Poisson Linear Dynamical System]\citep[PLDS;][]{mackeEtAl15}: for
+        smaller bin sizes, spike counts are better modeled as Poisson random
+        processes, rather than Gaussian ones. The algorithm described in
+        \citet{mackeEtAl15} can estimate the parameters of a LDS, and make
+        inferences about its states, from Poisson distributed observations.  We
+        propose to interface Bonsai with a Matlab implementation of
+        PLDS\footnote{\href{https://bitbucket.org/mackelab/pop\_spike\_dyn/src/master/}{https://bitbucket.org/mackelab/pop\_spike\_dyn/src/master/}}
+        that uses variational inference. PLDS does not provide online estimates
+        of the states, since data from a full trial is needed for state
+        inference.
+
+    \item[Hidden Markov Model]\citep[HMM;][]{rabiner89}: as LDSs, HMMs model a
+        time series of observations as random processes conditioned on hidden
+        states. However, in HMMs hidden states are discrete random variables,
+        while in LDSs they are continuous ones. In some application domains
+        (e.g., speech, epilepsy) discrete state assumptions are more pertinent
+        than continuous ones. We propose to use an R implementation of
+        HMMs\footnote{\href{https://github.com/joacorapela/hiddenMarkovModels}{https://github.com/joacorapela/hiddenMarkovModels}}.
+        As GLDSs, once trained, HMMs can be used online to infer the posterior
+        distribution of states given observations.
+
+    \item[Gaussian Processes Factor Analysis]\citep[GPFA;][]{yuEtAl09}: as
+        LDSs, GFPA models represent a time series of observation as random
+        processes conditioned on hidden states. However, in GPFA models the
+        state dynamics are nonlinear, while in LDS models they are linear.
+        Thus, GPFA models are more general than LDS ones. As GLDS models, GPFA
+        models assume that observations conditioned on states are Gaussian
+        random processes. We propose to use a Matlab implementation of
+        GPFA\footnote{\href{https://users.ece.cmu.edu/~byronyu/software/gpfa0203.tgz}{https://users.ece.cmu.edu/~byronyu/software/gpfa0203.tgz}}.
+        GPFA models do not provide online estimates of the states, since data
+        from the full trial are needed for state estimates.
+
+    \item[Sparse Variational Gaussian Processes Factor
+        Analysis]\citep[svGPFA;][]{dunckerAndSahani18}: is similar to GPFA, but
+        models point process observations (i.e., single spikes as opposed to
+        spike counts). We propose to use a Python implementation of
+        svGPFA\footnote{\href{https://github.com/joacorapela/svGPFA}{https://github.com/joacorapela/svGPFA}}.
+        svGPFA models do not provide online estimates of states, since data
+        from the full trial are needed for state estimates.
+
+    \item[Latent factor analysis through dynamical
+        systems]\citep[LFADS;][]{pandarinathEtAl18}: uses an autoencoder
+        framework, with recursive neural networks, to infer continuous states
+        conditioned on spike counts, similar to those inferred by GPFA. We
+        propose to use a Python implementation at
+        LFADS\footnote{\href{https://github.com/tensorflow/models/tree/master/research/lfads}{https://github.com/tensorflow/models/tree/master/research/lfads}.}.
+        As GPFA, LFADS do not provide online estimates of states.
+
+\end{description}
+
+\subsubsection{Low-dimensional representations of local-field potentials}
+\label{sec:lowDimensionalRepresentationsOfLFPs}
+
+Spikes are extracted from a higher-frequency range of extracellulary recorded
+voltages. Local field potentials (LFPs) are computed from a lower-frequency
+range and are another important signal to understand brain function. We propose
+to use states inferred from LFPs by GLDS, GPFA, HMM and LFADS
+(Section~\ref{sec:lowDimensionalRepresentationsOfSpikingData}) as
+low-dimensional representations of the LFP.
+
+\subsubsection{Neural decoding}
+\label{sec:neuralDecoding}
+
+In order to use low-dimensional representations of spiking activity and/or of
+LFPs to guide the control of experiments, we need a decoder to optimally map
+these low-dimensional representations to experimental controls.
+%
+We propose to implement in Bonsai several decoding/classification algorithms:
+k-nearest neighbor, linear discriminative analysis, support vector machines,
+random forests, artificial neural networks, naive Bayes and Gaussian process
+classifiers.
+
+\subsubsection{Comparing multiple data-analysis methods}
+\label{sec:comparisonOfMultipleMethods}
+
+We propose to add functionality to Bonsai to facilitate multi-method
+comparisons in user-supplied datasets. Below we describe a comparison that we
+will perform in Bonsai to asses the relative performance of the methods
+described in the previous sections with neural recordings from behaving rats.
+These recordings are currently being performed, to address scientific questions,
+at the laboratory of Prof.~Akrami in the SWC.
+%
+Details of the experimental task appear in~\citet{akramiEtAl18}.
+%
+Briefly, in this task there are three nose ports (left, center, right). Rats
+initiate a trial by inserting their nose in the center port for the duration of
+the fixation period, until they hear a Go stimulus. During the fixation period
+two stimuli $s_a$ and $s_b$ are presented. Rats should decide which stimuli is
+louder. If $s_a$ is louder than $s_b$ the correct action is to poke the nose
+into the right port in order to collect a liquid reward, and if $s_b$ is louder
+than $s_a$ the correct choice is the left port.
+
+We will use high-density Neuropixel spike and LFP recordings from the rats
+performing this task to learn low-dimensional representations of spiking
+activity and LFPs during the decision time period (between the presentation of
+the last auditory stimuli and the presentation of the Go stimulus), using the
+methods described in
+Sections~\ref{sec:lowDimensionalRepresentationsOfSpikingData}
+and~\ref{sec:lowDimensionalRepresentationsOfLFPs}. We will also learn the
+parameters of the classifiers described in Section~\ref{sec:neuralDecoding} to
+optimally classify animal decisions (poke the nose into the right/left port)
+based on the previous low-dimensional representations.
+
+We will compare the decoding accuracy of all decoders. This comparison will
+tells us, for the auditory working memory task, which neural data type (spikes
+or LFPs), which low-dimensional representation method (e.g., LDS or Gaussian
+processes), and which decoder (e.g., Bayesian or ANN) yields the best decodings
+in state-of-the-art neural recordings from behaving animals.
+
+\subsubsection{Interfacing Bonsai with Python, R and Matlab data analysis software}
+\label{sec:bonsaiPythonMatlabCommunication}
+
+Bonsai is written in C\# and most neural data-analysis methods are written in
+Python, R or Matlab. Fortunately, there exist software that allow C\# program
+to call and be called by programs in these other languages. For the
+communication of C\# programs with Python we will use
+Python.NET\footnote{\href{https://github.com/pythonnet/pythonnet}{https://github.com/pythonnet/pythonnet}},
+with R programs we will use
+R.Net\footnote{\href{https://github.com/rdotnet/rdotnet}{https://github.com/rdotnet/rdotnet}}
+and with Matlab we will use the built in Matlab .NET
+interface\footnote{\href{https://uk.mathworks.com/help/matlab/matlab\_external/using-net-from-matlab-an-overview.html}{https://uk.mathworks.com/help/matlab/matlab\_external/using-net-from-matlab-an-overview.html}}.
+In cases where these software cannot support some type of communication between
+C\# and another language we will use the messaging library
+ZeroMQ\footnote{\href{https://zeromq.org/}{https://zeromq.org/}}.
+
+\subsubsection{Reproducibility with Bonsai}
+\label{sec:reproducibility}
+
+The current version of Bonsai facilitates reproducibility of data acquisition
+and experimental control across laboratories. Most data acquisition and
+experimental control aspects of an experiment can be reproduced across
+different laboratories by sharing a Bonsai configuration file.
+%
+By adding data analysis functionality to Bonsai, besides reproducing my data
+acquisition and experimental control, Bonsai users will also be able to reproduce
+data analysis functionality.
+
+We will demonstrate each machine intelligence functionality added to Bonsai
+(Section~\ref{sec:toBeAddedMachineIntelligenceFunctionality}) and the multiple
+data-analysis methods comparisons
+(Section~\ref{sec:comparisonOfMultipleMethods}) in Bonsai experiments. We will
+then distribute Bonsai configuration files for users to reproduce these
+demonstrations.
+
+\subsubsection{Final remarks}
+\label{sec:finalRemarks}
+
+We proposed to add machine intelligence functionality to Bonsai to assist users
+in building close-loop neural experiments: online spike sorting, methods for
+low-dimensional representation of spiking activity and LFPs, and neural
+decoding algorithms. This functionality will allow neuroscience Bonsai users to
+build more sophisticated experiments and extract more information from their
+experimental recordings.
+
+We also proposed to add functionality to help Bonsai users compare the
+performance of multiple methods on their datasets. With this functionality
+Bonsai users will be able to select the best methods to model their own
+recordings. It will also motivate methods developers to interface their
+data-analysis functionality with Bonsai. By doing so, they will provide their
+software to the large community of Bonsai users, and will be able to easily
+compare the performance of their methods with that of preexisting methods in the
+Bonsai ecosystem. In this way method developers will become a new type of
+Bonsai users, which will guarantee the sustainability of developments
+proposed here.
+
+We focused this proposal on neuroscience applications of Bonsai. However, the
+addition of machine intelligence functionality is needed in many other
+experimental areas where Bonsai is or could be used. Therefore, the software
+abstractions that we will build to add to Bonsai machine intelligence
+functionality for neuroscience experiments (e.g., methods to perform
+multi-method comparisons) will translate to other application domains of
+Bonsai.
+
+Multiple groups at the SWC currently use Bonsai for data acquisition and
+experimental control. Carefully testing the functionality of new software is
+essential to ensure its correct functionality. We propose to use the SWC large
+internal Bonsai user base to test the functionality we will add to Bonsai,
+before distributing it to the general public.
 
@@
 
+\bibliographystyle{apalike}
+\bibliography{bonsai,epsrcSoftwareForResearchCommunities,monitoringBehavior,stateSpaceModels,machineLearning,gaussianProcesses}
+
 \end{document}

### 2021-11-10 caseForSupport.tex
Added

--- 2021-11-10 main.tex
@@
 
-%% Fonts, etc.
-
-%% Language and font encodings
-\usepackage[english]{babel}
-\usepackage[utf8x]{inputenc}
-\usepackage[T1]{fontenc}
-% \usepackage{microtype}      % fine pdf font control
-
-\usepackage[scaled=0.92]{helvet} \renewcommand{\familydefault}{\sfdefault}
-\usepackage{xcolor} % -- loaded by tikz
-% \usepackage{soul} % \ul \st \hl
-
-%% Maths
-% \usepackage{amssymb,amsmath}
-% \usepackage{algorithm,algpseudocode}
-% \usepackage{mathtools}
-% \usepackage{mathalfa}
-% \usepackage{bbm}
-% \usepackage{mathrsfs}
-% \usepackage{gensymb}
-% \input m.macros
-% \input m.symbols
-
-%% Other typsetting
-% \usepackage{booktabs}       % professional-quality tables
-% \usepackage{nicefrac}       % compact symbols for 1/2, etc.
-\usepackage{enumitem}
-
-%% Graphics
-% \usepackage{graphicx,tikz}
-% \usetikzlibrary{positioning}
-% \usetikzlibrary{fit}
-% \usetikzlibrary{calc}
-
-%% Floats
-% \usepackage{wrapfig}
-
-%% Margin bubbles
-% \usepackage{mbubble}
-% \mbubblestyle{ms}{font=\tiny,label=ms,color=blue}
-% \newcommand{\comment}[1]{}
-
-%% Sections
-% \setcounter{secnumdepth}{-2}
-% \usepackage[small]{titlesec} % format section titles
-% \titleformat{command}[shape]{format}{label}{sep}{before-code}[after-code]
-
-% \usepackage{caption,subcaption}
-% \DeclareCaptionSubType[Alph]{figure}
-% \captionsetup[subfigure]{labelformat=simple,labelfont={sf,bf,large},textfont={sf}}
-% \renewcommand{\thesubfigure}{{\sffamily\alph{subfigure}}}
-
-%% Spacing
-% \usepackage[doublespacing]{setspace}
-% \parindent 0pt
-% \parskip 1ex
-
-%% Citations, refs, hyperlinks
 \usepackage[colorlinks=true, allcolors=blue]{hyperref}
@@
 \usepackage{graphicx}
-% \usepackage{natbib}
-% \usepackage[capitalize]{cleveref}
-% \crefname{equation}{}{}
-% \crefname{Equation}{Equation}{Equations}
 
-%% Title
-
-\title{}
+\title{Case for Support}
 \author{}
-% \usepackage{filemod} % access to file modification times
-% \date{\Filemodtoday{\jobname}}  % replace \jobname if compiling from wrapper
-
-% \usepackage{titling} % modify title(page) appearance
-%\pretitle{\begin{center}\LARGE}
-%\posttitle{\par\end{center}\vskip 0.5em}
-%\preauthor{\begin{center} \large \lineskip 0.5em \begin{tabular}[t]{c}}
-%\postauthor{\end{tabular}\par\end{center}}
-%\predate{\begin{center}\large}
-%\postdate{\par\end{center}}
-
-\def\ii#1\par{{\color{blue!40}\sl #1}\par}
-\def\iibf#1\par{{\color{blue!40}\sl\bfseries #1}\par}
-\def\iitem#1\par{\ii\begin{itemize}[nosep]\item #1 \end{itemize}\par}
 
@@
 
-\noindent\textbf{Notes}
-
-(This page is not for submission)
-
-Required items:
-\begin{itemize}  
-  \item case for support (up to eight pages)
-  \item justification of resources (two pages)
-  \item data management plan (DMP) (up to three pages, data management plan template)
-        if you opt not to use the template for your plan, the topics listed in the template must be addressed in the DMP document you do provide
-  \item diagrammatic workplan (one page)
-  \item management structure (one page)
-  \item narrative CV (up to two pages per staff member)
-  \item community letters of support or demand (up to 10)  
-  \item proposal cover letter.
-\end{itemize}
+\maketitle
 
@@
 
-% \maketitle
-
-\section{Case for support}
-
-\ii Instructions and guidance:
-  \url{https://www.ukri.org/wp-content/uploads/2021/09/BBSRC-150921-FundingOpp-BioinformaticsBiologicalResourcesFund-GuidanceDocument.pdf}
-
-\iibf Assessment criteria.
-
-\ii \url{https://www.ukri.org/wp-content/uploads/2021/09/BBSRC-150921-FundingOpp-BioinformaticsBiologicalResourcesFund-AssessmentCriteria.pdf}
-
-\iibf Scientific quality and strategic relevance of the resource.
-
-\ii Including:
-\begin{itemize}[nosep]
-    \item the extent to which the resource meets the highest international standards in resource provision in its field
-    \item how well the resource is demonstrated to be unique or complementary to other similar existing resources
-    \item the extent to which the resource addresses the research and policy priority areas of BBSRC.
-\end{itemize}
-
-\iibf Cost effectiveness, particularly considerations for long-term
-sustainability beyond BBSRC funding.
-
-\ii The extent to which:
-\begin{itemize}[nosep]
-    \item the resource delivers value for money relative to the anticipated scientific gains it will provide
-    \item long term sustainability options have been considered, addressed, and planned where appropriate, particularly for existing resources.
-\end{itemize}
-
-\iibf Potential for economic and social impact beyond the academic
-community.
-
-\ii Including:
-\begin{itemize}
-    \item the extent to which the outputs from the resource will contribute to knowledge and potential for economic return or social impact
-    \item how well the proposal has outlined methods of engagement and measures of success in developing milestones and timelines of associated activities.
-\end{itemize}
-
-\iibf Fit to the scope.
-
-\ii How well the proposal addresses the scope of the opportunity.
-
-\iibf Assessment criteria adapted to new or existing resources
-
-\ii To allow for a more nuanced assessment between new and existing
-resources, the use of the following assessment criteria will be
-adapted accordingly.
-
-\ii For new resources, these criteria will assess the ‘plans, potential, and promise’ of the resources.
-
-\iibf Quality of the overall arrangements for resource management, advisory functions, as well as user access and engagement
-
-\ii Including:
-\begin{itemize}
-    \item the extent to which the proposal has evidenced or planned interaction with relevant users and the broader research community to ensure the aims of the resource are realised and there is sufficient uptake and continued development
-    \item the extent to which adequate user access arrangements have been discussed and considered
-    \item the set-up of project management and advisory structures of the resource to ensure longevity in delivering the resource to a broad user base.
-\end{itemize}
-
-\iibf Need or demand, and potential benefit to the UK academic research community
-
-\ii The extent to which:
-\begin{itemize}
-    \item the community has demonstrated demand for the proposed resource, relative to the total community size (in particular, proposals for new resources should have consulted their prospective community prior to application)
-    \item the proposed resource will deliver and benefit the wider BBSRC community indicating how the proposed resource will help to deliver high-quality research.
-\end{itemize}
-
-\newpage
-\ii Section headings taken from guide.  We can omit some if clearly not necessary, but should do so with caution.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\subsection{Background to the Resource}
-\ii Introduction of the proposed resource, including its academic and wider 
-economic and societal context.
-
-\ii Overview of past and current resource(s) in the subject area in both the UK 
-and abroad, including any alternative community resources currently 
-available. You should indicate the community size of the intended resource 
-and how this relates to the field in which it operates.
+\section{Background to the Resource}
 
@@
 
-\subsubsection{Bonsai}
+\subsection{Bonsai}
 \label{sec:bonsai}
@@
 
-\subsection{Details of Resource}
+\section{Details of Resource}
 
-\ii The case for support should outline the full details of the resource and 
-associated work packages presented in the proposal.
-
-\iitem Indication whether the project proposed to develop a new
-resource or is in support of an existing one.
-
-Existing if general Bonsai enhancements.  If specific to analysis we
-may be able to argue for "new" if helpful, while still making
-reference to Bonsai installed base.
-
-\iitem Objectives for the proposal should be detailed including
-individual measurable targets against which the outcome of the work
-will be assessed. This should refer to the objectives set out in the
-Je-S proposal form.
-
-\iitem{Significant technical details for the development, maintenance or 
-enhancement of the resource must be clearly outlined and indicated 
-how this is of internationally exceptional quality. }
-
-\iitem{If applicable, outline any proposed research efforts and how they 
-  directly facilitate development of the resource. }
-
-\iitem{For proposals looking to focus on maintaining status quo for an 
-existing resource instead of suggesting further development, you 
-should detail evidence of why significant upgrades are not required at 
-this time and detail why the resource needs continued support to 
-maintain world-leading functionality.}
-
-\ii{Additional questions that may be considered: }
-
-\iitem Does the facility begin or continue to support a growing field of 
-bioscience, what is the anticipated growth and does the proposal 
-adequately accommodate this?
-
-\iitem How will the resource accelerate science within its field and
-  beyond?
-
-\iitem What would be the impact on the scientific community if the resource 
-did not exist? How would this impact other, possibly dependent 
-resources?
-
-\subsubsection{Video-based behavioral analysis}
+\subsection{Video-based behavioral analysis}
 \label{sec:videoBasedBehavioralAnalysis}
@@
 
-\subsubsection{Spike sorting}
+\subsection{Spike sorting}
 \label{sec:spikeSorting}
@@
 
-\subsubsection{Low-dimensional representations of spiking data}
+\subsection{Low-dimensional representations of spiking data}
 \label{sec:lowDimensionalRepresentationsOfSpikingData}
@@
 
-\subsubsection{Low-dimensional representations of local-field potentials}
+\subsection{Low-dimensional representations of local-field potentials}
 \label{sec:lowDimensionalRepresentationsOfLFPs}
@@
 
-\subsubsection{Neural decoding}
+\subsection{Neural decoding}
 \label{sec:neuralDecoding}
@@
 
-\subsubsection{Comparing multiple data-analysis methods}
+\subsection{Comparing multiple data-analysis methods}
 \label{sec:comparisonOfMultipleMethods}
@@
 
-\subsubsection{Interfacing Bonsai with Python, R and Matlab data analysis software}
+\subsection{Interfacing Bonsai with Python, R and Matlab data analysis software}
 \label{sec:bonsaiPythonMatlabCommunication}
@@
 
-\subsubsection{Reproducibility with Bonsai}
+\subsection{Reproducibility with Bonsai}
 \label{sec:reproducibility}
@@
 
-\subsubsection{Final remarks}
+\subsection{Final remarks}
 \label{sec:finalRemarks}
@@
 
-\subsection{Community Demand}
-\ii Evidence of community demand should be primarily driven from UK 
-academic researchers working largely within BBSRC remit – see our 
-Forward look for UK Bioscience for more detail on research areas covered by 
-BBSRC. Demand from other users (such as academic communities outside 
-of BBSRC remit or industrial users) may be appropriate to provide additional 
-support, especially in highlighting the potential for economic, commercial or 
-societal impact, but should not be the focus of the demand demonstration. 
-Evidence provided should highlight examples of the high-quality science that 
-the resource will underpin or has underpinned. Where possible and relevant, 
-examples should be drawn from a wide research community to illustrate the 
-broad impact of the resource to support high-quality internationally excellent 
-science.
+\section{Community Demand}
 
-\ii The level of community demand should be benchmarked against other 
-relevant resources and/or the size of the community. This will allow the fair 
-assessment of resources with different user bases. The types of evidence 
-that may be appropriate to provide will be different for new and existing 
-resources.
+\section{User engagement}
 
-\ii Evidence of wider consultation of the prospective community (e.g community 
-surveys) is encouraged.
+\section{Resource management}
 
-\iibf New Resources
+\section{Long term sustainability planning}
 
-\ii New resources should estimate the number of researchers who may engage and benefit from the resource. Evidence, where possible, would be of benefit and may include: 
-\begin{itemize}
-    \item  Datasets (or samples) in public or private repositories 
-    \item Citations or acknowledgments 
-    \item Gap analysis with existing resources 
-    \item Pilot project uptake or feedback from potential users.  
-\end{itemize}
+\section{Potential for economic and societal impact}
 
-\ii In particular, proposals for new resources should have consulted their prospective community prior to application
-
-\iibf Existing Resources
-
-\ii For existing resources this should include usage data of the
-current resource. Data types may include:
-
-\iitem Access requests from independent users/ sites
-
-\iitem Citations or acknowledgements
-
-\iitem Other public resources providing links to the resource
-
-\iitem New major acquisitions captured by the resource
-
-\ii{Additionally, existing resources need 
-to evidence why this resource needs 
-to be maintained/updated by the 
-current grant. This could include: }
-
-\iitem Survey data from users on what upgrades are needed
-
-\iitem Evidence of an expanding 
-user base, which requires 
-additional resource
-
-\iitem Recent developments in the 
-field, which require upgrades 
-to be integrated into the 
-resource.
-
-\subsection{User engagement}
-\ii Discussion of user engagement provision should aim to answer the following 
-questions:
-
-\iitem Is there awareness of the resource within the user community?
-
-\iitem How do you plan to develop the engagement strategy within the 
-proposal timeframe to expand user engagement?
-
-\iitem How have access mechanisms to ensure usability of the resource 
-been considered?
-
-\iitem How have user needs been incorporated into this proposal to ensure 
-it is fit for purpose and will deliver on expectations?
-
-\iibf New Resources
-
-\ii Evidence should be provided as to how the resource plans to engage
-stakeholders and ensure that the resource meets their needs and is
-used by the community targeted by the resource.
-
-\iibf Existing Resources
-
-\ii Evidence should be provided as to whether the resource has
-achieved the level of engagement it originally anticipated, and
-consideration is given how the additional investment would change
-this.
-
-\subsection{Long term sustainability planning}
-\ii{
-In addition, the case for support should outline considerations for the long-
-term sustainability of the resource beyond UKRI-BBSRC funding, as well as 
-the true cost of running and maintaining the resource in question. 
-The proposal should include: }
-
-\iitem Cost recovery plans, where appropriate, or an explanation why
-not if not viable. Evidence of clear business planning with a focus on
-at least partial cost recovery is required, especially when applying
-as an existing resource.
-
-\iitem Details for alternative support plans, aside from UKRI-BBSRC 
-funding.
-
-\iitem The level of support the resource is projected to require for expected 
-maintenance and/or subsequent maturation/enhancement activities. 
-Clear arguments as to why UKRI-BBSRC should support the resource now 
-should be provided, if other cost recovery and support plans are deemed 
-unsuitable.
-
-\subsection{Potential for economic and societal impact}
-\ii{Outline how the outputs of the proposed resource will contribute to 
-knowledge and how this may have the potential for economic return or 
-societal benefits. Impact activities should be integrated into appropriate 
-sections of the case for support, not presented as an independent work 
-package. }
-
-\iitem All proposals are expected to demonstrate clear plans with recorded 
-milestones and timelines for associated activities to develop 
-economic, commercial and societal impacts.  
-August 2021 7
-
-\iitem Methods of engagement and measures of success should be outlined 
-including how these will be regularly reviewed throughout the project 
-in order to deliver the most impact.
-
-\iitem Any planned activities should be fully justified within the Justification 
-of Resources attachment
-
-\subsection{Track Record}
-\ii{
-The majority of the track record relevant to the project should be located 
-within the Narrative CV and should not be repeated within the case for 
-support. You may, however, want to describe: }
-
-\iitem Track record of the team working together
-
-\iitem Specific expertise available at the host organisation and any 
-proposed partner organisations to enable the successful delivery of 
-the project.
-
-\iitem The specific role of each applicant and collaborator in the project
+\section{Track Record}
 

### 2021-11-10 caseForSupport.tex
Section 1 Background to the resource completed

@@
 
-\usepackage[colorlinks=true, allcolors=blue]{hyperref}
+\usepackage[colorlinks=true]{hyperref}
 \usepackage{natbib}
@@
 
-\section{Background to the Resource}
+\section{Background to the resource}
 
@@
 
-A central goal of the software development proposed here is to address this
+A central goal of the software resource proposed here is to address this
 problem, by providing Bonsai~(an experimental control software ecosystem;
-Figure~\ref{fig:bonsai} and Section~\ref{sec:bonsai}) with state-of-the-art online (and offline) data
-analysis methods (Figure~\ref{fig:proposedBonsaiExtensions} and 
-Section~\ref{sec:toBeAddedMachineIntelligenceFunctionality}).
-%
-It is unfortunate that neuroscientists currently do not have software that
-allows them to easily compare the performance of different data analysis
-methods for the characterization of their datasets. The second central goal
-of this proposal is to address this limitation by adding to Bonsai a generic
-mechanism for comparing the performance of different methods for the analysis
-of specific datasets (Section~\ref{sec:comparisonOfMultipleMethods}).
+Figure~\ref{fig:bonsai} and Section~\ref{sec:bonsai}) with state-of-the-art
+online (and offline) data analysis methods
+(Figure~\ref{fig:proposedBonsaiExtensions} and
+Section~\ref{sec:toBeAddedMachineIntelligenceFunctionality}). Bonsai is
+\textbf{unique} among experimental control software in that it allows people
+with no programming experience to control sophisticated experiments. With the
+proposed functionality to be added to Bonsai, its users will be able to use
+sophisticated machine-learning methods to control their experiments and analyze
+their data. This popularization of machine learning functionality will allow
+scientists without computational skills to control sophisticated experiments
+and analyze their data in unprecedented ways, which will most probably yield
+important \textbf{advances in science}.
 
@@
 data analysis.
+
+Using the proposed Bonsai capabilities to communicate with software written in
+other languages, we will integrate into Bonsai an initial set of state-of-the
+art data analysis methods
+(Section~\ref{sec:toBeAddedMachineIntelligenceFunctionality}). Developers of
+advanced data-analysis methods should become interested in integrating their
+algorithm into de Bonsai ecosystem because, which will allow their methods to
+reach the wide Bonsai community. In addition this integration will enable
+method developers, to easily compare the performance of of their methods with
+that of other methods integrated into the Bonsai ecosystem. Method developers
+witll become a new type of Bonsai users. These new users, will contribute to
+the \textbf{sustainability} of the proposed resource beyond the period of
+BBSRC-BBR funding.
 
@@
 
-\section{Details of Resource}
+\section{Details of resource}
 
-\subsection{Video-based behavioral analysis}
+\subsection{Machine intelligence functionality to be added to Bonsai}
+\label{sec:toBeAddedMachineIntelligenceFunctionality}
+
+\subsubsection{Video-based behavioral analysis}
 \label{sec:videoBasedBehavioralAnalysis}
@@
 
-\subsection{Spike sorting}
+\subsubsection{Spike sorting}
 \label{sec:spikeSorting}
@@
 
-\subsection{Low-dimensional representations of spiking data}
+\subsubsection{Low-dimensional representations of spiking data}
 \label{sec:lowDimensionalRepresentationsOfSpikingData}
@@
 
-\subsection{Low-dimensional representations of local-field potentials}
+\subsubsection{Low-dimensional representations of local-field potentials}
 \label{sec:lowDimensionalRepresentationsOfLFPs}
@@
 
-\subsection{Neural decoding}
+\subsubsection{Neural decoding}
 \label{sec:neuralDecoding}
@@
 
-\subsection{Comparing multiple data-analysis methods}
+\subsubsection{Comparing multiple data-analysis methods}
 \label{sec:comparisonOfMultipleMethods}
@@
 software to the large community of Bonsai users, and will be able to easily
-compare the performance of their methods with that of preexisting methods in the
-Bonsai ecosystem. In this way method developers will become a new type of
+compare the performance of their methods with that of preexisting methods in
+the Bonsai ecosystem. In this way method developers will become a new type of
 Bonsai users, which will guarantee the sustainability of developments
@@
 
-\section{Community Demand}
+\section{Community demand}
 

### 2021-11-10 caseForSupport.tex
Sec 1 looking better and still working on Sec 2

@@
 \usepackage{graphicx}
+\usepackage{verbatim}
 
@@
 \textbf{unique} among experimental control software in that it allows people
-with no programming experience to control sophisticated experiments. With the
-proposed functionality to be added to Bonsai, its users will be able to use
-sophisticated machine-learning methods to control their experiments and analyze
-their data. This popularization of machine learning functionality will allow
-scientists without computational skills to control sophisticated experiments
-and analyze their data in unprecedented ways, which will most probably yield
-important \textbf{advances in science}.
+with no programming experience to control sophisticated experiments. Its large
+and rapidly growing user base demonstrates the need of this type of software
+resources. With the proposed functionality to be added to Bonsai, its users
+will be able to use sophisticated machine-learning methods to control their
+experiments and analyze their data. This popularization of machine learning
+functionality will allow scientists without computational skills to control
+sophisticated experiments and analyze their data in unprecedented ways, which
+will most probably yield important \textbf{advances in science}.
 
@@
 
-Using the proposed Bonsai capabilities to communicate with software written in
+Developers of advanced data-analysis methods should become interested in
+integrating them into de Bonsai ecosystem, which will allow their methods to
+reach the wide Bonsai community. In addition this integration will enable
+method developers, to easily compare the performance of their methods with
+that of many methods previously integrated into the Bonsai ecosystem. Method
+developers will thus become a new type of Bonsai users, which will contribute
+to the \textbf{sustainability} of the proposed resource beyond the period of
+BBSRC-BBR funding.
+
+Using the proposed capabilities to communicate Bonsai with software written in
 other languages, we will integrate into Bonsai an initial set of state-of-the
-art data analysis methods
-(Section~\ref{sec:toBeAddedMachineIntelligenceFunctionality}). Developers of
-advanced data-analysis methods should become interested in integrating their
-algorithm into de Bonsai ecosystem because, which will allow their methods to
-reach the wide Bonsai community. In addition this integration will enable
-method developers, to easily compare the performance of of their methods with
-that of other methods integrated into the Bonsai ecosystem. Method developers
-witll become a new type of Bonsai users. These new users, will contribute to
-the \textbf{sustainability} of the proposed resource beyond the period of
-BBSRC-BBR funding.
+art behavioral and neural data analysis methods
+(Section~\ref{sec:toBeAddedMachineIntelligenceFunctionality}).
 
@@
 (Section~\ref{sec:reproducibility}).
+
+\textcolor{red}{Missing reference to alternatives to Bonsai}.
+
+\textcolor{red}{\textit{Overview of past and current resource(s) in the subject
+area in both the UK and abroad, including any alternative community resources
+currently available. You should indicate the community size of the intended
+resource and how this relates to the field in which it operates.}}
 
@@
 \section{Details of resource}
+
+\subsection{Software infrastructure to incorporate machine intelligence
+functionality to Bonsai}
+\label{sec:softwareInfrastructure}
+
+\begin{comment}
+Bonsai does not currently provide machine learning functionality. For instance,
+the close-loop experimental control depicted in
+Figure~\ref{fig:bci} cannot be currently implemented in
+Bonsai. It represents the control of an experiment where a subject guides the
+mouse position in a computer screen with her brain activity (Section~\ref{}).
+Nodes highlighted in color require machine intelligence functionality currently
+not available in Bonsai.
+%
+\end{comment}
+
+A central aim of this proposal is to create software infrastructure to provide
+this machine intelligence functionality to Bonsai in extensible and general
+ways.
+
+\subsubsection{Generality}
+
+The number of current machine learning algorithms is very large and it is
+continuously increasing. We intend to build a small number of machine
+intelligence operators in Bonsai that can be configured to provide the
+functionality of most machine learning algorithms.
+
+For this, we will build on abstractions used in the \texttt{sickit-learn}
+library. The success of this library is, to a large extent, due to its simple
+and general application programming interface \citep[API;][]{buitinckEtAl13}.
+%
+This API consists of three complementary interfaces: an estimator interface for
+building and fitting models, a predictor interface for making predictions and a
+transformer interface for converting data.
+%
+Using only these three interfaces \texttt{scikit-learn} provides
+implementations of a large number of machine learning algorithms for
+dimensionality reduction, clustering, classification and regression.
+
+Inspired by \texttt{scikit-learn}, Bonsai will have three types of operators:
+estimators, predictors and transformers. Each of these operators will be
+instantiated by objects implementing the corresponding interface.
+%
+For example, consider the close-loop system for the control of a computer
+cursor with neural activity represented in Figure~\ref{fig:bci}a. The Bonsai
+graphical program for this system is given in Figure~\ref{fig:bci}b. The
+objects instantiating the SpikeThresholdDetector, the KiloSortSpikeSorter and
+GPFAlatentsExtractor will implement the transformer interface, and the object
+instantiating the TimeSeriesForestClassifier will implement the predictor
+interface (Figure~\ref{fig:bci}b, top).
+%
+The parameters for the SpikeThresholdDetector and the KiloSortSpikeSorter will
+be set manually by the user, as is commonly done in practice. The parameters of
+the GPFAlatentsExtractor will be learnt offline in an unsupervised way in a
+second pipeline (Figure~\ref{fig:bci}b, middle). Finally, the parameters of the
+TimeSeriesForestClassifier will also be learnt offline, but in a supervised
+way, in a third pipeline (Figure~\ref{fig:bci}b, bottom).
+
+\begin{figure}
+   \begin{center}
+       \includegraphics[width=4in]{figures/bci.png}
+       \caption{Close-loop system for controlling a computer cursor with brain activity}
+     \label{fig:bci}
+   \end{center}
+\end{figure}
+
+\begin{comment}
+    The spike detector transforms membrane
+potentials into spike sequences. The spike sorter transforms spikes squences
+into lists of spikes, one list per neuron. The latents extractor transforms
+multi-neuron lists of spikes into latent trajectories summarizing the
+multi-neuon spiking activity.  The neural decoder predicts the next cursor
+position from the latent trajectories.
+\end{comment}
+
+\begin{comment}
+middl, whose output will set
+As in this library,
+there will be three type of machine intelligence operators in Bonsai:
+%
+estimator operators for building and fitting models, predictor operators for
+making predictions and transformer operators for converting data. These
+operators will be implemented by objects exposing the corresponding interfaces.
+
+The estimator interface exposes a fit method for learning a model from training
+data.  This method is called with training data (e.g., supplied as two arrays
+X\_train and y\_train in supervised learning estimators). Its task is to run a
+learning algorithm and to determine model-specific parameters from the training
+data.  All supervised and unsupervised learning algorithms (e.g., for
+classification, regression or clustering) will be offered as objects
+implementing this interface. Machine learning tasks like feature extraction,
+feature selection or dimensionality reduction will also be provided as
+estimators.
+
+The predictor interface extends the notion of an estimator by adding a predict
+method that takes an array X\_test and produces predictions for X\_test, based
+on the learned parameters of the estimator (we call the input to predict
+“X\_test” in order to emphasize that predict generalizes to new data). In the
+case of supervised learning estimators, this method typically returns the
+predicted labels or values computed by the model.
+
+Since it is common to modify or filter data before feeding it to a learning
+algorithm, the transformer interface defines a transform method. It takes as
+input some new data X\_test and yields as output a transformed version of
+X\_test. Preprocessing, feature selection, feature extraction and
+dimensionality reduction algorithms will be provided as transformers in Bonsai.
+
+Bonsai will provide the ability to compose new
+estimators from several base estimators. Composition mechanisms can be used
+to combine typical machine learning workflows into a single object which is
+itself an estimator, and can be employed wherever usual estimators can be used.
+Composition of estimators will be done with Pipeline objects that chain multiple estimators into a single one. This is useful
+since a machine learning workflow typically involves a fixed sequence of processing steps (e.g., feature extraction, dimensionality reduction, learning and making
+predictions), many of which perform some kind of learning. A sequence of N such
+steps can be combined into a Pipeline if the first N-1 steps are transformers;
+the last can be either a predictor, a transformer or both.
+
+\end{comment}
+
+\subsubsection{Extensibility}
+
+\textcolor{red}{I will start this section by saying that we are not proposing
+to implement machine learning algorithms, as in \texttt{scipy.learn}, but to
+integrate into Bonsai existing machine intelligence algorithms.}
+
+Currently Bonsai has its own package manager, and its community of users have
+already extended Bonsai's functionality with several contributed packages
+(e.g., \ldots). However, these packages need to be written in C\#, while most
+current data analysis methods are written in Python, R or Matlab. We propose to
+add to Bonsai capabilities to communicate with software written in these
+languages.
+
+Fortunately, there exist software that allow C\# program
+to call and be called by programs in these other languages. For the
+communication of C\# programs with Python we will use
+Python.NET\footnote{\href{https://github.com/pythonnet/pythonnet}{https://github.com/pythonnet/pythonnet}},
+with R programs we will use
+R.Net\footnote{\href{https://github.com/rdotnet/rdotnet}{https://github.com/rdotnet/rdotnet}}
+and with Matlab we will use the built in Matlab .NET
+interface\footnote{\href{https://uk.mathworks.com/help/matlab/matlab\_external/using-net-from-matlab-an-overview.html}{https://uk.mathworks.com/help/matlab/matlab\_external/using-net-from-matlab-an-overview.html}}.
+In cases where these software cannot support some type of communication between
+C\# and another language we will use the messaging library
+ZeroMQ\footnote{\href{https://zeromq.org/}{https://zeromq.org/}}.
+
+Using the previous software, we will build C\# wrappers around machine
+intellgence programs written in other languages, so that these programs
+implement the Estimator, Predictor or Transformer interfaces and can be
+integrated into Bonsai as machine intelligence operators.
 
@@
 
+\begin{comment}
+
 \subsection{Interfacing Bonsai with Python, R and Matlab data analysis software}
@@
 ZeroMQ\footnote{\href{https://zeromq.org/}{https://zeromq.org/}}.
+
+\end{comment}
 

### 2021-11-11 caseForSupport.tex
caseForSupport.tex in progress

@@
 \usepackage{verbatim}
+
+\makeatletter
+\renewcommand\paragraph{\@startsection{paragraph}{4}{\z@}%
+            {-2.5ex\@plus -1ex \@minus -.25ex}%
+            {1.25ex \@plus .25ex}%
+            {\normalfont\normalsize\bfseries}}
+\makeatother
+\setcounter{secnumdepth}{4} % how many sectioning levels to assign numbers to
+\setcounter{tocdepth}{4}    % how many sectioning levels to show in ToC
 
@@
 (Figure~\ref{fig:proposedBonsaiExtensions} and
-Section~\ref{sec:toBeAddedMachineIntelligenceFunctionality}). Bonsai is
+Section~\ref{sec:functionality}). Bonsai is
 \textbf{unique} among experimental control software in that it allows people
@@
 will be able to use sophisticated machine-learning methods to control their
-experiments and analyze their data. This popularization of machine learning
+experiments and analyze their data. This
+\textbf{democratisation} of machine learning
 functionality will allow scientists without computational skills to control
@@
 art behavioral and neural data analysis methods
-(Section~\ref{sec:toBeAddedMachineIntelligenceFunctionality}).
+(Section~\ref{sec:functionality}).
 
@@
 
-\subsection{Software infrastructure to incorporate machine intelligence
-functionality to Bonsai}
+We intend to improve an \textbf{existing resource}, the Bonsai ecosystem. We
+propose to create software infrastructure to facilitate the integration into it
+of existing advanced data analysis methods
+(Section~\ref{sec:softwareInfrastructure}).
+%
+To demonstate the use of this infrastructure, and to provide Bonsai users with
+an initial set of advanced machine intelligence neural data analysis methods,
+we will add to the Bonsai ecosystem machine learning functionality to control
+neuroscience experiments and to analyze data generated by them
+(Section~\ref{sec:functionality}).
+
+This initial set of methods should demonstrate to Bonsai users, and to
+machine-learning methods developers, the potential of embedding
+machine-learning functionality into the experimental control loop. We propose
+user engagement activities to expose both of these groups of people to this
+potential (Section~\ref{sec:userEngagement}).  When data-analysis method
+developers recognize this potentials, they will become interested in
+contributing their methods to the Bonsai ecosystem, ensuring the
+\textbf{long-term sustainability} of the proposed resource.
+
+\subsection{Software infrastructure}
 \label{sec:softwareInfrastructure}
@@
 
-A central aim of this proposal is to create software infrastructure to provide
-this machine intelligence functionality to Bonsai in extensible and general
-ways.
+We propose to create software infrastructure to add machine intelligence
+functionality to the Bonsai ecosystem in general
+(Section~\ref{sec:generality}) and extensible ways
+(Section~\ref{sec:extensibility}).
 
 \subsubsection{Generality}
+\label{sec:generality}
 
@@
 \subsubsection{Extensibility}
+\label{sec:extensibility}
 
@@
 \subsection{Machine intelligence functionality to be added to Bonsai}
-\label{sec:toBeAddedMachineIntelligenceFunctionality}
+\label{sec:functionality}
+
+We propose to initially add to Bonsai machine intelligence functionality to
+support neuroscience experiments because, currently, neuroscience is the
+research area with the largest number of Bonsai users. Within neuroscience, we
+propose to add functionality for video-based behavioral analysis
+(Section~\ref{sec:videoBasedBehavioralAnalysis}) and for running
+brain-computer-interface experiments (Section~\ref{sec:bci}), since these tasks
+are relevant to many Bonsai users, and recent advances in machine learning
+will greatly improve their implementation in Bonsai.
 
@@
 
-\subsubsection{Spike sorting}
-\label{sec:spikeSorting}
+\subsubsection{Brain-computer-interface applications}
+\label{sec:bci}
+
+\paragraph{Spike sorting}
+% \label{sec:spikeSorting}
 
@@
 
-\subsubsection{Low-dimensional representations of spiking data}
-\label{sec:lowDimensionalRepresentationsOfSpikingData}
+\paragraph{Low-dimensional representations of spiking data}
+% \label{sec:lowDimensionalRepresentationsOfSpikingData}
 
@@
 
-\subsubsection{Low-dimensional representations of local-field potentials}
-\label{sec:lowDimensionalRepresentationsOfLFPs}
+\paragraph{Low-dimensional representations of local-field potentials}
+% \label{sec:lowDimensionalRepresentationsOfLFPs}
 
@@
 
-\subsubsection{Neural decoding}
-\label{sec:neuralDecoding}
+\paragraph{Neural decoding}
+% \label{sec:neuralDecoding}
 
@@
 
-\subsubsection{Comparing multiple data-analysis methods}
-\label{sec:comparisonOfMultipleMethods}
+\paragraph{Comparing multiple data-analysis methods}
+% \label{sec:comparisonOfMultipleMethods}
 
@@
 We will demonstrate each machine intelligence functionality added to Bonsai
-(Section~\ref{sec:toBeAddedMachineIntelligenceFunctionality}) and the multiple
+(Section~\ref{sec:functionality}) and the multiple
 data-analysis methods comparisons
@@
 demonstrations.
+
+\subsection{Measurable targets}
+\label{sec:measurableTargets}
+
+To assess the progress of the proposed project, we propose the following
+measurable targets:
+
+\begin{description}
+
+    \item[implementation of software infrastructure]
+        (Section~\ref{sec:softwareInfrastructure}) includes the
+        implementation of the machine abstractions described in Section~\ref{}
+        and of the capabilities to be added to Bonsai to communicate with
+        software written in other languages described in Section~\ref{}. We
+        will consider this target completed when we can implement a simple BCI
+        application, with the architecture depicted in Figure~\ref{}, with
+        machine learning functionality provided by software in Python, Matlab
+        and R.
+
+    \item[implementation of functionality] (Sections~\ref{sec:functionality})
+        includes the integration into Bonsai of machine learning software
+        for video-based behavioral analysis (Section~\ref{}) and for
+        brain-computer-interface applications (Section~\ref{}). Completion of
+        this target requires that Bonsai can perform these tasks interacting
+        with software written in other languages.
+
+    \item[internal testing] multiple groups at the SWC use Bonsai to control
+        their experiments.
+
+        Bonsai Github repository and the success use by Bonsai users
+
+    \item Bonsai users use the implemented machine learning functionality. We
+        propose to evaluate this target by checking in the Bonsai user forum
+        the frequency with which a novel machine learning functionality is
+        discussed, and by tracking publications using this functionality.
+
+    \item method developers integrate their methods into the Bonsai ecosystem.
+        The integration of machine learning functionality into Bonsai will use
+        Bonsai's package manager. We propose to use this package manaer to tack
+        the number of machine intelligence methods integrated into Bonsai.
+
+\end{description}
 
@@
 \section{Community demand}
+\label{sec:communityDemand}
 
 \section{User engagement}
+\label{sec:userEngagement}
 

### 2021-11-12 caseForSupport.tex
Updates from Overleaf

--- 2021-11-10 caseForSupport.tex
@@
 with no programming experience to control sophisticated experiments. Its large
-and rapidly growing user base demonstrates the need of this type of software
+and rapidly growing user base demonstrates the need for this type of software
 resources. With the proposed functionality to be added to Bonsai, its users
@@
 be written in C\#, while most current neuroscience data analysis methods are
-written in Python, R or Matlab. We propose to add to Bonsai capabilities to
+written using programming languages such as Python, R or Matlab. We propose to add to Bonsai capabilities to
 communicate with software written in these languages
-(Section~\ref{sec:bonsaiPythonMatlabCommunication}).  These capabilities will
-allow a large number of neuroscience data analysis methods to be easily
+(Section~\ref{sec:bonsaiPythonMatlabCommunication}), as this would increase our reach to hundreds more experiments. Moreover, these novel capabilities will
+increase the number of neuroscience data analysis methods to be easily
 integrated into the Bonsai ecosystem and will provide Bonsai users a large
-repertoire of advance machine learning methods for their behavioral and neural
+repertoire of advanced machine learning methods for their behavioural and neural
 data analysis.
 
-Developers of advanced data-analysis methods should become interested in
-integrating them into de Bonsai ecosystem, which will allow their methods to
-reach the wide Bonsai community. In addition this integration will enable
-method developers, to easily compare the performance of their methods with
-that of many methods previously integrated into the Bonsai ecosystem. Method
-developers will thus become a new type of Bonsai users, which will contribute
-to the \textbf{sustainability} of the proposed resource beyond the period of
-BBSRC-BBR funding.
+Extending the functionalities of Bonsai to other programming languages is of interest to many developers of advanced data-analysis methods, who will likely integrate their methods into the existing Bonsai ecosystem, allowing them to reach the wide Bonsai community.
+In addition this integration will enable
+method developers to easily compare the performance of their methods with
+that of others previously integrated into the Bonsai ecosystem. Data-analysis
+developers will thus become a new type of Bonsai users. We anticipate this to contribute to the long-term \textbf{sustainability} of the proposed resource beyond the period of BBSRC-BBR funding.
 
@@
 other languages, we will integrate into Bonsai an initial set of state-of-the
-art behavioral and neural data analysis methods
+art behavioural and neural data analysis methods
 (Section~\ref{sec:toBeAddedMachineIntelligenceFunctionality}).
@@
 Bonsai is an excellent tool for reproducible data acquisition and
-experimental control. An experiment implemented in Bonsai can be replicated in
+experimental control. For example, an experiment implemented in Bonsai can be replicated in
 a new laboratory by just sharing a Bonsai configuration file. With the addition
@@
 also allow researchers to stay in control and change their experimental
-paradigms.  Many open-source software tools are either inaccessible to
+paradigms online as results are obtained.  Many open-source software tools are either inaccessible to
 non-programmers, or too constrained to be of general use outside their narrow
-domain of application. Bonsai has been successful because it has straddled this
-gap to some extent.
+domain of application. Bonsai has been successful because it identified a gap in the field and was capable of providing a meaningful solution to an existing problem.
 
@@
 International Brain
-Laboratory\footnote{\href{https://www.internationalbrainlab.com/}{https://www.internationalbrainlab.com/}}
+Laboratory\footnote{\href{https://www.internationalbrainlab.com/}{https://www.internationalbrainlab.com/}}, a global collaboration comprising of 22 research groups, 
 and the Allen Institute for Brain

### 2021-11-12 caseForSupport.tex
Still in progress

--- 2021-11-11 caseForSupport.tex
@@
 Experimental control and neural data analysis are currently highly
-disassociated.  Most neuroscientists collect their data and only afterward
-analyze it. This split has had dramatic consequences for the progress of
+disassociated.  Most neuroscientists collect their data and only after ward
+analyse it. This split has had dramatic consequences for the progress of
 neuroscience. First, most experimental control methods currently do not use
@@
 experiments that are currently possible, and generate new demands for neural
-data analysis methods that will generate advances in the field.
+data analysis methods that will trigger advanced in machine learning.
 
@@
 will be able to use sophisticated machine-learning methods to control their
-experiments and analyze their data. This
+experiments and analyse their data. This
 \textbf{democratisation} of machine learning
 functionality will allow scientists without computational skills to control
-sophisticated experiments and analyze their data in unprecedented ways, which
+sophisticated experiments and analyse their data in unprecedented ways, which
 will most probably yield important \textbf{advances in science}.
@@
 integrated into the Bonsai ecosystem and will provide Bonsai users a large
-repertoire of advance machine learning methods for their behavioral and neural
+repertoire of advance machine learning methods for their behavioural and neural
 data analysis.
@@
 Developers of advanced data-analysis methods should become interested in
-integrating them into de Bonsai ecosystem, which will allow their methods to
+integrating them into the Bonsai ecosystem, which will allow their methods to
 reach the wide Bonsai community. In addition this integration will enable
@@
 other languages, we will integrate into Bonsai an initial set of state-of-the
-art behavioral and neural data analysis methods
+art behavioural and neural data analysis methods
 (Section~\ref{sec:functionality}).
@@
 Bonsai~\citep{lopesEtAl15,lopesAndMonteiro21} is a free and open-source visual
-programming language that emphasizes performance, flexibility, and ease-of-use,
+programming language that emphasises performance, flexibility, and ease-of-use,
 allowing scientists with no previous programming experience to quickly develop
@@
 Bonsai has a large user base in the systems neuroscience community.
-Quantifications of this user base are provided in Figures 1d-1f.  In the last
+Quantification of this user base are provided in Figures 1d-1f.  In the last
 year alone, we estimate more than 1,000 new users have started to incorporate
@@
 %
-To demonstate the use of this infrastructure, and to provide Bonsai users with
+To demonstrate the use of this infrastructure, and to provide Bonsai users with
 an initial set of advanced machine intelligence neural data analysis methods,
 we will add to the Bonsai ecosystem machine learning functionality to control
-neuroscience experiments and to analyze data generated by them
+neuroscience experiments and to analyse data generated by them
 (Section~\ref{sec:functionality}).
@@
 potential (Section~\ref{sec:userEngagement}).  When data-analysis method
-developers recognize this potentials, they will become interested in
+developers recognise this potentials, they will become interested in
 contributing their methods to the Bonsai ecosystem, ensuring the
 \textbf{long-term sustainability} of the proposed resource.
+
+Adding machine learning functionality to the experimental control loop will
+make possible the implementation of experiments with unprecedented levels of
+control. Moreover, equipping experimental scientists without programming
+experience with machine learning tools will considerably increase the
+application of machine learning to the biosciences, which most probably will
+translate into groundbreaking scientific discoveries.
 
@@
 mouse position in a computer screen with her brain activity (Section~\ref{}).
-Nodes highlighted in color require machine intelligence functionality currently
+Nodes highlighted in colour require machine intelligence functionality currently
 not available in Bonsai.
@@
     The spike detector transforms membrane
-potentials into spike sequences. The spike sorter transforms spikes squences
+potentials into spike sequences. The spike sorter transforms spikes sequences
 into lists of spikes, one list per neuron. The latents extractor transforms
-multi-neuron lists of spikes into latent trajectories summarizing the
-multi-neuon spiking activity.  The neural decoder predicts the next cursor
+multi-neuron lists of spikes into latent trajectories summarising the
+multi-neuron spiking activity.  The neural decoder predicts the next cursor
 position from the latent trajectories.
@@
 \begin{comment}
-middl, whose output will set
+middle, whose output will set
 As in this library,
@@
 on the learned parameters of the estimator (we call the input to predict
-“X\_test” in order to emphasize that predict generalizes to new data). In the
+“X\_test” in order to emphasise that predict generalises to new data). In the
 case of supervised learning estimators, this method typically returns the
@@
 Using the previous software, we will build C\# wrappers around machine
-intellgence programs written in other languages, so that these programs
+intelligence programs written in other languages, so that these programs
 implement the Estimator, Predictor or Transformer interfaces and can be
@@
 research area with the largest number of Bonsai users. Within neuroscience, we
-propose to add functionality for video-based behavioral analysis
+propose to add functionality for video-based behavioural analysis
 (Section~\ref{sec:videoBasedBehavioralAnalysis}) and for running
@@
 
-Precisely quantifying animal behavior is an essential step toward understanding
+Precisely quantifying animal behaviour is an essential step toward understanding
 brain function.  Deeplabcut is a Python-based software for tracking animal body
@@
     \item[Motion sequencing]\citep[MoSeq;][]{wiltschkoEtAl15}: extracts patterns
-        of behavior that repeat over time (i.e., syllables of behavior) from
+        of behaviour that repeat over time (i.e., syllables of behaviour) from
         video data. For instance, it parses (in an unsupervised way) the
-        behavior of a mouse in an arena into segments of time where a mouse
+        behaviour of a mouse in an arena into segments of time where a mouse
         was running, rearing and grooming. It uses a hidden Markov model and it
@@
 
-    \item[Decoding behavior from neural
+    \item[Decoding behaviour from neural
         activity]\citep[BehaveNet;][]{battyEtAl19} combines hidden Markov
@@
         System]\citep[GLDS;][]{andersonAndMoore12}: with sufficiently large
-        bin sizes, spike counts can be modeled as Gaussian random processes.
+        bin sizes, spike counts can be modelled as Gaussian random processes.
         This Gaussianity assumption greatly simplifies the estimation of
@@
     \item[Poisson Linear Dynamical System]\citep[PLDS;][]{mackeEtAl15}: for
-        smaller bin sizes, spike counts are better modeled as Poisson random
+        smaller bin sizes, spike counts are better modelled as Poisson random
         processes, rather than Gaussian ones. The algorithm described in
@@
 
-Spikes are extracted from a higher-frequency range of extracellulary recorded
+Spikes are extracted from a higher-frequency range of extracellularly recorded
 voltages. Local field potentials (LFPs) are computed from a lower-frequency
@@
 We propose to implement in Bonsai several decoding/classification algorithms:
-k-nearest neighbor, linear discriminative analysis, support vector machines,
+k-nearest neighbour, linear discriminative analysis, support vector machines,
 random forests, artificial neural networks, naive Bayes and Gaussian process
@@
 %
-Briefly, in this task there are three nose ports (left, center, right). Rats
-initiate a trial by inserting their nose in the center port for the duration of
+Briefly, in this task there are three nose ports (left, centre, right). Rats
+initiate a trial by inserting their nose in the centre port for the duration of
 the fixation period, until they hear a Go stimulus. During the fixation period
@@
 
-To assess the progress of the proposed project, we propose the following
+We propose to assess the progress of the proposed project using the following
 measurable targets:
+
+\begin{comment}
 
@@
         (Section~\ref{sec:softwareInfrastructure}) includes the
-        implementation of the machine abstractions described in Section~\ref{}
-        and of the capabilities to be added to Bonsai to communicate with
-        software written in other languages described in Section~\ref{}. We
+        implementation of the machine abstractions, described in
+        Section~\ref{},
+        and of the capabilities to allow Bonsai communicate with
+        software written in other languages, described in Section~\ref{}. We
         will consider this target completed when we can implement a simple BCI
@@
     \item[implementation of functionality] (Sections~\ref{sec:functionality})
-        includes the integration into Bonsai of machine learning software
-        for video-based behavioral analysis (Section~\ref{}) and for
+        includes the integration into Bonsai of machine learning software for
+        video-based behavioral analysis (Section~\ref{}) and for
         brain-computer-interface applications (Section~\ref{}). Completion of
-        this target requires that Bonsai can perform these tasks interacting
-        with software written in other languages.
+        this target requires that Bonsai can perform each of these tasks
+        separately, in conjunction with software written in Python, R and
+        Matlab.
 
@@
         The integration of machine learning functionality into Bonsai will use
-        Bonsai's package manager. We propose to use this package manaer to tack
+        Bonsai's package manager. We propose to use this package manager to tack
         the number of machine intelligence methods integrated into Bonsai.
@@
 \end{description}
+
+\end{comment}
 

### 2021-11-12 caseForSupport.tex

@@
 Experimental control and neural data analysis are currently highly
-disassociated.  Most neuroscientists collect their data and only after ward
+disassociated.  Most neuroscientists collect their data and only afterward
 analyse it. This split has had dramatic consequences for the progress of
@@
 integrated into the Bonsai ecosystem and will provide Bonsai users a large
-repertoire of advance machine learning methods for their behavioural and neural
+repertoire of advanced machine learning methods for their behavioural and neural
 data analysis.
@@
 Bonsai~\citep{lopesEtAl15,lopesAndMonteiro21} is a free and open-source visual
-programming language that emphasises performance, flexibility, and ease-of-use,
+programming language that emphasizes performance, flexibility, and ease-of-use,
 allowing scientists with no previous programming experience to quickly develop
@@
 Bonsai has a large user base in the systems neuroscience community.
-Quantification of this user base are provided in Figures 1d-1f.  In the last
+Quantifications of this user base are provided in Figures 1d-1f.  In the last
 year alone, we estimate more than 1,000 new users have started to incorporate

### 2021-11-12 caseForSupport.tex
Splitted caseForSupport.tex into several files

--- 2021-11-12 caseForSupport.tex
@@
 
+\setcounter{page}{1}
+
 \section{Background to the resource}
 
-Experimental control and neural data analysis are currently highly
-disassociated.  Most neuroscientists collect their data and only afterward
-analyze it. This split has had dramatic consequences for the progress of
-neuroscience. First, most experimental control methods currently do not use
-advanced data analysis methods to enhance their functionality. Second, the vast
-majority of neural data analysis methods are offline and do not provide key
-functionality needed for online experimental control. Using advanced data
-analysis methods in experimental control will improve the sophistication of
-experiments that are currently possible, and generate new demands for neural
-data analysis methods that will generate advances in the field.
-
-A central goal of the software resource proposed here is to address this
-problem, by providing Bonsai~(an experimental control software ecosystem;
-Figure~\ref{fig:bonsai} and Section~\ref{sec:bonsai}) with state-of-the-art
-online (and offline) data analysis methods
-(Figure~\ref{fig:proposedBonsaiExtensions} and
-Section~\ref{sec:toBeAddedMachineIntelligenceFunctionality}). Bonsai is
-\textbf{unique} among experimental control software in that it allows people
-with no programming experience to control sophisticated experiments. Its large
-and rapidly growing user base demonstrates the need for this type of software
-resources. With the proposed functionality to be added to Bonsai, its users
-will be able to use sophisticated machine-learning methods to control their
-experiments and analyze their data. This popularization of machine learning
-functionality will allow scientists without computational skills to control
-sophisticated experiments and analyze their data in unprecedented ways, which
-will most probably yield important \textbf{advances in science}.
-
-\begin{figure}
-    \begin{center}
-        \includegraphics[width=7in]{figures/bonsai.pdf}
-        \label{fig:bonsai}
-    \end{center}
-\end{figure}
-
-\setcounter{figure}{1}
-
-\begin{figure}
-    \begin{center}
-        \includegraphics[width=4in]{figures/proposed_bonsai_extensions.png}
-        \caption{Proposed extensions to Bonsai}
-        \label{fig:proposedBonsaiExtensions}
-    \end{center}
-\end{figure}
-
-Currently Bonsai has its own package manager, and its community of users have
-already extended Bonsai's functionality with several contributed packages
-(e.g., \ldots \textcolor{red}{Goncalo please reference the best contributed
-packages developed by the Bonsai community}). However, these packages need to
-be written in C\#, while most current neuroscience data analysis methods are
-written using programming languages such as Python, R or Matlab. We propose to add to Bonsai capabilities to
-communicate with software written in these languages
-(Section~\ref{sec:bonsaiPythonMatlabCommunication}), as this would increase our reach to hundreds more experiments. Moreover, these novel capabilities will
-increase the number of neuroscience data analysis methods to be easily
-integrated into the Bonsai ecosystem and will provide Bonsai users a large
-repertoire of advanced machine learning methods for their behavioural and neural
-data analysis.
-
-Extending the functionalities of Bonsai to other programming languages is of interest to many developers of advanced data-analysis methods, who will likely integrate their methods into the existing Bonsai ecosystem, allowing them to reach the wide Bonsai community.
-In addition this integration will enable
-method developers to easily compare the performance of their methods with
-that of others previously integrated into the Bonsai ecosystem. Data-analysis
-developers will thus become a new type of Bonsai users. We anticipate this to contribute to the long-term \textbf{sustainability} of the proposed resource beyond the period of BBSRC-BBR funding.
-
-Using the proposed capabilities to communicate Bonsai with software written in
-other languages, we will integrate into Bonsai an initial set of state-of-the
-art behavioural and neural data analysis methods
-(Section~\ref{sec:toBeAddedMachineIntelligenceFunctionality}).
-
-Bonsai is an excellent tool for reproducible data acquisition and
-experimental control. For example, an experiment implemented in Bonsai can be replicated in
-a new laboratory by just sharing a Bonsai configuration file. With the addition
-of the proposed machine learning functionality Bonsai will extend this
-reproducibility to the domain of data analysis
-(Section~\ref{sec:reproducibility}).
-
-\textcolor{red}{Missing reference to alternatives to Bonsai}.
-
-\textcolor{red}{\textit{Overview of past and current resource(s) in the subject
-area in both the UK and abroad, including any alternative community resources
-currently available. You should indicate the community size of the intended
-resource and how this relates to the field in which it operates.}}
-
-\subsection{Bonsai}
-\label{sec:bonsai}
-
-Bonsai~\citep{lopesEtAl15,lopesAndMonteiro21} is a free and open-source visual
-programming language that emphasizes performance, flexibility, and ease-of-use,
-allowing scientists with no previous programming experience to quickly develop
-their own high-performance data acquisition and experimental control systems
-(Figure 1c). Bonsai combines a high-level event algebra for data streams with
-an integrated development environment (IDE) and a strong library of plugins
-supporting multiple hardware and software packages used by the biomedical
-research community (Figures 1a, 1b).
-
-Bonsai has a large user base in the systems neuroscience community.
-Quantifications of this user base are provided in Figures 1d-1f.  In the last
-year alone, we estimate more than 1,000 new users have started to incorporate
-Bonsai into their experimental protocols across the world. The surprising rate
-of adoption of Bonsai in non-programmer experimental labs highlights the need
-for accessible programming tools that enable state-of-the-art technology but
-also allow researchers to stay in control and change their experimental
-paradigms online as results are obtained.  Many open-source software tools are either inaccessible to
-non-programmers, or too constrained to be of general use outside their narrow
-domain of application. Bonsai has been successful because it identified a gap in the field and was capable of providing a meaningful solution to an existing problem.
-
-The language has also helped to potentiate the growing wave of foundational
-open hardware initiatives, such as the OpenEphys \citep{siegleEtAl17} and UCLA
-Miniscope~\citep{caiEtAl16}, making it possible to quickly combine and
-integrate these tools into new experiments~\citep{buccinoEtAl18}.
-%
-Bonsai has been adopted in large neuroscience undertakings like the
-International Brain
-Laboratory\footnote{\href{https://www.internationalbrainlab.com/}{https://www.internationalbrainlab.com/}}, a global collaboration comprising of 22 research groups, 
-and the Allen Institute for Brain
-Science\footnote{\href{https://alleninstitute.org/what-we-do/brain-science/}{https://alleninstitute.org/what-we-do/brain-science/}}.
-%
-More recently, Bonsai has started to expand outside the domain of neuroscience
-into biomedical research and biotechnology tool development, and even outside
-academia into public outreach and education programs.
+\input{backgroundToTheResource}
 
@@
 
-\subsection{Software infrastructure to incorporate machine intelligence
-functionality to Bonsai}
-\label{sec:softwareInfrastructure}
-
-\begin{comment}
-Bonsai does not currently provide machine learning functionality. For instance,
-the close-loop experimental control depicted in
-Figure~\ref{fig:bci} cannot be currently implemented in
-Bonsai. It represents the control of an experiment where a subject guides the
-mouse position in a computer screen with her brain activity (Section~\ref{}).
-Nodes highlighted in color require machine intelligence functionality currently
-not available in Bonsai.
-%
-\end{comment}
-
-A central aim of this proposal is to create software infrastructure to provide
-this machine intelligence functionality to Bonsai in extensible and general
-ways.
-
-\subsubsection{Generality}
-
-The number of current machine learning algorithms is very large and it is
-continuously increasing. We intend to build a small number of machine
-intelligence operators in Bonsai that can be configured to provide the
-functionality of most machine learning algorithms.
-
-For this, we will build on abstractions used in the \texttt{sickit-learn}
-library. The success of this library is, to a large extent, due to its simple
-and general application programming interface \citep[API;][]{buitinckEtAl13}.
-%
-This API consists of three complementary interfaces: an estimator interface for
-building and fitting models, a predictor interface for making predictions and a
-transformer interface for converting data.
-%
-Using only these three interfaces \texttt{scikit-learn} provides
-implementations of a large number of machine learning algorithms for
-dimensionality reduction, clustering, classification and regression.
-
-Inspired by \texttt{scikit-learn}, Bonsai will have three types of operators:
-estimators, predictors and transformers. Each of these operators will be
-instantiated by objects implementing the corresponding interface.
-%
-For example, consider the close-loop system for the control of a computer
-cursor with neural activity represented in Figure~\ref{fig:bci}a. The Bonsai
-graphical program for this system is given in Figure~\ref{fig:bci}b. The
-objects instantiating the SpikeThresholdDetector, the KiloSortSpikeSorter and
-GPFAlatentsExtractor will implement the transformer interface, and the object
-instantiating the TimeSeriesForestClassifier will implement the predictor
-interface (Figure~\ref{fig:bci}b, top).
-%
-The parameters for the SpikeThresholdDetector and the KiloSortSpikeSorter will
-be set manually by the user, as is commonly done in practice. The parameters of
-the GPFAlatentsExtractor will be learnt offline in an unsupervised way in a
-second pipeline (Figure~\ref{fig:bci}b, middle). Finally, the parameters of the
-TimeSeriesForestClassifier will also be learnt offline, but in a supervised
-way, in a third pipeline (Figure~\ref{fig:bci}b, bottom).
-
-\begin{figure}
-   \begin{center}
-       \includegraphics[width=4in]{figures/bci.png}
-       \caption{Close-loop system for controlling a computer cursor with brain activity}
-     \label{fig:bci}
-   \end{center}
-\end{figure}
-
-\begin{comment}
-    The spike detector transforms membrane
-potentials into spike sequences. The spike sorter transforms spikes squences
-into lists of spikes, one list per neuron. The latents extractor transforms
-multi-neuron lists of spikes into latent trajectories summarizing the
-multi-neuon spiking activity.  The neural decoder predicts the next cursor
-position from the latent trajectories.
-\end{comment}
-
-\begin{comment}
-middl, whose output will set
-As in this library,
-there will be three type of machine intelligence operators in Bonsai:
-%
-estimator operators for building and fitting models, predictor operators for
-making predictions and transformer operators for converting data. These
-operators will be implemented by objects exposing the corresponding interfaces.
-
-The estimator interface exposes a fit method for learning a model from training
-data.  This method is called with training data (e.g., supplied as two arrays
-X\_train and y\_train in supervised learning estimators). Its task is to run a
-learning algorithm and to determine model-specific parameters from the training
-data.  All supervised and unsupervised learning algorithms (e.g., for
-classification, regression or clustering) will be offered as objects
-implementing this interface. Machine learning tasks like feature extraction,
-feature selection or dimensionality reduction will also be provided as
-estimators.
-
-The predictor interface extends the notion of an estimator by adding a predict
-method that takes an array X\_test and produces predictions for X\_test, based
-on the learned parameters of the estimator (we call the input to predict
-“X\_test” in order to emphasize that predict generalizes to new data). In the
-case of supervised learning estimators, this method typically returns the
-predicted labels or values computed by the model.
-
-Since it is common to modify or filter data before feeding it to a learning
-algorithm, the transformer interface defines a transform method. It takes as
-input some new data X\_test and yields as output a transformed version of
-X\_test. Preprocessing, feature selection, feature extraction and
-dimensionality reduction algorithms will be provided as transformers in Bonsai.
-
-Bonsai will provide the ability to compose new
-estimators from several base estimators. Composition mechanisms can be used
-to combine typical machine learning workflows into a single object which is
-itself an estimator, and can be employed wherever usual estimators can be used.
-Composition of estimators will be done with Pipeline objects that chain multiple estimators into a single one. This is useful
-since a machine learning workflow typically involves a fixed sequence of processing steps (e.g., feature extraction, dimensionality reduction, learning and making
-predictions), many of which perform some kind of learning. A sequence of N such
-steps can be combined into a Pipeline if the first N-1 steps are transformers;
-the last can be either a predictor, a transformer or both.
-
-\end{comment}
-
-\subsubsection{Extensibility}
-
-\textcolor{red}{I will start this section by saying that we are not proposing
-to implement machine learning algorithms, as in \texttt{scipy.learn}, but to
-integrate into Bonsai existing machine intelligence algorithms.}
-
-Currently Bonsai has its own package manager, and its community of users have
-already extended Bonsai's functionality with several contributed packages
-(e.g., \ldots). However, these packages need to be written in C\#, while most
-current data analysis methods are written in Python, R or Matlab. We propose to
-add to Bonsai capabilities to communicate with software written in these
-languages.
-
-Fortunately, there exist software that allow C\# program
-to call and be called by programs in these other languages. For the
-communication of C\# programs with Python we will use
-Python.NET\footnote{\href{https://github.com/pythonnet/pythonnet}{https://github.com/pythonnet/pythonnet}},
-with R programs we will use
-R.Net\footnote{\href{https://github.com/rdotnet/rdotnet}{https://github.com/rdotnet/rdotnet}}
-and with Matlab we will use the built in Matlab .NET
-interface\footnote{\href{https://uk.mathworks.com/help/matlab/matlab\_external/using-net-from-matlab-an-overview.html}{https://uk.mathworks.com/help/matlab/matlab\_external/using-net-from-matlab-an-overview.html}}.
-In cases where these software cannot support some type of communication between
-C\# and another language we will use the messaging library
-ZeroMQ\footnote{\href{https://zeromq.org/}{https://zeromq.org/}}.
-
-Using the previous software, we will build C\# wrappers around machine
-intellgence programs written in other languages, so that these programs
-implement the Estimator, Predictor or Transformer interfaces and can be
-integrated into Bonsai as machine intelligence operators.
-
-\subsection{Machine intelligence functionality to be added to Bonsai}
-\label{sec:toBeAddedMachineIntelligenceFunctionality}
-
-\subsubsection{Video-based behavioral analysis}
-\label{sec:videoBasedBehavioralAnalysis}
-
-Precisely quantifying animal behavior is an essential step toward understanding
-brain function.  Deeplabcut is a Python-based software for tracking animal body
-parts \citep{mathisEtAl18}, which it is currently well integrated with
-Bonsai~\citep{kaneEtAl20}. Here we propose extensions to Bonsai to extract
-other informative features from video recordings.
-
-\begin{description}
-
-    \item[Motion sequencing]\citep[MoSeq;][]{wiltschkoEtAl15}: extracts patterns
-        of behavior that repeat over time (i.e., syllables of behavior) from
-        video data. For instance, it parses (in an unsupervised way) the
-        behavior of a mouse in an arena into segments of time where a mouse
-        was running, rearing and grooming. It uses a hidden Markov model and it
-        is implemented in Python\footnote{Code for MoSeq can be requested from
-        the Datta laboratory, as indicated at
-        \href{http://datta.hms.harvard.edu/research/behavioral-analysis/}{http://datta.hms.harvard.edu/research/behavioral-analysis/}.}.
-        After trained MoSeq can be used to detect behavioral syllables online.
-
-    \item[Decoding behavior from neural
-        activity]\citep[BehaveNet;][]{battyEtAl19} combines hidden Markov
-        models with convolutional autoencoders and discriminative models to
-        decode video data from neural recordings. It is implemented in
-        Python\footnote{\href{https://github.com/themattinthehatt/behavenet}{https://github.com/themattinthehatt/behavenet}}.
-        After trained it can be used online to decode video data and detect
-        behavioral syllables.
-
-    \item[Interpretable latents for behavioral videos]\citep[Partitioned
-        Subspace Variational Autoencoder, PS-VAE;][]{whitewayEtAl21}: produces
-        interpretable low-dimensional representations of behavioral videos by
-        combining the output of pose-estimation algorithms (e.g., DeepLabCut)
-        with unsupervised dimensionality reduction methods. These
-        low-dimensional representations facilitate downstream behavioral and
-        neural analyses. It is based on autoencoders and is implemented in
-        Python\footnote{code for PS-VAE is embedded in the BehaveNet code
-        \href{https://github.com/themattinthehatt/behavenet}{https://github.com/themattinthehatt/behavenet}.
-        Please refer to class \texttt{PSVAE} in 
-        \texttt{behavenet.models.vaes.py}.}. After trained it can be used
-        online to extract low-dimensional features and perform downstream
-        processing.
-
-\end{description}
-
-\subsubsection{Spike sorting}
-\label{sec:spikeSorting}
-
-\noindent\textcolor{red}{TODO}
-
-\subsubsection{Low-dimensional representations of spiking data}
-\label{sec:lowDimensionalRepresentationsOfSpikingData}
-
-Bonsai is well integrated with OpenEphys~\citep{siegleEtAl17}, which allows
-scientists to record neural data from a large number of devices. However, it
-lacks functionality to process these recordings. Here we describe software that
-we propose to integrate into Bonsai to extract interpretable summary statistics
-(i.e., latent variables) from neural spiking activity.
-
-\begin{description}
-
-    \item[Gaussian Linear Dynamical
-        System]\citep[GLDS;][]{andersonAndMoore12}: with sufficiently large
-        bin sizes, spike counts can be modeled as Gaussian random processes.
-        This Gaussianity assumption greatly simplifies the estimation of
-        parameters of linear dynamical system (LDS) models, as well as
-        inferences about their states. After models parameters have been
-        learned, GLDS can be used online. A unique feature of GLDS is that the
-        posterior distribution of states given all observation up to the
-        present can be calculated efficiently. This estimate of the posterior
-        distribution can be used online for experimental control, as we
-        propose in Section~\ref{sec:comparisonOfMultipleMethods}. We will use an R implementation of GLDS
-        which allows the use of external
-        inputs\footnote{\href{https://github.com/joacorapela/kalmanFilter}{https://github.com/joacorapela/kalmanFilter}}.
-
-    \item[Poisson Linear Dynamical System]\citep[PLDS;][]{mackeEtAl15}: for
-        smaller bin sizes, spike counts are better modeled as Poisson random
-        processes, rather than Gaussian ones. The algorithm described in
-        \citet{mackeEtAl15} can estimate the parameters of a LDS, and make
-        inferences about its states, from Poisson distributed observations.  We
-        propose to interface Bonsai with a Matlab implementation of
-        PLDS\footnote{\href{https://bitbucket.org/mackelab/pop\_spike\_dyn/src/master/}{https://bitbucket.org/mackelab/pop\_spike\_dyn/src/master/}}
-        that uses variational inference. PLDS does not provide online estimates
-        of the states, since data from a full trial is needed for state
-        inference.
-
-    \item[Hidden Markov Model]\citep[HMM;][]{rabiner89}: as LDSs, HMMs model a
-        time series of observations as random processes conditioned on hidden
-        states. However, in HMMs hidden states are discrete random variables,
-        while in LDSs they are continuous ones. In some application domains
-        (e.g., speech, epilepsy) discrete state assumptions are more pertinent
-        than continuous ones. We propose to use an R implementation of
-        HMMs\footnote{\href{https://github.com/joacorapela/hiddenMarkovModels}{https://github.com/joacorapela/hiddenMarkovModels}}.
-        As GLDSs, once trained, HMMs can be used online to infer the posterior
-        distribution of states given observations.
-
-    \item[Gaussian Processes Factor Analysis]\citep[GPFA;][]{yuEtAl09}: as
-        LDSs, GFPA models represent a time series of observation as random
-        processes conditioned on hidden states. However, in GPFA models the
-        state dynamics are nonlinear, while in LDS models they are linear.
-        Thus, GPFA models are more general than LDS ones. As GLDS models, GPFA
-        models assume that observations conditioned on states are Gaussian
-        random processes. We propose to use a Matlab implementation of
-        GPFA\footnote{\href{https://users.ece.cmu.edu/~byronyu/software/gpfa0203.tgz}{https://users.ece.cmu.edu/~byronyu/software/gpfa0203.tgz}}.
-        GPFA models do not provide online estimates of the states, since data
-        from the full trial are needed for state estimates.
-
-    \item[Sparse Variational Gaussian Processes Factor
-        Analysis]\citep[svGPFA;][]{dunckerAndSahani18}: is similar to GPFA, but
-        models point process observations (i.e., single spikes as opposed to
-        spike counts). We propose to use a Python implementation of
-        svGPFA\footnote{\href{https://github.com/joacorapela/svGPFA}{https://github.com/joacorapela/svGPFA}}.
-        svGPFA models do not provide online estimates of states, since data
-        from the full trial are needed for state estimates.
-
-    \item[Latent factor analysis through dynamical
-        systems]\citep[LFADS;][]{pandarinathEtAl18}: uses an autoencoder
-        framework, with recursive neural networks, to infer continuous states
-        conditioned on spike counts, similar to those inferred by GPFA. We
-        propose to use a Python implementation at
-        LFADS\footnote{\href{https://github.com/tensorflow/models/tree/master/research/lfads}{https://github.com/tensorflow/models/tree/master/research/lfads}.}.
-        As GPFA, LFADS do not provide online estimates of states.
-
-\end{description}
-
-\subsubsection{Low-dimensional representations of local-field potentials}
-\label{sec:lowDimensionalRepresentationsOfLFPs}
-
-Spikes are extracted from a higher-frequency range of extracellulary recorded
-voltages. Local field potentials (LFPs) are computed from a lower-frequency
-range and are another important signal to understand brain function. We propose
-to use states inferred from LFPs by GLDS, GPFA, HMM and LFADS
-(Section~\ref{sec:lowDimensionalRepresentationsOfSpikingData}) as
-low-dimensional representations of the LFP.
-
-\subsubsection{Neural decoding}
-\label{sec:neuralDecoding}
-
-In order to use low-dimensional representations of spiking activity and/or of
-LFPs to guide the control of experiments, we need a decoder to optimally map
-these low-dimensional representations to experimental controls.
-%
-We propose to implement in Bonsai several decoding/classification algorithms:
-k-nearest neighbor, linear discriminative analysis, support vector machines,
-random forests, artificial neural networks, naive Bayes and Gaussian process
-classifiers.
-
-\subsubsection{Comparing multiple data-analysis methods}
-\label{sec:comparisonOfMultipleMethods}
-
-We propose to add functionality to Bonsai to facilitate multi-method
-comparisons in user-supplied datasets. Below we describe a comparison that we
-will perform in Bonsai to asses the relative performance of the methods
-described in the previous sections with neural recordings from behaving rats.
-These recordings are currently being performed, to address scientific questions,
-at the laboratory of Prof.~Akrami in the SWC.
-%
-Details of the experimental task appear in~\citet{akramiEtAl18}.
-%
-Briefly, in this task there are three nose ports (left, center, right). Rats
-initiate a trial by inserting their nose in the center port for the duration of
-the fixation period, until they hear a Go stimulus. During the fixation period
-two stimuli $s_a$ and $s_b$ are presented. Rats should decide which stimuli is
-louder. If $s_a$ is louder than $s_b$ the correct action is to poke the nose
-into the right port in order to collect a liquid reward, and if $s_b$ is louder
-than $s_a$ the correct choice is the left port.
-
-We will use high-density Neuropixel spike and LFP recordings from the rats
-performing this task to learn low-dimensional representations of spiking
-activity and LFPs during the decision time period (between the presentation of
-the last auditory stimuli and the presentation of the Go stimulus), using the
-methods described in
-Sections~\ref{sec:lowDimensionalRepresentationsOfSpikingData}
-and~\ref{sec:lowDimensionalRepresentationsOfLFPs}. We will also learn the
-parameters of the classifiers described in Section~\ref{sec:neuralDecoding} to
-optimally classify animal decisions (poke the nose into the right/left port)
-based on the previous low-dimensional representations.
-
-We will compare the decoding accuracy of all decoders. This comparison will
-tells us, for the auditory working memory task, which neural data type (spikes
-or LFPs), which low-dimensional representation method (e.g., LDS or Gaussian
-processes), and which decoder (e.g., Bayesian or ANN) yields the best decodings
-in state-of-the-art neural recordings from behaving animals.
-
-\begin{comment}
-
-\subsection{Interfacing Bonsai with Python, R and Matlab data analysis software}
-\label{sec:bonsaiPythonMatlabCommunication}
-
-Bonsai is written in C\# and most neural data-analysis methods are written in
-Python, R or Matlab. Fortunately, there exist software that allow C\# program
-to call and be called by programs in these other languages. For the
-communication of C\# programs with Python we will use
-Python.NET\footnote{\href{https://github.com/pythonnet/pythonnet}{https://github.com/pythonnet/pythonnet}},
-with R programs we will use
-R.Net\footnote{\href{https://github.com/rdotnet/rdotnet}{https://github.com/rdotnet/rdotnet}}
-and with Matlab we will use the built in Matlab .NET
-interface\footnote{\href{https://uk.mathworks.com/help/matlab/matlab\_external/using-net-from-matlab-an-overview.html}{https://uk.mathworks.com/help/matlab/matlab\_external/using-net-from-matlab-an-overview.html}}.
-In cases where these software cannot support some type of communication between
-C\# and another language we will use the messaging library
-ZeroMQ\footnote{\href{https://zeromq.org/}{https://zeromq.org/}}.
-
-\end{comment}
-
-\subsection{Reproducibility with Bonsai}
-\label{sec:reproducibility}
-
-The current version of Bonsai facilitates reproducibility of data acquisition
-and experimental control across laboratories. Most data acquisition and
-experimental control aspects of an experiment can be reproduced across
-different laboratories by sharing a Bonsai configuration file.
-%
-By adding data analysis functionality to Bonsai, besides reproducing my data
-acquisition and experimental control, Bonsai users will also be able to reproduce
-data analysis functionality.
-
-We will demonstrate each machine intelligence functionality added to Bonsai
-(Section~\ref{sec:toBeAddedMachineIntelligenceFunctionality}) and the multiple
-data-analysis methods comparisons
-(Section~\ref{sec:comparisonOfMultipleMethods}) in Bonsai experiments. We will
-then distribute Bonsai configuration files for users to reproduce these
-demonstrations.
-
-\subsection{Final remarks}
-\label{sec:finalRemarks}
-
-We proposed to add machine intelligence functionality to Bonsai to assist users
-in building close-loop neural experiments: online spike sorting, methods for
-low-dimensional representation of spiking activity and LFPs, and neural
-decoding algorithms. This functionality will allow neuroscience Bonsai users to
-build more sophisticated experiments and extract more information from their
-experimental recordings.
-
-We also proposed to add functionality to help Bonsai users compare the
-performance of multiple methods on their datasets. With this functionality
-Bonsai users will be able to select the best methods to model their own
-recordings. It will also motivate methods developers to interface their
-data-analysis functionality with Bonsai. By doing so, they will provide their
-software to the large community of Bonsai users, and will be able to easily
-compare the performance of their methods with that of preexisting methods in
-the Bonsai ecosystem. In this way method developers will become a new type of
-Bonsai users, which will guarantee the sustainability of developments
-proposed here.
-
-We focused this proposal on neuroscience applications of Bonsai. However, the
-addition of machine intelligence functionality is needed in many other
-experimental areas where Bonsai is or could be used. Therefore, the software
-abstractions that we will build to add to Bonsai machine intelligence
-functionality for neuroscience experiments (e.g., methods to perform
-multi-method comparisons) will translate to other application domains of
-Bonsai.
-
-Multiple groups at the SWC currently use Bonsai for data acquisition and
-experimental control. Carefully testing the functionality of new software is
-essential to ensure its correct functionality. We propose to use the SWC large
-internal Bonsai user base to test the functionality we will add to Bonsai,
-before distributing it to the general public.
+\input{detailsOfResource}
 
 \section{Community demand}
+\label{sec:communityDemand}
+
+\input{communityDemand}
 
 \section{User engagement}
+\label{sec:userEngagement}
+
+\input{userEngagement}
 

### 2021-11-12 backgroundToTheResource.tex
Added

Experimental control and neural data analysis are currently highly
disassociated.  Most neuroscientists collect their data and only afterward
analyse it. This split has had dramatic consequences for the progress of
neuroscience. First, most experimental control methods currently do not use
advanced data analysis methods to enhance their functionality. Second, the vast
majority of neural data analysis methods are offline and do not provide key
functionality needed for online experimental control. Using advanced data
analysis methods in experimental control will improve the sophistication of
experiments that are currently possible, and generate new demands for neural
data analysis methods that will trigger advances in machine learning.

A central goal of the software resource proposed here is to address this
problem, by providing Bonsai~(an experimental control software ecosystem;
Figure~\ref{fig:bonsai} and Section~\ref{sec:bonsai}) with state-of-the-art
online (and offline) data analysis methods
(Figure~\ref{fig:proposedBonsaiExtensions} and
Section~\ref{sec:functionality}). Bonsai is
\textbf{unique} among experimental control software in that it allows people
with no programming experience to control sophisticated experiments. Its large
and rapidly growing user base demonstrates the need for this type of software
resources. With the proposed functionality to be added to Bonsai, its users
will be able to use sophisticated machine-learning methods to control their
experiments and analyse their data. This
\textbf{democratisation} of machine learning
functionality will allow scientists without computational skills to control
sophisticated experiments and analyse their data in unprecedented ways, which
will most probably yield important \textbf{advances in science}.

\begin{figure}
    \begin{center}
        \includegraphics[width=7in]{figures/bonsai.pdf}
        \label{fig:bonsai}
    \end{center}
\end{figure}

\setcounter{figure}{1}

\begin{figure}
    \begin{center}
        \includegraphics[width=4in]{figures/proposed_bonsai_extensions.png}
        \caption{Proposed extensions to Bonsai}
        \label{fig:proposedBonsaiExtensions}
    \end{center}
\end{figure}

Currently Bonsai has its own package manager, and its community of users have
already extended Bonsai's functionality with several contributed packages
(e.g., \ldots \textcolor{red}{Goncalo please reference the best contributed
packages developed by the Bonsai community}). However, these packages need to
be written in C\#, while most current neuroscience data analysis methods are
written in Python, R or Matlab. We propose to add to Bonsai capabilities to
communicate with software written in these languages
(Section~\ref{sec:softwareInfrastructure}).  These capabilities will allow a
large number of neuroscience data analysis methods to be easily integrated into
the Bonsai ecosystem and will provide Bonsai users a large repertoire of
advanced machine learning methods for their behavioural and neural data
analysis.
%
Using these capabilities we will integrate into Bonsai an initial set of
state-of-the art behavioural and neural data analysis methods
(Section~\ref{sec:functionality}).

Developers of advanced data-analysis methods should become interested in
integrating them into the Bonsai ecosystem, which will allow their methods to
reach the wide Bonsai community. In addition this integration will enable
method developers to easily compare the performance of their methods with
that of many methods previously integrated into the Bonsai ecosystem. Method
developers will thus become a new type of Bonsai users, which will contribute
to the \textbf{sustainability} of the proposed resource beyond the period of
BBSRC-BBR funding.

Bonsai is an excellent tool for reproducible data acquisition and
experimental control. For example, an experiment implemented in Bonsai can be replicated in
a new laboratory by just sharing a Bonsai configuration file. With the addition
of the proposed machine learning functionality Bonsai will extend this
reproducibility to the domain of data analysis
(Section~\ref{sec:reproducibility}).

\subsection{Bonsai}
\label{sec:bonsai}

\input{bonsai.tex}

### 2021-11-12 bonsai.tex
Added

--- 2021-11-09 main.tex
@@
-\documentclass[a4paper,11point]{article}
-\usepackage[margin=2cm]{geometry}
+Bonsai~\citep{lopesEtAl15,lopesAndMonteiro21} is a free and open-source visual
+programming language that emphasizes performance, flexibility, and ease-of-use,
+allowing scientists with no previous programming experience to quickly develop
+their own high-performance data acquisition and experimental control systems
+(Figure 1c). Bonsai combines a high-level event algebra for data streams with
+an integrated development environment (IDE) and a strong library of plugins
+supporting multiple hardware and software packages used by the biomedical
+research community (Figures 1a, 1b).
 
-%% Fonts, etc.
+Standard software tools for data analysis (ImageJ, MATLAB, R, etc.) have been
+transformative to the progress of increasingly “data-rich” sciences. However,
+the equivalent standardized software tools for data acquisition and control of
+animal behaviour experiments are still lacking. Life science experiments demand
+a combination of multiple instrumentation and control technologies, in both
+physiology and behaviour studies. The growing complexity on both the amount of
+data that is collected, and the rich conditions under which behaviour must be
+explored, place an increasing burden on the experimenter to integrate highly
+specialized equipment in unique configurations, while often lacking expertise
+in the relevant engineering fields.
 
-%% Language and font encodings
-\usepackage[english]{babel}
-\usepackage[utf8x]{inputenc}
-\usepackage[T1]{fontenc}
-% \usepackage{microtype}      % fine pdf font control
+Bonsai has a large user base in the systems neuroscience community.
+Quantifications of this user base are provided in Figures 1d-1f.  In the last
+year alone, we estimate more than 1,000 new users have started to incorporate
+Bonsai into their experimental protocols across the world. The surprising rate
+of adoption of Bonsai in non-programmer experimental labs highlights the need
+for accessible programming tools that enable state-of-the-art technology but
+also allow researchers to stay in control and change their experimental
+paradigms.  Many open-source software tools are either inaccessible to
+non-programmers, or too constrained to be of general use outside their narrow
+domain of application. Bonsai has been successful because it has straddled this
+gap to some extent.
 
-\usepackage[scaled=0.92]{helvet} \renewcommand{\familydefault}{\sfdefault}
-\usepackage{xcolor} % -- loaded by tikz
-% \usepackage{soul} % \ul \st \hl
+The language has also helped to potentiate the growing wave of foundational
+open hardware initiatives, such as the OpenEphys \citep{siegleEtAl17} and UCLA
+Miniscope~\citep{caiEtAl16}, making it possible to quickly combine and
+integrate these tools into new experiments~\citep{buccinoEtAl18}.
+%
+Bonsai has been adopted in large neuroscience undertakings like the
+International Brain
+Laboratory\footnote{\href{https://www.internationalbrainlab.com/}{https://www.internationalbrainlab.com/}}
+and the Allen Institute for Brain
+Science\footnote{\href{https://alleninstitute.org/what-we-do/brain-science/}{https://alleninstitute.org/what-we-do/brain-science/}}.
+%
+More recently, Bonsai has started to expand outside the domain of neuroscience
+into biomedical research and biotechnology tool development, and even outside
+academia into public outreach and education programs.
 
-%% Maths
-% \usepackage{amssymb,amsmath}
-% \usepackage{algorithm,algpseudocode}
-% \usepackage{mathtools}
-% \usepackage{mathalfa}
-% \usepackage{bbm}
-% \usepackage{mathrsfs}
-% \usepackage{gensymb}
-% \input m.macros
-% \input m.symbols
+\subsubsection*{Other resources in the subject area}
 
-%% Other typsetting
-% \usepackage{booktabs}       % professional-quality tables
-% \usepackage{nicefrac}       % compact symbols for 1/2, etc.
-\usepackage{enumitem}
-
-%% Graphics
-% \usepackage{graphicx,tikz}
-% \usetikzlibrary{positioning}
-% \usetikzlibrary{fit}
-% \usetikzlibrary{calc}
-
-%% Floats
-% \usepackage{wrapfig}
-
-%% Margin bubbles
-% \usepackage{mbubble}
-% \mbubblestyle{ms}{font=\tiny,label=ms,color=blue}
-% \newcommand{\comment}[1]{}
-
-%% Sections
-\setcounter{secnumdepth}{-2}
-\usepackage[small]{titlesec} % format section titles
-% \titleformat{command}[shape]{format}{label}{sep}{before-code}[after-code]
-
-% \usepackage{caption,subcaption}
-% \DeclareCaptionSubType[Alph]{figure}
-% \captionsetup[subfigure]{labelformat=simple,labelfont={sf,bf,large},textfont={sf}}
-% \renewcommand{\thesubfigure}{{\sffamily\alph{subfigure}}}
-
-%% Spacing
-% \usepackage[doublespacing]{setspace}
-\parindent 0pt
-\parskip 1ex
-
-%% Citations, refs, hyperlinks
-\usepackage[colorlinks=true, allcolors=blue]{hyperref}
-% \usepackage{natbib}
-% \usepackage[capitalize]{cleveref}
-% \crefname{equation}{}{}
-% \crefname{Equation}{Equation}{Equations}
-
-%% Title
-
-\title{}
-\author{}
-% \usepackage{filemod} % access to file modification times
-% \date{\Filemodtoday{\jobname}}  % replace \jobname if compiling from wrapper
-
-% \usepackage{titling} % modify title(page) appearance
-%\pretitle{\begin{center}\LARGE}
-%\posttitle{\par\end{center}\vskip 0.5em}
-%\preauthor{\begin{center} \large \lineskip 0.5em \begin{tabular}[t]{c}}
-%\postauthor{\end{tabular}\par\end{center}}
-%\predate{\begin{center}\large}
-%\postdate{\par\end{center}}
-
-\def\ii#1\par{{\color{blue!40}\sl #1}\par}
-\def\iibf#1\par{{\color{blue!40}\sl\bfseries #1}\par}
-\def\iitem#1\par{\ii\begin{itemize}[nosep]\item #1 \end{itemize}\par}
-
-\begin{document}
-
-\section{Notes}
-
-(This page is not for submission)
-
-Required items:
-\begin{itemize}  
-  \item case for support (up to eight pages)
-  \item justification of resources (two pages)
-  \item data management plan (DMP) (up to three pages, data management plan template)
-        if you opt not to use the template for your plan, the topics listed in the template must be addressed in the DMP document you do provide
-  \item diagrammatic workplan (one page)
-  \item management structure (one page)
-  \item narrative CV (up to two pages per staff member)
-  \item community letters of support or demand (up to 10)  
-  \item proposal cover letter.
-\end{itemize}
-
-\newpage
-
-% \maketitle
-
-\section{Case for support}
-
-\ii Instructions and guidance:
-  \url{https://www.ukri.org/wp-content/uploads/2021/09/BBSRC-150921-FundingOpp-BioinformaticsBiologicalResourcesFund-GuidanceDocument.pdf}
-
-\iibf Assessment criteria.
-
-\ii \url{https://www.ukri.org/wp-content/uploads/2021/09/BBSRC-150921-FundingOpp-BioinformaticsBiologicalResourcesFund-AssessmentCriteria.pdf}
-
-\iibf Scientific quality and strategic relevance of the resource.
-
-\ii Including:
-\begin{itemize}[nosep]
-    \item the extent to which the resource meets the highest international standards in resource provision in its field
-    \item how well the resource is demonstrated to be unique or complementary to other similar existing resources
-    \item the extent to which the resource addresses the research and policy priority areas of BBSRC.
-\end{itemize}
-
-\iibf Cost effectiveness, particularly considerations for long-term
-sustainability beyond BBSRC funding.
-
-\ii The extent to which:
-\begin{itemize}[nosep]
-    \item the resource delivers value for money relative to the anticipated scientific gains it will provide
-    \item long term sustainability options have been considered, addressed, and planned where appropriate, particularly for existing resources.
-\end{itemize}
-
-\iibf Potential for economic and social impact beyond the academic
-community.
-
-\ii Including:
-\begin{itemize}
-    \item the extent to which the outputs from the resource will contribute to knowledge and potential for economic return or social impact
-    \item how well the proposal has outlined methods of engagement and measures of success in developing milestones and timelines of associated activities.
-\end{itemize}
-
-\iibf Fit to the scope.
-
-\ii How well the proposal addresses the scope of the opportunity.
-
-\iibf Assessment criteria adapted to new or existing resources
-
-\ii To allow for a more nuanced assessment between new and existing
-resources, the use of the following assessment criteria will be
-adapted accordingly.
-
-\ii For new resources, these criteria will assess the ‘plans, potential, and promise’ of the resources.
-
-\iibf Quality of the overall arrangements for resource management, advisory functions, as well as user access and engagement
-
-\ii Including:
-\begin{itemize}
-    \item the extent to which the proposal has evidenced or planned interaction with relevant users and the broader research community to ensure the aims of the resource are realised and there is sufficient uptake and continued development
-    \item the extent to which adequate user access arrangements have been discussed and considered
-    \item the set-up of project management and advisory structures of the resource to ensure longevity in delivering the resource to a broad user base.
-\end{itemize}
-
-\iibf Need or demand, and potential benefit to the UK academic research community
-
-\ii The extent to which:
-\begin{itemize}
-    \item the community has demonstrated demand for the proposed resource, relative to the total community size (in particular, proposals for new resources should have consulted their prospective community prior to application)
-    \item the proposed resource will deliver and benefit the wider BBSRC community indicating how the proposed resource will help to deliver high-quality research.
-\end{itemize}
-
-\newpage
-\ii Section headings taken from guide.  We can omit some if clearly not necessary, but should do so with caution.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\subsection{Background to the Resource}
-\ii Introduction of the proposed resource, including its academic and wider 
-economic and societal context.
-
-\ii Overview of past and current resource(s) in the subject area in both the UK 
-and abroad, including any alternative community resources currently 
-available. You should indicate the community size of the intended resource 
-and how this relates to the field in which it operates.
-
-\subsection{Details of Resource}
-
-\ii The case for support should outline the full details of the resource and 
-associated work packages presented in the proposal.
-
-\iitem Indication whether the project proposed to develop a new
-resource or is in support of an existing one.
-
-Existing if general Bonsai enhancements.  If specific to analysis we
-may be able to argue for "new" if helpful, while still making
-reference to Bonsai installed base.
-
-\iitem Objectives for the proposal should be detailed including
-individual measurable targets against which the outcome of the work
-will be assessed. This should refer to the objectives set out in the
-Je-S proposal form.
-
-\iitem{Significant technical details for the development, maintenance or 
-enhancement of the resource must be clearly outlined and indicated 
-how this is of internationally exceptional quality. }
-
-\iitem{If applicable, outline any proposed research efforts and how they 
-  directly facilitate development of the resource. }
-
-\iitem{For proposals looking to focus on maintaining status quo for an 
-existing resource instead of suggesting further development, you 
-should detail evidence of why significant upgrades are not required at 
-this time and detail why the resource needs continued support to 
-maintain world-leading functionality.}
-
-\ii{Additional questions that may be considered: }
-
-\iitem Does the facility begin or continue to support a growing field of 
-bioscience, what is the anticipated growth and does the proposal 
-adequately accommodate this?
-
-\iitem How will the resource accelerate science within its field and
-  beyond?
-
-\iitem What would be the impact on the scientific community if the resource 
-did not exist? How would this impact other, possibly dependent 
-resources?
-
-\subsection{Community Demand}
-\ii Evidence of community demand should be primarily driven from UK 
-academic researchers working largely within BBSRC remit – see our 
-Forward look for UK Bioscience for more detail on research areas covered by 
-BBSRC. Demand from other users (such as academic communities outside 
-of BBSRC remit or industrial users) may be appropriate to provide additional 
-support, especially in highlighting the potential for economic, commercial or 
-societal impact, but should not be the focus of the demand demonstration. 
-Evidence provided should highlight examples of the high-quality science that 
-the resource will underpin or has underpinned. Where possible and relevant, 
-examples should be drawn from a wide research community to illustrate the 
-broad impact of the resource to support high-quality internationally excellent 
-science.
-
-\ii The level of community demand should be benchmarked against other 
-relevant resources and/or the size of the community. This will allow the fair 
-assessment of resources with different user bases. The types of evidence 
-that may be appropriate to provide will be different for new and existing 
-resources.
-
-\ii Evidence of wider consultation of the prospective community (e.g community 
-surveys) is encouraged.
-
-\iibf New Resources
-
-\ii New resources should estimate the number of researchers who may engage and benefit from the resource. Evidence, where possible, would be of benefit and may include: 
-\begin{itemize}
-    \item  Datasets (or samples) in public or private repositories 
-    \item Citations or acknowledgments 
-    \item Gap analysis with existing resources 
-    \item Pilot project uptake or feedback from potential users.  
-\end{itemize}
-
-\ii In particular, proposals for new resources should have consulted their prospective community prior to application
-
-\iibf Existing Resources
-
-\ii For existing resources this should include usage data of the
-current resource. Data types may include:
-
-\iitem Access requests from independent users/ sites
-
-\iitem Citations or acknowledgements
-
-\iitem Other public resources providing links to the resource
-
-\iitem New major acquisitions captured by the resource
-
-\ii{Additionally, existing resources need 
-to evidence why this resource needs 
-to be maintained/updated by the 
-current grant. This could include: }
-
-\iitem Survey data from users on what upgrades are needed
-
-\iitem Evidence of an expanding 
-user base, which requires 
-additional resource
-
-\iitem Recent developments in the 
-field, which require upgrades 
-to be integrated into the 
-resource.
-
-\subsection{User engagement}
-\ii Discussion of user engagement provision should aim to answer the following 
-questions:
-
-\iitem Is there awareness of the resource within the user community?
-
-\iitem How do you plan to develop the engagement strategy within the 
-proposal timeframe to expand user engagement?
-
-\iitem How have access mechanisms to ensure usability of the resource 
-been considered?
-
-\iitem How have user needs been incorporated into this proposal to ensure 
-it is fit for purpose and will deliver on expectations?
-
-\iibf New Resources
-
-\ii Evidence should be provided as to how the resource plans to engage
-stakeholders and ensure that the resource meets their needs and is
-used by the community targeted by the resource.
-
-\iibf Existing Resources
-
-\ii Evidence should be provided as to whether the resource has
-achieved the level of engagement it originally anticipated, and
-consideration is given how the additional investment would change
-this.
-
-\subsection{Long term sustainability planning}
-\ii{
-In addition, the case for support should outline considerations for the long-
-term sustainability of the resource beyond UKRI-BBSRC funding, as well as 
-the true cost of running and maintaining the resource in question. 
-The proposal should include: }
-
-\iitem Cost recovery plans, where appropriate, or an explanation why
-not if not viable. Evidence of clear business planning with a focus on
-at least partial cost recovery is required, especially when applying
-as an existing resource.
-
-\iitem Details for alternative support plans, aside from UKRI-BBSRC 
-funding.
-
-\iitem The level of support the resource is projected to require for expected 
-maintenance and/or subsequent maturation/enhancement activities. 
-Clear arguments as to why UKRI-BBSRC should support the resource now 
-should be provided, if other cost recovery and support plans are deemed 
-unsuitable.
-
-\subsection{Potential for economic and societal impact}
-\ii{Outline how the outputs of the proposed resource will contribute to 
-knowledge and how this may have the potential for economic return or 
-societal benefits. Impact activities should be integrated into appropriate 
-sections of the case for support, not presented as an independent work 
-package. }
-
-\iitem All proposals are expected to demonstrate clear plans with recorded 
-milestones and timelines for associated activities to develop 
-economic, commercial and societal impacts.  
-August 2021 7
-
-\iitem Methods of engagement and measures of success should be outlined 
-including how these will be regularly reviewed throughout the project 
-in order to deliver the most impact.
-
-\iitem Any planned activities should be fully justified within the Justification 
-of Resources attachment
-
-\subsection{Track Record}
-\ii{
-The majority of the track record relevant to the project should be located 
-within the Narrative CV and should not be repeated within the case for 
-support. You may, however, want to describe: }
-
-\iitem Track record of the team working together
-
-\iitem Specific expertise available at the host organisation and any 
-proposed partner organisations to enable the successful delivery of 
-the project.
-
-\iitem The specific role of each applicant and collaborator in the project
-
-\end{document}
+The space of technologies serving experimental control and behaviour monitoring
+is large, and is traditionally occupied either by domain-specific graphical
+user interfaces for control and recording of specific devices and experiment
+types (e.g. Open Ephys
+GUI\footnote{\href{https://open-ephys.org/gui/}{https://open-ephys.org/gui/}},
+Miniscope DAQ
+Software\footnote{\href{https://github.com/Aharoni-Lab/Miniscope-DAQ-QT-Software}{https://github.com/Aharoni-Lab/Miniscope-DAQ-QT-Software}})
+or by real-time control frameworks for specifying task logic using state
+machine or similar formalisms (e.g. NIMH ML
+(nih.gov)\footnote{\href{https://monkeylogic.nimh.nih.gov/}{https://monkeylogic.nimh.nih.gov/}},
+pyControl\footnote{\href{https://pycontrol.readthedocs.io/en/latest/}{https://pycontrol.readthedocs.io/en/latest/}},
+Autopilot\footnote{\href{https://docs.auto-pi-lot.com/en/latest/}{https://docs.auto-pi-lot.com/en/latest/}},
+Sanworks\footnote{\href{https://sanworks.io/index.php}{https://sanworks.io/index.php}}).
+These dedicated interfaces are typically very comfortable for experimenters
+operating within the specific domain that the tool is designed for but tend to
+become unwieldy when there is a need to introduce a new device or variation of
+a task which is outside the scope of the framework. The alternative is to use a
+more general programming language like Python or MATLAB, with the disadvantage
+of the code being harder to understand, maintain, and change. Programming
+languages like LabVIEW straddle the middle ground and provide a high-level and
+flexible visual interface for composing data acquisition and control systems
+themselves. Unlike Bonsai, however, the graphical elements of LabVIEW are
+heterogeneous and very fine grained, creating the need for long and complicated
+logical structures to implement even a simple experimental control system. By
+providing an extremely simple, yet flexible, visual syntax, Bonsai opens the
+opportunity even for complete non-programmers to design and successfully
+customize relatively complex experiments from the ground up. It is mostly this
+capability which has made Bonsai so attractive as a standard tool in
+experimental neuroscience.

### 2021-11-12 communityDemand.tex
Added

% Bonsai has directly enabled high-impact research in neuroscience~\citep[e.g.,][]{soaresEtAl16,grundemannEtAl17,netoEtAl16,hagiharEtAla21,anguillonRodriguezEtAl21} and contributed to accelerating technical progress in the field by functioning as an integrator of open-source, community developed technologies \citep{itskovEtAl14,lopesEtAl21,kaneEtAl20}.
The lack of a strong and
flexible machine learning package capable of being easily integrated with
real-time experimental control workflows has been a recurrent concern in the
Bonsai user community. This has been brought up spontaneously multiple times in
the user community forums, both in the form of specific problems related to the
tracking and interpretation of online animal behaviour, and in the form of
explicit proposals and suggestions for the creation of a machine learning
package\footnote{\href{https://groups.google.com/g/bonsai-users/c/BZ3zOOdv_1c/m/x6OP75frBQAJ?utm_medium=email&utm_source=footer&pli=1}{[Speculative
Enchancement] Deeplabcut + bonsai (google.com)}}. Even
relatively modest contributions such as integrating support for specific models
in Bonsai, such as DeepLabCut, or even the more specific zebrafish feature
tracking package BonZeb, has attracted tremendous attention and positive
feedback from the community \citep[e.g.,][]{kaneEtAl20,guilbeaultEtAl21}.

That there is a need for an easier interface into machine learning algorithms
is evidenced by the development of multiple graphical user interfaces for
tracking animal behaviour over the last several years, especially in the much
larger Python user community
\citep[e.g.,][]{walterAndCouzin21,guilbeaultEtAl21}. However, these graphical
user interfaces have no interact with real-time experimental control, and
their support libraries require expertise with the Python programming language
to use flexibly. Bonsai provides a way of addressing this gap, using its
flexible visual programming language to combine ease-of-use and full
flexibility for experimental control.

The ongoing AI revolution makes it extremely timely to integrate these
technologies into Bonsai. The proposed software infrastructure will make very
easy the integration of novel machine learning methods into the Bonsai
ecosystem. This integration will be critical to unlock new understanding, value
and scientific leads from the rich and complex behavioral and neural
experimetal data that Bonsai makes possible to collect. In addition, this
integration will open to non-programmer experimentalists in the biological
sciences unprecedented opportunities to combine real-time machine learning with
experimental control in neuroscience, which has to the best of our knowledge
not really been done before in such a broad and comprehensive scale. We believe
addressing this bottleneck and enabling these technologies to non-programmers
will make this resource into a transformative and fundamental technology for
the next generation studies in animal behaviour.

### 2021-11-12 detailsOfResource.tex
Added

--- 2021-11-10 main.tex
@@
-\documentclass[a4paper,11point]{article}
-\usepackage[margin=2cm]{geometry}
+We intend to improve an \textbf{existing resource}, the Bonsai ecosystem. We
+propose to create software infrastructure to facilitate the integration into it
+of existing advanced data analysis methods
+(Section~\ref{sec:softwareInfrastructure}).
+%
+To demonstrate the use of this infrastructure, and to provide Bonsai users with
+an initial set of advanced machine intelligence neural data analysis methods,
+we will add to the Bonsai ecosystem machine learning functionality to control
+neuroscience experiments and to analyse data generated by them
+(Section~\ref{sec:functionality}).
 
-%% Fonts, etc.
+This initial set of methods should demonstrate to Bonsai users, and to
+machine-learning methods developers, the potential of embedding
+machine-learning functionality into the experimental control loop. We propose
+user engagement activities to expose both of these groups of people to this
+potential (Section~\ref{sec:userEngagement}).  When data-analysis method
+developers recognise this potentials, they will become interested in
+contributing their methods to the Bonsai ecosystem, ensuring the
+\textbf{long-term sustainability} of the proposed resource.
 
-%% Language and font encodings
-\usepackage[english]{babel}
-\usepackage[utf8x]{inputenc}
-\usepackage[T1]{fontenc}
-% \usepackage{microtype}      % fine pdf font control
+Adding machine learning functionality to the experimental control loop will
+make possible the implementation of experiments with unprecedented levels of
+control. Moreover, equipping experimental scientists without programming
+experience with machine learning tools will considerably increase the
+application of machine learning to the biosciences, which most probably will
+translate into groundbreaking scientific discoveries.
 
-\usepackage[scaled=0.92]{helvet} \renewcommand{\familydefault}{\sfdefault}
-\usepackage{xcolor} % -- loaded by tikz
-% \usepackage{soul} % \ul \st \hl
+\subsection{Software infrastructure}
+\label{sec:softwareInfrastructure}
 
-%% Maths
-% \usepackage{amssymb,amsmath}
-% \usepackage{algorithm,algpseudocode}
-% \usepackage{mathtools}
-% \usepackage{mathalfa}
-% \usepackage{bbm}
-% \usepackage{mathrsfs}
-% \usepackage{gensymb}
-% \input m.macros
-% \input m.symbols
+\begin{comment}
+Bonsai does not currently provide machine learning functionality. For instance,
+the close-loop experimental control depicted in
+Figure~\ref{fig:bci} cannot be currently implemented in
+Bonsai. It represents the control of an experiment where a subject guides the
+mouse position in a computer screen with her brain activity (Section~\ref{}).
+Nodes highlighted in colour require machine intelligence functionality currently
+not available in Bonsai.
+%
+\end{comment}
 
-%% Other typsetting
-% \usepackage{booktabs}       % professional-quality tables
-% \usepackage{nicefrac}       % compact symbols for 1/2, etc.
-\usepackage{enumitem}
+We propose to create software infrastructure to add machine intelligence
+functionality to the Bonsai ecosystem in general
+(Section~\ref{sec:generality}) and extensible ways
+(Section~\ref{sec:extensibility}).
 
-%% Graphics
-% \usepackage{graphicx,tikz}
-% \usetikzlibrary{positioning}
-% \usetikzlibrary{fit}
-% \usetikzlibrary{calc}
+\subsubsection{Generality}
+\label{sec:generality}
 
-%% Floats
-% \usepackage{wrapfig}
+The number of current machine learning algorithms is very large and it is
+continuously increasing. We intend to build a small number of machine
+intelligence operators in Bonsai that can be configured to provide the
+functionality of most machine learning algorithms.
 
-%% Margin bubbles
-% \usepackage{mbubble}
-% \mbubblestyle{ms}{font=\tiny,label=ms,color=blue}
-% \newcommand{\comment}[1]{}
+For this, we will build on abstractions used in the \texttt{sickit-learn}
+library. The success of this library is, to a large extent, due to its simple
+and general application programming interface \citep[API;][]{buitinckEtAl13}.
+%
+This API consists of three complementary interfaces: an estimator interface for
+building and fitting models, a predictor interface for making predictions and a
+transformer interface for converting data.
+%
+Using only these three interfaces \texttt{scikit-learn} provides
+implementations of a large number of machine learning algorithms for
+dimensionality reduction, clustering, classification and regression.
 
-%% Sections
-% \setcounter{secnumdepth}{-2}
-% \usepackage[small]{titlesec} % format section titles
-% \titleformat{command}[shape]{format}{label}{sep}{before-code}[after-code]
-
-% \usepackage{caption,subcaption}
-% \DeclareCaptionSubType[Alph]{figure}
-% \captionsetup[subfigure]{labelformat=simple,labelfont={sf,bf,large},textfont={sf}}
-% \renewcommand{\thesubfigure}{{\sffamily\alph{subfigure}}}
-
-%% Spacing
-% \usepackage[doublespacing]{setspace}
-% \parindent 0pt
-% \parskip 1ex
-
-%% Citations, refs, hyperlinks
-\usepackage[colorlinks=true, allcolors=blue]{hyperref}
-\usepackage{natbib}
-\usepackage{apalike}
-\usepackage{graphicx}
-% \usepackage{natbib}
-% \usepackage[capitalize]{cleveref}
-% \crefname{equation}{}{}
-% \crefname{Equation}{Equation}{Equations}
-
-%% Title
-
-\title{}
-\author{}
-% \usepackage{filemod} % access to file modification times
-% \date{\Filemodtoday{\jobname}}  % replace \jobname if compiling from wrapper
-
-% \usepackage{titling} % modify title(page) appearance
-%\pretitle{\begin{center}\LARGE}
-%\posttitle{\par\end{center}\vskip 0.5em}
-%\preauthor{\begin{center} \large \lineskip 0.5em \begin{tabular}[t]{c}}
-%\postauthor{\end{tabular}\par\end{center}}
-%\predate{\begin{center}\large}
-%\postdate{\par\end{center}}
-
-\def\ii#1\par{{\color{blue!40}\sl #1}\par}
-\def\iibf#1\par{{\color{blue!40}\sl\bfseries #1}\par}
-\def\iitem#1\par{\ii\begin{itemize}[nosep]\item #1 \end{itemize}\par}
-
-\begin{document}
-
-\noindent\textbf{Notes}
-
-(This page is not for submission)
-
-Required items:
-\begin{itemize}  
-  \item case for support (up to eight pages)
-  \item justification of resources (two pages)
-  \item data management plan (DMP) (up to three pages, data management plan template)
-        if you opt not to use the template for your plan, the topics listed in the template must be addressed in the DMP document you do provide
-  \item diagrammatic workplan (one page)
-  \item management structure (one page)
-  \item narrative CV (up to two pages per staff member)
-  \item community letters of support or demand (up to 10)  
-  \item proposal cover letter.
-\end{itemize}
-
-\tableofcontents
-
-\newpage
-
-% \maketitle
-
-\section{Case for support}
-
-\ii Instructions and guidance:
-  \url{https://www.ukri.org/wp-content/uploads/2021/09/BBSRC-150921-FundingOpp-BioinformaticsBiologicalResourcesFund-GuidanceDocument.pdf}
-
-\iibf Assessment criteria.
-
-\ii \url{https://www.ukri.org/wp-content/uploads/2021/09/BBSRC-150921-FundingOpp-BioinformaticsBiologicalResourcesFund-AssessmentCriteria.pdf}
-
-\iibf Scientific quality and strategic relevance of the resource.
-
-\ii Including:
-\begin{itemize}[nosep]
-    \item the extent to which the resource meets the highest international standards in resource provision in its field
-    \item how well the resource is demonstrated to be unique or complementary to other similar existing resources
-    \item the extent to which the resource addresses the research and policy priority areas of BBSRC.
-\end{itemize}
-
-\iibf Cost effectiveness, particularly considerations for long-term
-sustainability beyond BBSRC funding.
-
-\ii The extent to which:
-\begin{itemize}[nosep]
-    \item the resource delivers value for money relative to the anticipated scientific gains it will provide
-    \item long term sustainability options have been considered, addressed, and planned where appropriate, particularly for existing resources.
-\end{itemize}
-
-\iibf Potential for economic and social impact beyond the academic
-community.
-
-\ii Including:
-\begin{itemize}
-    \item the extent to which the outputs from the resource will contribute to knowledge and potential for economic return or social impact
-    \item how well the proposal has outlined methods of engagement and measures of success in developing milestones and timelines of associated activities.
-\end{itemize}
-
-\iibf Fit to the scope.
-
-\ii How well the proposal addresses the scope of the opportunity.
-
-\iibf Assessment criteria adapted to new or existing resources
-
-\ii To allow for a more nuanced assessment between new and existing
-resources, the use of the following assessment criteria will be
-adapted accordingly.
-
-\ii For new resources, these criteria will assess the ‘plans, potential, and promise’ of the resources.
-
-\iibf Quality of the overall arrangements for resource management, advisory functions, as well as user access and engagement
-
-\ii Including:
-\begin{itemize}
-    \item the extent to which the proposal has evidenced or planned interaction with relevant users and the broader research community to ensure the aims of the resource are realised and there is sufficient uptake and continued development
-    \item the extent to which adequate user access arrangements have been discussed and considered
-    \item the set-up of project management and advisory structures of the resource to ensure longevity in delivering the resource to a broad user base.
-\end{itemize}
-
-\iibf Need or demand, and potential benefit to the UK academic research community
-
-\ii The extent to which:
-\begin{itemize}
-    \item the community has demonstrated demand for the proposed resource, relative to the total community size (in particular, proposals for new resources should have consulted their prospective community prior to application)
-    \item the proposed resource will deliver and benefit the wider BBSRC community indicating how the proposed resource will help to deliver high-quality research.
-\end{itemize}
-
-\newpage
-\ii Section headings taken from guide.  We can omit some if clearly not necessary, but should do so with caution.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\subsection{Background to the Resource}
-\ii Introduction of the proposed resource, including its academic and wider 
-economic and societal context.
-
-\ii Overview of past and current resource(s) in the subject area in both the UK 
-and abroad, including any alternative community resources currently 
-available. You should indicate the community size of the intended resource 
-and how this relates to the field in which it operates.
-
-Experimental control and neural data analysis are currently highly
-disassociated.  Most neuroscientists collect their data and only afterward
-analyze it. This split has had dramatic consequences for the progress of
-neuroscience. First, most experimental control methods currently do not use
-advanced data analysis methods to enhance their functionality. Second, the vast
-majority of neural data analysis methods are offline and do not provide key
-functionality needed for online experimental control. Using advanced data
-analysis methods in experimental control will improve the sophistication of
-experiments that are currently possible, and generate new demands for neural
-data analysis methods that will generate advances in the field.
-
-A central goal of the software development proposed here is to address this
-problem, by providing Bonsai~(an experimental control software ecosystem;
-Figure~\ref{fig:bonsai} and Section~\ref{sec:bonsai}) with state-of-the-art online (and offline) data
-analysis methods (Figure~\ref{fig:proposedBonsaiExtensions} and 
-Section~\ref{sec:toBeAddedMachineIntelligenceFunctionality}).
+Inspired by \texttt{scikit-learn}, Bonsai will have three types of operators:
+estimators, predictors and transformers. Each of these operators will be
+instantiated by objects implementing the corresponding interface.
 %
-It is unfortunate that neuroscientists currently do not have software that
-allows them to easily compare the performance of different data analysis
-methods for the characterization of their datasets. The second central goal
-of this proposal is to address this limitation by adding to Bonsai a generic
-mechanism for comparing the performance of different methods for the analysis
-of specific datasets (Section~\ref{sec:comparisonOfMultipleMethods}).
+For example, consider the close-loop system for the control of a computer
+cursor with neural activity represented in Figure~\ref{fig:bci}a. The Bonsai
+graphical program for this system is given in Figure~\ref{fig:bci}b. The
+objects instantiating the SpikeThresholdDetector, the KiloSortSpikeSorter and
+GPFAlatentsExtractor will implement the transformer interface, and the object
+instantiating the TimeSeriesForestClassifier will implement the predictor
+interface (Figure~\ref{fig:bci}b, top).
+%
+The parameters for the SpikeThresholdDetector and the KiloSortSpikeSorter will
+be set manually by the user, as is commonly done in practice. The parameters of
+the GPFAlatentsExtractor will be learnt offline in an unsupervised way in a
+second pipeline (Figure~\ref{fig:bci}b, middle). Finally, the parameters of the
+TimeSeriesForestClassifier will also be learnt offline, but in a supervised
+way, in a third pipeline (Figure~\ref{fig:bci}b, bottom).
 
@@
    \begin{center}
-        \includegraphics[width=7in]{figures/bonsai.pdf}
-        \label{fig:bonsai}
+       \includegraphics[width=4in]{figures/bci.png}
+       \caption{Close-loop system for controlling a computer cursor with brain activity}
+     \label{fig:bci}
    \end{center}
@@
 
-\setcounter{figure}{1}
+\begin{comment}
+    The spike detector transforms membrane
+potentials into spike sequences. The spike sorter transforms spikes sequences
+into lists of spikes, one list per neuron. The latents extractor transforms
+multi-neuron lists of spikes into latent trajectories summarising the
+multi-neuron spiking activity.  The neural decoder predicts the next cursor
+position from the latent trajectories.
+\end{comment}
 
-\begin{figure}
-    \begin{center}
-        \includegraphics[width=4in]{figures/proposed_bonsai_extensions.png}
-        \caption{Proposed extensions to Bonsai}
-        \label{fig:proposedBonsaiExtensions}
-    \end{center}
-\end{figure}
+\begin{comment}
+middle, whose output will set
+As in this library,
+there will be three type of machine intelligence operators in Bonsai:
+%
+estimator operators for building and fitting models, predictor operators for
+making predictions and transformer operators for converting data. These
+operators will be implemented by objects exposing the corresponding interfaces.
+
+The estimator interface exposes a fit method for learning a model from training
+data.  This method is called with training data (e.g., supplied as two arrays
+X\_train and y\_train in supervised learning estimators). Its task is to run a
+learning algorithm and to determine model-specific parameters from the training
+data.  All supervised and unsupervised learning algorithms (e.g., for
+classification, regression or clustering) will be offered as objects
+implementing this interface. Machine learning tasks like feature extraction,
+feature selection or dimensionality reduction will also be provided as
+estimators.
+
+The predictor interface extends the notion of an estimator by adding a predict
+method that takes an array X\_test and produces predictions for X\_test, based
+on the learned parameters of the estimator (we call the input to predict
+“X\_test” in order to emphasise that predict generalises to new data). In the
+case of supervised learning estimators, this method typically returns the
+predicted labels or values computed by the model.
+
+Since it is common to modify or filter data before feeding it to a learning
+algorithm, the transformer interface defines a transform method. It takes as
+input some new data X\_test and yields as output a transformed version of
+X\_test. Preprocessing, feature selection, feature extraction and
+dimensionality reduction algorithms will be provided as transformers in Bonsai.
+
+Bonsai will provide the ability to compose new
+estimators from several base estimators. Composition mechanisms can be used
+to combine typical machine learning workflows into a single object which is
+itself an estimator, and can be employed wherever usual estimators can be used.
+Composition of estimators will be done with Pipeline objects that chain multiple estimators into a single one. This is useful
+since a machine learning workflow typically involves a fixed sequence of processing steps (e.g., feature extraction, dimensionality reduction, learning and making
+predictions), many of which perform some kind of learning. A sequence of N such
+steps can be combined into a Pipeline if the first N-1 steps are transformers;
+the last can be either a predictor, a transformer or both.
+
+\end{comment}
+
+\subsubsection{Extensibility}
+\label{sec:extensibility}
+
+\textcolor{red}{I will start this section by saying that we are not proposing
+to implement machine learning algorithms, as in \texttt{scipy.learn}, but to
+integrate into Bonsai existing machine intelligence algorithms.}
 
@@
 already extended Bonsai's functionality with several contributed packages
-(e.g., \ldots \textcolor{red}{Goncalo please reference the best contributed
-packages developed by the Bonsai community}). However, these packages need to
-be written in C\#, while most current neuroscience data analysis methods are
-written in Python, R or Matlab.  We propose to add to Bonsai capabilities to
-communicate with software written in these languages
-(Section~\ref{sec:bonsaiPythonMatlabCommunication}).  These capabilities will
-allow a large number of neuroscience data analysis methods to be easily
-integrated into the Bonsai ecosystem and will provide Bonsai users a large
-repertoire of advance machine learning methods for their behavioral and neural
-data analysis.
+(e.g., \ldots). However, these packages need to be written in C\#, while most
+current data analysis methods are written in Python, R or Matlab. We propose to
+add to Bonsai capabilities to communicate with software written in these
+languages.
 
-Bonsai is an excellent tool for reproducible data acquisition and
-experimental control. An experiment implemented in Bonsai can be replicated in
-a new laboratory by just sharing a Bonsai configuration file. With the addition
-of the proposed machine learning functionality Bonsai will extend this
-reproducibility to the domain of data analysis
-(Section~\ref{sec:reproducibility}).
+Fortunately, there exist software that allow C\# program
+to call and be called by programs in these other languages. For the
+communication of C\# programs with Python we will use
+Python.NET\footnote{\href{https://github.com/pythonnet/pythonnet}{https://github.com/pythonnet/pythonnet}},
+with R programs we will use
+R.Net\footnote{\href{https://github.com/rdotnet/rdotnet}{https://github.com/rdotnet/rdotnet}}
+and with Matlab we will use the built in Matlab .NET
+interface\footnote{\href{https://uk.mathworks.com/help/matlab/matlab\_external/using-net-from-matlab-an-overview.html}{https://uk.mathworks.com/help/matlab/matlab\_external/using-net-from-matlab-an-overview.html}}.
+In cases where these software cannot support some type of communication between
+C\# and another language we will use the messaging library
+ZeroMQ\footnote{\href{https://zeromq.org/}{https://zeromq.org/}}.
 
-\subsubsection{Bonsai}
-\label{sec:bonsai}
+Using the previous software, we will build C\# wrappers around machine
+intelligence programs written in other languages, so that these programs
+implement the Estimator, Predictor or Transformer interfaces and can be
+integrated into Bonsai as machine intelligence operators.
 
-Bonsai~\citep{lopesEtAl15,lopesAndMonteiro21} is a free and open-source visual
-programming language that emphasizes performance, flexibility, and ease-of-use,
-allowing scientists with no previous programming experience to quickly develop
-their own high-performance data acquisition and experimental control systems
-(Figure 1c). Bonsai combines a high-level event algebra for data streams with
-an integrated development environment (IDE) and a strong library of plugins
-supporting multiple hardware and software packages used by the biomedical
-research community (Figures 1a, 1b).
+\subsection{Machine intelligence functionality to be added to Bonsai}
+\label{sec:functionality}
 
-Bonsai has a large user base in the systems neuroscience community.
-Quantifications of this user base are provided in Figures 1d-1f.  In the last
-year alone, we estimate more than 1,000 new users have started to incorporate
-Bonsai into their experimental protocols across the world. The surprising rate
-of adoption of Bonsai in non-programmer experimental labs highlights the need
-for accessible programming tools that enable state-of-the-art technology but
-also allow researchers to stay in control and change their experimental
-paradigms.  Many open-source software tools are either inaccessible to
-non-programmers, or too constrained to be of general use outside their narrow
-domain of application. Bonsai has been successful because it has straddled this
-gap to some extent.
-
-The language has also helped to potentiate the growing wave of foundational
-open hardware initiatives, such as the OpenEphys \citep{siegleEtAl17} and UCLA
-Miniscope~\citep{caiEtAl16}, making it possible to quickly combine and
-integrate these tools into new experiments~\citep{buccinoEtAl18}.
-%
-Bonsai has been adopted in large neuroscience undertakings like the
-International Brain
-Laboratory\footnote{\href{https://www.internationalbrainlab.com/}{https://www.internationalbrainlab.com/}}
-and the Allen Institute for Brain
-Science\footnote{\href{https://alleninstitute.org/what-we-do/brain-science/}{https://alleninstitute.org/what-we-do/brain-science/}}.
-%
-More recently, Bonsai has started to expand outside the domain of neuroscience
-into biomedical research and biotechnology tool development, and even outside
-academia into public outreach and education programs.
-
-\subsection{Details of Resource}
-
-\ii The case for support should outline the full details of the resource and 
-associated work packages presented in the proposal.
-
-\iitem Indication whether the project proposed to develop a new
-resource or is in support of an existing one.
-
-Existing if general Bonsai enhancements.  If specific to analysis we
-may be able to argue for "new" if helpful, while still making
-reference to Bonsai installed base.
-
-\iitem Objectives for the proposal should be detailed including
-individual measurable targets against which the outcome of the work
-will be assessed. This should refer to the objectives set out in the
-Je-S proposal form.
-
-\iitem{Significant technical details for the development, maintenance or 
-enhancement of the resource must be clearly outlined and indicated 
-how this is of internationally exceptional quality. }
-
-\iitem{If applicable, outline any proposed research efforts and how they 
-  directly facilitate development of the resource. }
-
-\iitem{For proposals looking to focus on maintaining status quo for an 
-existing resource instead of suggesting further development, you 
-should detail evidence of why significant upgrades are not required at 
-this time and detail why the resource needs continued support to 
-maintain world-leading functionality.}
-
-\ii{Additional questions that may be considered: }
-
-\iitem Does the facility begin or continue to support a growing field of 
-bioscience, what is the anticipated growth and does the proposal 
-adequately accommodate this?
-
-\iitem How will the resource accelerate science within its field and
-  beyond?
-
-\iitem What would be the impact on the scientific community if the resource 
-did not exist? How would this impact other, possibly dependent 
-resources?
+We propose to initially add to Bonsai machine intelligence functionality to
+support neuroscience experiments because, currently, neuroscience is the
+research area with the largest number of Bonsai users. Within neuroscience, we
+propose to add functionality for video-based behavioural analysis
+(Section~\ref{sec:videoBasedBehavioralAnalysis}) and for running
+brain-computer-interface experiments (Section~\ref{sec:bci}), since these tasks
+are relevant to many Bonsai users, and recent advances in machine learning
+will greatly improve their implementation in Bonsai.
 
@@
 
-\subsubsection{Low-dimensional representations of local-field potentials}
-\label{sec:lowDimensionalRepresentationsOfLFPs}
+\paragraph{Low-dimensional representations of local-field potentials}
+% \label{sec:lowDimensionalRepresentationsOfLFPs}
 
-Spikes are extracted from a higher-frequency range of extracellulary recorded
+Spikes are extracted from a higher-frequency range of extracellularly recorded
 voltages. Local field potentials (LFPs) are computed from a lower-frequency
@@
 
-\subsubsection{Interfacing Bonsai with Python, R and Matlab data analysis software}
+\begin{comment}
+
+\subsection{Interfacing Bonsai with Python, R and Matlab data analysis software}
 \label{sec:bonsaiPythonMatlabCommunication}
@@
 
-\subsubsection{Reproducibility with Bonsai}
+\end{comment}
+
+\subsection{Reproducibility with Bonsai}
 \label{sec:reproducibility}
@@
 
-\subsubsection{Final remarks}
+\subsection{Measurable targets}
+\label{sec:measurableTargets}
+
+We propose to assess the progress of the proposed project using the following
+measurable targets:
+
+\begin{comment}
+
+\begin{description}
+
+    \item[implementation of software infrastructure]
+        (Section~\ref{sec:softwareInfrastructure}) includes the
+        implementation of the machine abstractions, described in
+        Section~\ref{},
+        and of the capabilities to allow Bonsai communicate with
+        software written in other languages, described in Section~\ref{}. We
+        will consider this target completed when we can implement a simple BCI
+        application, with the architecture depicted in Figure~\ref{}, with
+        machine learning functionality provided by software in Python, Matlab
+        and R.
+
+    \item[implementation of functionality] (Sections~\ref{sec:functionality})
+        includes the integration into Bonsai of machine learning software for
+        video-based behavioral analysis (Section~\ref{}) and for
+        brain-computer-interface applications (Section~\ref{}). Completion of
+        this target requires that Bonsai can perform each of these tasks
+        separately, in conjunction with software written in Python, R and
+        Matlab.
+
+    \item[internal testing] multiple groups at the SWC use Bonsai to control
+        their experiments.
+
+        Bonsai Github repository and the success use by Bonsai users
+
+    \item Bonsai users use the implemented machine learning functionality. We
+        propose to evaluate this target by checking in the Bonsai user forum
+        the frequency with which a novel machine learning functionality is
+        discussed, and by tracking publications using this functionality.
+
+    \item method developers integrate their methods into the Bonsai ecosystem.
+        The integration of machine learning functionality into Bonsai will use
+        Bonsai's package manager. We propose to use this package manager to tack
+        the number of machine intelligence methods integrated into Bonsai.
+
+\end{description}
+
+\end{comment}
+
+\subsection{Final remarks}
 \label{sec:finalRemarks}
@@
 before distributing it to the general public.
-
-\subsection{Community Demand}
-\ii Evidence of community demand should be primarily driven from UK 
-academic researchers working largely within BBSRC remit – see our 
-Forward look for UK Bioscience for more detail on research areas covered by 
-BBSRC. Demand from other users (such as academic communities outside 
-of BBSRC remit or industrial users) may be appropriate to provide additional 
-support, especially in highlighting the potential for economic, commercial or 
-societal impact, but should not be the focus of the demand demonstration. 
-Evidence provided should highlight examples of the high-quality science that 
-the resource will underpin or has underpinned. Where possible and relevant, 
-examples should be drawn from a wide research community to illustrate the 
-broad impact of the resource to support high-quality internationally excellent 
-science.
-
-\ii The level of community demand should be benchmarked against other 
-relevant resources and/or the size of the community. This will allow the fair 
-assessment of resources with different user bases. The types of evidence 
-that may be appropriate to provide will be different for new and existing 
-resources.
-
-\ii Evidence of wider consultation of the prospective community (e.g community 
-surveys) is encouraged.
-
-\iibf New Resources
-
-\ii New resources should estimate the number of researchers who may engage and benefit from the resource. Evidence, where possible, would be of benefit and may include: 
-\begin{itemize}
-    \item  Datasets (or samples) in public or private repositories 
-    \item Citations or acknowledgments 
-    \item Gap analysis with existing resources 
-    \item Pilot project uptake or feedback from potential users.  
-\end{itemize}
-
-\ii In particular, proposals for new resources should have consulted their prospective community prior to application
-
-\iibf Existing Resources
-
-\ii For existing resources this should include usage data of the
-current resource. Data types may include:
-
-\iitem Access requests from independent users/ sites
-
-\iitem Citations or acknowledgements
-
-\iitem Other public resources providing links to the resource
-
-\iitem New major acquisitions captured by the resource
-
-\ii{Additionally, existing resources need 
-to evidence why this resource needs 
-to be maintained/updated by the 
-current grant. This could include: }
-
-\iitem Survey data from users on what upgrades are needed
-
-\iitem Evidence of an expanding 
-user base, which requires 
-additional resource
-
-\iitem Recent developments in the 
-field, which require upgrades 
-to be integrated into the 
-resource.
-
-\subsection{User engagement}
-\ii Discussion of user engagement provision should aim to answer the following 
-questions:
-
-\iitem Is there awareness of the resource within the user community?
-
-\iitem How do you plan to develop the engagement strategy within the 
-proposal timeframe to expand user engagement?
-
-\iitem How have access mechanisms to ensure usability of the resource 
-been considered?
-
-\iitem How have user needs been incorporated into this proposal to ensure 
-it is fit for purpose and will deliver on expectations?
-
-\iibf New Resources
-
-\ii Evidence should be provided as to how the resource plans to engage
-stakeholders and ensure that the resource meets their needs and is
-used by the community targeted by the resource.
-
-\iibf Existing Resources
-
-\ii Evidence should be provided as to whether the resource has
-achieved the level of engagement it originally anticipated, and
-consideration is given how the additional investment would change
-this.
-
-\subsection{Long term sustainability planning}
-\ii{
-In addition, the case for support should outline considerations for the long-
-term sustainability of the resource beyond UKRI-BBSRC funding, as well as 
-the true cost of running and maintaining the resource in question. 
-The proposal should include: }
-
-\iitem Cost recovery plans, where appropriate, or an explanation why
-not if not viable. Evidence of clear business planning with a focus on
-at least partial cost recovery is required, especially when applying
-as an existing resource.
-
-\iitem Details for alternative support plans, aside from UKRI-BBSRC 
-funding.
-
-\iitem The level of support the resource is projected to require for expected 
-maintenance and/or subsequent maturation/enhancement activities. 
-Clear arguments as to why UKRI-BBSRC should support the resource now 
-should be provided, if other cost recovery and support plans are deemed 
-unsuitable.
-
-\subsection{Potential for economic and societal impact}
-\ii{Outline how the outputs of the proposed resource will contribute to 
-knowledge and how this may have the potential for economic return or 
-societal benefits. Impact activities should be integrated into appropriate 
-sections of the case for support, not presented as an independent work 
-package. }
-
-\iitem All proposals are expected to demonstrate clear plans with recorded 
-milestones and timelines for associated activities to develop 
-economic, commercial and societal impacts.  
-August 2021 7
-
-\iitem Methods of engagement and measures of success should be outlined 
-including how these will be regularly reviewed throughout the project 
-in order to deliver the most impact.
-
-\iitem Any planned activities should be fully justified within the Justification 
-of Resources attachment
-
-\subsection{Track Record}
-\ii{
-The majority of the track record relevant to the project should be located 
-within the Narrative CV and should not be repeated within the case for 
-support. You may, however, want to describe: }
-
-\iitem Track record of the team working together
-
-\iitem Specific expertise available at the host organisation and any 
-proposed partner organisations to enable the successful delivery of 
-the project.
-
-\iitem The specific role of each applicant and collaborator in the project
-
-\bibliographystyle{apalike}
-\bibliography{bonsai,epsrcSoftwareForResearchCommunities,monitoringBehavior,stateSpaceModels,machineLearning,gaussianProcesses}
-
-\end{document}

### 2021-11-14 backgroundToTheResource.tex
Minor changes

@@
 (e.g., \ldots \textcolor{red}{Goncalo please reference the best contributed
-packages developed by the Bonsai community}). However, these packages need to
+packages developed by the Bonsai user community}). However, these packages need to
 be written in C\#, while most current neuroscience data analysis methods are
@@
 Developers of advanced data-analysis methods should become interested in
-integrating them into the Bonsai ecosystem, which will allow their methods to
-reach the wide Bonsai community. In addition this integration will enable
+integrating them into the Bonsai ecosystem, as it will allow their methods to
+reach the wide Bonsai user community. In addition this integration will enable
 method developers to easily compare the performance of their methods with

### 2021-11-14 communityDemand.tex
Added reference to focus on neuro applications

@@
 \citep[e.g.,][]{walterAndCouzin21,guilbeaultEtAl21}. However, these graphical
-user interfaces have no interact with real-time experimental control, and
+user interfaces have no interaction with real-time experimental control, and
 their support libraries require expertise with the Python programming language
@@
 flexibility for experimental control.
+
+We focused this proposal on neuroscience applications of Bonsai. However, the
+addition of machine intelligence functionality is needed in many other
+experimental areas where Bonsai is or could be used. Therefore, the software
+abstractions that we will build to add to Bonsai machine intelligence
+functionality for neuroscience experiments (e.g., methods to perform
+multi-method comparisons) will translate to other application domains of
+Bonsai.
 

### 2021-11-14 detailsOfResource.tex
A few relevant updates (e.g.; added targets, relocated final remarks, mentioned two alternatives to incorporate ML functionality into Bonsai)

@@
 We intend to improve an \textbf{existing resource}, the Bonsai ecosystem. We
-propose to create software infrastructure to facilitate the integration into it
+propose to create software infrastructure to facilitate the integration into
+Bonsai
 of existing advanced data analysis methods
-(Section~\ref{sec:softwareInfrastructure}).
+(Section~\ref{sec:infrastructure}).
 %
@@
 \subsection{Software infrastructure}
-\label{sec:softwareInfrastructure}
+\label{sec:infrastructure}
 
@@
 continuously increasing. We intend to build a small number of machine
-intelligence operators in Bonsai that can be configured to provide the
-functionality of most machine learning algorithms.
+intelligence operators in Bonsai (Section~\ref{sec:bonsai},
+Figure~\ref{fig:bonsai}c) that can be configured to provide the functionality
+of all machine learning algorithms.
 
@@
 %
-This API consists of three complementary interfaces: an estimator interface for
-building and fitting models, a predictor interface for making predictions and a
-transformer interface for converting data.
+This API consists of three complementary interfaces: an \texttt{Estimator}
+interface for building and fitting models, a \texttt{Predictor} interface for
+making predictions and a \texttt{Transformer} interface for converting data.
 %
@@
 Inspired by \texttt{scikit-learn}, Bonsai will have three types of operators:
-estimators, predictors and transformers. Each of these operators will be
-instantiated by objects implementing the corresponding interface.
+\texttt{Estimators}, \texttt{Predictors} and \texttt{Transformers}. Each of
+these operators will be instantiated by objects implementing the corresponding
+interface.
 %
-For example, consider the close-loop system for the control of a computer
+As an example, consider the close-loop system for the control of a computer
 cursor with neural activity represented in Figure~\ref{fig:bci}a. The Bonsai
@@
 objects instantiating the SpikeThresholdDetector, the KiloSortSpikeSorter and
-GPFAlatentsExtractor will implement the transformer interface, and the object
-instantiating the TimeSeriesForestClassifier will implement the predictor
-interface (Figure~\ref{fig:bci}b, top).
+GPFAlatentsExtractor will implement the \texttt{Transformer} interface (blue),
+and the object instantiating the TimeSeriesForestClassifier will implement the
+\texttt{Predictor} interface (green; Figure~\ref{fig:bci}b, top).
 %
@@
 be set manually by the user, as is commonly done in practice. The parameters of
-the GPFAlatentsExtractor will be learnt offline in an unsupervised way in a
-second pipeline (Figure~\ref{fig:bci}b, middle). Finally, the parameters of the
-TimeSeriesForestClassifier will also be learnt offline, but in a supervised
-way, in a third pipeline (Figure~\ref{fig:bci}b, bottom).
+the GPFAlatentsExtractor will be learnt offline in an unsupervised way by an
+\texttt{Estimator} (red) in a second pipeline (Figure~\ref{fig:bci}b, middle).
+Finally, the parameters of the TimeSeriesForestClassifier will also be learnt
+offline, but in a supervised way, by another \texttt{Estimator} in a third
+pipeline (Figure~\ref{fig:bci}b, bottom).
 
@@
 
-\textcolor{red}{I will start this section by saying that we are not proposing
-to implement machine learning algorithms, as in \texttt{scipy.learn}, but to
-integrate into Bonsai existing machine intelligence algorithms.}
+To incorporate machine learning functionality into the Bonsai ecosystem, we
+could
+implement this functionality directly into Bonsai, as done in
+\texttt{scikit-learn}. Alternatively, or in addition, we could build software
+infrastructure to facilitate the integration of existing machine learning
+software into Bonsai. We chose to begin with the second alternative because
+high-quality implementations of machine learning functionality already exist,
+which can be quickly integrated into the Bonsai ecosystem. In addition, the
+second alternative is more sustainable, as it will allow the continued addition
+of new machine learning functionality to Bonsai after the completion of the
+BBSRC-BBR funding period. Later we may decide to implement some widely-used
+machine learning functionality directly into Bonsai (e.g., for performance or
+robustness issues).
 
@@
 intelligence programs written in other languages, so that these programs
-implement the Estimator, Predictor or Transformer interfaces and can be
-integrated into Bonsai as machine intelligence operators.
+implement the \texttt{Estimator}, \texttt{Predictor} or \texttt{Transformer}
+interfaces and can be integrated into Bonsai as machine intelligence operators.
 
@@
 \paragraph{Spike sorting}
-% \label{sec:spikeSorting}
+\label{sec:spikeSorting}
 
@@
 \paragraph{Low-dimensional representations of spiking data}
-% \label{sec:lowDimensionalRepresentationsOfSpikingData}
+\label{sec:lowDimensionalRepresentationsOfSpikingData}
 
@@
 \paragraph{Low-dimensional representations of local-field potentials}
-% \label{sec:lowDimensionalRepresentationsOfLFPs}
+\label{sec:lowDimensionalRepresentationsOfLFPs}
 
@@
 \paragraph{Neural decoding}
-% \label{sec:neuralDecoding}
+\label{sec:neuralDecoding}
 
@@
 \paragraph{Comparing multiple data-analysis methods}
-% \label{sec:comparisonOfMultipleMethods}
+\label{sec:comparisonOfMultipleMethods}
 
-We propose to add functionality to Bonsai to facilitate multi-method
-comparisons in user-supplied datasets. Below we describe a comparison that we
+% We propose to add functionality to Bonsai to facilitate multi-method
+% comparisons in user-supplied datasets.
+
+Below we describe a comparison that we
 will perform in Bonsai to asses the relative performance of the methods
-described in the previous sections with neural recordings from behaving rats.
+described in the previous sections, using neural recordings from behaving rats.
 These recordings are currently being performed, to address scientific questions,
@@
 
+\subsection{Testing with neuroscience data}
+\label{sec:testingWithNeuroscienceData}
+
+Multiple groups at the SWC currently use Bonsai for data acquisition and
+experimental control. Carefully testing the functionality of new software is
+essential to ensure its correct functionality. We propose to use the SWC large
+internal Bonsai user base to test the functionality we will add to Bonsai,
+before distributing it to the general public.
+
 \subsection{Reproducibility with Bonsai}
@@
 
-We propose to assess the progress of the proposed project using the following
-measurable targets:
+We propose to assess the progress of the proposed project using three different
+types of targets.
+%
+First, we will have targets associated with each release to the general public
+of the software components described in Sections~\ref{sec:infrastructure}
+and~\ref{sec:functionality}.
+%
+Second, we will monitor user activity in the Bonsai
+forum\footnote{\href{https://groups.google.com/g/bonsai-users}{https://groups.google.com/g/bonsai-users}},
+and publications citing Bonsai that use th new machine larning functionality,
+to estimate the number of Bonsai users using this functionality.
+%
+Third, we will count the number of machine learning methods that are integrated
+into the Bonsai ecosystem by us and by other method developers.
 
@@
     \item[implementation of software infrastructure]
-        (Section~\ref{sec:softwareInfrastructure}) includes the
+        (Section~\ref{sec:infrastructure}) includes the
         implementation of the machine abstractions, described in
@@
 
+\begin{comment}
+
 \subsection{Final remarks}
@@
 
-We focused this proposal on neuroscience applications of Bonsai. However, the
-addition of machine intelligence functionality is needed in many other
-experimental areas where Bonsai is or could be used. Therefore, the software
-abstractions that we will build to add to Bonsai machine intelligence
-functionality for neuroscience experiments (e.g., methods to perform
-multi-method comparisons) will translate to other application domains of
-Bonsai.
-
-Multiple groups at the SWC currently use Bonsai for data acquisition and
-experimental control. Carefully testing the functionality of new software is
-essential to ensure its correct functionality. We propose to use the SWC large
-internal Bonsai user base to test the functionality we will add to Bonsai,
-before distributing it to the general public.
+\end{comment}

### 2021-11-15 detailsOfResource.tex
Started to fill the spike sorting section

@@
 
+We propose to add functionality into Bonsai for the implementation of
+brain-computer-interface applications illustrated in Figure~\ref{fig:bci}. In
+this implementation voltage recordings from the brain will be converted into
+spikes fired by single neurons by a \texttt{SpikeSorter}
+(Section~\ref{sec:spikeSorting}), or to local field potentials (LFP) by an
+\texttt{LFPextractor} (Section~\ref{sec:lfpCalculation}). Next, neural activity
+(spikes or LFPs) will be represented in a low-dimensional latent space by a
+\texttt{LatentsExtractor}. Finally, a \texttt{Decoder} will extract the
+intended behavior of the animal from this low-dimensional representation.
+
 \paragraph{Spike sorting}
@@
 
-\noindent\textcolor{red}{TODO}
+Spike sorting is an essential computational step needed to isolate the activity
+of individual neurons, or units, within extracellular recordings which combine
+noisy signals from many neurons. Most spike sorting algorithms are designed to
+work offline (i.e., identify individual neurons from a complete recording
+session, after this session has been completed). We propose to integrate into Bonsai
+
+\paragraph{LFP calculation}
+\label{sec:lfpCalculation}
+
+Spikes are extracted from a higher-frequency range of extracellularly recorded
+voltages. Local field potentials (LFPs) 
+are another important signal to understand brain function, which is extracted
+from a lower-frequency range of these voltages. We propose to use in-house
+functions implemented in Python to compute LFPs.
 

### 2021-11-15 detailsOfResource.tex
Updates with spike sorting

--- 2021-11-14 detailsOfResource.tex
@@
 (Section~\ref{sec:videoBasedBehavioralAnalysis}) and for running
-brain-computer-interface experiments (Section~\ref{sec:bci}), since these tasks
+brain-computer-interface (BCI) experiments (Section~\ref{sec:bci}), since these tasks
 are relevant to many Bonsai users, and recent advances in machine learning
@@
 
-\paragraph{Spike sorting}
-\label{sec:spikeSorting}
+We propose to add functionality into Bonsai for the implementation of
+brain-computer-interface (BCI) applications illustrated in Figure~\ref{fig:bci}. In
+this implementation voltage recordings from the brain will be converted into
+spikes fired by single neurons by a \texttt{SpikeDetector}
+(Section~\ref{sec:spikeDetection}), or to local field potentials (LFP) by an
+\texttt{LFPextractor} (Section~\ref{sec:lfpExtraction}). Next, neural activity
+(spikes or LFPs) will be represented in a low-dimensional latent space by a
+\texttt{LatentsCalculator}. Finally, a \texttt{Decoder} will extract the
+intended behavior of the animal from this low-dimensional representation.
 
-\noindent\textcolor{red}{TODO}
+\paragraph{Spike detection}
+\label{sec:spikeDetection}
 
-\paragraph{Low-dimensional representations of spiking data}
-\label{sec:lowDimensionalRepresentationsOfSpikingData}
+Spikes from multiple neurons can be detected in recordings from a single
+electrode inserted in neural tissue. Spike sorting is the computational step
+used to assign each spike to the neuron that generated it. Most spike sorting
+algorithms are designed to work offline (i.e., to use a complete recording
+session, after this session terminated). Online spike sorting (i.e., the task
+of assigning spikes to individual neurons as recordings are being aquired) is a
+still unsolved task, specially for recording devices with large number of
+electrodes.
+%
+However, for the type of BCI application proposed here, previous studies have
+shown that spike sorting is not beneficial and high performance can be retained
+by just detecting spikes, without assigning them to invidivual
+neurons~\citep{trautmannEtAl19,todorovaEtAl14}. Thus, for the BCI application
+proposed here we will only detect spikes with a simple zero-corssing method,
+and not assign them to individual neurons.  We will, however, integrate into
+Bonsai two offline spike sorting methods, Kilosort~\citep[][written in Matlab]{pachitariuEtAl16} and MountainSort~\citep[][written in
+Python]{chungEtAl17}, to provide Bonsai users spike sorting functionality for
+applications that benefit from it.
+
+\paragraph{LFP extraction}
+\label{sec:lfpExtractor}
+
+Spikes are extracted from a higher-frequency range of extracellularly recorded
+voltages. Local field potentials (LFPs) 
+are another important signal to understand brain function, which is extracted
+from a lower-frequency range of these voltages. We propose to use in-house
+functions implemented in Python to compute LFPs.
+
+\paragraph{Low-dimensional representations of neural recordings}
+\label{sec:lowDimensionalRepresentation}
 
@@
 to use states inferred from LFPs by GLDS, GPFA, HMM and LFADS
-(Section~\ref{sec:lowDimensionalRepresentationsOfSpikingData}) as
+(Section~\ref{sec:lowDimensionalRepresentation}) as
 low-dimensional representations of the LFP.
@@
 methods described in
-Sections~\ref{sec:lowDimensionalRepresentationsOfSpikingData}
+Sections~\ref{sec:lowDimensionalRepresentation}
 and~\ref{sec:lowDimensionalRepresentationsOfLFPs}. We will also learn the
@@
         video-based behavioral analysis (Section~\ref{}) and for
-        brain-computer-interface applications (Section~\ref{}). Completion of
+        BCI applications (Section~\ref{}). Completion of
         this target requires that Bonsai can perform each of these tasks

### 2021-11-15 application/bonsai.tex
Moved files to application dir

--- 2021-11-12 bonsai.tex
@@
 allowing scientists with no previous programming experience to quickly develop
-their own high-performance data acquisition and experimental control systems
-(Figure 1c). Bonsai combines a high-level event algebra for data streams with
-an integrated development environment (IDE) and a strong library of plugins
-supporting multiple hardware and software packages used by the biomedical
-research community (Figures 1a, 1b).
+their own high-performance data acquisition and experimental control systems.
+Bonsai combines a high-level event algebra for data streams with an integrated
+development environment (IDE) and a strong library of plugins supporting
+multiple hardware and software packages used by the biomedical research
+community (Figures~\ref{fig:bonsai}a and~\ref{fig:bonsai}b).
+%
+A Bonsai graphical program consists of one or more source data streams
+and several interconnected operators that transform input to output
+datastreams (Figure~\ref{fig:bonsai}c).
 
@@
 animal behaviour experiments are still lacking. Life science experiments demand
-a combination of multiple instrumentation and control technologies, in both
-physiology and behaviour studies. The growing complexity on both the amount of
-data that is collected, and the rich conditions under which behaviour must be
-explored, place an increasing burden on the experimenter to integrate highly
-specialized equipment in unique configurations, while often lacking expertise
-in the relevant engineering fields.
+a combination of multiple instrumentation and control technologies, for both
+behavioral and physiological investigations. The growing complexity on both the
+amount of data that is collected, and the rich conditions under which behaviour
+must be explored, place an increasing burden on experimenters to integrate
+highly specialized equipment in unique configurations, while often lacking
+expertise in the relevant engineering fields.
 
-Bonsai has a large user base in the systems neuroscience community.
-Quantifications of this user base are provided in Figures 1d-1f.  In the last
-year alone, we estimate more than 1,000 new users have started to incorporate
-Bonsai into their experimental protocols across the world. The surprising rate
-of adoption of Bonsai in non-programmer experimental labs highlights the need
-for accessible programming tools that enable state-of-the-art technology but
-also allow researchers to stay in control and change their experimental
-paradigms.  Many open-source software tools are either inaccessible to
-non-programmers, or too constrained to be of general use outside their narrow
-domain of application. Bonsai has been successful because it has straddled this
-gap to some extent.
+Bonsai has a large user base in the systems neuroscience community
+(Figures~\ref{fig:bonsai}d-\ref{fig:bonsai}f).  In the last year alone, we estimate more than 1,000 new users have
+started to incorporate Bonsai into their experimental protocols across the
+world. The surprising rate of adoption of Bonsai in non-programmer experimental
+labs highlights the need for accessible programming tools that enable
+state-of-the-art technology but also allow researchers to stay in control and
+change their experimental paradigms.  Many open-source software tools are
+either inaccessible to non-programmers, or too constrained to be of general use
+outside their narrow domain of application. Bonsai has been successful because
+it has straddled this gap to some extent.
 
@@
 These dedicated interfaces are typically very comfortable for experimenters
-operating within the specific domain that the tool is designed for but tend to
+operating within the specific domain that the tool is designed for, but tend to
 become unwieldy when there is a need to introduce a new device or variation of

### 2021-11-15 application/userEngagement.tex
Moved files to application dir

In the past four years we have developed the bonsai-rx.org website with
documentation on how to use the language, and as a central hub for learning
resources, such as courses, video tutorials and examples. We have also engaged
with leading training efforts in the field such as the CAJAL Neuroscience
Training Programme (Bonsai 0121) and the Transylvania Experimental Neuroscience
Summer School (TENSS)  to increase awareness of the availability of Bonsai and
its packages throughout the neuroscience community. Together with participation
at smaller invited venues, there has been an average of 5 courses in Bonsai
each year, with at least 20 students each, so about 100 neuroscience students
every year get their introduction to experimental tools directly through a
course in the Bonsai programming language.

The development of this new machine learning package will be deeply integrated
into the Bonsai ecosystem which will complete its 10th anniversary in 2022, and
will thus leverage planned venues, events, and infrastructure to disseminate
awareness throughout the community, including presentations at conferences,
workshops, and training sessions, as well as electronic newsletter, forums and
social media articles.

The new infrastructure that we will build on Bonsai to support the machine
learning package will facilitate interfacing with other programming languages
such as Python, R, and MATLAB, which will therefore attract a whole new
community of users to Bonsai with expertise in software development. To take
advantage of this new community and provide to them the full power of Bonsai,
we plan to organize hackathons to accelerate integration of their existing data
analysis and machine learning algorithms and methods into Bonsai using the new
machine learning package.

 Given the modular and integrative nature of the machine learning package, we
 plan to advertise its development early in the project, so that interested
 partners have the opportunity to contribute to the requirements of key
 components of the platform, such as the Python and MATLAB integrations. We
 also want to promote the co-development of algorithms by early adopters of the
 package, in true collaborative fashion. Given its modular nature and built-in
 package manager capable of handling dependency resolution and curation of
 community contributed content, Bonsai is well positioned for such an approach.

### 2021-11-16 application/workPlan.tex
Added

\documentclass[a4paper,11point]{article}
\usepackage[margin=2cm]{geometry}
\usepackage{graphicx}

\makeatletter
\setlength{\@fptop}{0pt}
\makeatother

\author{}
\date{}

\begin{document}

\begin{figure}[!t]
   \begin{center}
       \includegraphics[width=7in]{figures/workPlan_coarse.png}

       \caption{Gnatt chart for major project tasks. The label next to the task
       name refers to the section of \emph{Case for Support} document where the
       task is described}

     \label{fig:gnattChart}
   \end{center}
\end{figure}

\end{document}

### 2021-11-16 application/caseForSupport.tex
Moved files to application dir
Updates from Overleaf

--- 2021-11-12 caseForSupport.tex
@@
 \usepackage[margin=2cm]{geometry}
+%% Language and font encodings
+\usepackage[english]{babel}
+\usepackage[utf8x]{inputenc}
+\usepackage[T1]{fontenc}
+% \usepackage{microtype}      % fine pdf font control
+
+\usepackage[scaled=0.92]{helvet} \renewcommand{\familydefault}{\sfdefault}
+\usepackage{xcolor} % -- loaded by tikz
 
@@
 \usepackage{verbatim}
+
+\usepackage{wrapfig}
 
@@
 \setcounter{tocdepth}{4}    % how many sectioning levels to show in ToC
+\usepackage[small]{titlesec} % format section titles
 
@@
 \setcounter{page}{1}
-
-\section{Background to the resource}
 
@@
 
+\input resourceManagement
+
 \section{Long term sustainability planning}
@@
 \section{Potential for economic and societal impact}
+\textbf{Understanding the brain.} Systems neuroscientists aim to uncover how adaptive behaviour arises
+from computation in neural circuits in different brain regions.
+Extending the Bonsai ecosystem will facilitate experiments that connect neural circuits with elemental behavioural processes.
+Long-term, this research will help identify the missing links between circuit dysfunction and behavioural deficits in neurological or mental health disorders, and will pave the way for the development of new neural circuit biomarkers and therapies.
+The Bonsai ecosystem will also enable the experimental testing of emerging theories of brain function and learning, and thus contribute to the neuro-inspired machine learning algorithms that fuel intelligent technologies, increasingly important in our society.
+
+\textbf{Improve reproducibility in behavioural and brain sciences.} Lack of reproducible results slows down scientific progress and incurs personal and economic costs to scientists and funders.
+In neuroscience,  many behavioural and neural animal studies have been to be difficult to reproduce, and experiments are rarely replicated across laboratories.
+By creating a resource that standardises and enables reproducible experimental design/control, as well as analysis of neural and behavioural data, the Bonsai ecosystem will allow multiple laboratories to implement and align experimental and analytical protocols as closely as possible.
+
+\textbf{Improve efficiency of research.} Most neuroscience laboratories worldwide develop custom, unique solutions to their experimental control and analysis needs.
+Some laboratories also lack the technical know-how to either develop their own software for experimental hardware control or write software for the state-of-the-art analysis of rich neural and behavioural data, increasingly reliant on latest machine learning algorithms.
+The proposed upgrading of the open-source and community empowered Bonsai ecosystem with a toolbox of online and offline analytical tools, that can dovetail with closed-loop experimental requirements, will (i) allow the deployment of such tools into laboratories that lack the expertise, (ii) reduce the need to reinvent the same tools in multiple labs (cost/time efficiency) and (iii) concomitantly streamline and standardise the experimental data processing streams in systems neuroscience.
 
@@
 \bibliographystyle{apalike}
-\bibliography{bonsai,epsrcSoftwareForResearchCommunities,monitoringBehavior,stateSpaceModels,machineLearning,gaussianProcesses}
+\bibliography{bonsai,epsrcSoftwareForResearchCommunities,monitoringBehavior,stateSpaceModels,machineLearning,gaussianProcesses,spikeSorting}
 

### 2021-11-16 application/detailsOfResource.tex

--- 2021-11-15 detailsOfResource.tex
@@
 We intend to improve an \textbf{existing resource}, the Bonsai ecosystem. We
-propose to create software infrastructure to facilitate the integration into
-Bonsai
-of existing advanced data analysis methods
-(Section~\ref{sec:infrastructure}).
+propose to create software infrastructure to enable the\textbf{ integration of advanced data analysis methods} into
+Bonsai (Section~\ref{sec:infrastructure}).
 %
@@
 machine-learning functionality into the experimental control loop. We propose
-user engagement activities to expose both of these groups of people to this
-potential (Section~\ref{sec:userEngagement}).  When data-analysis method
-developers recognise this potentials, they will become interested in
+user engagement activities to expose both of these groups to this
+potential (Section~\ref{sec:userEngagement}).  As data-analysis method
+developers recognise this opportunity, they will become interested in
 contributing their methods to the Bonsai ecosystem, ensuring the
@@
 
-For this, we will build on abstractions used in the \texttt{sickit-learn}
+For this, we will build on abstractions used in the \texttt{scikit-learn}
 library. The success of this library is, to a large extent, due to its simple

### 2021-11-16 application/resourceManagement.tex

\textcolor{red}{I'm making this up -- Goncalo please comment/correct.}
%
As an open source environment developed over almost 10 years, Bonsai 
has well-established user-community-based management structures.
%
The core system and the majority of extension packages are hosted on a
publicly accessible website at (\url{bonsai-rx.org}.
%
Package installation and updates are supported by standalone package
management software included within the core Bonsai distribution.
%
This ensures that bugfixes and enhancements are easily available but
under full user control (to ensure working experiments are not
disrupted by changes in the system).

Best-practice code review procedures are followed prior to including
package updates.  \textcolor{red}{Goncalo/Joaquin -- could you expand
  on this?}

\textcolor{red}{Testing}

The software engineers recruited to this project will work with our
collaborators at Neurogears Ltd, to strengthen and maintain these
practices.  Once fully familiar with the Bonsai core archtecture they
will participate in review of pull requests and other code
contributions.

Long-term sustainability

relies on distributed peer-reviewed management.  The collaborator

To complement the management structure document, proposals may outline
how the resource management and advisory structures will operate including
outlining any review procedures.
Additionally, opportunities for staff training and support to ensure their
continuous development should be identified to safeguard the successful
delivery of the resource.

open source

user testing

workshops, user panels

### 2021-11-16 application/Je-S_sections.tex

List the main objectives of the proposal in order of priority [up to
  4000 chars]

The overall goal of this proposal is to extend, establish, distribute and support Bonsai, a fully integrated software
environment to enable cutting-edge reproducible laboratory systems neuroscience experiments in animal models.

Specifically, we will:

1. Develop, test and deploy a new software pipeline to incorporate external packages into Bonsai's C# environment.

2. Incorporate a suite of machine-learning-driven analysis tools into the Bonsai environment to implement:
2a. video-based  behavioural analysis
2b. real-time interfacing of population-scale activity with laboratory control  
2c. evaluate competing data analytic approaches within a single experimental framework

3. Enhance the quality and reproducibility of systems neuroscience through:
3a. fully open-source and public-facing development
3b. rapid release of, and community engagement around, new tools
3c. provision of standard implementations of analysis and control methods, with the ability to distribute automated
descriptions of analysis pipelines for reproducibility.

4. Extend and support the Bonsai user community through code and forum engagement and development workshops.

[2825 chars]

Describe the Bioinformatic and/or Biological Resource in simple terms in a way that could be publicised to a general audience [up to 4000 chars]

Describe the Bioinformatic and/or Biological Resource in a manner suitable for a specialist reader. This summary will be made publicly available if the proposal is funded. [up to 2000 characters]

Describe who will benefit from this Bioinformatic and/or Biological resource(s) being funded [up to 4000 characters]

### 2021-11-16 application/backgroundToTheResource.tex

\section{Background: real-time neuroinformatics and control in behavioural systems
  neuroscience}

Modern systems neuroscience depends on data from carefully designed behavioural experiments in
animal models such as mice frequently combined with invasive electrophysiology and/or
cellular-resolution imaging.  Experimental design might require behavioural feedback (e.g. reward,
lights, sounds, actuated devices\dots) or neural pertubation (e.g. using light-sensitive polarising
molecules) that depends on real-time behaviour or neural state.
%
Furthermore, it has become increasingly apparent that the neural signals that drive complex
behaviour are manifest in a broad population of neurons rather than in single cells.  So current
neural recording emphasises high-dimensional signals from 100s to 1000s of neurons.

This complexity of experiments limits the potential for database-driven science.  The number of
possible experiments is astronomical, and a new question must usually be answered by designing and
conducting a new experiment rather than mining the data from an old one.
%
At present, the field is held back by our ability to efficiently design and analyse new experiments,
and to integrate state-of-the-art data processing and neuroinformatics into experimental designs.
%
\textbf{Our proposal is to enhance, distribute and support a neuroinformatics resource which \dots}
(simplifies data collection and analysis, something about sharing data??), potentially creating the most powerful experimental control and analysis tool available to neuroscientists, psychologists and ethologists worldwide.

\setlength{\columnsep}{1em}
\begin{wrapfigure}{r}{0.5\textwidth}
  \includegraphics[width=0.5\columnwidth,viewport=65 50 500 800,clip]{figures/bonsai.pdf}
  \label{fig:bonsai}
\end{wrapfigure}

\setcounter{figure}{1}

This resource will be built on the platform of Bonsai, an established experimental control software
ecosystem with a long development history and a large user base in systems neuroscience
(Figure~\ref{fig:bonsai}).

A central goal of the software resource proposed here is to address this
problem, by extending Bonsai~(an experimental control software ecosystem;
Figure~\ref{fig:bonsai} and Section~\ref{sec:bonsai}) with state-of-the-art
online (and offline) data analysis methods
(Figure~\ref{fig:proposedBonsaiExtensions} and
Section~\ref{sec:functionality}). \textbf{Bonsai is}
\textbf{unique} among experimental control software in that it allows people
with no programming experience to control sophisticated experiments. Its large
and rapidly growing user base demonstrates the need for this type of software
resources. With the proposed functionality added to Bonsai, its users
will be able to use sophisticated \textbf{machine-learning methods} to control their
experiments and analyse their data. This
\textbf{democratisation} of machine learning
functionality will allow scientists without computational skills to control
sophisticated experiments and analyse their data in unprecedented ways, which
will facilitate \textbf{advances in explaining how the brain gives rise to behaviour}.

Currently Bonsai has its own package manager, and its community of users have
already extended Bonsai's functionality with several contributed packages
(e.g., \ldots \textcolor{red}{Goncalo please reference the best contributed
packages developed by the Bonsai user community}). However, these packages need to
be written in C\#, while most current neuroscience data analysis methods are
written in Python, R or Matlab. We propose to add to Bonsai capabilities to
communicate with software written in these languages
(Section~\ref{sec:softwareInfrastructure}).  These capabilities will allow a
large number of neuroscience data analysis methods to be easily integrated into
the Bonsai ecosystem and will provide Bonsai users a large repertoire of
advanced machine learning methods for their behavioural and neural data
analysis.
%
Using these capabilities we will integrate into Bonsai an initial set of
state-of-the art behavioural and neural data analysis methods
(Section~\ref{sec:functionality}).

Developers of advanced data-analysis methods should become interested in
integrating them into the Bonsai ecosystem, as it will allow their methods to
reach the wide Bonsai user community. In addition this integration will enable
method developers to easily compare the performance of their methods with
that of many methods previously integrated into the Bonsai ecosystem. Method
developers will thus become a new type of Bonsai users, which will contribute
to the \textbf{sustainability} of the proposed resource beyond the period of
BBSRC-BBR funding.

Scientific advances rely on \textbf{reproducibility}, which in turn depends on standardized tools for experimental control and analysis, shared between laboratories (Baker, 2016; Ioannidis, 2005). Although such tools exist in genetics, astronomy, physics and medicine (Fish et al., 2016; Abdalla et al., 2018, CERN Education, Communications and Outreach Group, 2018, Dickinson et al., 2016, Bycroft et al., 2018), they are mostly lacking in neuroscience, which in itself suffers from a paucity of reproducible discoveries (Baker, 2016; Botvinik-Nezer et al., 2020; Button et al., 2013). Bonsai is an excellent tool for reproducible data acquisition and experimental control. For example, an experiment implemented in Bonsai can be replicated in
any laboratory by just sharing a Bonsai configuration file. With the addition of the proposed machine learning functionality Bonsai will extend this
reproducibility to the domain of data analysis
(Section~\ref{sec:reproducibility}).

\begin{figure}
  \includegraphics[width=4in]{figures/proposed_bonsai_extensions.png}
  \caption{Proposed extensions to Bonsai}
  \label{fig:proposedBonsaiExtensions}
\end{figure}

\subsection{Bonsai}
\label{sec:bonsai}

\input{bonsai.tex}

### 2021-11-16 application/bonsai.tex
Updates from Overleaf

@@
 %
-A Bonsai graphical program consists of one or more source data streams
+A Bonsai graphical program consists of one or more source data streams (e.g. neural activity, video, audio, sensors, etc)
 and several interconnected operators that transform input to output
@@
 the equivalent standardized software tools for data acquisition and control of
-animal behaviour experiments are still lacking. Life science experiments demand
+animal behaviour neuroscience experiments are still lacking. Life science experiments demand
 a combination of multiple instrumentation and control technologies, for both
-behavioral and physiological investigations. The growing complexity on both the
+behavioural and physiological investigations. The growing complexity on both the
 amount of data that is collected, and the rich conditions under which behaviour
 must be explored, place an increasing burden on experimenters to integrate
-highly specialized equipment in unique configurations, while often lacking
+highly specialised equipment in unique configurations, while often lacking
 expertise in the relevant engineering fields.
 
-Bonsai has a large user base in the systems neuroscience community
-(Figures~\ref{fig:bonsai}d-\ref{fig:bonsai}f).  In the last year alone, we estimate more than 1,000 new users have
-started to incorporate Bonsai into their experimental protocols across the
-world. The surprising rate of adoption of Bonsai in non-programmer experimental
+Bonsai has been in adopted in hundreds of laboratories worldwide and has the largest user base in the systems neuroscience community (Figures~\ref{fig:bonsai}d-\ref{fig:bonsai}f).
+In the last year alone, more than 1,000 new users incorporated Bonsai into their experimental protocols.
+The surprising rate of adoption of Bonsai in non-programmer experimental
 labs highlights the need for accessible programming tools that enable
-state-of-the-art technology but also allow researchers to stay in control and
+state-of-the-art technology but also allow researchers to stay in control and flexibly
 change their experimental paradigms.  Many open-source software tools are
@@
 outside their narrow domain of application. Bonsai has been successful because
-it has straddled this gap to some extent.
+it has straddled this gap.
 

### 2021-11-16 application/Je-S_sections.tex
Updates from Overleaf

@@
 
-The overall goal of this proposal is to extend, establish, distribute and support Bonsai, a fully integrated software
+The overall goal of this proposal is to extend, enhance, maintain and support Bonsai, a fully integrated software
 environment to enable cutting-edge reproducible laboratory systems neuroscience experiments in animal models.
@@
 
-1. Develop, test and deploy a new software pipeline to incorporate external packages into Bonsai's C# environment.
+1. Develop, test and deploy a new software workflow to incorporate external packages into Bonsai's C# environment.
 
@@
 
-[2825 chars]
+[2829 chars remaining]
 
@@
 
+Understanding the brain is a major scientific challenge of our era.
+To succeed, scientists must be able to explain how the rich repertoire of animal behaviour relates to the activity in neural circuits across different brain regions.
+This necessitates careful design and control of behavioural experiments wherein experimental animals (e.g.
+non-human primates, rodents, fish, insects) perform particular tasks while experimenters either record or manipulate neural activity in candidate neural circuits.
+Experiments in behavioural and brain science laboratories require software that integrates and controls hardware from multiple recording devices (video, electrodes for neural activity measurement, sensors), and analysis tools that can interpret large and complex behavioural and neural datasets.
+
+Scientists studying brain and behaviour dedicate the majority of their time to the design of experimental software protocols that control hardware and to data analysis, with least amount spent on data acquisition itself.
+Moreover, hundreds of neuroscience research groups worldwide develop their own software tools for experimental control and data analysis in different programming languages, leading to inefficiencies resulting from time-loss spent on writing software (often by individuals lacking programming expertise), the re-invention of software with the same functionality and concomitant lack of experimental reproducibility across laboratories.
+
+Here we propose to provide a community software tool that will dramatically increase the efficiency of experimental control and data analysis by providing a new set of functionalities to an existing software platform, Bonsai.
+Bonsai is an open-source reactive visual programming environment that naturally orchestrates complex networks of sensors and hardware into workflows and allows integration of multiple programming languages into its framework.
+Thus far, Bonsai has been adopted by hundreds of scientists worldwide to provide interactive experimental control in behavioural and brain sciences.
+In this proposal, we aim to extend Bonsai’s functionality with a toolbox of online and offline Machine Intelligence tools for analysis of behavioural and neural datasets, and to create an open-access platform for software sharing.
+Bonsai’s enhanced functionality will enable new types of research via online experimental control, speed up discovery and improve efficiency by (i) providing access to such tools into laboratories lacking expertise, (ii) reducing the need to reinvent the same tools in multiple labs and (iii) standardising the data processing streams, thus increasing reproducibility across laboratories.
+We believe this effort will enable and accelerate new discoveries in how the brain generates behaviour.
+
 Describe the Bioinformatic and/or Biological Resource in a manner suitable for a specialist reader. This summary will be made publicly available if the proposal is funded. [up to 2000 characters]
 
+To understand the brain, scientists aim to explain how the rich repertoire of animal behaviour relates to the activity in neural circuits across different brain regions.
+This necessitates careful design and precise control of behavioural experiments wherein experimental animals perform particular tasks while experimenters either record or manipulate neural activity in candidate neural circuits.
+Such experiments require software that integrates and controls hardware from multiple recording devices (cameras, electrodes, sensors), and analysis tools that can interpret large and complex behavioural and neural datasets.
+The lack of standardised tools for experimental control and analysis across the brain-and-behaviour community is incurring major inefficiencies, with disproportionate time allocated to software engineering and writing analysis code, and efforts often replicated across laboratories.
+Similarly, disparate experimental and analysis protocols makes is almost impossible to replicate experiments between research groups.
+Bonsai is an open-source reactive visual programming environment that naturally orchestrates complex networks of sensors and hardware into workflows.
+Bonsai has been adopted by hundreds of scientists worldwide to provide interactive experimental control in behavioural and brain sciences.
+Here we aim to extend Bonsai’s functionality with a toolbox of online and offline Machine Intelligence tools for analysis of behavioural and neural data (video-based analysis of behavioural motifs, real-time and off-line analysis of neural signals), and create an open-access platform for software sharing and integration with multiple programming languages.
+Bonsai’s enhanced functionality will enable new types of research, accelerate discovery and improve efficiency by (i) providing access to such tools for scientists lacking expertise, (ii) reducing duplication of effort, and (iii) standardising the data processing streams, thus increasing reproducibility across laboratories.
+[12 chars over]
+
 Describe who will benefit from this Bioinformatic and/or Biological resource(s) being funded [up to 4000 characters]

### 2021-11-16 application/backgroundToTheResource.tex
Updates from Overleaf

@@
 
-Modern systems neuroscience depends on data from carefully designed behavioural experiments in
-animal models such as mice frequently combined with invasive electrophysiology and/or
-cellular-resolution imaging.  Experimental design might require behavioural feedback (e.g. reward,
-lights, sounds, actuated devices\dots) or neural pertubation (e.g. using light-sensitive polarising
-molecules) that depends on real-time behaviour or neural state.
+To understand the neural basis of behaviour, experiments need to control and quantify animal behaviour whilst recording, and often manipulating, neural activity. 
 %
-Furthermore, it has become increasingly apparent that the neural signals that drive complex
-behaviour are manifest in a broad population of neurons rather than in single cells.  So current
-neural recording emphasises high-dimensional signals from 100s to 1000s of neurons.
+Powerful experimental designs require behavioural feedback (e.g.\ reward, lights, sounds, actuated
+devices) or neural perturbation (e.g.\ using light-sensitive polarising
+molecules) that depend on real-time behaviour or neural state.
+%
+Furthermore, it has become increasingly apparent that the neural signals that
+drive complex behaviour are manifest in a broad population of neurons rather
+than in single cells.  So current neural recording emphasises high-dimensional
+signals from 100s to 1000s of neurons monitored using multi-electrode arrays or cellular-resolution imaging, with a wide array of probe configurations and neural targets.
+%
+This complexity of experiments limits the potential for database-driven progress in systems neuroscience.
+%
+The number of possible experiments is astronomical, and a new question must
+usually be answered by designing and conducting a new experiment rather than
+mining the data from an old one.
 
-This complexity of experiments limits the potential for database-driven science.  The number of
-possible experiments is astronomical, and a new question must usually be answered by designing and
-conducting a new experiment rather than mining the data from an old one.
+At present, progress in the field is held back by the challenge of rapidly designing and implementing new experimental protocols, and the difficulty of integrating state-of-the-art data processing and
+neuroinformatics into custom experimental designs.
 %
-At present, the field is held back by our ability to efficiently design and analyse new experiments,
-and to integrate state-of-the-art data processing and neuroinformatics into experimental designs.
-%
-\textbf{Our proposal is to enhance, distribute and support a neuroinformatics resource which \dots}
-(simplifies data collection and analysis, something about sharing data??), potentially creating the most powerful experimental control and analysis tool available to neuroscientists, psychologists and ethologists worldwide.
+\textbf{Reproducibility}, in particular, depends on the availability of standardized tools for experimental control and analysis, shared between laboratories (Baker, 2016; Ioannidis, 2005).
+Although such tools exist in genetics, astronomy, physics and medicine (Fish et al., 2016; Abdalla et al., 2018, CERN Education, Communications and Outreach Group, 2018, Dickinson et al., 2016, Bycroft et al., 2018), they are mostly lacking in neuroscience, with documented consequences for the reliability of ``discoveries'' (Baker, 2016; Botvinik-Nezer et al., 2020; Button et al., 2013).
 
@@
 
-\setcounter{figure}{1}
+\input{bonsai.tex}
 
-This resource will be built on the platform of Bonsai, an established experimental control software
-ecosystem with a long development history and a large user base in systems neuroscience
-(Figure~\ref{fig:bonsai}).
+Our proposed enhancements will \textbf{enhance, reinforce and support this key neuroinformatics
+  resource to streamline reproducible experimental design, data collection and analysis, while making standardised implementations of cutting-edge machine-learning methodologies widely available to the community}, potentially creating the most powerful experimental
+control and analysis tool available to neuroscientists, psychologists and
+ethologists worldwide.
 
-A central goal of the software resource proposed here is to address this
-problem, by extending Bonsai~(an experimental control software ecosystem;
-Figure~\ref{fig:bonsai} and Section~\ref{sec:bonsai}) with state-of-the-art
-online (and offline) data analysis methods
-(Figure~\ref{fig:proposedBonsaiExtensions} and
-Section~\ref{sec:functionality}). \textbf{Bonsai is}
-\textbf{unique} among experimental control software in that it allows people
-with no programming experience to control sophisticated experiments. Its large
-and rapidly growing user base demonstrates the need for this type of software
-resources. With the proposed functionality added to Bonsai, its users
-will be able to use sophisticated \textbf{machine-learning methods} to control their
-experiments and analyse their data. This
-\textbf{democratisation} of machine learning
-functionality will allow scientists without computational skills to control
-sophisticated experiments and analyse their data in unprecedented ways, which
-will facilitate \textbf{advances in explaining how the brain gives rise to behaviour}.
+\input{otherResources.tex}
 
-Currently Bonsai has its own package manager, and its community of users have
-already extended Bonsai's functionality with several contributed packages
-(e.g., \ldots \textcolor{red}{Goncalo please reference the best contributed
-packages developed by the Bonsai user community}). However, these packages need to
-be written in C\#, while most current neuroscience data analysis methods are
-written in Python, R or Matlab. We propose to add to Bonsai capabilities to
-communicate with software written in these languages
-(Section~\ref{sec:softwareInfrastructure}).  These capabilities will allow a
-large number of neuroscience data analysis methods to be easily integrated into
-the Bonsai ecosystem and will provide Bonsai users a large repertoire of
-advanced machine learning methods for their behavioural and neural data
-analysis.
 %
-Using these capabilities we will integrate into Bonsai an initial set of
-state-of-the art behavioural and neural data analysis methods
-(Section~\ref{sec:functionality}).
 
-Developers of advanced data-analysis methods should become interested in
-integrating them into the Bonsai ecosystem, as it will allow their methods to
-reach the wide Bonsai user community. In addition this integration will enable
-method developers to easily compare the performance of their methods with
-that of many methods previously integrated into the Bonsai ecosystem. Method
-developers will thus become a new type of Bonsai users, which will contribute
-to the \textbf{sustainability} of the proposed resource beyond the period of
-BBSRC-BBR funding.
+%
 
-Scientific advances rely on \textbf{reproducibility}, which in turn depends on standardized tools for experimental control and analysis, shared between laboratories (Baker, 2016; Ioannidis, 2005).
-Although such tools exist in genetics, astronomy, physics and medicine (Fish et al., 2016; Abdalla et al., 2018, CERN Education, Communications and Outreach Group, 2018, Dickinson et al., 2016, Bycroft et al., 2018), they are mostly lacking in neuroscience, which in itself suffers from a paucity of reproducible discoveries (Baker, 2016; Botvinik-Nezer et al., 2020; Button et al., 2013).
-Bonsai is an excellent tool for reproducible data acquisition and experimental control.
-For example, an experiment implemented in Bonsai can be replicated in
-any laboratory by just sharing a Bonsai configuration file. With the addition of the proposed machine learning functionality Bonsai will extend this
-reproducibility to the domain of data analysis
-(Section~\ref{sec:reproducibility}).
+% This resource will be built on the platform established by \textbf{Bonsai}.
+%
 
-\begin{figure}
-  \includegraphics[width=4in]{figures/proposed_bonsai_extensions.png}
-  \caption{Proposed extensions to Bonsai}
-  \label{fig:proposedBonsaiExtensions}
-\end{figure}
+% an established
+% experimental control software ecosystem with a long development history and a
+% large user base in systems neuroscience (Figure~\ref{fig:bonsai}).
 
-\subsection{Bonsai}
-\label{sec:bonsai}
+% \subsection{Bonsai}
+% \label{sec:bonsai}
+%
+% We will 
+% 
+% 
+% 
+% A central goal of the software resource proposed here is to address this
+% problem, by extending Bonsai~(an experimental control software ecosystem;
+% Figure~\ref{fig:bonsai} and Section~\ref{sec:bonsai}) with state-of-the-art
+% online (and offline) data analysis methods
+% (Figure~\ref{fig:proposedBonsaiExtensions} and
+% Section~\ref{sec:functionality}).
+% Bonsai is unique amongst experimental control systems in that
+% it allows users with no programming expertise to design and implement
+% sophisticated experiments.
+%
+% The latest Its large and rapidly growing community of users demonstrates the demand for such software resources.
+%
 
-\input{bonsai.tex}
+% With the proposed enhancements, users of Bonsai will be able to use sophisticated \textbf{machine-learning methods} to
+% control their experiments and analyse their data. This \textbf{democratisation}
+% of machine learning functionality will allow scientists without computational
+% skills to control sophisticated experiments and analyse their data in
+% unprecedented ways, which will facilitate \textbf{advances in explaining how the
+%   brain gives rise to behaviour}.
+
+% \subsection{Bonsai}
+% \label{sec:bonsai}

### 2021-11-16 application/caseForSupport.tex
Added track record
Updates from Overleaf

@@
 \setcounter{tocdepth}{4}    % how many sectioning levels to show in ToC
-\usepackage[small]{titlesec} % format section titles
+\usepackage[tiny]{titlesec} % format section titles
+\def\headfmt{\color{blue!50!black!50}\bfseries}
+\titleformat*{\section}{\headfmt}
+\titlespacing*{\section}{0pt}{*1}{*0}[-4em]
+\titleformat*{\subsection}{\headfmt}
+\titlespacing*{\subsection}{0pt}{*0.5}{*0}
+\titleformat*{\subsubsection}{\headfmt}
+\titlespacing*{\subsubsection}{0pt}{*0}{*0}
+\titleformat*{\paragraph}{\headfmt}
+\titlespacing*{\paragraph}{0pt}{*0.5}{*0}
 
@@
 
-\maketitle
+% \maketitle
 
@@
 
+\section*{\large Case for support}
+\vskip 0.5ex
+We propose to extend, enhance, maintain and support \textbf{Bonsai}, a fully
+integrated software environment to \textbf{enable cutting-edge reproducible
+  laboratory systems neuroscience experiments in animal models}, with a
+particular emphasis on \textbf{machine-intelligence-enabled, real-time
+  neuroinformatics} methods.
+
 \input{backgroundToTheResource}
-
-\section{Details of resource}
 

### 2021-11-16 application/bonsai.tex

@@
-Bonsai~\citep{lopesEtAl15,lopesAndMonteiro21} is a free and open-source visual
-programming language that emphasizes performance, flexibility, and ease-of-use,
+\textbf{Bonsai}~\citep{lopesEtAl15,lopesAndMonteiro21} is a free and open-source visual programming language developed 
+in response to these challenges.
+% ~\citep{lopesEtAl15,lopesAndMonteiro21} is a free and open-source visual
+% programming language that 
+Its design emphasizes performance, flexibility, and ease-of-use,
 allowing scientists with no previous programming experience to quickly develop
@@
 Bonsai combines a high-level event algebra for data streams with an integrated
-development environment (IDE) and a strong library of plugins supporting
-multiple hardware and software packages used by the biomedical research
-community (Figures~\ref{fig:bonsai}a and~\ref{fig:bonsai}b).
+development environment (IDE) and an extensive library of plugins supporting
+multiple hardware and software packages used by the neuroscience research
+community (Figure~\ref{fig:bonsai}A-B).
 %
@@
 and several interconnected operators that transform input to output
-datastreams (Figure~\ref{fig:bonsai}c).
+datastreams (Figure~\ref{fig:bonsai}C).
 
@@
 
-Bonsai has been in adopted in hundreds of laboratories worldwide and has the largest user base in the systems neuroscience community (Figures~\ref{fig:bonsai}d-\ref{fig:bonsai}f).
+Reproducibility is a key element of Bonsai's design.  Data acquisition and experimental control protocols can be 
+replicated in any laboratory by just sharing a Bonsai configuration file.
+
+Bonsai has been in adopted in hundreds of laboratories worldwide and has the largest user base in the systems neuroscience community (Figure~\ref{fig:bonsai}D-F).
 In the last year alone, more than 1,000 new users incorporated Bonsai into their experimental protocols.
 The surprising rate of adoption of Bonsai in non-programmer experimental
-labs highlights the need for accessible programming tools that enable
+labs highlights the need for accessible design tools that enable
 state-of-the-art technology but also allow researchers to stay in control and flexibly
@@
 %
-More recently, Bonsai has started to expand outside the domain of neuroscience
-into biomedical research and biotechnology tool development, and even outside
-academia into public outreach and education programs.
-
-\subsubsection*{Other resources in the subject area}
-
-The space of technologies serving experimental control and behaviour monitoring
-is large, and is traditionally occupied either by domain-specific graphical
-user interfaces for control and recording of specific devices and experiment
-types (e.g. Open Ephys
-GUI\footnote{\href{https://open-ephys.org/gui/}{https://open-ephys.org/gui/}},
-Miniscope DAQ
-Software\footnote{\href{https://github.com/Aharoni-Lab/Miniscope-DAQ-QT-Software}{https://github.com/Aharoni-Lab/Miniscope-DAQ-QT-Software}})
-or by real-time control frameworks for specifying task logic using state
-machine or similar formalisms (e.g. NIMH ML
-(nih.gov)\footnote{\href{https://monkeylogic.nimh.nih.gov/}{https://monkeylogic.nimh.nih.gov/}},
-pyControl\footnote{\href{https://pycontrol.readthedocs.io/en/latest/}{https://pycontrol.readthedocs.io/en/latest/}},
-Autopilot\footnote{\href{https://docs.auto-pi-lot.com/en/latest/}{https://docs.auto-pi-lot.com/en/latest/}},
-Sanworks\footnote{\href{https://sanworks.io/index.php}{https://sanworks.io/index.php}}).
-These dedicated interfaces are typically very comfortable for experimenters
-operating within the specific domain that the tool is designed for, but tend to
-become unwieldy when there is a need to introduce a new device or variation of
-a task which is outside the scope of the framework. The alternative is to use a
-more general programming language like Python or MATLAB, with the disadvantage
-of the code being harder to understand, maintain, and change. Programming
-languages like LabVIEW straddle the middle ground and provide a high-level and
-flexible visual interface for composing data acquisition and control systems
-themselves. Unlike Bonsai, however, the graphical elements of LabVIEW are
-heterogeneous and very fine grained, creating the need for long and complicated
-logical structures to implement even a simple experimental control system. By
-providing an extremely simple, yet flexible, visual syntax, Bonsai opens the
-opportunity even for complete non-programmers to design and successfully
-customize relatively complex experiments from the ground up. It is mostly this
-capability which has made Bonsai so attractive as a standard tool in
-experimental neuroscience.
+% LEAVE OUT FOR THIS GRANT:
+% More recently, Bonsai has started to expand outside the domain of neuroscience
+% into biomedical research and biotechnology tool development, and even outside
+% academia into public outreach and education programs.

### 2021-11-16 application/trackRecord.tex
Added

The team responsible for the proposed project comprises two academic
institutes, the Gatsby Computational Neuroscience Unit (GCNU) and the Sainsbury
Wellcome Centre for Neural Circuits and Behavior (SWC), and a private
collaborator, NeuroGEARS Ltd.

The GCNU is a world-class research center for computational neuroscience and
machine learning, while the Sainsbury Wellcome Center is a top-notch experimental
neuroscience research center. They share a common building and collaborative
activities are at the heart of both institutes.
%
NeuroGEARS is a spinoff of the SWC and is currently contributing to a major
collaborative research project between the SWC and the GCNU. It has developed
Bonsai, a very unique ecosystem for experimental control and data acquisition,
it has offered courses around the globe instructing students on advanced
experimental control, and has used Bonsai in numerous research and private
experimental projects. Thus, NeuroGEARS has extensive experience on the
implementation of neuroscience experiments.

In summary, the partners in this project have outstanding expertise on their
focus areas (computational neuroscience and machine learning, experimental
neuroscience, implementation of experiments), as well as a rich collaborative
track record. The proposed project will merge the unique expertise of its
partner institutions to create a resource that will be essential for the
advanced of biological sciences.

### 2021-11-16 application/detailsOfResource.tex

@@
+% 1. Develop, test and deploy a new software workflow to incorporate external packages into Bonsai's C# environment.
+
+% 2. Incorporate a suite of machine-learning-driven analysis tools into the Bonsai environment to implement:
+% 2a. video-based  behavioural analysis
+% 2b. real-time interfacing of population-scale activity with laboratory control  
+% 2c. evaluate competing data analytic approaches within a single experimental framework
+
+\section{Details of the proposed resource enhancements}
+
 We intend to improve an \textbf{existing resource}, the Bonsai ecosystem. We
@@
 translate into groundbreaking scientific discoveries.
+
+Currently Bonsai has its own package manager, and its community of users have
+already extended Bonsai's functionality with several contributed packages.
+%
+However, these packages need to
+be written in C\#, while most current neuroscience data analysis methods are
+written in Python, R or Matlab. We propose to add to Bonsai capabilities to
+communicate with software written in these languages
+(Section~\ref{sec:softwareInfrastructure}).  These capabilities will allow a
+large number of neuroscience data analysis methods to be easily integrated into
+the Bonsai ecosystem and will provide Bonsai users a large repertoire of
+advanced machine learning methods for their behavioural and neural data
+analysis.
+%
+Using these capabilities we will integrate into Bonsai an initial set of
+state-of-the art behavioural and neural data analysis methods
+(Section~\ref{sec:functionality}).
+
+Developers of advanced data-analysis methods should become interested in
+integrating them into the Bonsai ecosystem, as it will allow their methods to
+reach the wide Bonsai user community. In addition this integration will enable
+method developers to easily compare the performance of their methods with
+that of many methods previously integrated into the Bonsai ecosystem. Method
+developers will thus become a new type of Bonsai users, which will contribute
+to the \textbf{sustainability} of the proposed resource beyond the period of
+BBSRC-BBR funding.
+
+% Scientific advances rely on \textbf{reproducibility}, which in turn depends on standardized tools for experimental control and analysis, shared between laboratories (Baker, 2016; Ioannidis, 2005).
+Although such tools exist in genetics, astronomy, physics and medicine (Fish et al., 2016; Abdalla et al., 2018, CERN Education, Communications and Outreach Group, 2018, Dickinson et al., 2016, Bycroft et al., 2018), they are mostly lacking in neuroscience, which in itself suffers from a paucity of reproducible discoveries (Baker, 2016; Botvinik-Nezer et al., 2020; Button et al., 2013).
+Bonsai is an excellent tool for reproducible data acquisition and experimental control.
+For example, an experiment implemented in Bonsai can be replicated in
+% any laboratory by just sharing a Bonsai configuration file. With the addition of the proposed machine learning functionality Bonsai will extend this
+% reproducibility to the domain of data analysis
+% (Section~\ref{sec:reproducibility}).
+
+\begin{figure}
+  \includegraphics[width=4in]{figures/proposed_bonsai_extensions.png}
+  \caption{Proposed extensions to Bonsai}
+  \label{fig:proposedBonsaiExtensions}
+\end{figure}
 

### 2021-11-16 application/narrativeCVSahani.tex

% Instead of submitting a traditional academic CV, we invite applicants to submit a two-page 
% Narrative CV. This should act as a narrative which explains how the individual’s relevant 
% experience and expertise demonstrates their ability to successfully deliver the proposal. A 
% CV which simply lists past positions, publications, and funding will not adequately support an 
% application.
% 
% 
% Applicants should draw on a breadth of examples which illustrate of they how they have 
% contributed to new ideas, hypotheses and tools, as well as how they have contributed to 
% teams and collaborations, the research community, and to wider society. Applicants should 
% describe only a selection of their outputs and, in each case, clearly explain the relevance to 
% their ability to deliver the proposed project.
% 
% 
% All individuals should describe contributions across the levels below. The relative size 
% of each section will vary depending on the relevant skills and
% experience of each applicant.
% 
%

\section{Prof.\ Maneesh Sahani -- narrative CV}

% 1) Eligibility criteria: State your current position, indicating how you meet the 
% eligibility criteria as outlined within the UKRI-BBSRC Grants Guide.

\subsection{Eligibility criteria}

Prof.\ Sahani is a Professor of Theoretical Neuroscience and Machine
Learning, and director of the Gatsby Computational Neuroscience Unit
at UCL.
%
As such, he is resident in the UK, and holds an open-ended academic
employment contract with UCL.

% 2) How have you contributed to the generation and flow of new ideas, hypotheses, 
% tools or knowledge?
% 
% Examples might include: contributions to and skills acquired from past research projects, and 
% key outputs such as data sets, software, and research and policy publications. In each case 
% the relevance to delivering the proposed project should be summarised.

\subsection{Contributions to ideas}

Prof.\ Sahani has authored over 100 peer-reviewed scientific papers,
with an h-index (computed by Google scholar) of 47.

GPFA

PLDS

Grant funding?

% 3) How have you contributed to research teams and the development of
% others?

% From QQR:
As a PI, I have supervised a total of \textbf{X} students and \textbf{X} postdocs.  
In the last five years, 4 of students have moved to postdoctoral positions (at
Aalto, Cambridge, CMU and Stanford), 1 works at DeepMind, and 1 holds a faculty offer at Shanghai Tech.  In all, 8 previous students have now secured faculty (or group leader) positions at Cambridge, CMU, Columbia, Copenhagen, Janelia (2), Pittsburgh and Shanghai.
%
Of the postdoctoral fellows I have supervised in the last five years, 1 is doing further postdoc work in Bonn; 1 holds an independent
Research Fellowship at Radboud; and 2 have faculty positions: at Edinburgh and at UBC.

DARPA repair

Gatsby directorship

% Examples might include: project management, supervision, mentoring or line management 
% contributions critical to the success of a team or team members, or where you exerted 
% strategic leadership in shaping the direction of a team, organisation, company or institution.

% 4) How have you contributed to the wider research community?
% 
% Examples might include: how you have contributed to wider collaborations and networks 
% across disciplines, institutions, and / or countries, commitments such as editing, reviewing 
% and committee work, positions of responsibility, and activities which have contributed to the 
% improvement of research integrity or culture, or examples where you have shown visionary 
% strategic leadership in influencing a research agenda.

I have been an active member of the academic community. I co-founded the Neural Coding, Computation and Dynamics meeting in 2007, and was in the organising committee in 2015 and 2017.

I participated in the selection committee for the Swartz Prize for Theoretical and Computational Neuroscience in 2014-2016, and was in the programme committee for the Computational and Systems Neuroscience (COSYNE) conference in 2007, becoming a  Programme Chair in 2009 and General Chair in 2010.

I was also in the Programme Committee for NIPS (now NeurIPS) in 2004 and 2006, becoming a Workshops Chair in 2008. I was also a Workshops Chair for the Computational Neuroscience Meeting from 1999 - 2003.

I was a member of the Board of Directors of the Computational Neuroscience Organization between 2003 - 2006. I am currently a member of the Society for Neuroscience and IEEE, and in 2021 became a Fellow of the European Laboratory for Learning and Intelligent Systems (ELLIS).

I currently have editorial roles in various journals including, Current Opinion in Neurobiology (Special issue on Big Data and Neuroscience), Neural Computation, Neural Systems and Circuits, Network:
Computation in Neural Systems, and Faculty of
1000, "Machine learning: life sciences" collection.

I have taken various advisory roles for international review and award panels. I reviewed the Donders Institute for Brain, Cognition and Behaviour (Netherlands), The Edmond and Lily Safra Center for Brain Sciences, Hebrew
University (Israel), and the Barcelona Summer School on Advanced Modelling of Behaviour (Spain). I have advised on awards such as the Swartz Prize for Theoretical and Computational Neuroscience (SFN, US), the Bernstein Prize (BMBF, Germany), and COSYNE.

Recently, I have acted as an advisor to a new established cross-disciplinary effort at the Indian Institute of Science in Bangalore; visiting the site a
number of times and hosting a joint workshop here at UCL.

I have sat on various grant award panels for the BMBF and DFG, and the (US) NSF/NIH collaborative research
in computational neuroscience (CRCNS) programme. I have also acted as a reviewer for proposals for the Bernstein prize, Einstein Foundation, EPSRC, MRC, Novo Nordisk Foundation (Denmark) and NWO (Netherlands).

% 5) How have you contributed to broader society? 
% Examples might include: engagement across the public and / or private sectors or with the 
% wider public, past research which has contributed to policy development or public 
% understanding, and other impacts across research, policy, practice and business, and other 
% examples of and how you have ensured your research reaches and influences relevant 
% audiences.

% 6) Additional information  
% Any additional relevant information you wish to include in support of your capability to deliver, 
% which may include further information about key qualifications and relevant positions, 
% secondments, volunteering, or other relevant experience such as time spent in different 
% sectors. There is no need to provide information about career breaks, part-time working etc, 
% however if there are any details you do wish panel members to consider in their assessment 
% of the proposal they may be included here.
%

### 2021-11-16 application/otherResources.tex

--- 2021-11-12 bonsai.tex
@@
-Bonsai~\citep{lopesEtAl15,lopesAndMonteiro21} is a free and open-source visual
-programming language that emphasizes performance, flexibility, and ease-of-use,
-allowing scientists with no previous programming experience to quickly develop
-their own high-performance data acquisition and experimental control systems
-(Figure 1c). Bonsai combines a high-level event algebra for data streams with
-an integrated development environment (IDE) and a strong library of plugins
-supporting multiple hardware and software packages used by the biomedical
-research community (Figures 1a, 1b).
-
-Standard software tools for data analysis (ImageJ, MATLAB, R, etc.) have been
-transformative to the progress of increasingly “data-rich” sciences. However,
-the equivalent standardized software tools for data acquisition and control of
-animal behaviour experiments are still lacking. Life science experiments demand
-a combination of multiple instrumentation and control technologies, in both
-physiology and behaviour studies. The growing complexity on both the amount of
-data that is collected, and the rich conditions under which behaviour must be
-explored, place an increasing burden on the experimenter to integrate highly
-specialized equipment in unique configurations, while often lacking expertise
-in the relevant engineering fields.
-
-Bonsai has a large user base in the systems neuroscience community.
-Quantifications of this user base are provided in Figures 1d-1f.  In the last
-year alone, we estimate more than 1,000 new users have started to incorporate
-Bonsai into their experimental protocols across the world. The surprising rate
-of adoption of Bonsai in non-programmer experimental labs highlights the need
-for accessible programming tools that enable state-of-the-art technology but
-also allow researchers to stay in control and change their experimental
-paradigms.  Many open-source software tools are either inaccessible to
-non-programmers, or too constrained to be of general use outside their narrow
-domain of application. Bonsai has been successful because it has straddled this
-gap to some extent.
-
-The language has also helped to potentiate the growing wave of foundational
-open hardware initiatives, such as the OpenEphys \citep{siegleEtAl17} and UCLA
-Miniscope~\citep{caiEtAl16}, making it possible to quickly combine and
-integrate these tools into new experiments~\citep{buccinoEtAl18}.
-%
-Bonsai has been adopted in large neuroscience undertakings like the
-International Brain
-Laboratory\footnote{\href{https://www.internationalbrainlab.com/}{https://www.internationalbrainlab.com/}}
-and the Allen Institute for Brain
-Science\footnote{\href{https://alleninstitute.org/what-we-do/brain-science/}{https://alleninstitute.org/what-we-do/brain-science/}}.
-%
-More recently, Bonsai has started to expand outside the domain of neuroscience
-into biomedical research and biotechnology tool development, and even outside
-academia into public outreach and education programs.
-
-\subsubsection*{Other resources in the subject area}
+\subsection*{Other resources in the subject area}
 

### 2021-11-16 application/trackRecord.tex
Cosmetic change

@@
 neuroscience, implementation of experiments), as well as a rich collaborative
-track record. The proposed project will merge the unique expertise of its
-partner institutions to create a resource that will be essential for the
-advanced of biological sciences.
+track record. The proposed project will merge this unique expertise to create a
+resource that will be essential for the advanced of biological sciences.

### 2021-11-16 application/Je-S_sections.tex
11/16 changes Maneesh

@@
 Understanding the brain is a major scientific challenge of our era.
-To succeed, scientists must be able to explain how the rich repertoire of animal behaviour relates to the activity in neural circuits across different brain regions.
-This necessitates careful design and control of behavioural experiments wherein experimental animals (e.g.
-non-human primates, rodents, fish, insects) perform particular tasks while experimenters either record or manipulate neural activity in candidate neural circuits.
+To succeed, scientists must be able to explain how animal behaviour relates to neural activity across different brain regions.
+This requires careful design and manipulation of behavioural experiments, where experimenters either record or manipulate neural activity while the animals (e.g.
+non-human primates, rodents, fish, insects) perform particular tasks.
+
 Experiments in behavioural and brain science laboratories require software that integrates and controls hardware from multiple recording devices (video, electrodes for neural activity measurement, sensors), and analysis tools that can interpret large and complex behavioural and neural datasets.
@@
 
-To understand the brain, scientists aim to explain how the rich repertoire of animal behaviour relates to the activity in neural circuits across different brain regions.
-This necessitates careful design and precise control of behavioural experiments wherein experimental animals perform particular tasks while experimenters either record or manipulate neural activity in candidate neural circuits.
-Such experiments require software that integrates and controls hardware from multiple recording devices (cameras, electrodes, sensors), and analysis tools that can interpret large and complex behavioural and neural datasets.
-The lack of standardised tools for experimental control and analysis across the brain-and-behaviour community is incurring major inefficiencies, with disproportionate time allocated to software engineering and writing analysis code, and efforts often replicated across laboratories.
-Similarly, disparate experimental and analysis protocols makes is almost impossible to replicate experiments between research groups.
-Bonsai is an open-source reactive visual programming environment that naturally orchestrates complex networks of sensors and hardware into workflows.
-Bonsai has been adopted by hundreds of scientists worldwide to provide interactive experimental control in behavioural and brain sciences.
-Here we aim to extend Bonsai’s functionality with a toolbox of online and offline Machine Intelligence tools for analysis of behavioural and neural data (video-based analysis of behavioural motifs, real-time and off-line analysis of neural signals), and create an open-access platform for software sharing and integration with multiple programming languages.
-Bonsai’s enhanced functionality will enable new types of research, accelerate discovery and improve efficiency by (i) providing access to such tools for scientists lacking expertise, (ii) reducing duplication of effort, and (iii) standardising the data processing streams, thus increasing reproducibility across laboratories.
-[12 chars over]
+To understand the brain, scientists aim to explain how animal behaviour relates to the activity in neural circuits.
+This necessitates design and precise control of behavioural experiments wherein experimental animals perform particular tasks while experimenters either record or manipulate neural activity in candidate neural circuits.
+Such experiments require data acquisition software that integrates and controls hardware from multiple recording devices (cameras, electrodes, sensors), and analysis tools that can interpret large and complex behavioural and neural datasets.
+Progress is held back by the lack of standardised tools for rapid design and implementation of experimental protocols, and the difficulty of integrating state-of-the-art data processing and neuroinformatics into custom experimental designs.
+The fields of behavioural and brain sciences have consequently suffered from both inefficiency and poor reproducibility, due to disparate data acquisition and analysis solutions created independently across laboratories.
+To address these challenges, we propose  to  extend,  enhance,  maintain  and  support \textbf{Bonsai},  a  fully  integrated  software  environment  to enable cutting-edge reproducible laboratory systems neuroscience experiments in animal models, with a particular emphasis on machine-intelligence-enabled, real-time neuroinformatics methods.
+While Bonsai is already adopted by hundreds of scientists worldwide to provide interactive experimental control, we aim to extend Bonsai’s functionality with a toolbox of online and offline Machine Intelligence tools for analysis of behavioural and neural data (video-based analysis of behavioural motifs, real-time and off-line analysis of neural signals), and create an open-access platform for software sharing and integration with multiple programming languages.
+These to the Bonsai ecosystem will enable new types of research, increase user base, and dramatically improve efficiency and reproducibility of research across behavioural and brain sciences.
+[44 chars over]
 
 Describe who will benefit from this Bioinformatic and/or Biological resource(s) being funded [up to 4000 characters]
+
+Resource is Bonsai.
+Monitor animal welfare

### 2021-11-16 application/backgroundToTheResource.tex
11/16 changes Maneesh

@@
 %
-\textbf{Reproducibility}, in particular, depends on the availability of standardized tools for experimental control and analysis, shared between laboratories (Baker, 2016; Ioannidis, 2005).
+\textbf{Reproducibility}, in particular, depends on the availability of standardised tools for experimental control and analysis, shared between laboratories (Baker, 2016; Ioannidis, 2005).
 Although such tools exist in genetics, astronomy, physics and medicine (Fish et al., 2016; Abdalla et al., 2018, CERN Education, Communications and Outreach Group, 2018, Dickinson et al., 2016, Bycroft et al., 2018), they are mostly lacking in neuroscience, with documented consequences for the reliability of ``discoveries'' (Baker, 2016; Botvinik-Nezer et al., 2020; Button et al., 2013).
@@
 \end{wrapfigure}
+\stepcounter{figure}
 
 \input{bonsai.tex}
+
+Bonsai achieves robust behavioural control, hardware interaction and signal acquisition while ensuring remarkable ease of use.
+The proposed project will provide valuable enhancements.
+We will develop, test and deploy a new software workflow making it possible to incorporate external packages into Bonsai's C\#-based environment.
+Building on this new extensibility, we will then incorporate a suite of machine-learning-driven analysis tools into the Bonsai environment.
+These added capabilities will allow  video-based behavioural analysis as well as real-time interfacing of population-scale activity with laboratory control.
+In parallel we will develop the infrastructure needed to evaluate competing data analytic approaches within a single experimental framework
 
@@
 
-%
-
-%
-
-% This resource will be built on the platform established by \textbf{Bonsai}.
-%
-
-% an established
-% experimental control software ecosystem with a long development history and a
-% large user base in systems neuroscience (Figure~\ref{fig:bonsai}).
-
-% \subsection{Bonsai}
-% \label{sec:bonsai}
-%
-% We will 
-% 
-% 
-% 
 % A central goal of the software resource proposed here is to address this

### 2021-11-16 application/narrativeCVSahani.tex
11/16 changes Maneesh

@@
 % 3) How have you contributed to research teams and the development of
-% others?
+% others? % Examples might include: project management, supervision, mentoring or line management 
+% contributions critical to the success of a team or team members, or where you exerted 
+% strategic leadership in shaping the direction of a team, organisation, company or institution.
 
@@
 %
-Of the postdoctoral fellows I have supervised in the last five years, 1 is doing further postdoc work in Bonn; 1 holds an independent
+Of the postdoctoral fellows I have supervised in the last five years, one is doing further postdoc work in Bonn; another holds an independent
 Research Fellowship at Radboud; and 2 have faculty positions: at Edinburgh and at UBC.
@@
 
-Gatsby directorship
-
-% Examples might include: project management, supervision, mentoring or line management 
-% contributions critical to the success of a team or team members, or where you exerted 
-% strategic leadership in shaping the direction of a team, organisation, company or institution.
+Since 2017, I have been Director of the Gatsby Computational Neuroscience Unit, and have spearheaded its new strategic research direction.
+I created two new roles to support and develop the new strategy, and have recruited 2 new faculty members who have added new and highly relevant research strands to our research portfolio.
+As Director, I sit in the Executive Leadership Committee of the Faculty of Life Sciences at UCL.
 
@@
 
-I have been an active member of the academic community. I co-founded the Neural Coding, Computation and Dynamics meeting in 2007, and was in the organising committee in 2015 and 2017.
+Current collaborations/networks?
+
+I have been an active member of the academic community by contributing the development of programme content for many scientific meetings, critically assessing the intellectual content of research proposals and awards applications, and reviewing the research and strategic direction of various world-class research institutions.
+
+I co-founded the Neural Coding, Computation and Dynamics meeting in 2007, and was in the organising committee in 2015 and 2017.
 
@@
 
-I have sat on various grant award panels for the BMBF and DFG, and the (US) NSF/NIH collaborative research
+I have sat on various grant award panels for the BMBF and DFG (Germany), and the (US) NSF/NIH collaborative research
 in computational neuroscience (CRCNS) programme.
-I have also acted as a reviewer for proposals for the Bernstein prize, Einstein Foundation, EPSRC, MRC, Novo Nordisk Foundation (Denmark) and NWO (Netherlands).
+I have also acted as a reviewer for proposals for the Bernstein prize, the Einstein Foundation, EPSRC, MRC, Novo Nordisk Foundation (Denmark) and NWO (Netherlands).
 

### 2021-11-16 application/resourceManagementandSustainability.tex
11/16 changes Maneesh

\section{Resource management}
\textcolor{red}{I'm making this up -- Goncalo please comment/correct.}
%
As an open source environment developed over almost 10 years, Bonsai 
has well-established user-community-based management structures.
%
The core system and the majority of extension packages are hosted on a
publicly accessible website at \url{bonsai-rx.org}.
%
Package installation and updates are supported by standalone package
management software included within the core Bonsai distribution.
%
This ensures that bugfixes and enhancements are easily available but
under full user control (to ensure working experiments are not
disrupted by changes in the system).

Best-practice code review procedures are followed prior to including
package updates.  \textcolor{red}{Goncalo/Joaquin -- could you expand
  on this?}

\textcolor{red}{Testing}

The software engineers recruited to this project will work with our
collaborators at Neurogears Ltd, to strengthen and maintain these
practices.  Once fully familiar with the Bonsai core architecture they
will participate in review of pull requests and other code
contributions.

% To complement the management structure document, proposals may outline
% how the resource management and advisory structures will operate including
% outlining any review procedures.

% Additionally, opportunities for staff training and support to ensure their
% continuous development should be identified to safeguard the successful
% delivery of the resource.

\section{Long-term sustainability planning}

Bonsai is a freely available open-source resource, already widely adopted within the systems neuroscience and behavioural science communities (see Fig~\ref{fig:bonsai}).  
%
Current contributors include other large open projects (OpenEphys, \textcolor{red}{\dots}), as well as companies developing and supplying new hardware devices  (\textcolor{red}{\dots}).
%
This level of community uptake, engagement, and commitment provides a foundation for long-term sustainability of the resource.

Neurogears Ltd is a keystone of this community and a key a collaborator in the current proposal.  As a company founded by the original developer of Bonsai, it helps to support and expand the user base well beyond the life of the current proposal.  Neurogears provides quality control and curation for core Bonsai code.  \textcolor{red}{other ways NG contributes ...}

The specific contributions developed as part of the current project will form new components within the core Bonsai package. 
%
They will be developed with input from the advisory panel, early public distribution, user testing within the SWC, and planned workshops, ensuring close user engagement from the beginning of the development cycle and so building momentum for their use, support and dissemination.
%
\textcolor{red}{Can we say that NG will take over responsibility for maintenance and support beyond the project?}

### 2021-11-16 application/detailsOfResource.tex
11/16 changes Maneesh
Merged Figs 2 and 3

@@
 
-We intend to improve an \textbf{existing resource}, the Bonsai ecosystem. We
-propose to create software infrastructure to enable the\textbf{ integration of advanced data analysis methods} into
+We propose to create software infrastructure to enable the\textbf{ integration of advanced data analysis methods} into
 Bonsai (Section~\ref{sec:infrastructure}).
@@
 \begin{figure}
-  \includegraphics[width=4in]{figures/proposed_bonsai_extensions.png}
+    \begin{center}
+        \includegraphics[width=5in]{figures/proposed_bonsai_extensions_combined.png}
         \caption{Proposed extensions to Bonsai}
         \label{fig:proposedBonsaiExtensions}
+    \end{center}
 \end{figure}
@@
 the close-loop experimental control depicted in
-Figure~\ref{fig:bci} cannot be currently implemented in
+Figure~\ref{fig:proposedBonsaiExtensions}b cannot be currently implemented in
 Bonsai. It represents the control of an experiment where a subject guides the
@@
 As an example, consider the close-loop system for the control of a computer
-cursor with neural activity represented in Figure~\ref{fig:bci}a. The Bonsai
-graphical program for this system is given in Figure~\ref{fig:bci}b. The
+cursor with neural activity represented in
+Figure~\ref{fig:proposedBonsaiExtensions}b. The Bonsai
+graphical program for this system is given in
+Figure~\ref{fig:proposedBonsaiExtensions}c. The
 objects instantiating the SpikeThresholdDetector, the KiloSortSpikeSorter and
@@
 and the object instantiating the TimeSeriesForestClassifier will implement the
-\texttt{Predictor} interface (green; Figure~\ref{fig:bci}b, top).
+\texttt{Predictor} interface (green; Figure~\ref{fig:proposedBonsaiExtensions}c, top).
 %
@@
 the GPFAlatentsExtractor will be learnt offline in an unsupervised way by an
-\texttt{Estimator} (red) in a second pipeline (Figure~\ref{fig:bci}b, middle).
+\texttt{Estimator} (red) in a second pipeline
+(Figure~\ref{fig:proposedBonsaiExtensions}c, middle).
 Finally, the parameters of the TimeSeriesForestClassifier will also be learnt
 offline, but in a supervised way, by another \texttt{Estimator} in a third
-pipeline (Figure~\ref{fig:bci}b, bottom).
+pipeline (Figure~\ref{fig:proposedBonsaiExtensions}c, bottom).
 
+\begin{comment}
 \begin{figure}
@@
 
-\begin{comment}
     The spike detector transforms membrane
@@
 We propose to add functionality into Bonsai for the implementation of
-brain-computer-interface (BCI) applications illustrated in Figure~\ref{fig:bci}. In
+brain-computer-interface (BCI) applications illustrated in
+Figure~\ref{fig:proposedBonsaiExtensions}b. In
 this implementation voltage recordings from the brain will be converted into

### 2021-11-16 application/Je-S_sections.tex
Updates from Overleaf

@@
 While Bonsai is already adopted by hundreds of scientists worldwide to provide interactive experimental control, we aim to extend Bonsai’s functionality with a toolbox of online and offline Machine Intelligence tools for analysis of behavioural and neural data (video-based analysis of behavioural motifs, real-time and off-line analysis of neural signals), and create an open-access platform for software sharing and integration with multiple programming languages.
-These to the Bonsai ecosystem will enable new types of research, increase user base, and dramatically improve efficiency and reproducibility of research across behavioural and brain sciences.
+Enhancing Bonsai's ecosystem will enable new types of research, increase user base, and dramatically improve efficiency and reproducibility of research across behavioural and brain sciences.
 [44 chars over]

### 2021-11-16 application/detailsOfResource.tex
Updates from Overleaf

--- 2021-11-16 application/detailsOfResource.tex
@@
 
-We intend to improve an \textbf{existing resource}, the Bonsai ecosystem. We
-propose to create software infrastructure to enable the\textbf{ integration of advanced data analysis methods} into
+We intend to create software infrastructure to enable the\textbf{ integration of advanced data analysis methods} into
 Bonsai (Section~\ref{sec:infrastructure}).

### 2021-11-16 application/backgroundToTheResource.tex

@@
 
-Our proposed enhancements will \textbf{enhance, reinforce and support this key neuroinformatics
-  resource to streamline reproducible experimental design, data collection and analysis, while making standardised implementations of cutting-edge machine-learning methodologies widely available to the community}, potentially creating the most powerful experimental
+The proposed project will \textbf{enhance, reinforce and support} this key neuroinformatics
+  resource to streamline \textbf{reproducible experimental design, data collection and analysis}, while making \textbf{standardised implementations of cutting-edge machine-learning methodologies widely available to the community}, potentially creating the most powerful experimental
 control and analysis tool available to neuroscientists, psychologists and

### 2021-11-16 application/detailsOfResource.tex

--- 2021-11-16 application/detailsOfResource.tex
@@
 
-We propose to create software infrastructure to enable the\textbf{ integration of advanced data analysis methods} into
+We intend to create software infrastructure to enable the\textbf{ integration of advanced data analysis methods} into
 Bonsai (Section~\ref{sec:infrastructure}).

### 2021-11-16 application/backgroundToTheResource.tex
Added caption to wrapfigure

--- 2021-11-16 application/backgroundToTheResource.tex
@@
 \setlength{\columnsep}{1em}
-\begin{wrapfigure}{r}{0.5\textwidth}
-  \includegraphics[width=0.5\columnwidth,viewport=65 50 500 800,clip]{figures/bonsai.pdf}
+\begin{wrapfigure}{R}{0.5\textwidth}
+  % \includegraphics[width=0.5\columnwidth,viewport=65 50 500 800,clip]{figures/bonsai_noCaption.pdf}
+    \includegraphics[width=0.48\textwidth]{figures/roadmap-bbsrc_noCaption.png}
+
+  \caption{Bonsai development timmeline and community adoption. (A) Timeline of
+    milestones and landmark package releases. (B) Example neuroscience
+    experiments developed in Bonsai over the years. (C) Example Bonsai workflow
+    for tracking the positioon of objects in a video. (D) Worldwide
+    distribution of visitors to the bonsai-rx.org website. (E)
+    Number of downloads for each Bonsai version. (F) Cumulative number of
+    registered forum users, citations of the main publication, and number of
+    training events, by year (bottom panels), versus release and development
+    cycle (top panel), measured in commits to main repository.}
   \label{fig:bonsai}
+
 \end{wrapfigure}
-\stepcounter{figure}
+% \stepcounter{figure}
 

### 2021-11-16 application/detailsOfResource.tex
Improved caption to fig:proposedBonsaiExtensions

--- 2021-11-16 application/detailsOfResource.tex
@@
 
-\begin{figure}
-    \begin{center}
+\begin{figure} \begin{center}
     \includegraphics[width=5in]{figures/proposed_bonsai_extensions_combined.png}
-        \caption{Proposed extensions to Bonsai}
+    \caption{Extensions to Bonsai. (A) we propose to integrate machine learning
+    functionality into the experimental control loop by building software
+    infrastructure to allow Bonsai to communicate with existing machine
+    learning software written in Python, R and Matlab. (B) flow chart
+    representing the BCI application that we propose to implement with the
+    machine learning functionality to be added to Bonsai. (C) Bonsai graphic
+    program implementing the BCI application in (B).}
+
         \label{fig:proposedBonsaiExtensions}

### 2021-11-16 application/Je-S_sections.tex
Updates from Overleaf

@@
 
-Scientists studying brain and behaviour dedicate the majority of their time to the design of experimental software protocols that control hardware and to data analysis, with least amount spent on data acquisition itself.
-Moreover, hundreds of neuroscience research groups worldwide develop their own software tools for experimental control and data analysis in different programming languages, leading to inefficiencies resulting from time-loss spent on writing software (often by individuals lacking programming expertise), the re-invention of software with the same functionality and concomitant lack of experimental reproducibility across laboratories.
-
-Here we propose to provide a community software tool that will dramatically increase the efficiency of experimental control and data analysis by providing a new set of functionalities to an existing software platform, Bonsai.
-Bonsai is an open-source reactive visual programming environment that naturally orchestrates complex networks of sensors and hardware into workflows and allows integration of multiple programming languages into its framework.
+Scientists studying brain and behaviour dedicate the majority of their time designing experiments and analysing the data, with least amount spent on data acquisition itself, which may impact the quality of data.
+Moreover, hundreds of neuroscience research groups worldwide develop their own experimental and analytical tools, most using different programming languages, leading to  inefficiencies in data sharing and analysis, and impacting reproducibility (i.e.
+how easy it is for someone else to do the same experiment).
+Here we propose to provide the neuroscience community a software tool that will dramatically increase the efficiency of experimental control and data analysis.
+We will do so by developing a new set of functionalities to an existing software platform, Bonsai.
+Bonsai is an open-source visual programming environment that naturally orchestrates complex networks of sensors and hardware into workflows and allows integration of multiple programming languages into its framework.
 Thus far, Bonsai has been adopted by hundreds of scientists worldwide to provide interactive experimental control in behavioural and brain sciences.
@@
 While Bonsai is already adopted by hundreds of scientists worldwide to provide interactive experimental control, we aim to extend Bonsai’s functionality with a toolbox of online and offline Machine Intelligence tools for analysis of behavioural and neural data (video-based analysis of behavioural motifs, real-time and off-line analysis of neural signals), and create an open-access platform for software sharing and integration with multiple programming languages.
-Enhancing Bonsai's ecosystem will enable new types of research, increase user base, and dramatically improve efficiency and reproducibility of research across behavioural and brain sciences.
+Enhancing Bonsai's ecosystem will enable new types of research, increase and diversify user base, and dramatically improve efficiency and reproducibility of research across behavioural and brain sciences.
 [44 chars over]

### 2021-11-16 application/backgroundToTheResource.tex
Updates from Overleaf

--- 2021-11-16 application/backgroundToTheResource.tex
@@
 The proposed project will \textbf{enhance, reinforce and support} this key neuroinformatics
-  resource to streamline \textbf{reproducible experimental design, data collection and analysis}, while making \textbf{standardised implementations of cutting-edge machine-learning methodologies widely available to the community}, potentially creating the most powerful experimental
+resource to streamline \textbf{reproducible experimental design, data collection and analysis}, while making \textbf{standardised implementations of cutting-edge machine-learning methodologies} widely available to the community, potentially creating the most powerful experimental
 control and analysis tool available to neuroscientists, psychologists and

### 2021-11-16 application/caseForSupport.tex
11/16 changes Maneesh
Added package float and fixed problem with 11point
Updates from Overleaf

@@
 
-\section{Resource management}
+% \section{Resource management}
 
-\input resourceManagement
+\input resourceManagementandSustainability
 
-\section{Long term sustainability planning}
+% \section{Long term sustainability planning}
 
 \section{Potential for economic and societal impact}
-\textbf{Understanding the brain.} Systems neuroscientists aim to uncover how adaptive behaviour arises
+\textbf{Understanding the brain and behaviour.} Systems neuroscientists aim to uncover how adaptive behaviour arises
 from computation in neural circuits in different brain regions.
-Extending the Bonsai ecosystem will facilitate experiments that connect neural circuits with elemental behavioural processes.
+Extending the Bonsai ecosystem will facilitate new types of experiment that causally connect neural circuits with elemental behavioural processes.
 Long-term, this research will help identify the missing links between circuit dysfunction and behavioural deficits in neurological or mental health disorders, and will pave the way for the development of new neural circuit biomarkers and therapies.
 The Bonsai ecosystem will also enable the experimental testing of emerging theories of brain function and learning, and thus contribute to the neuro-inspired machine learning algorithms that fuel intelligent technologies, increasingly important in our society.
+The enhanced Bonsai platform will also support purely behavioural experiments, in which the animals' posture, body parts or location are tracked in real-time, allowing investigation of environmental impacts and captivit the impact of the envuronmental variables in behaviour but also enhace monitoring of intaractive experimental designs animals or other
 
@@
 
+\input{trackRecord}
+
 \bibliographystyle{apalike}

### 2021-11-16 application/detailsOfResource.tex
Updates from Overleaf

--- 2021-11-16 application/detailsOfResource.tex
@@
 
-\section{Details of the proposed resource enhancements}
+\section{Details of proposed resource enhancements}
 
-We intend to create software infrastructure to enable the\textbf{ integration of advanced data analysis methods} into
-Bonsai (Section~\ref{sec:infrastructure}).
+The core proposed developments will enhance Bonsai's extensibility in two ways.
 %
-To demonstrate the use of this infrastructure, and to provide Bonsai users with
-an initial set of advanced machine intelligence neural data analysis methods,
-we will add to the Bonsai ecosystem machine learning functionality to control
-neuroscience experiments and to analyse data generated by them
+First, we will create a new package infrastructure to enable the integration of software toolboxes implemented in non-native languages into Bonsai.
+%
+Second, we will create a new general abstraction framework within Bonsai suited to the implementation and application of data analytic algorithms.
+%
+Both frameworks will expose public interfaces available to all developers within the user community, thus simplifying and accelerating community-driven development.
+
+Furthermore, we will build on these extensibility developments ourselves to implement a suite of advanced machine intelligence neural data analysis methods that will be made available as one or more core Bonsai packages.
+
+\subsection{Extensions to incorporate non-native code packages into Bonsai}
+
+As it stands, Bonsai is designed to be extensible and is distributed with a package management framework that controls installation of both optional core packages and user-contributed extensions.
+%
+However, the present framework requires extensions to be written in Bonsai's native C\# programming language.
+%
+This specificity restricts the number of potential contributors to the ecosystem.
+%
+It also hinders easy incorporation of published analytic tools, many of which are implemented for interpreted language systems such as Python, R, or MATLAB.
+
+We will extend the Bonsai package interface to allow direct interaction with the runtime execution environments that implement these common languages.  
+%
+Where possible, the extensions will be based on existing pipeline tools that allow C\#-based programs to interact with a number of alternative run-time engines.
+%
+These include:
+%
+Python.NET\footnote{\href{https://github.com/pythonnet/pythonnet}{https://github.com/pythonnet/pythonnet}},
+%
+R.Net\footnote{\href{https://github.com/rdotnet/rdotnet}{https://github.com/rdotnet/rdotnet}}
+%
+and the native .NET interface shipped with MATLAB\footnote{\href{https://uk.mathworks.com/help/matlab/matlab\_external/using-net-from-matlab-an-overview.html}{https://uk.mathworks.com/help/matlab/matlab\_external/using-net-from-matlab-an-overview.html}}.
+%
+If necessary, we will also incorporate direct inter-runtime communication using the messaging library ZeroMQ\footnote{\href{https://zeromq.org/}{https://zeromq.org/}}.
+
+\subsection{General abstraction framework for data processing}
+\label{sec:generality}
+
+The number of current machine learning algorithms is very large and it is
+continuously increasing. We intend to build a small number of machine
+intelligence operators in Bonsai (Section~\ref{sec:bonsai},
+Figure~\ref{fig:bonsai}c) that can be configured to provide the functionality
+of all machine learning algorithms.
+
+For this, we will build on abstractions used in the \texttt{scikit-learn}
+library. The success of this library is, to a large extent, due to its simple
+and general application programming interface \citep[API;][]{buitinckEtAl13}.
+%
+This API consists of three complementary interfaces: an \texttt{Estimator}
+interface for building and fitting models, a \texttt{Predictor} interface for
+making predictions and a \texttt{Transformer} interface for converting data.
+%
+Using only these three interfaces \texttt{scikit-learn} provides
+implementations of a large number of machine learning algorithms for
+dimensionality reduction, clustering, classification and regression.
+
+Inspired by \texttt{scikit-learn}, Bonsai will have three types of operators:
+\texttt{Estimators}, \texttt{Predictors} and \texttt{Transformers}. Each of
+these operators will be instantiated by objects implementing the corresponding
+interface.
+%
+As an example, consider the close-loop system for the control of a computer
+cursor with neural activity represented in
+Figure~\ref{fig:proposedBonsaiExtensions}b. The Bonsai
+graphical program for this system is given in
+Figure~\ref{fig:proposedBonsaiExtensions}c. The
+objects instantiating the SpikeThresholdDetector, the KiloSortSpikeSorter and
+GPFAlatentsExtractor will implement the \texttt{Transformer} interface (blue),
+and the object instantiating the TimeSeriesForestClassifier will implement the
+\texttt{Predictor} interface (green; Figure~\ref{fig:proposedBonsaiExtensions}c, top).
+%
+The parameters for the SpikeThresholdDetector and the KiloSortSpikeSorter will
+be set manually by the user, as is commonly done in practice. The parameters of
+the GPFAlatentsExtractor will be learnt offline in an unsupervised way by an
+\texttt{Estimator} (red) in a second pipeline
+(Figure~\ref{fig:proposedBonsaiExtensions}c, middle).
+Finally, the parameters of the TimeSeriesForestClassifier will also be learnt
+offline, but in a supervised way, by another \texttt{Estimator} in a third
+pipeline (Figure~\ref{fig:proposedBonsaiExtensions}c, bottom).
+
+Using the previous software, we will build C\# wrappers around machine
+intelligence programs written in other languages, so that these programs
+implement the \texttt{Estimator}, \texttt{Predictor} or \texttt{Transformer}
+interfaces and can be integrated into Bonsai as machine intelligence operators.
+
+a package management , and its community of users have
+already extended Bonsai's functionality with several contributed packages.
+%
+However, these packages need to
+be written in C\#, while most current neuroscience data analysis methods are
+written in Python, R or Matlab. We propose to add to Bonsai capabilities to
+communicate with software written in these languages
+(Section~\ref{sec:softwareInfrastructure}).  These capabilities will allow a
+large number of neuroscience data analysis methods to be easily integrated into
+the Bonsai ecosystem and will provide Bonsai users a large repertoire of
+advanced machine learning methods for their behavioural and neural data
+analysis.
+%
+Using these capabilities we will integrate into Bonsai an initial set of
+state-of-the art behavioural and neural data analysis methods
 (Section~\ref{sec:functionality}).
+
+% \subsubsection{Extensibility}
+% \label{sec:extensibility}
+
+% To incorporate machine learning functionality into the Bonsai ecosystem, we
+% could
+% implement this functionality directly into Bonsai, as done in
+% \texttt{scikit-learn}. Alternatively, or in addition, we could build software
+% infrastructure to facilitate the integration of existing machine learning
+% software into Bonsai. We chose to begin with the second alternative because
+% high-quality implementations of machine learning functionality already exist,
+% which can be quickly integrated into the Bonsai ecosystem. In addition, the
+% second alternative is more sustainable, as it will allow the continued addition
+% of new machine learning functionality to Bonsai after the completion of the
+% BBSRC-BBR funding period. Later we may decide to implement some widely-used
+% machine learning functionality directly into Bonsai (e.g., for performance or
+% robustness issues).
 
@@
 translate into groundbreaking scientific discoveries.
-
-Currently Bonsai has its own package manager, and its community of users have
-already extended Bonsai's functionality with several contributed packages.
-%
-However, these packages need to
-be written in C\#, while most current neuroscience data analysis methods are
-written in Python, R or Matlab. We propose to add to Bonsai capabilities to
-communicate with software written in these languages
-(Section~\ref{sec:softwareInfrastructure}).  These capabilities will allow a
-large number of neuroscience data analysis methods to be easily integrated into
-the Bonsai ecosystem and will provide Bonsai users a large repertoire of
-advanced machine learning methods for their behavioural and neural data
-analysis.
-%
-Using these capabilities we will integrate into Bonsai an initial set of
-state-of-the art behavioural and neural data analysis methods
-(Section~\ref{sec:functionality}).
 
@@
 (Section~\ref{sec:extensibility}).
-
-\subsubsection{Generality}
-\label{sec:generality}
-
-The number of current machine learning algorithms is very large and it is
-continuously increasing. We intend to build a small number of machine
-intelligence operators in Bonsai (Section~\ref{sec:bonsai},
-Figure~\ref{fig:bonsai}c) that can be configured to provide the functionality
-of all machine learning algorithms.
-
-For this, we will build on abstractions used in the \texttt{scikit-learn}
-library. The success of this library is, to a large extent, due to its simple
-and general application programming interface \citep[API;][]{buitinckEtAl13}.
-%
-This API consists of three complementary interfaces: an \texttt{Estimator}
-interface for building and fitting models, a \texttt{Predictor} interface for
-making predictions and a \texttt{Transformer} interface for converting data.
-%
-Using only these three interfaces \texttt{scikit-learn} provides
-implementations of a large number of machine learning algorithms for
-dimensionality reduction, clustering, classification and regression.
-
-Inspired by \texttt{scikit-learn}, Bonsai will have three types of operators:
-\texttt{Estimators}, \texttt{Predictors} and \texttt{Transformers}. Each of
-these operators will be instantiated by objects implementing the corresponding
-interface.
-%
-As an example, consider the close-loop system for the control of a computer
-cursor with neural activity represented in
-Figure~\ref{fig:proposedBonsaiExtensions}b. The Bonsai
-graphical program for this system is given in
-Figure~\ref{fig:proposedBonsaiExtensions}c. The
-objects instantiating the SpikeThresholdDetector, the KiloSortSpikeSorter and
-GPFAlatentsExtractor will implement the \texttt{Transformer} interface (blue),
-and the object instantiating the TimeSeriesForestClassifier will implement the
-\texttt{Predictor} interface (green; Figure~\ref{fig:proposedBonsaiExtensions}c, top).
-%
-The parameters for the SpikeThresholdDetector and the KiloSortSpikeSorter will
-be set manually by the user, as is commonly done in practice. The parameters of
-the GPFAlatentsExtractor will be learnt offline in an unsupervised way by an
-\texttt{Estimator} (red) in a second pipeline
-(Figure~\ref{fig:proposedBonsaiExtensions}c, middle).
-Finally, the parameters of the TimeSeriesForestClassifier will also be learnt
-offline, but in a supervised way, by another \texttt{Estimator} in a third
-pipeline (Figure~\ref{fig:proposedBonsaiExtensions}c, bottom).
 
@@
 \end{comment}
-
-\subsubsection{Extensibility}
-\label{sec:extensibility}
-
-To incorporate machine learning functionality into the Bonsai ecosystem, we
-could
-implement this functionality directly into Bonsai, as done in
-\texttt{scikit-learn}. Alternatively, or in addition, we could build software
-infrastructure to facilitate the integration of existing machine learning
-software into Bonsai. We chose to begin with the second alternative because
-high-quality implementations of machine learning functionality already exist,
-which can be quickly integrated into the Bonsai ecosystem. In addition, the
-second alternative is more sustainable, as it will allow the continued addition
-of new machine learning functionality to Bonsai after the completion of the
-BBSRC-BBR funding period. Later we may decide to implement some widely-used
-machine learning functionality directly into Bonsai (e.g., for performance or
-robustness issues).
-
-Currently Bonsai has its own package manager, and its community of users have
-already extended Bonsai's functionality with several contributed packages
-(e.g., \ldots). However, these packages need to be written in C\#, while most
-current data analysis methods are written in Python, R or Matlab. We propose to
-add to Bonsai capabilities to communicate with software written in these
-languages.
-
-Fortunately, there exist software that allow C\# program
-to call and be called by programs in these other languages. For the
-communication of C\# programs with Python we will use
-Python.NET\footnote{\href{https://github.com/pythonnet/pythonnet}{https://github.com/pythonnet/pythonnet}},
-with R programs we will use
-R.Net\footnote{\href{https://github.com/rdotnet/rdotnet}{https://github.com/rdotnet/rdotnet}}
-and with Matlab we will use the built in Matlab .NET
-interface\footnote{\href{https://uk.mathworks.com/help/matlab/matlab\_external/using-net-from-matlab-an-overview.html}{https://uk.mathworks.com/help/matlab/matlab\_external/using-net-from-matlab-an-overview.html}}.
-In cases where these software cannot support some type of communication between
-C\# and another language we will use the messaging library
-ZeroMQ\footnote{\href{https://zeromq.org/}{https://zeromq.org/}}.
-
-Using the previous software, we will build C\# wrappers around machine
-intelligence programs written in other languages, so that these programs
-implement the \texttt{Estimator}, \texttt{Predictor} or \texttt{Transformer}
-interfaces and can be integrated into Bonsai as machine intelligence operators.
 

### 2021-11-16 application/trackRecord.tex
Updates from Overleaf

@@
 The team responsible for the proposed project comprises two academic
-institutes, the Gatsby Computational Neuroscience Unit (GCNU) and the Sainsbury
+institutes at University College London, the Gatsby Computational Neuroscience Unit (GCNU) and the Sainsbury
 Wellcome Centre for Neural Circuits and Behavior (SWC), and a private

### 2021-11-16 application/backgroundToTheResource.tex

--- 2021-11-16 application/backgroundToTheResource.tex
@@
 
-Our proposed enhancements will \textbf{enhance, reinforce and support this key neuroinformatics
-  resource to streamline reproducible experimental design, data collection and analysis, while making standardised implementations of cutting-edge machine-learning methodologies widely available to the community}, potentially creating the most powerful experimental
+The proposed project will \textbf{enhance, reinforce and support} this key neuroinformatics
+resource to streamline \textbf{reproducible experimental design, data collection and analysis}, while making \textbf{standardised implementations of cutting-edge machine-learning methodologies} widely available to the community, potentially creating the most powerful experimental
 control and analysis tool available to neuroscientists, psychologists and

### 2021-11-16 application/backgroundToTheResource.tex
Added href to online figures

@@
 \begin{wrapfigure}{R}{0.5\textwidth}
-  % \includegraphics[width=0.5\columnwidth,viewport=65 50 500 800,clip]{figures/bonsai_noCaption.pdf}
-    \includegraphics[width=0.48\textwidth]{figures/roadmap-bbsrc_noCaption.png}
+    \href{http://www.gatsby.ucl.ac.uk/~rapela/bbsrc-bbr21/roadmap-bbsrc_noCaption.png}{\includegraphics[width=0.48\textwidth]{figures/roadmap-bbsrc_noCaption.png}}
 
@@
     training events, by year (bottom panels), versus release and development
-    cycle (top panel), measured in commits to main repository.}
+    cycle (top panel), measured in commits to main repository. Click on the
+    image to see its online version.}
+
   \label{fig:bonsai}

### 2021-11-16 application/detailsOfResource.tex
Added href to online figures

@@
 \begin{figure} \begin{center}
-    \includegraphics[width=5in]{figures/proposed_bonsai_extensions_combined.png}
+
+    \href{http://www.gatsby.ucl.ac.uk/~rapela/bbsrc-bbr21/proposed_bonsai_extensions_combined.png}{\includegraphics[width=5in]{figures/proposed_bonsai_extensions_combined.png}}
+
     \caption{Extensions to Bonsai. (A) we propose to integrate machine learning
@@
     machine learning functionality to be added to Bonsai. (C) Bonsai graphic
-    program implementing the BCI application in (B).}
+    program implementing the BCI application in (B). Click on the image to see
+    its online version.}
 

### 2021-11-16 application/caseForSupport.tex

@@
-\documentclass[a4paper,11point]{article}
+\documentclass[a4paper,11pt]{article}
 \usepackage[margin=2cm]{geometry}
@@
 
+\usepackage{float}
 \usepackage{wrapfig}
@@
 The Bonsai ecosystem will also enable the experimental testing of emerging theories of brain function and learning, and thus contribute to the neuro-inspired machine learning algorithms that fuel intelligent technologies, increasingly important in our society.
-The enhanced Bonsai platform will also support purely behavioural experiments, in which the animals' posture, body parts or location are tracked in real-time, allowing investigation of environmental impacts and captivit the impact of the envuronmental variables in behaviour but also enhace monitoring of intaractive experimental designs animals or other
+The enhanced Bonsai platform will also support purely behavioural experiments, in which the animals' posture, body parts or location are tracked in real-time, allowing investigation of environmental variables or captivity on their behaviour, especially important for assessment of animal welfare.
 

### 2021-11-16 application/Je-S_sections.tex

@@
 how easy it is for someone else to do the same experiment).
+
 Here we propose to provide the neuroscience community a software tool that will dramatically increase the efficiency of experimental control and data analysis.
@@
 Bonsai is an open-source visual programming environment that naturally orchestrates complex networks of sensors and hardware into workflows and allows integration of multiple programming languages into its framework.
+Importantly, Bonsai users require minimal programming experience.
+
 Thus far, Bonsai has been adopted by hundreds of scientists worldwide to provide interactive experimental control in behavioural and brain sciences.
@@
 To understand the brain, scientists aim to explain how animal behaviour relates to the activity in neural circuits.
-This necessitates design and precise control of behavioural experiments wherein experimental animals perform particular tasks while experimenters either record or manipulate neural activity in candidate neural circuits.
+This requires design and precise control of behavioural experiments, wherein animals perform particular tasks while experimenters either record or manipulate neural activity in specific neural circuits.
 Such experiments require data acquisition software that integrates and controls hardware from multiple recording devices (cameras, electrodes, sensors), and analysis tools that can interpret large and complex behavioural and neural datasets.
@@
 The fields of behavioural and brain sciences have consequently suffered from both inefficiency and poor reproducibility, due to disparate data acquisition and analysis solutions created independently across laboratories.
-To address these challenges, we propose  to  extend,  enhance,  maintain  and  support \textbf{Bonsai},  a  fully  integrated  software  environment  to enable cutting-edge reproducible laboratory systems neuroscience experiments in animal models, with a particular emphasis on machine-intelligence-enabled, real-time neuroinformatics methods.
+
+To address these challenges, we propose  to  extend,  enhance,  maintain  and  support \textbf{Bonsai},  a  fully  integrated  software  environment  to enable cutting-edge reproducible systems neuroscience experiments using animal models, with a particular emphasis on machine-intelligence-enabled, real-time neuroinformatics methods.
 While Bonsai is already adopted by hundreds of scientists worldwide to provide interactive experimental control, we aim to extend Bonsai’s functionality with a toolbox of online and offline Machine Intelligence tools for analysis of behavioural and neural data (video-based analysis of behavioural motifs, real-time and off-line analysis of neural signals), and create an open-access platform for software sharing and integration with multiple programming languages.

### 2021-11-16 application/caseForSupport.tex
Added \description in section Potential for economic and societal impact

@@
 \section{Potential for economic and societal impact}
-\textbf{Understanding the brain and behaviour.} Systems neuroscientists aim to uncover how adaptive behaviour arises
-from computation in neural circuits in different brain regions.
-Extending the Bonsai ecosystem will facilitate new types of experiment that causally connect neural circuits with elemental behavioural processes.
-Long-term, this research will help identify the missing links between circuit dysfunction and behavioural deficits in neurological or mental health disorders, and will pave the way for the development of new neural circuit biomarkers and therapies.
-The Bonsai ecosystem will also enable the experimental testing of emerging theories of brain function and learning, and thus contribute to the neuro-inspired machine learning algorithms that fuel intelligent technologies, increasingly important in our society.
-The enhanced Bonsai platform will also support purely behavioural experiments, in which the animals' posture, body parts or location are tracked in real-time, allowing investigation of environmental variables or captivity on their behaviour, especially important for assessment of animal welfare.
 
-\textbf{Improve reproducibility in behavioural and brain sciences.} Lack of reproducible results slows down scientific progress and incurs personal and economic costs to scientists and funders.
-In neuroscience,  many behavioural and neural animal studies have been to be difficult to reproduce, and experiments are rarely replicated across laboratories.
-By creating a resource that standardises and enables reproducible experimental design/control, as well as analysis of neural and behavioural data, the Bonsai ecosystem will allow multiple laboratories to implement and align experimental and analytical protocols as closely as possible.
+\begin{description}
 
-\textbf{Improve efficiency of research.} Most neuroscience laboratories worldwide develop custom, unique solutions to their experimental control and analysis needs.
-Some laboratories also lack the technical know-how to either develop their own software for experimental hardware control or write software for the state-of-the-art analysis of rich neural and behavioural data, increasingly reliant on latest machine learning algorithms.
-The proposed upgrading of the open-source and community empowered Bonsai ecosystem with a toolbox of online and offline analytical tools, that can dovetail with closed-loop experimental requirements, will (i) allow the deployment of such tools into laboratories that lack the expertise, (ii) reduce the need to reinvent the same tools in multiple labs (cost/time efficiency) and (iii) concomitantly streamline and standardise the experimental data processing streams in systems neuroscience.
+    \item[Understanding the brain and behaviour] Systems neuroscientists aim to
+        uncover how adaptive behaviour arises from computation in neural
+        circuits in different brain regions. Extending the Bonsai ecosystem
+        will facilitate new types of experiment that causally connect neural
+        circuits with elemental behavioural processes. Long-term, this research
+        will help identify the missing links between circuit dysfunction and
+        behavioural deficits in neurological or mental health disorders, and
+        will pave the way for the development of new neural circuit biomarkers
+        and therapies. The Bonsai ecosystem will also enable the experimental
+        testing of emerging theories of brain function and learning, and thus
+        contribute to the neuro-inspired machine learning algorithms that fuel
+        intelligent technologies, increasingly important in our society.  The
+        enhanced Bonsai platform will also support purely behavioural
+        experiments, in which the animals' posture, body parts or location are
+        tracked in real-time, allowing investigation of environmental variables
+        or captivity on their behaviour, especially important for assessment of
+        animal welfare.
+
+    \item[Improve reproducibility in behavioural and brain sciences] Lack of
+        reproducible results slows down scientific progress and incurs personal
+        and economic costs to scientists and funders.  In neuroscience,  many
+        behavioural and neural animal studies have been to be difficult to
+        reproduce, and experiments are rarely replicated across laboratories.
+        By creating a resource that standardises and enables reproducible
+        experimental design/control, as well as analysis of neural and
+        behavioural data, the Bonsai ecosystem will allow multiple laboratories
+        to implement and align experimental and analytical protocols as closely
+        as possible.
+
+    \item[Improve efficiency of research] Most neuroscience laboratories
+        worldwide develop custom, unique solutions to their experimental
+        control and analysis needs. Some laboratories also lack the technical
+        know-how to either develop their own software for experimental hardware
+        control or write software for the state-of-the-art analysis of rich
+        neural and behavioural data, increasingly reliant on latest machine
+        learning algorithms. The proposed upgrading of the open-source and
+        community empowered Bonsai ecosystem with a toolbox of online and
+        offline analytical tools, that can dovetail with closed-loop
+        experimental requirements, will (i) allow the deployment of such tools
+        into laboratories that lack the expertise, (ii) reduce the need to
+        reinvent the same tools in multiple labs (cost/time efficiency) and
+        (iii) concomitantly streamline and standardise the experimental data
+        processing streams in systems neuroscience.
+
+\end{description}
 

### 2021-11-16 application/backgroundToTheResource.tex

@@
 drive complex behaviour are manifest in a broad population of neurons rather
-than in single cells.  So current neural recording emphasises high-dimensional
+than in single cells. Current neural recording emphasises high-dimensional
 signals from 100s to 1000s of neurons monitored using multi-electrode arrays or cellular-resolution imaging, with a wide array of probe configurations and neural targets.
 %
-This complexity of experiments limits the potential for database-driven progress in systems neuroscience.
-%
-The number of possible experiments is astronomical, and a new question must
-usually be answered by designing and conducting a new experiment rather than
-mining the data from an old one.
+\textbf{This complexity of experiments limits the potential for database-driven progress in systems neuroscience}.
+The number of possible experiments is astronomical, and new questions are typically answered by designing and conducting new experiments rather than mining the data from an old one.
 
@@
 \setlength{\columnsep}{1em}
-\begin{wrapfigure}{R}{0.5\textwidth}
-    \href{http://www.gatsby.ucl.ac.uk/~rapela/bbsrc-bbr21/roadmap-bbsrc_noCaption.png}{\includegraphics[width=0.48\textwidth]{figures/roadmap-bbsrc_noCaption.png}}
+\begin{wrapfigure}{R}{0.6\textwidth}
+    \href{http://www.gatsby.ucl.ac.uk/~rapela/bbsrc-bbr21/roadmap-bbsrc_noCaption.png}{\includegraphics[width=0.6\textwidth]{figures/roadmap-bbsrc_noCaption.png}}
 
-  \caption{Bonsai development timmeline and community adoption. (A) Timeline of
-    milestones and landmark package releases. (B) Example neuroscience
-    experiments developed in Bonsai over the years. (C) Example Bonsai workflow
-    for tracking the positioon of objects in a video. (D) Worldwide
+  \caption{\small Bonsai development timeline and community adoption. (A) Timeline of
+    milestones and landmark package releases. (B) Example capabilities 
+    within Bonsai. (C) Bonsai workflow for tracking the position of objects
+    in a video. (D) Worldwide
     distribution of visitors to the bonsai-rx.org website. (E)
@@
 \end{wrapfigure}
-% \stepcounter{figure}
 

### 2021-11-16 application/Je-S_sections.tex

@@
 We will do so by developing a new set of functionalities to an existing software platform, Bonsai.
-Bonsai is an open-source visual programming environment that naturally orchestrates complex networks of sensors and hardware into workflows and allows integration of multiple programming languages into its framework.
+Bonsai is a fully integrated software environment for cutting-edge reproducible experiments using animal models, which provides machine-intelligence, real-time neuroinformatics methods.
 Importantly, Bonsai users require minimal programming experience.
@@
 In this proposal, we aim to extend Bonsai’s functionality with a toolbox of online and offline Machine Intelligence tools for analysis of behavioural and neural datasets, and to create an open-access platform for software sharing.
-Bonsai’s enhanced functionality will enable new types of research via online experimental control, speed up discovery and improve efficiency by (i) providing access to such tools into laboratories lacking expertise, (ii) reducing the need to reinvent the same tools in multiple labs and (iii) standardising the data processing streams, thus increasing reproducibility across laboratories.
+Bonsai’s enhanced functionality will enable new types of research, and speed up discovery and improve efficiency by (i) providing access to such tools to laboratories lacking expertise, (ii) reducing the need to reinvent the same tools in multiple labs and (iii) standardising the data processing streams, thus increasing reproducibility across laboratories.
 We believe this effort will enable and accelerate new discoveries in how the brain generates behaviour.
@@
 
-To understand the brain, scientists aim to explain how animal behaviour relates to the activity in neural circuits.
-This requires design and precise control of behavioural experiments, wherein animals perform particular tasks while experimenters either record or manipulate neural activity in specific neural circuits.
-Such experiments require data acquisition software that integrates and controls hardware from multiple recording devices (cameras, electrodes, sensors), and analysis tools that can interpret large and complex behavioural and neural datasets.
-Progress is held back by the lack of standardised tools for rapid design and implementation of experimental protocols, and the difficulty of integrating state-of-the-art data processing and neuroinformatics into custom experimental designs.
+To understand the brain, scientists aim to explain how animal behaviour relates to neural activity.
+This requires the design and precise control of behavioural experiments, wherein animals perform particular tasks while experimenters either record or manipulate neural activity in specific neural circuits.
+Such experiments require data acquisition software that integrates and controls hardware from multiple recording devices (cameras, electrodes, sensors), and analysis tools that can interpret large and complex datasets.
+Progress is held back by the lack of standardised tools for design and implementation of experimental protocols, and the difficulty of integrating state-of-the-art data processing and neuroinformatics into custom experimental designs.
 The fields of behavioural and brain sciences have consequently suffered from both inefficiency and poor reproducibility, due to disparate data acquisition and analysis solutions created independently across laboratories.
@@
 To address these challenges, we propose to extend, enhance, maintain and support \textbf{Bonsai}, a fully integrated software environment to enable cutting-edge reproducible systems neuroscience experiments using animal models, with a particular emphasis on machine-intelligence-enabled, real-time neuroinformatics methods.
-While Bonsai is already adopted by hundreds of scientists worldwide to provide interactive experimental control, we aim to extend Bonsai’s functionality with a toolbox of online and offline Machine Intelligence tools for analysis of behavioural and neural data (video-based analysis of behavioural motifs, real-time and off-line analysis of neural signals), and create an open-access platform for software sharing and integration with multiple programming languages.
-Enhancing Bonsai's ecosystem will enable new types of research, increase and diversify user base, and dramatically improve efficiency and reproducibility of research across behavioural and brain sciences.
-[44 chars over]
+While Bonsai is already adopted by hundreds of scientists worldwide, we aim to extend Bonsai’s functionality with a toolbox of online and offline Machine Intelligence tools for analysis of behavioural and neural data (video-based analysis of behavioural motifs, real-time and offline analysis of neural signals), and create an open-access platform for software sharing and integration with multiple programming languages.
+Enhancing Bonsai's ecosystem will be a game-changer for behavioural and brain science experiments by enabling new types of research, increasing and diversifying user base, and dramatically improve efficiency and reproducibility of research.
 

### 2021-11-16 application/backgroundToTheResource.tex
Remove from the caption the links to the online images.

@@
     training events, by year (bottom panels), versus release and development
-    cycle (top panel), measured in commits to main repository. Click on the
-    image to see its online version.}
+    cycle (top panel), measured in commits to main repository.}
 

### 2021-11-16 application/bonsai.tex
Updates from Overleaf

@@
 % programming language that 
-Its design emphasizes performance, flexibility, and ease-of-use,
+Its design emphasises performance, flexibility, and ease-of-use,
 allowing scientists with no previous programming experience to quickly develop

### 2021-11-16 application/backgroundToTheResource.tex

--- 2021-11-16 application/backgroundToTheResource.tex
@@
   neuroscience}
+
+\setlength{\columnsep}{1em}
+\begin{wrapfigure}{R}{0.5\textwidth}
+    \href{http://www.gatsby.ucl.ac.uk/~rapela/bbsrc-bbr21/roadmap-bbsrc_noCaption.png}{\includegraphics[width=0.5\textwidth]{figures/roadmap-bbsrc_noCaption.png}}
+
+  \caption{\footnotesize Bonsai development timeline and community adoption. (A) Timeline of
+    milestones and landmark package releases. (B) Example capabilities 
+    within Bonsai. (C) Bonsai workflow for tracking the position of objects
+    in a video. (D) Worldwide
+    distribution of visitors to the bonsai-rx.org website. (E)
+    Number of downloads for each Bonsai version. (F) Cumulative number of
+    registered forum users, citations of the main publication, and number of
+    training events, by year (bottom panels), versus release and development
+    cycle (top panel), measured in commits to main repository. 
+    % Click on the image to see its online version.
+    }
+
+  \label{fig:bonsai}
+\end{wrapfigure}
 
@@
 Although such tools exist in genetics, astronomy, physics and medicine (Fish et al., 2016; Abdalla et al., 2018, CERN Education, Communications and Outreach Group, 2018, Dickinson et al., 2016, Bycroft et al., 2018), they are mostly lacking in neuroscience, with documented consequences for the reliability of ``discoveries'' (Baker, 2016; Botvinik-Nezer et al., 2020; Button et al., 2013).
-
-\setlength{\columnsep}{1em}
-\begin{wrapfigure}{R}{0.6\textwidth}
-    \href{http://www.gatsby.ucl.ac.uk/~rapela/bbsrc-bbr21/roadmap-bbsrc_noCaption.png}{\includegraphics[width=0.6\textwidth]{figures/roadmap-bbsrc_noCaption.png}}
-
-  \caption{\small Bonsai development timeline and community adoption. (A) Timeline of
-    milestones and landmark package releases. (B) Example capabilities 
-    within Bonsai. (C) Bonsai workflow for tracking the position of objects
-    in a video. (D) Worldwide
-    distribution of visitors to the bonsai-rx.org website. (E)
-    Number of downloads for each Bonsai version. (F) Cumulative number of
-    registered forum users, citations of the main publication, and number of
-    training events, by year (bottom panels), versus release and development
-    cycle (top panel), measured in commits to main repository. Click on the
-    image to see its online version.}
-
-  \label{fig:bonsai}
-
-\end{wrapfigure}
 

### 2021-11-16 application/bonsai.tex
Removed live links

@@
 International Brain
-Laboratory\footnote{\href{https://www.internationalbrainlab.com/}{https://www.internationalbrainlab.com/}}
+Laboratory\footnote{https://www.internationalbrainlab.com/}
 and the Allen Institute for Brain
-Science\footnote{\href{https://alleninstitute.org/what-we-do/brain-science/}{https://alleninstitute.org/what-we-do/brain-science/}}.
+Science\footnote{https://alleninstitute.org/what-we-do/brain-science/}.
 %

### 2021-11-16 application/communityDemand.tex
Removed live links

--- 2021-11-14 communityDemand.tex
@@
 explicit proposals and suggestions for the creation of a machine learning
-package\footnote{\href{https://groups.google.com/g/bonsai-users/c/BZ3zOOdv_1c/m/x6OP75frBQAJ?utm_medium=email&utm_source=footer&pli=1}{[Speculative
-Enchancement] Deeplabcut + bonsai (google.com)}}. Even
+package\footnote{https://groups.google.com/g/bonsai-users/c/BZ3zOOdv\_1c/m/x6OP75frBQAJ}. Even
 relatively modest contributions such as integrating support for specific models

### 2021-11-16 application/detailsOfResource.tex
Remove from the caption the links to the online images.
Removed live links

@@
 %
-Python.NET\footnote{\href{https://github.com/pythonnet/pythonnet}{https://github.com/pythonnet/pythonnet}},
+Python.NET\footnote{https://github.com/pythonnet/pythonnet},
 %
-R.Net\footnote{\href{https://github.com/rdotnet/rdotnet}{https://github.com/rdotnet/rdotnet}}
+R.Net\footnote{https://github.com/rdotnet/rdotnet}
 %
-and the native .NET interface shipped with MATLAB\footnote{\href{https://uk.mathworks.com/help/matlab/matlab\_external/using-net-from-matlab-an-overview.html}{https://uk.mathworks.com/help/matlab/matlab\_external/using-net-from-matlab-an-overview.html}}.
+and the native .NET interface shipped with MATLAB\footnote{https://uk.mathworks.com/help/matlab/matlab\_external/using-net-from-matlab-an-overview.html}.
 %
-If necessary, we will also incorporate direct inter-runtime communication using the messaging library ZeroMQ\footnote{\href{https://zeromq.org/}{https://zeromq.org/}}.
+If necessary, we will also incorporate direct inter-runtime communication using the messaging library ZeroMQ\footnote{https://zeromq.org/}.
 
@@
     machine learning functionality to be added to Bonsai. (C) Bonsai graphic
-    program implementing the BCI application in (B). Click on the image to see
-    its online version.}
+    program implementing the BCI application in (B).}
 
@@
         the Datta laboratory, as indicated at
-        \href{http://datta.hms.harvard.edu/research/behavioral-analysis/}{http://datta.hms.harvard.edu/research/behavioral-analysis/}.}.
+        http://datta.hms.harvard.edu/research/behavioral-analysis/.}.
         After trained MoSeq can be used to detect behavioral syllables online.
@@
         decode video data from neural recordings. It is implemented in
-        Python\footnote{\href{https://github.com/themattinthehatt/behavenet}{https://github.com/themattinthehatt/behavenet}}.
+        Python\footnote{https://github.com/themattinthehatt/behavenet}.
         After trained it can be used online to decode video data and detect
@@
         Python\footnote{code for PS-VAE is embedded in the BehaveNet code
-        \href{https://github.com/themattinthehatt/behavenet}{https://github.com/themattinthehatt/behavenet}.
+        https://github.com/themattinthehatt/behavenet.
         Please refer to class \texttt{PSVAE} in 
@@
         which allows the use of external
-        inputs\footnote{\href{https://github.com/joacorapela/kalmanFilter}{https://github.com/joacorapela/kalmanFilter}}.
+        inputs\footnote{https://github.com/joacorapela/kalmanFilter}.
 
@@
         propose to interface Bonsai with a Matlab implementation of
-        PLDS\footnote{\href{https://bitbucket.org/mackelab/pop\_spike\_dyn/src/master/}{https://bitbucket.org/mackelab/pop\_spike\_dyn/src/master/}}
+        PLDS\footnote{https://bitbucket.org/mackelab/pop\_spike\_dyn/src/master/}
         that uses variational inference. PLDS does not provide online estimates
@@
         than continuous ones. We propose to use an R implementation of
-        HMMs\footnote{\href{https://github.com/joacorapela/hiddenMarkovModels}{https://github.com/joacorapela/hiddenMarkovModels}}.
+        HMMs\footnote{https://github.com/joacorapela/hiddenMarkovModels}.
         As GLDSs, once trained, HMMs can be used online to infer the posterior
@@
         random processes. We propose to use a Matlab implementation of
-        GPFA\footnote{\href{https://users.ece.cmu.edu/~byronyu/software/gpfa0203.tgz}{https://users.ece.cmu.edu/~byronyu/software/gpfa0203.tgz}}.
+        GPFA\footnote{https://users.ece.cmu.edu/~byronyu/software/gpfa0203.tgz}.
         GPFA models do not provide online estimates of the states, since data
@@
         spike counts). We propose to use a Python implementation of
-        svGPFA\footnote{\href{https://github.com/joacorapela/svGPFA}{https://github.com/joacorapela/svGPFA}}.
+        svGPFA\footnote{https://github.com/joacorapela/svGPFA}.
         svGPFA models do not provide online estimates of states, since data
@@
         propose to use a Python implementation at
-        LFADS\footnote{\href{https://github.com/tensorflow/models/tree/master/research/lfads}{https://github.com/tensorflow/models/tree/master/research/lfads}.}.
+        LFADS\footnote{https://github.com/tensorflow/models/tree/master/research/lfads.}.
         As GPFA, LFADS do not provide online estimates of states.
@@
 communication of C\# programs with Python we will use
-Python.NET\footnote{\href{https://github.com/pythonnet/pythonnet}{https://github.com/pythonnet/pythonnet}},
+Python.NET\footnote{https://github.com/pythonnet/pythonnet},
 with R programs we will use
-R.Net\footnote{\href{https://github.com/rdotnet/rdotnet}{https://github.com/rdotnet/rdotnet}}
+R.Net\footnote{https://github.com/rdotnet/rdotnet}
 and with Matlab we will use the built in Matlab .NET
-interface\footnote{\href{https://uk.mathworks.com/help/matlab/matlab\_external/using-net-from-matlab-an-overview.html}{https://uk.mathworks.com/help/matlab/matlab\_external/using-net-from-matlab-an-overview.html}}.
+interface\footnote{https://uk.mathworks.com/help/matlab/matlab\_external/using-net-from-matlab-an-overview.html}.
 In cases where these software cannot support some type of communication between
 C\# and another language we will use the messaging library
-ZeroMQ\footnote{\href{https://zeromq.org/}{https://zeromq.org/}}.
+ZeroMQ\footnote{https://zeromq.org/}.
 
@@
 Second, we will monitor user activity in the Bonsai
-forum\footnote{\href{https://groups.google.com/g/bonsai-users}{https://groups.google.com/g/bonsai-users}},
+forum\footnote{https://groups.google.com/g/bonsai-users},
 and publications citing Bonsai that use th new machine larning functionality,

### 2021-11-16 application/otherResources.tex
Removed live links

@@
 types (e.g. Open Ephys
-GUI\footnote{\href{https://open-ephys.org/gui/}{https://open-ephys.org/gui/}},
+GUI\footnote{https://open-ephys.org/gui/},
 Miniscope DAQ
-Software\footnote{\href{https://github.com/Aharoni-Lab/Miniscope-DAQ-QT-Software}{https://github.com/Aharoni-Lab/Miniscope-DAQ-QT-Software}})
+Software\footnote{https://github.com/Aharoni-Lab/Miniscope-DAQ-QT-Software})
 or by real-time control frameworks for specifying task logic using state
 machine or similar formalisms (e.g. NIMH ML
-(nih.gov)\footnote{\href{https://monkeylogic.nimh.nih.gov/}{https://monkeylogic.nimh.nih.gov/}},
-pyControl\footnote{\href{https://pycontrol.readthedocs.io/en/latest/}{https://pycontrol.readthedocs.io/en/latest/}},
-Autopilot\footnote{\href{https://docs.auto-pi-lot.com/en/latest/}{https://docs.auto-pi-lot.com/en/latest/}},
-Sanworks\footnote{\href{https://sanworks.io/index.php}{https://sanworks.io/index.php}}).
+(nih.gov)\footnote{https://monkeylogic.nimh.nih.gov/},
+pyControl\footnote{https://pycontrol.readthedocs.io/en/latest/},
+Autopilot\footnote{https://docs.auto-pi-lot.com/en/latest/},
+Sanworks\footnote{https://sanworks.io/index.php}).
 These dedicated interfaces are typically very comfortable for experimenters

### 2021-11-16 application/Je-S_sections.tex

@@
 We will do so by developing a new set of functionalities to an existing software platform, Bonsai.
-Bonsai is a fully integrated software environment for cutting-edge reproducible experiments using animal models, which provides machine-intelligence, real-time neuroinformatics methods.
-Importantly, Bonsai users require minimal programming experience.
+Bonsai is a fully integrated software environment that emphasises performance, flexibility, and ease-of-use, allowing scientists with no previous programming experience to quickly develop their own high-performance data acquisition and experimental control systems.
 

### 2021-11-16 application/detailsOfResource.tex
Updates from Overleaf

--- 2021-11-16 application/detailsOfResource.tex
@@
 
-The core proposed developments will enhance Bonsai's extensibility in two ways.
+We will enhance Bonsai's extensibility in two ways.
 %
-First, we will create a new package infrastructure to enable the integration of software toolboxes implemented in non-native languages into Bonsai.
+First, we will create a new package infrastructure to enable the integration of software objects implemented in non-native languages into Bonsai.
 %
@@
 
-Furthermore, we will build on these extensibility developments ourselves to implement a suite of advanced machine intelligence neural data analysis methods that will be made available as one or more core Bonsai packages.
+Furthermore, we will build on these extensibility developments to implement a suite of advanced machine intelligence neural data analysis methods that will be made available as one or more core Bonsai packages.
 
@@
 
-As it stands, Bonsai is designed to be extensible and is distributed with a package management framework that controls installation of both optional core packages and user-contributed extensions.
+As it stands, Bonsai is designed to be extensible and is distributed with a package-management framework that controls installation of optional core packages as well as user-contributed extensions.
 %
-However, the present framework requires extensions to be written in Bonsai's native C\# programming language.
+However extensions must be written in Bonsai's native C\# programming language.
 %
-This specificity restricts the number of potential contributors to the ecosystem.
+This specificity restricts the range of potential contributors to the ecosystem.
 %
-It also hinders easy incorporation of published analytic tools, many of which are implemented for interpreted language systems such as Python, R, or MATLAB.
+It also hinders easy incorporation of published analytic tools, many of which are implemented within other interpreted language systems such as Python, R, or MATLAB.
 
@@
 %
-Python.NET\footnote{\href{https://github.com/pythonnet/pythonnet}{https://github.com/pythonnet/pythonnet}},
+Python.NET\footnote{\url{https://github.com/pythonnet/pythonnet}},
 %
-R.Net\footnote{\href{https://github.com/rdotnet/rdotnet}{https://github.com/rdotnet/rdotnet}}
+R.Net\footnote{\url{https://github.com/rdotnet/rdotnet}}
 %
-and the native .NET interface shipped with MATLAB\footnote{\href{https://uk.mathworks.com/help/matlab/matlab\_external/using-net-from-matlab-an-overview.html}{https://uk.mathworks.com/help/matlab/matlab\_external/using-net-from-matlab-an-overview.html}}.
+and the native .NET interface shipped with MATLAB\footnote{\url{https://uk.mathworks.com/help/matlab/matlab\_external/using-net-from-matlab-an-overview.html}}.
 %
-If necessary, we will also incorporate direct inter-runtime communication using the messaging library ZeroMQ\footnote{\href{https://zeromq.org/}{https://zeromq.org/}}.
+If necessary, we will also incorporate direct inter-runtime communication using the messaging library ZeroMQ\footnote{\url{https://zeromq.org/}}.
 
@@
 
-The number of current machine learning algorithms is very large and it is
-continuously increasing. We intend to build a small number of machine
-intelligence operators in Bonsai (Section~\ref{sec:bonsai},
-Figure~\ref{fig:bonsai}c) that can be configured to provide the functionality
-of all machine learning algorithms.
-
-For this, we will build on abstractions used in the \texttt{scikit-learn}
-library. The success of this library is, to a large extent, due to its simple
-and general application programming interface \citep[API;][]{buitinckEtAl13}.
+To avoid the need to re-implement a single data analysis algorithm in multiple different contexts we will create a set of abstract machine-intelligence operators in Bonsai that can be configured to map between data substrates and machine-learning algorithms.
 %
-This API consists of three complementary interfaces: an \texttt{Estimator}
-interface for building and fitting models, a \texttt{Predictor} interface for
-making predictions and a \texttt{Transformer} interface for converting data.
+The planned abstraction layer is based on the successful \texttt{scikit-learn} Python
+library \citep{buitinckEtAl13}.
 %
-Using only these three interfaces \texttt{scikit-learn} provides
-implementations of a large number of machine learning algorithms for
-dimensionality reduction, clustering, classification and regression.
-
-Inspired by \texttt{scikit-learn}, Bonsai will have three types of operators:
-\texttt{Estimators}, \texttt{Predictors} and \texttt{Transformers}. Each of
-these operators will be instantiated by objects implementing the corresponding
-interface.
+By analogy to the interfaces provided by this library we will introduce three types of Bonsai operator: \texttt{Estimator} objects for building and fitting models, \texttt{Predictor} objects for making predictions and \texttt{Transformer} objects for converting data.
+%
+Th
 %

### 2021-11-16 application/trackRecord.tex

--- 2021-11-16 application/trackRecord.tex
@@
 The team responsible for the proposed project comprises two academic
-institutes, the Gatsby Computational Neuroscience Unit (GCNU) and the Sainsbury
-Wellcome Centre for Neural Circuits and Behavior (SWC), and a private
+institutes at University College London, the Gatsby Computational Neuroscience Unit (GCNU) and the Sainsbury
+Wellcome Centre for Neural Circuits and Behaviour (SWC), and a private
 collaborator, NeuroGEARS Ltd.
 
-The GCNU is a world-class research center for computational neuroscience and
-machine learning, while the Sainsbury Wellcome Center is a top-notch experimental
-neuroscience research center. They share a common building and collaborative
-activities are at the heart of both institutes.
+The GCNU is an authority in computational neuroscience and
+machine learning, while the Sainsbury Wellcome Centre is a world-leading experimental
+neuroscience research institute.
+They share a common building, which facilitates regular interaction and collaboration across various research groups, including the two named PIs in this project, and also collaborate in many joint activities.
 %
 NeuroGEARS is a spinoff of the SWC and is currently contributing to a major
-collaborative research project between the SWC and the GCNU. It has developed
-Bonsai, a very unique ecosystem for experimental control and data acquisition,
-it has offered courses around the globe instructing students on advanced
-experimental control, and has used Bonsai in numerous research and private
-experimental projects. Thus, NeuroGEARS has extensive experience on the
+collaborative research project between the SWC and the GCNU. It developed
+Bonsai, a unique ecosystem for experimental control and data acquisition,
+and offers courses around the globe instructing students on advanced
+experimental control. NeuroGEARS has used Bonsai in numerous research and private experimental projects, and has extensive experience on the
 implementation of neuroscience experiments.
 
-In summary, the partners in this project have outstanding expertise on their
-focus areas (computational neuroscience and machine learning, experimental
+In summary, the partners in this project have outstanding expertise on their focus areas (computational neuroscience and machine learning, experimental
 neuroscience, implementation of experiments), as well as a rich collaborative

### 2021-11-16 application/detailsOfResource.tex
Resolved conflict with hrefs

--- 2021-11-16 application/detailsOfResource.tex
@@
 %
-Th
+Interactions between these different object types will allow a single algorithm class (implemented by \texttt{Estimator} and \texttt{Predictor} operators) to act on many different data streams.
 %
-As an example, consider the close-loop system for the control of a computer
-cursor with neural activity represented in
+An example closed-loop system allowing  control of a computer cursor with neural activity is shown represented in
 Figure~\ref{fig:proposedBonsaiExtensions}b. The Bonsai

### 2021-11-17 application/backgroundToTheResource.tex
Updates from Overleaf
Updates from Goncalo

@@
 
-  \caption{\footnotesize Bonsai development timeline and community adoption. (A) Timeline of
+  \caption{Bonsai development timeline and community adoption. (A) Timeline of
     milestones and landmark package releases. (B) Example capabilities 
@@
 %
-Furthermore, it has become increasingly apparent that the neural signals that
+It has also become increasingly apparent that the neural signals that
 drive complex behaviour are manifest in a broad population of neurons rather
@@
 \textbf{This complexity of experiments limits the potential for database-driven progress in systems neuroscience}.
-The number of possible experiments is astronomical, and new questions are typically answered by designing and conducting new experiments rather than mining the data from an old one.
+The number of possible experiments is astronomical, and new questions are typically answered by designing and conducting new experiments rather than mining data from old ones.
 
-At present, progress in the field is held back by the challenge of rapidly designing and implementing new experimental protocols, and the difficulty of integrating state-of-the-art data processing and
-neuroinformatics into custom experimental designs.
+At present, progress in the field is held back by the challenge of designing and implementing new experimental protocols, and the difficulty of integrating state-of-the-art data processing and neuroinformatics into custom experimental designs.
 %
@@
 Bonsai achieves robust behavioural control, hardware interaction and signal acquisition while ensuring remarkable ease of use.
-The proposed project will provide valuable enhancements.
+The proposed project will provide valuable enhancements to the existing codebase.
 We will develop, test and deploy a new software workflow making it possible to incorporate external packages into Bonsai's C\#-based environment.

### 2021-11-17 application/bonsai.tex
Updates from Goncalo

@@
 
-Standard software tools for data analysis (ImageJ, MATLAB, R, etc.) have been
-transformative to the progress of increasingly “data-rich” sciences. However,
-the equivalent standardized software tools for data acquisition and control of
-animal behaviour neuroscience experiments are still lacking. Life science experiments demand
-a combination of multiple instrumentation and control technologies, for both
-behavioural and physiological investigations. The growing complexity on both the
-amount of data that is collected, and the rich conditions under which behaviour
-must be explored, place an increasing burden on experimenters to integrate
-highly specialised equipment in unique configurations, while often lacking
-expertise in the relevant engineering fields.
-
-Reproducibility is a key element of Bonsai's design.  Data acquisition and experimental control protocols can be 
-replicated in any laboratory by just sharing a Bonsai configuration file.
+% I think we've said something similar in the first para  - m.
+%
+% Standard software tools for data analysis (ImageJ, MATLAB, R, etc.) have
+% transformed progress inf increasingly “data-rich” sciences. However,
+% equivalent standardised software tools for data acquisition and control of
+% animal behaviour in neuroscience experiments are still lacking. Life science experiments demand
+% a combination of multiple instrumentation and control technologies, for both
+% behavioural and physiological investigations. The growing complexity on both the
+% amount of data that is collected, and the rich conditions under which behaviour
+% must be explored, place an increasing burden on experimenters to integrate
+% highly specialised equipment in unique configurations, while often lacking
+% expertise in the relevant engineering fields.
 
@@
 %
+Key to the wide adoption is the reproducibility that Bonsai ensures.  Data acquisition and experimental control protocols can be 
+replicated in any laboratory by just sharing a Bonsai configuration file. 
+%
 % LEAVE OUT FOR THIS GRANT:

### 2021-11-17 application/caseForSupport.tex
Removed live links
Updates from Overleaf
Updates from Goncalo

@@
 
-\usepackage[colorlinks=true]{hyperref}
+% \usepackage[colorlinks=true]{hyperref}
+\usepackage[colorlinks = true,
+            linkcolor = blue!50!black!50,
+            urlcolor  = blue!50!black!50,
+            citecolor = blue!50!black!50,
+            anchorcolor = blue!50!black!50]{hyperref}
 \usepackage{natbib}
@@
 
+\usepackage{tikz}
+\usetikzlibrary{positioning}
+\usetikzlibrary{fit}
+\usetikzlibrary{calc}
+
 \usepackage{float}
 \usepackage{wrapfig}
+
+\usepackage[font={footnotesize}]{caption}
 
@@
 \vskip 0.5ex
-We propose to extend, enhance, maintain and support \textbf{Bonsai}, a fully
-integrated software environment to \textbf{enable cutting-edge reproducible
+We propose to extend, enhance, maintain and support \textbf{Bonsai}, an integrated software environment to enable \textbf{cutting-edge reproducible
   laboratory systems neuroscience experiments in animal models}, with a

### 2021-11-17 application/detailsOfResource.tex
Updates from Goncalo

@@
 
+\begin{wrapfigure}{R}{0.4\textwidth}
+  \includegraphics[width=0.4\textwidth]{application/figures/extensions.pdf}
+%   \input{figures/extensions}
+  \caption{Proposed extensions to Bonsai}
+  \label{fig:proposedBonsaiExtensions}
+\end{wrapfigure}
+
 We will extend the Bonsai package interface to allow direct interaction with the runtime execution environments that implement these common languages.  
@@
 
-To avoid the need to re-implement a single data analysis algorithm in multiple different contexts we will create a set of abstract machine-intelligence operators in Bonsai that can be configured to map between data substrates and machine-learning algorithms.
+To avoid the need to re-implement a single data analysis algorithm to handle multiple different data types we will create a set of abstract machine-intelligence operators in Bonsai that can be configured to map between data substrates and machine-learning algorithms.
 %
@@
 %
-An example closed-loop system allowing  control of a computer cursor with neural activity is shown represented in
-Figure~\ref{fig:proposedBonsaiExtensions}b. The Bonsai
-graphical program for this system is given in
-Figure~\ref{fig:proposedBonsaiExtensions}c. The
-objects instantiating the SpikeThresholdDetector, the KiloSortSpikeSorter and
-GPFAlatentsExtractor will implement the \texttt{Transformer} interface (blue),
-and the object instantiating the TimeSeriesForestClassifier will implement the
-\texttt{Predictor} interface (green; Figure~\ref{fig:proposedBonsaiExtensions}c, top).
-%
-The parameters for the SpikeThresholdDetector and the KiloSortSpikeSorter will
-be set manually by the user, as is commonly done in practice. The parameters of
-the GPFAlatentsExtractor will be learnt offline in an unsupervised way by an
-\texttt{Estimator} (red) in a second pipeline
-(Figure~\ref{fig:proposedBonsaiExtensions}c, middle).
-Finally, the parameters of the TimeSeriesForestClassifier will also be learnt
-offline, but in a supervised way, by another \texttt{Estimator} in a third
-pipeline (Figure~\ref{fig:proposedBonsaiExtensions}c, bottom).
+% An example closed-loop architecture allowing online control of a compu|ter cursor with neural activity is shown in Figure~\ref{fig:proposedBonsaiExtensions}b. The Bonsai
+% graphical program for this system is given in Figure~\ref{fig:proposedBonsaiExtensions}c. The
+% objects instantiating the SpikeThresholdDetector, the KiloSortSpikeSorter and
+% GPFAlatentsExtractor implement the \texttt{Transformer} interface (blue),
+% and the object instantiating the TimeSeriesForestClassifier implements the
+% \texttt{Predictor} interface (green; Figure~\ref{fig:proposedBonsaiExtensions}c, top).
+% %
+% While the parameters for the SpikeThresholdDetector and the KiloSortSpikeSorter may
+% be set manually by the user, the parameters of
+% the GPFAlatentsExtractor can be learnt offline in an unsupervised way using an
+% \texttt{Estimator} (red) in a second pipeline
+% (Figure~\ref{fig:proposedBonsaiExtensions}c, middle).
+% Finally, the parameters of the TimeSeriesForestClassifier can also be learnt
+% offline, but in a supervised way, using another \texttt{Estimator} in a third
+% pipeline (Figure~\ref{fig:proposedBonsaiExtensions}c, bottom).
 
-Using the previous software, we will build C\# wrappers around machine
-intelligence programs written in other languages, so that these programs
-implement the \texttt{Estimator}, \texttt{Predictor} or \texttt{Transformer}
-interfaces and can be integrated into Bonsai as machine intelligence operators.
+% This initial set of methods should demonstrate to Bonsai users, and to
+% machine-learning methods developers, the potential of embedding
+% machine-learning functionality into the experimental control loop. We propose
+% user engagement activities to expose both of these groups to this
+% potential (Section~\ref{sec:userEngagement}).  As data-analysis method
+% developers recognise this opportunity, they will become interested in
+% contributing their methods to the Bonsai ecosystem, ensuring the
+% \textbf{long-term sustainability} of the proposed resource.
 
-a package management , and its community of users have
-already extended Bonsai's functionality with several contributed packages.
-%
-However, these packages need to
-be written in C\#, while most current neuroscience data analysis methods are
-written in Python, R or Matlab. We propose to add to Bonsai capabilities to
-communicate with software written in these languages
-(Section~\ref{sec:softwareInfrastructure}).  These capabilities will allow a
-large number of neuroscience data analysis methods to be easily integrated into
-the Bonsai ecosystem and will provide Bonsai users a large repertoire of
-advanced machine learning methods for their behavioural and neural data
-analysis.
-%
-Using these capabilities we will integrate into Bonsai an initial set of
-state-of-the art behavioural and neural data analysis methods
-(Section~\ref{sec:functionality}).
+% Adding machine learning functionality to the experimental control loop will
+% make possible the implementation of experiments with unprecedented levels of
+% control. Moreover, equipping experimental scientists without programming
+% experience with machine learning tools will considerably increase the
+% application of machine learning to the biosciences, which most probably will
+% translate into groundbreaking scientific discoveries.
 
-% \subsubsection{Extensibility}
-% \label{sec:extensibility}
-
-% To incorporate machine learning functionality into the Bonsai ecosystem, we
-% could
-% implement this functionality directly into Bonsai, as done in
-% \texttt{scikit-learn}. Alternatively, or in addition, we could build software
-% infrastructure to facilitate the integration of existing machine learning
-% software into Bonsai. We chose to begin with the second alternative because
-% high-quality implementations of machine learning functionality already exist,
-% which can be quickly integrated into the Bonsai ecosystem. In addition, the
-% second alternative is more sustainable, as it will allow the continued addition
-% of new machine learning functionality to Bonsai after the completion of the
-% BBSRC-BBR funding period. Later we may decide to implement some widely-used
-% machine learning functionality directly into Bonsai (e.g., for performance or
-% robustness issues).
-
-This initial set of methods should demonstrate to Bonsai users, and to
-machine-learning methods developers, the potential of embedding
-machine-learning functionality into the experimental control loop. We propose
-user engagement activities to expose both of these groups to this
-potential (Section~\ref{sec:userEngagement}).  As data-analysis method
-developers recognise this opportunity, they will become interested in
-contributing their methods to the Bonsai ecosystem, ensuring the
-\textbf{long-term sustainability} of the proposed resource.
-
-Adding machine learning functionality to the experimental control loop will
-make possible the implementation of experiments with unprecedented levels of
-control. Moreover, equipping experimental scientists without programming
-experience with machine learning tools will considerably increase the
-application of machine learning to the biosciences, which most probably will
-translate into groundbreaking scientific discoveries.
-
-Developers of advanced data-analysis methods should become interested in
-integrating them into the Bonsai ecosystem, as it will allow their methods to
-reach the wide Bonsai user community. In addition this integration will enable
-method developers to easily compare the performance of their methods with
-that of many methods previously integrated into the Bonsai ecosystem. Method
-developers will thus become a new type of Bonsai users, which will contribute
-to the \textbf{sustainability} of the proposed resource beyond the period of
-BBSRC-BBR funding.
+% Developers of advanced data-analysis methods should become interested in
+% integrating them into the Bonsai ecosystem, as it will allow their methods to
+% reach the wide Bonsai user community. In addition this integration will enable
+% method developers to easily compare the performance of their methods with
+% that of many methods previously integrated into the Bonsai ecosystem. Method
+% developers will thus become a new type of Bonsai users, which will contribute
+% to the \textbf{sustainability} of the proposed resource beyond the period of
+% BBSRC-BBR funding.
 
@@
 
-\begin{figure} \begin{center}
+% \begin{figure} \begin{center}
 
-    \href{http://www.gatsby.ucl.ac.uk/~rapela/bbsrc-bbr21/proposed_bonsai_extensions_combined.png}{\includegraphics[width=5in]{figures/proposed_bonsai_extensions_combined.png}}
+%     \href{http://www.gatsby.ucl.ac.uk/~rapela/bbsrc-bbr21/proposed_bonsai_extensions_combined.png}{\includegraphics[width=5in]{figures/proposed_bonsai_extensions_combined.png}}
 
-    \caption{Extensions to Bonsai. (A) we propose to integrate machine learning
-    functionality into the experimental control loop by building software
-    infrastructure to allow Bonsai to communicate with existing machine
-    learning software written in Python, R and Matlab. (B) flow chart
-    representing the BCI application that we propose to implement with the
-    machine learning functionality to be added to Bonsai. (C) Bonsai graphic
-    program implementing the BCI application in (B).}
+%     \caption{Extensions to Bonsai. (A) we propose to integrate machine learning
+%     functionality into the experimental control loop by building software
+%     infrastructure to allow Bonsai to communicate with existing machine
+%     learning software written in Python, R and Matlab. (B) flow chart
+%     representing the BCI application that we propose to implement with the
+%     machine learning functionality to be added to Bonsai. (C) Bonsai graphic
+%     program implementing the BCI application in (B).}
 
-        \label{fig:proposedBonsaiExtensions}
-    \end{center}
-\end{figure}
+%         
+%     \end{center}
+% \end{figure}
 
-\subsection{Software infrastructure}
-\label{sec:infrastructure}
-
-\begin{comment}
-Bonsai does not currently provide machine learning functionality. For instance,
-the close-loop experimental control depicted in
-Figure~\ref{fig:proposedBonsaiExtensions}b cannot be currently implemented in
-Bonsai. It represents the control of an experiment where a subject guides the
-mouse position in a computer screen with her brain activity (Section~\ref{}).
-Nodes highlighted in colour require machine intelligence functionality currently
-not available in Bonsai.
-%
-\end{comment}
-
-We propose to create software infrastructure to add machine intelligence
-functionality to the Bonsai ecosystem in general
-(Section~\ref{sec:generality}) and extensible ways
-(Section~\ref{sec:extensibility}).
-
-\begin{comment}
-\begin{figure}
-   \begin{center}
-       \includegraphics[width=4in]{figures/bci.png}
-       \caption{Close-loop system for controlling a computer cursor with brain activity}
-     \label{fig:bci}
-   \end{center}
-\end{figure}
-
-    The spike detector transforms membrane
-potentials into spike sequences. The spike sorter transforms spikes sequences
-into lists of spikes, one list per neuron. The latents extractor transforms
-multi-neuron lists of spikes into latent trajectories summarising the
-multi-neuron spiking activity.  The neural decoder predicts the next cursor
-position from the latent trajectories.
-\end{comment}
-
-\begin{comment}
-middle, whose output will set
-As in this library,
-there will be three type of machine intelligence operators in Bonsai:
-%
-estimator operators for building and fitting models, predictor operators for
-making predictions and transformer operators for converting data. These
-operators will be implemented by objects exposing the corresponding interfaces.
-
-The estimator interface exposes a fit method for learning a model from training
-data.  This method is called with training data (e.g., supplied as two arrays
-X\_train and y\_train in supervised learning estimators). Its task is to run a
-learning algorithm and to determine model-specific parameters from the training
-data.  All supervised and unsupervised learning algorithms (e.g., for
-classification, regression or clustering) will be offered as objects
-implementing this interface. Machine learning tasks like feature extraction,
-feature selection or dimensionality reduction will also be provided as
-estimators.
-
-The predictor interface extends the notion of an estimator by adding a predict
-method that takes an array X\_test and produces predictions for X\_test, based
-on the learned parameters of the estimator (we call the input to predict
-“X\_test” in order to emphasise that predict generalises to new data). In the
-case of supervised learning estimators, this method typically returns the
-predicted labels or values computed by the model.
-
-Since it is common to modify or filter data before feeding it to a learning
-algorithm, the transformer interface defines a transform method. It takes as
-input some new data X\_test and yields as output a transformed version of
-X\_test. Preprocessing, feature selection, feature extraction and
-dimensionality reduction algorithms will be provided as transformers in Bonsai.
-
-Bonsai will provide the ability to compose new
-estimators from several base estimators. Composition mechanisms can be used
-to combine typical machine learning workflows into a single object which is
-itself an estimator, and can be employed wherever usual estimators can be used.
-Composition of estimators will be done with Pipeline objects that chain multiple estimators into a single one. This is useful
-since a machine learning workflow typically involves a fixed sequence of processing steps (e.g., feature extraction, dimensionality reduction, learning and making
-predictions), many of which perform some kind of learning. A sequence of N such
-steps can be combined into a Pipeline if the first N-1 steps are transformers;
-the last can be either a predictor, a transformer or both.
-
-\end{comment}
-
-\subsection{Machine intelligence functionality to be added to Bonsai}
+\subsection{Machine intelligence functionality}
 \label{sec:functionality}
 
-We propose to initially add to Bonsai machine intelligence functionality to
-support neuroscience experiments because, currently, neuroscience is the
-research area with the largest number of Bonsai users. Within neuroscience, we
-propose to add functionality for video-based behavioural analysis
-(Section~\ref{sec:videoBasedBehavioralAnalysis}) and for running
-brain-computer-interface (BCI) experiments (Section~\ref{sec:bci}), since these tasks
-are relevant to many Bonsai users, and recent advances in machine learning
-will greatly improve their implementation in Bonsai.
+We will use the frameworks developed above to integrate existing machine-intelligence programs within one or more Bonsai packages.
+Based on user input, we have identified initial domains of video-based behavioural analysis (Section~\ref{sec:videoBasedBehavioralAnalysis}) and
+brain-computer-interface (BCI) experiments (Section~\ref{sec:bci}).
 
@@
 
-\paragraph{Comparing multiple data-analysis methods}
+\subsection{Comparing multiple data-analysis methods}
 \label{sec:comparisonOfMultipleMethods}
 
-% We propose to add functionality to Bonsai to facilitate multi-method
-% comparisons in user-supplied datasets.
+We propose to add functionality to Bonsai to facilitate multi-method
+comparisons in user-supplied datasets.
 
-Below we describe a comparison that we
-will perform in Bonsai to asses the relative performance of the methods
-described in the previous sections, using neural recordings from behaving rats.
-These recordings are currently being performed, to address scientific questions,
-at the laboratory of Prof.~Akrami in the SWC.
-%
-Details of the experimental task appear in~\citet{akramiEtAl18}.
-%
-Briefly, in this task there are three nose ports (left, centre, right). Rats
-initiate a trial by inserting their nose in the centre port for the duration of
-the fixation period, until they hear a Go stimulus. During the fixation period
-two stimuli $s_a$ and $s_b$ are presented. Rats should decide which stimuli is
-louder. If $s_a$ is louder than $s_b$ the correct action is to poke the nose
-into the right port in order to collect a liquid reward, and if $s_b$ is louder
-than $s_a$ the correct choice is the left port.
+% Below we describe a comparison that we
+% will perform in Bonsai to asses the relative performance of the methods
+% described in the previous sections, using neural recordings from behaving rats.
+% These recordings are currently being performed, to address scientific questions,
+% at the laboratory of Prof.~Akrami in the SWC.
+% %
+% Details of the experimental task appear in~\citet{akramiEtAl18}.
+% %
+% Briefly, in this task there are three nose ports (left, centre, right). Rats
+% initiate a trial by inserting their nose in the centre port for the duration of
+% the fixation period, until they hear a Go stimulus. During the fixation period
+% two stimuli $s_a$ and $s_b$ are presented. Rats should decide which stimuli is
+% louder. If $s_a$ is louder than $s_b$ the correct action is to poke the nose
+% into the right port in order to collect a liquid reward, and if $s_b$ is louder
+% than $s_a$ the correct choice is the left port.
 
-We will use high-density Neuropixel spike and LFP recordings from the rats
-performing this task to learn low-dimensional representations of spiking
-activity and LFPs during the decision time period (between the presentation of
-the last auditory stimuli and the presentation of the Go stimulus), using the
-methods described in
-Sections~\ref{sec:lowDimensionalRepresentation}
-and~\ref{sec:lowDimensionalRepresentationsOfLFPs}. We will also learn the
-parameters of the classifiers described in Section~\ref{sec:neuralDecoding} to
-optimally classify animal decisions (poke the nose into the right/left port)
-based on the previous low-dimensional representations.
+% We will use high-density Neuropixel spike and LFP recordings from the rats
+% performing this task to learn low-dimensional representations of spiking
+% activity and LFPs during the decision time period (between the presentation of
+% the last auditory stimuli and the presentation of the Go stimulus), using the
+% methods described in
+% Sections~\ref{sec:lowDimensionalRepresentation}
+% and~\ref{sec:lowDimensionalRepresentationsOfLFPs}. We will also learn the
+% parameters of the classifiers described in Section~\ref{sec:neuralDecoding} to
+% optimally classify animal decisions (poke the nose into the right/left port)
+% based on the previous low-dimensional representations.
 
-We will compare the decoding accuracy of all decoders. This comparison will
-tells us, for the auditory working memory task, which neural data type (spikes
-or LFPs), which low-dimensional representation method (e.g., LDS or Gaussian
-processes), and which decoder (e.g., Bayesian or ANN) yields the best decodings
-in state-of-the-art neural recordings from behaving animals.
+% We will compare the decoding accuracy of all decoders. This comparison will
+% tells us, for the auditory working memory task, which neural data type (spikes
+% or LFPs), which low-dimensional representation method (e.g., LDS or Gaussian
+% processes), and which decoder (e.g., Bayesian or ANN) yields the best decodings
+% in state-of-the-art neural recordings from behaving animals.
 

### 2021-11-17 application/figures/extensions.tex
Updates from Goncalo

% \documentclass{standalone}
% %% Fonts, etc.

% %% Language and font encodings
% \usepackage[english]{babel}
% \usepackage[utf8x]{inputenc}
% \usepackage[T1]{fontenc}

%  \usepackage[scaled=0.92]{helvet} \renewcommand{\familydefault}{\sfdefault}
%  \usepackage{xcolor}
% % \usepackage{soul} % \ul \st \hl

% %% Maths

% % \usepackage{amssymb,amsmath}
% % \input m.macros
% % \input m.symbols

% %% Graphics

% \usepackage{graphicx,tikz}
% \usetikzlibrary{positioning}
% \usetikzlibrary{fit}
% \usetikzlibrary{calc}

% %% Document
% \begin{document}

\begin{tikzpicture}[
    ML/.style={text width=10em, rounded corners,
      fill=green!50!black!30, draw=green!50!black, very thick},
    Ext/.style={text width=10em, rounded corners,
      fill=blue!50!black!30, draw=blue!50!black, very thick}
    ]

  \node[text width=10em, font={\bf}, align=center]
       (bonsai)
       {\includegraphics[width=7em]{bonsai-lettering}\\
         Bonsai core (C\#)};

 \coordinate [below = 5em of bonsai.south] (A);

  \node[below left=0 and 1em of A, Ext]
       (lang)
       {\textbf{Language interface:}\\
         C\# $\leftrightarrow$ Python\\
         $\qquad \leftrightarrow$ R\\
         $\qquad \leftrightarrow$ MATLAB
       };

  \node[below right=0 and 1em of A, Ext]
       (abstr)
       {\textbf{Analysis layers:}\\
         \texttt{Estimator}\\
         \texttt{Predictor}\\
         \texttt{Tranformer}};

  \node[above=1ex of A, text=blue!50!black, font={\bf}, inner sep=0]  (intr) {Interface layers};

  \node[fit=(lang) (abstr) (intr), line width=1mm, draw=blue!50!black!50, rounded
    corners, inner sep=1ex] (intr box) {};

 \coordinate [below = 12em of A] (B);

 \node[below=0 of B, ML]
       (signal)
       {\textbf{Signal analysis:}\\
         Spike detection\\
         LFP extraction\\
         Decoding};

  \node[below left=0 and 1em of signal.north west, ML]
       (video)
       {\textbf{Video analysis:}\\
         Motion sequencing\\
         Behaviour decoding\\
         Interpretable latents};

  \node[below right=0 and 1em of signal.north east, ML]
       (dimred)
       {\textbf{Population analysis:}\\
         Gaussian LDS\\
         Poisson LDS\\
         HMM\\
         GPFA\\
         svGPFA\\
         LFADS};

  \node[above=1ex of B, text=green!50!black, font={\bf}, inner sep=0] (ml)
       {Machine-intelligence algorithms};

  \node[fit=(video) (signal) (dimred) (ml), line width=1mm, draw=green!50!black!50, rounded
    corners, inner sep=1ex] (ml box) {};

  \draw[<->, line width=1mm, gray] (bonsai) -- (intr box);

  \draw[<->, line width=1mm, gray] (intr box) -- (ml box);

  %% \node[below = 15em of bonsai, ML]
  %% (compare)
  %%      {\textbf{Method comparison}};

\end{tikzpicture}

% \end{document}

% Local Variables:
% mode: latex
% eval: (set-fill-column 80)
% End:

### 2021-11-17 application/resourceManagementandSustainability.tex
Updates from Goncalo

@@
 %
-As an open source environment developed over almost 10 years, Bonsai 
-has well-established user-community-based management structures.
+As an open source project developed over almost 10 years, Bonsai 
+has well-established management practices built around the user community.
 %
-The core system and the majority of extension packages are hosted on a
-publicly accessible website at \url{bonsai-rx.org}.
+The source code for the core system and the majority of extension packages
+is hosted on GitHub at \url{github.com/bonsai-rx}.
+%
+The documentation, including learning materials and all relevant community
+links are published in a publicly accessible website at \url{bonsai-rx.org}.
+%
+A community user forum is available as a public Google Group with an active
+user base of more than 700 users (Figure~\ref{fig:bonsai}f) and several
+new questions and answers exchanged every day.
 %
 Package installation and updates are supported by standalone package
-management software included within the core Bonsai distribution.
+management software included within the core Bonsai distribution and hosted
+in a shared centralized and curated repository.
 %
-This ensures that bugfixes and enhancements are easily available but
-under full user control (to ensure working experiments are not
-disrupted by changes in the system).
-
-Best-practice code review procedures are followed prior to including
-package updates.  \textcolor{red}{Goncalo/Joaquin -- could you expand
-  on this?}
+This ensures that bugfixes and enhancements are made easily available but
+also importantly allows for reproducible deployment of user environments
+to ensure working experiments are not disrupted by changes in the system.
+Large-scale experimental projects such as the International Brain Laboratory
+have leveraged this capability of Bonsai to standardize data acquisition and
+control software across dozens of laboratories around the world.
+%
+Community technology previews are made available through the package manager,
+providing early adopters the possibility of opting-in to the latest features
+prior to making them generally available. This allows for the community
+to help identify bugs and report issues with new features on GitHub, thereby
+distributing bug reporting and triage and overall improving the quality of
+stable releases.
 
@@
 
+The core compiler and IDE is extensively tested, and also takes advantage of
+a strong type system to greatly reduce bugs at compile-time. A set of best
+practices for leveraging this infrastructure for building new packages is
+in place.
+
 The software engineers recruited to this project will work with our
-collaborators at Neurogears Ltd, to strengthen and maintain these
+collaborators at NeuroGEARS Ltd, to strengthen and maintain these
 practices.  Once fully familiar with the Bonsai core architecture they
-will participate in review of pull requests and other code
-contributions.
+will participate in review of pull requests and other code contributions.
 
@@
 
-Bonsai is a freely available open-source resource, already widely adopted within the systems neuroscience and behavioural science communities (see Fig~\ref{fig:bonsai}).  
+Bonsai is a freely available open-source resource, already widely adopted within
+the systems neuroscience and behavioural science communities (see Fig~\ref{fig:bonsai}).  
 %
-Current contributors include other large open projects (OpenEphys, \textcolor{red}{\dots}), as well as companies developing and supplying new hardware devices  (\textcolor{red}{\dots}).
+Current contributors include other large open projects (OpenEphys, DeepLabCut),
+as well as companies developing and supplying new hardware devices  (Neurophotometrics).
 %
-This level of community uptake, engagement, and commitment provides a foundation for long-term sustainability of the resource.
+This level of community uptake, engagement, and commitment provides a foundation for
+long-term sustainability of the resource.
 
-Neurogears Ltd is a keystone of this community and a key a collaborator in the current proposal.
-As a company founded by the original developer of Bonsai, it helps to support and expand the user base well beyond the life of the current proposal.
-Neurogears provides quality control and curation for core Bonsai code.
-\textcolor{red}{other ways NG contributes ...}
+NeuroGEARS Ltd is a keystone of this community and a key collaborator in the current proposal.
+As a company founded by the original developer of Bonsai, it helps to support and expand the
+user base well beyond the life of the current proposal. NeuroGEARS currently provides
+quality control and curation for the core Bonsai code, as well as providing training and support
+services for research institutions around the world who want to start using Bonsai.
+Finally, NeuroGEARS also facilitates knowledge transfer across users and institutions adopting
+Bonsai, by moderating the community user forums and collaborating with organisations
+who want to build new innovative research projects on top of Bonsai, as is the case with the
+current proposal.
 
-The specific contributions developed as part of the current project will form new components within the core Bonsai package. 
+The specific contributions developed as part of the current project will form new core
+components and packages within the Bonsai ecosystem. The new language integration packages are
+of particular relevance to the long-term sustainability of Bonsai, as they will dramatically
+increase the interoperability of Bonsai with existing software development communities which are
+already well established in basic science (MATLAB, Python, R).
 %
-They will be developed with input from the advisory panel, early public distribution, user testing within the SWC, and planned workshops, ensuring close user engagement from the beginning of the development cycle and so building momentum for their use, support and dissemination.
+The opportunity to develop these new packages with input from the core advisory panel of Bonsai,
+and early public distribution and user testing within the SWC and planned workshops, will ensure
+solid and broad user engagement from the beginning of the development cycle and will therefore
+build momentum for their use, support and dissemination.
 %
-\textcolor{red}{Can we say that NG will take over responsibility for maintenance and support beyond the project?}
+The close collaboration with the core maintainers of the Bonsai ecosystem at NeuroGEARS will
+allow them to take over responsibility for maintenance and support of these packages beyond the
+end of the project, following the same principles of open-source and free sharing of knowledge
+that characterises the entire Bonsai ecosystem.

### 2021-11-17 application/narrativeCVRapela.tex
Added

--- 2021-11-16 application/narrativeCVSahani.tex
@@
 
-\section{Prof.\ Maneesh Sahani -- narrative CV}
+\section{Dr.\ Joaquin Rapela -- narrative CV}
 
@@
 
-Prof.\ Sahani is a Professor of Theoretical Neuroscience and Machine
-Learning, and director of the Gatsby Computational Neuroscience Unit
-at UCL.
+Dr.~Joaquin Rapela is a Research Engineer Fellow at the Gatsby Computational
+Neuroscience Unit (GCNU), University College London.
 %
-As such, he is resident in the UK, and holds an open-ended academic
-employment contract with UCL.
+He has made a substantial contribution to the formulation and development of
+the proposed project, and will be engaged with the ensuing research. He is a
+Research Co-Investigator in the current application.
 
@@
 
-Prof.\ Sahani has authored over 100 peer-reviewed scientific papers,
-with an h-index (computed by Google scholar) of 47.
+Dr.~Rapela has acquired unique skills for the successful implementation of the
+proposed project. He completed his undergraduate degree in Computer Sciences,
+worked as a staff software engineer at the IBM Almaden Research Center, before
+pursuing his doctoral degree in Electrical Engineer, with specialisation on
+Signal Processing, at the University of Southern California.  His doctoral
+dissertation, as well as his two postdoctoral positions at University of
+California San Diego (UCSD) and at Brown University, focused on the use of
+advances mathematical and statistical tools to characterise
+electrophysiological neural signals. His
+publications\footnote{https://scholar.google.com/citations?user=eXkDg2UAAAAJ\&hl=en}
+demonstrate unique skills in signal processing, statistics and computational
+neuroscience.
 
-GPFA
+Dr.~Rapela is a strong supporter of open source software, and all his
+publications are accompanied by free
+code\footnote{http://www.gatsby.ucl.ac.uk/~rapela/software.htm}. While at UCSD,
+he contributed to the development of EEGLAB, a large open-source project for
+the characterisation of the electroencephalogram. Currently, as an Engineer
+Research Fellow at the GCNU, he distributes high-quality implementations of
+algorithms developed at unit (e.g., sparse variational Gaussian process factor
+analysis\footnote{https://github.com/joacorapela/svGPFA}).
 
-PLDS
+Another key responsibility of Dr.~Rapela at the GCNU, is to establish
+collaborations with experimental neuroscientists at the SWC and help them on
+the use of advanced statistical methods to process their state-of-the-art
+neural recordings.
 
-Grant funding?
+Therefore, Dr.~Rapela is excellently trained to implement all aspects
+(neuroscience, statistics, signal processing, and software development) of the
+proposed project.
 
 % 3) How have you contributed to research teams and the development of
-% others?
-
-% From QQR:
-As a PI, I have supervised a total of \textbf{X} students and \textbf{X} postdocs.  
-In the last five years, 4 of students have moved to postdoctoral positions (at
-Aalto, Cambridge, CMU and Stanford), 1 works at DeepMind, and 1 holds a faculty offer at Shanghai Tech.
-In all, 8 previous students have now secured faculty (or group leader) positions at Cambridge, CMU, Columbia, Copenhagen, Janelia (2), Pittsburgh and Shanghai.
-%
-Of the postdoctoral fellows I have supervised in the last five years, 1 is doing further postdoc work in Bonn; 1 holds an independent
-Research Fellowship at Radboud; and 2 have faculty positions: at Edinburgh and at UBC.
-
-DARPA repair
-
-Gatsby directorship
-
-% Examples might include: project management, supervision, mentoring or line management 
+% others? % Examples might include: project management, supervision, mentoring or line management 
 % contributions critical to the success of a team or team members, or where you exerted 
 % strategic leadership in shaping the direction of a team, organisation, company or institution.
+
+\subsection{Contributions to the development of others}
+
+Dr.~Rapela has mentored Mr.~Tsong-Yan Lin (master student at University
+California San Diego) on signal processing for processing
+electroencephalographic and electrooculographic time
+series~\footnote{Joaquin Rapela, Tsong-Yan Lin, Marissa Westerfield, Tzyy-Ping
+Jung, and Jeanne Townsend (2012b). Assisting autistic children with wireless
+EOG technology. Proceedings of the 34th Annual International Conference of the
+IEEE EMBS, San Diego, California.}.
+
+Currently, Dr.~Rapela is mentoring Ms.~Aishah Qureshi (undergraduate student at
+Queens College, London), as part of the Simons Foundation Undergraduate Research
+Program, on the use of linear dynamical systems to understand long-range
+cortical communication in rodents.
+
+% \subsection{Contributions to teams and development}
 
@@
 
-I have been an active member of the academic community. I co-founded the Neural Coding, Computation and Dynamics meeting in 2007, and was in the organising committee in 2015 and 2017.
+\subsection{Contributions to wider research community}
 
-I participated in the selection committee for the Swartz Prize for Theoretical and Computational Neuroscience in 2014-2016, and was in the programme committee for the Computational and Systems Neuroscience (COSYNE) conference in 2007, becoming a  Programme Chair in 2009 and General Chair in 2010.
-
-I was also in the Programme Committee for NIPS (now NeurIPS) in 2004 and 2006, becoming a Workshops Chair in 2008.
-I was also a Workshops Chair for the Computational Neuroscience Meeting from 1999 - 2003.
-
-I was a member of the Board of Directors of the Computational Neuroscience Organization between 2003 - 2006.
-I am currently a member of the Society for Neuroscience and IEEE, and in 2021 became a Fellow of the European Laboratory for Learning and Intelligent Systems (ELLIS).
-
-I currently have editorial roles in various journals including, Current Opinion in Neurobiology (Special issue on Big Data and Neuroscience), Neural Computation, Neural Systems and Circuits, Network:
-Computation in Neural Systems, and Faculty of
-1000, "Machine learning: life sciences" collection.
-
-I have taken various advisory roles for international review and award panels.
-I reviewed the Donders Institute for Brain, Cognition and Behaviour (Netherlands), The Edmond and Lily Safra Center for Brain Sciences, Hebrew
-University (Israel), and the Barcelona Summer School on Advanced Modelling of Behaviour (Spain).
-I have advised on awards such as the Swartz Prize for Theoretical and Computational Neuroscience (SFN, US), the Bernstein Prize (BMBF, Germany), and COSYNE.
-
-Recently, I have acted as an advisor to a new established cross-disciplinary effort at the Indian Institute of Science in Bangalore; visiting the site a
-number of times and hosting a joint workshop here at UCL.
-
-I have sat on various grant award panels for the BMBF and DFG, and the (US) NSF/NIH collaborative research
-in computational neuroscience (CRCNS) programme.
-I have also acted as a reviewer for proposals for the Bernstein prize, Einstein Foundation, EPSRC, MRC, Novo Nordisk Foundation (Denmark) and NWO (Netherlands).
+Dr.~Rapela has reviewed articles for Vision Research, Frontiers in Neuroscience
+and the Journal of Perceptual Imaging. He currently contributes to the SWC
+Research Culture Working
+Group~\footnote{https://sainsburywellcomecentre.github.io/RCWG/research} to
+improve the research culture at the SWC and GCNU.
 

### 2021-11-17 application/Je-S_sections.tex

@@
 
-The overall goal of this proposal is to extend, enhance, maintain and support Bonsai, a fully integrated software
-environment to enable cutting-edge reproducible laboratory systems neuroscience experiments in animal models.
+The overall goal of this proposal is to extend, enhance, maintain and support Bonsai, a fully integrated software environment to enable cutting-edge reproducible laboratory systems neuroscience experiments in animal models.
 
@@
 
-1. Develop, test and deploy a new software workflow to incorporate external packages into Bonsai's C# environment.
+1.
+Develop, test and deploy a new software workflow to incorporate external packages into Bonsai's C# environment.
+\textcolor{red}{I believe this is already done.
+All existing bonsai community packages already depend on the existence of the extendable plugin architecture -- perhaps you mean "languages" rather than "packages"?
+i.e.
+MATLAB, Python, R, in that case I would rewrite as below}
 
-2. Incorporate a suite of machine-learning-driven analysis tools into the Bonsai environment to implement:
-2a. video-based  behavioural analysis
-2b. real-time interfacing of population-scale activity with laboratory control  
-2c. evaluate competing data analytic approaches within a single experimental framework
+1. Develop, test and deploy new foreign language interfaces to allow extending the Bonsai environment with new packages developed in languages other than C#, namely MATLAB, Python and R.
 
-3. Enhance the quality and reproducibility of systems neuroscience through:
-3a. fully open-source and public-facing development
-3b. rapid release of, and community engagement around, new tools
-3c. provision of standard implementations of analysis and control methods, with the ability to distribute automated
-descriptions of analysis pipelines for reproducibility.
+2. Define a standard API and set of operators appropriate for composing machine-learning functionality into a Bonsai workflow, specifically to enable:
+2a. learning
+2b. inference
+2c. visualisation
+2d. model-comparison and validation
+2e. exchange of models and hyper-parameters
 
-4. Extend and support the Bonsai user community through code and forum engagement and development workshops.
+3. Incorporate a suite of machine-learning-driven analysis tools into the Bonsai environment using the standard API specified above, to implement:
+3a. video-based behavioural analysis
+3b. real-time interfacing of population-scale extracellular activity with laboratory control  
+3c. evaluation of competing data analytic approaches within a single experimental framework
+
+4. Enhance the quality and reproducibility of systems neuroscience through:
+4a. fully open-source and public-facing development
+4b. rapid release of, and community engagement around, new tools
+4c. provision of standard implementations of analysis and control methods, with the ability to distribute fully automated and reproducible analysis pipelines.
+
+5. Extend and support the Bonsai user community through code and forum engagement and development workshops.
 
@@
 
-Here we propose to provide the neuroscience community a software tool that will dramatically increase the efficiency of experimental control and data analysis.
+Here we propose to provide the neuroscience community with a software tool that will dramatically increase the efficiency of experimental control and data analysis.
 We will do so by developing a new set of functionalities to an existing software platform, Bonsai.
@@
 
-Resource is Bonsai.
-Monitor animal welfare
+The main beneficiaries of this proposal will be experimental labs worldwide who currently lack the highly technical expertise required to apply Machine Intelligence to their experiments.
+While more and more automated Machine Learning platforms are becoming generally available across many data science fields, they either focus on making available completely general-purpose tools which require deep expertise and training to construct and train new algorithms and models from scratch, or they encapsulate some pre-trained and domain-specific model which is useful only for very specific classes of problems (e.g.
+face recognition, pose estimation).
+
+A middle ground where existing algorithms can be easily composed by non-experts in a flexible way to design entirely novel Machine Intelligence applications tailor-made for specific research questions is currently entirely absent.
+In fields like systems neuroscience, where progress is made essentially by exploring the vast unknown space of animal behaviour and wide-scale brain physiology, the need for new bespoke algorithms and analysis procedures is constant and critical.
+If successful, this proposal will entirely transform the landscape of Machine Intelligence applied to experimental lab research in biology, by empowering non-technical researchers to recombine and repurpose Machine Intelligence tools using a flexible visual programming environment and providing seamless integration with all the measurement, instrumentation and control packages already available in Bonsai.
+
+In addition to the researchers themselves, the standard analysis pipelines will allow everyone who is interested in asking questions about the behaviour and physiology of animals to analyse and understand their data.
+As an example, increasingly in research institutions worldwide Bonsai is being used to monitor animal welfare by in-house technicians who are concerned about a broader range of environmental variables which might impact animal housing quality but which are beyond the ones which are available in standard vivarium solutions.
+They use Bonsai as it allows them to prototype and design entirely new monitoring systems on-site, without the need to engage an expensive software development process to answer simple questions.
+With access to the neuroinformatics packages developed for this proposal, staff at these institutions would be immediately empowered to address more sophisticated questions regarding interactions between housing conditions and animal physiology, which at the moment are entirely unthinkable, but would lead to instantly improving the welfare of hundreds of thousands of research animals in biological research facilities.

### 2021-11-17 application/trackRecord.tex

@@
 %
-NeuroGEARS is a spinoff of the SWC and is currently contributing to a major
-collaborative research project between the SWC and the GCNU. It developed
-Bonsai, a unique ecosystem for experimental control and data acquisition,
-and offers courses around the globe instructing students on advanced
-experimental control. NeuroGEARS has used Bonsai in numerous research and private experimental projects, and has extensive experience on the
-implementation of neuroscience experiments.
+NeuroGEARS is a consulting company started by the creators of Bonsai, a unique ecosystem
+for experimental control and data acquisition. It is currently contributing to a major
+collaborative research project between the SWC and the GCNU. It continues to develop Bonsai,
+and offers courses around the globe instructing students on advanced experimental control.
+NeuroGEARS has used Bonsai in numerous research and private experimental projects,
+and has extensive experience on the implementation of neuroscience experiments.
 

### 2021-11-17 application/backgroundToTheResource.tex
Added new Fig1 from Goncalo
Working on references

@@
 \begin{wrapfigure}{R}{0.5\textwidth}
-    \href{http://www.gatsby.ucl.ac.uk/~rapela/bbsrc-bbr21/roadmap-bbsrc_noCaption.png}{\includegraphics[width=0.5\textwidth]{figures/roadmap-bbsrc_noCaption.png}}
+    \includegraphics[width=0.5\textwidth]{figures/roadmap-bbsrc-2x_cropped.pdf}
 
@@
 %
-\textbf{Reproducibility}, in particular, depends on the availability of standardised tools for experimental control and analysis, shared between laboratories (Baker, 2016; Ioannidis, 2005).
+\textbf{Reproducibility}, in particular, depends on the availability of
+standardised tools for experimental control and analysis, shared between
+laboratories \citep{baker16,ioannidis05}.
 Although such tools exist in genetics, astronomy, physics and medicine (Fish et al., 2016; Abdalla et al., 2018, CERN Education, Communications and Outreach Group, 2018, Dickinson et al., 2016, Bycroft et al., 2018), they are mostly lacking in neuroscience, with documented consequences for the reliability of ``discoveries'' (Baker, 2016; Botvinik-Nezer et al., 2020; Button et al., 2013).

### 2021-11-17 application/backgroundToTheResource.tex
Updates from Overleaf

--- 2021-11-17 application/backgroundToTheResource.tex
@@
 
-\setlength{\columnsep}{1em}
-\begin{wrapfigure}{R}{0.5\textwidth}
-    \href{http://www.gatsby.ucl.ac.uk/~rapela/bbsrc-bbr21/roadmap-bbsrc_noCaption.png}{\includegraphics[width=0.5\textwidth]{figures/roadmap-bbsrc_noCaption.png}}
-
+\setlength{\columnsep}{0.5em}
+\setlength{\intextsep}{0pt}
+\begin{wrapfigure}{R}{0.65\textwidth}
+    \includegraphics[width=0.65\textwidth]{figures/roadmap-bbsrc-3x.pdf}%
   \caption{Bonsai development timeline and community adoption. (A) Timeline of
@@
     training events, by year (bottom panels), versus release and development
-    cycle (top panel), measured in commits to main repository. 
-    % Click on the image to see its online version.
-    }
-
-  \label{fig:bonsai}
+    cycle (top panel), measured in commits to main repository.}%
+  \label{fig:bonsai}%
 \end{wrapfigure}
@@
 We will develop, test and deploy a new software workflow making it possible to incorporate external packages into Bonsai's C\#-based environment.
-Building on this new extensibility, we will then incorporate a suite of machine-learning-driven analysis tools into the Bonsai environment.
+Building on this new extensibility, we will then incorporate a suite of machine learning (ML)-driven analysis tools into the Bonsai environment.
 These added capabilities will allow  video-based behavioural analysis as well as real-time interfacing of population-scale activity with laboratory control.

### 2021-11-17 application/bonsai.tex
Updates from Overleaf

@@
 %
-A Bonsai graphical program consists of one or more source data streams (e.g. neural activity, video, audio, sensors, etc)
+A Bonsai graphical program consists of one or more source data streams (e.g.\ neural activity, video, audio, sensors, etc)
 and several interconnected operators that transform input to output

### 2021-11-17 application/caseForSupport.tex
Working on references
Updates from Overleaf

@@
             anchorcolor = blue!50!black!50]{hyperref}
-\usepackage{natbib}
-\usepackage{apalike}
+\usepackage[super]{natbib}
+\renewcommand{\citet}[1]{[\citenum{#1}]}
+% \usepackage{apalike}
 \usepackage{graphicx}
@@
 \titlespacing*{\paragraph}{0pt}{*0.5}{*0}
+
+\usepackage{enumitem}
+\setlist[description]{
+    topsep=0pt,
+    itemsep=0pt,
+    partopsep=0pt,
+    parsep=0pt,
+    font=\headfmt,
+    leftmargin=1em}
+
+\renewcommand{\footnote}[1]{ [#1]}
 
@@
         circuits with elemental behavioural processes. Long-term, this research
-        will help identify the missing links between circuit dysfunction and
+        will help identify missing links between circuit dysfunction and
         behavioural deficits in neurological or mental health disorders, and
-        will pave the way for the development of new neural circuit biomarkers
-        and therapies. The Bonsai ecosystem will also enable the experimental
+        will support the development of new neural circuit biomarkers
+        and therapies. The Bonsai ecosystem will also enable  experimental
         testing of emerging theories of brain function and learning, and thus
@@
         enhanced Bonsai platform will also support purely behavioural
-        experiments, in which the animals' posture, body parts or location are
-        tracked in real-time, allowing investigation of environmental variables
-        or captivity on their behaviour, especially important for assessment of
+        experiments, which track animals' posture, body parts or location in real-time, to investigate impacts of environmental variables
+        or captivity on behaviour - crucial for
         animal welfare.
@@
         and economic costs to scientists and funders.  In neuroscience,  many
-        behavioural and neural animal studies have been to be difficult to
+        behavioural and neural animal studies have been difficult to
         reproduce, and experiments are rarely replicated across laboratories.
@@
         know-how to either develop their own software for experimental hardware
-        control or write software for the state-of-the-art analysis of rich
+        control or write software for state-of-the-art analysis of rich
         neural and behavioural data, increasingly reliant on latest machine
         learning algorithms. The proposed upgrading of the open-source and
-        community empowered Bonsai ecosystem with a toolbox of online and
-        offline analytical tools, that can dovetail with closed-loop
+        community-empowered Bonsai ecosystem with a toolbox of online and
+        offline analytical tools, which can dovetail with closed-loop
         experimental requirements, will (i) allow the deployment of such tools
-        into laboratories that lack the expertise, (ii) reduce the need to
+        into laboratories that lack the necessary expertise, (ii) reduce the need to
         reinvent the same tools in multiple labs (cost/time efficiency) and
@@
 
-\bibliographystyle{apalike}
-\bibliography{bonsai,epsrcSoftwareForResearchCommunities,monitoringBehavior,stateSpaceModels,machineLearning,gaussianProcesses,spikeSorting}
+\bibliographystyle{unsrtveryabbrv}
+\bibliography{journalsabbrvnodots,bonsai,epsrcSoftwareForResearchCommunities,monitoringBehavior,stateSpaceModels,machineLearning,gaussianProcesses,spikeSorting}
 

### 2021-11-17 application/otherResources.tex
Updates from Overleaf

@@
 user interfaces for control and recording of specific devices and experiment
-types (e.g. Open Ephys
+types (e.g.\ Open Ephys
 GUI\footnote{https://open-ephys.org/gui/},
@@
 or by real-time control frameworks for specifying task logic using state
-machine or similar formalisms (e.g. NIMH ML
-(nih.gov)\footnote{https://monkeylogic.nimh.nih.gov/},
+machine or similar formalisms (e.g. NIMH ML\footnote{https://monkeylogic.nimh.nih.gov/},
 pyControl\footnote{https://pycontrol.readthedocs.io/en/latest/},

### 2021-11-17 application/caseForSupport.tex

@@
 \bibliographystyle{unsrtveryabbrv}
-\bibliography{journalsabbrvnodots,bonsai,epsrcSoftwareForResearchCommunities,monitoringBehavior,stateSpaceModels,machineLearning,gaussianProcesses,spikeSorting}
+\bibliography{journalsabbrvnodots,bonsai,epsrcSoftwareForResearchCommunities,monitoringBehavior,stateSpaceModels,machineLearning,gaussianProcesses,spikeSorting,tomsCitations}
 

### 2021-11-17 application/communityDemand.tex

@@
 % Bonsai has directly enabled high-impact research in neuroscience~\citep[e.g.,][]{soaresEtAl16,grundemannEtAl17,netoEtAl16,hagiharEtAla21,anguillonRodriguezEtAl21} and contributed to accelerating technical progress in the field by functioning as an integrator of open-source, community developed technologies \citep{itskovEtAl14,lopesEtAl21,kaneEtAl20}.
-The lack of a strong and
-flexible machine learning package capable of being easily integrated with
-real-time experimental control workflows has been a recurrent concern in the
-Bonsai user community. This has been brought up spontaneously multiple times in
-the user community forums, both in the form of specific problems related to the
-tracking and interpretation of online animal behaviour, and in the form of
-explicit proposals and suggestions for the creation of a machine learning
+The need for a strong and
+flexible ML package that can be easily integrated with
+real-time experimental control workflows has been raised multiple times by the
+Bonsai user community, both as specific problems around 
+tracking and interpretation of animal behaviour, and as 
+explicit proposals to create a ML
 package\footnote{https://groups.google.com/g/bonsai-users/c/BZ3zOOdv\_1c/m/x6OP75frBQAJ}. Even
-relatively modest contributions such as integrating support for specific models
-in Bonsai, such as DeepLabCut, or even the more specific zebrafish feature
-tracking package BonZeb, has attracted tremendous attention and positive
-feedback from the community \citep[e.g.,][]{kaneEtAl20,guilbeaultEtAl21}.
+relatively modest contributions such as integrating support for specific models, such as DeepLabCut, or even the more specific zebrafish feature
+tracking package BonZeb, have attracted positive
+feedback from the community, \citep[e.g.,][]{kaneEtAl20,guilbeaultEtAl21}.
 
-That there is a need for an easier interface into machine learning algorithms
+The need for an easier interface into ML algorithms
 is evidenced by the development of multiple graphical user interfaces for
@@
 their support libraries require expertise with the Python programming language
-to use flexibly. Bonsai provides a way of addressing this gap, using its
+to use flexibly. Bonsai addresses this gap, using its
 flexible visual programming language to combine ease-of-use and full
@@
 
-We focused this proposal on neuroscience applications of Bonsai. However, the
-addition of machine intelligence functionality is needed in many other
-experimental areas where Bonsai is or could be used. Therefore, the software
-abstractions that we will build to add to Bonsai machine intelligence
-functionality for neuroscience experiments (e.g., methods to perform
-multi-method comparisons) will translate to other application domains of
+We focused this proposal on neuroscience applications of Bonsai. However, machine intelligence functionality is needed in many other
+experimental areas where Bonsai is or could be used. The software
+abstractions that will add ML
+functionality to Bonsai for neuroscience experiments (e.g., methods to perform
+multi-method comparisons) will also translate to other application domains of
 Bonsai.
 
-The ongoing AI revolution makes it extremely timely to integrate these
-technologies into Bonsai. The proposed software infrastructure will make very
-easy the integration of novel machine learning methods into the Bonsai
-ecosystem. This integration will be critical to unlock new understanding, value
-and scientific leads from the rich and complex behavioral and neural
-experimetal data that Bonsai makes possible to collect. In addition, this
-integration will open to non-programmer experimentalists in the biological
-sciences unprecedented opportunities to combine real-time machine learning with
-experimental control in neuroscience, which has to the best of our knowledge
-not really been done before in such a broad and comprehensive scale. We believe
-addressing this bottleneck and enabling these technologies to non-programmers
-will make this resource into a transformative and fundamental technology for
-the next generation studies in animal behaviour.
+The ongoing AI revolution makes integration of these
+technologies into Bonsai extremely timely. The proposed software infrastructure will facilitate the integration of novel ML methods into the Bonsai
+ecosystem. This integration will unlock new value
+and scientific directions from the rich behavioral and neural
+experimental data that Bonsai already allows scientists to collect. This
+integration will also provide experimentalists in the biological
+sciences who lack programming skills the opportunity to combine real-time ML with
+experimental control in neuroscience. To our knowledge
+this has not been done before on such a comprehensive scale, and will, we believe, make Bonsai into a fundamental and transformative technology for
+the next generation of animal behavioural research.

### 2021-11-17 application/detailsOfResource.tex
Fixed problem with application/figures/extensions.pdf

@@
 It also hinders easy incorporation of published analytic tools, many of which are implemented within other interpreted language systems such as Python, R, or MATLAB.  
-
-\begin{wrapfigure}{R}{0.4\textwidth}
-  \includegraphics[width=0.4\textwidth]{application/figures/extensions.pdf}
+\begin{wrapfigure}{R}{0pt}
+  \includegraphics[width=0.5\textwidth]{figures/extensions.pdf}
 %   \input{figures/extensions}
@@
 %
-R.Net\footnote{https://github.com/rdotnet/rdotnet}
+R.Net\break \footnote{https://github.com/rdotnet/rdotnet}
 %
-and the native .NET interface shipped with MATLAB\footnote{https://uk.mathworks.com/help/matlab/matlab\_external/using-net-from-matlab-an-overview.html}.
+and the native .NET interface shipped with MATLAB.
+% \footnote{https://uk.mathworks.com/help/matlab/matlab\_external/using-net-from-matlab-an-overview.html}.  TOO LONG
 %
@@
         Python\footnote{code for PS-VAE is embedded in the BehaveNet code
-        https://github.com/themattinthehatt/behavenet.
-        Please refer to class \texttt{PSVAE} in 
-        \texttt{behavenet.models.vaes.py}.}. After trained it can be used
+        https://github.com/themattinthehatt/behavenet.%
+        % Please refer to class \texttt{PSVAE} in texttt{behavenet.models.vaes.py}.
+        }. After trained it can be used
         online to extract low-dimensional features and perform downstream
@@
     \item[Gaussian Linear Dynamical
-        System]\citep[GLDS;][]{andersonAndMoore12}: with sufficiently large
+        System]\citep[GLDS][]{andersonAndMoore12}: with sufficiently large
         bin sizes, spike counts can be modelled as Gaussian random processes.
@@
 
-    \item[Poisson Linear Dynamical System]\citep[PLDS;][]{mackeEtAl15}: for
+    \item[Poisson Linear Dynamical System]\citep[PLDS][]{mackeEtAl15}: for
         smaller bin sizes, spike counts are better modelled as Poisson random
@@
 First, we will have targets associated with each release to the general public
-of the software components described in Sections~\ref{sec:infrastructure}
-and~\ref{sec:functionality}.
+of the software components described above. 
+% in Sections~\ref{sec:infrastructure} and~\ref{sec:functionality}.
 %

### 2021-11-17 application/figures/extensions.tex

@@
-% \documentclass{standalone}
-% %% Fonts, etc.
+\documentclass{standalone}
+%% Fonts, etc.
 
-% %% Language and font encodings
-% \usepackage[english]{babel}
-% \usepackage[utf8x]{inputenc}
-% \usepackage[T1]{fontenc}
+%% Language and font encodings
+\usepackage[english]{babel}
+\usepackage[utf8x]{inputenc}
+\usepackage[T1]{fontenc}
 
-%  \usepackage[scaled=0.92]{helvet} \renewcommand{\familydefault}{\sfdefault}
-%  \usepackage{xcolor}
-% % \usepackage{soul} % \ul \st \hl
+ \usepackage[scaled=0.92]{helvet} \renewcommand{\familydefault}{\sfdefault}
+ \usepackage{xcolor}
+% \usepackage{soul} % \ul \st \hl
 
-% %% Maths
+%% Maths
 
-% % \usepackage{amssymb,amsmath}
-% % \input m.macros
-% % \input m.symbols
+% \usepackage{amssymb,amsmath}
+% \input m.macros
+% \input m.symbols
 
-% %% Graphics
+%% Graphics
 
-% \usepackage{graphicx,tikz}
-% \usetikzlibrary{positioning}
-% \usetikzlibrary{fit}
-% \usetikzlibrary{calc}
+\usepackage{graphicx,tikz}
+\usetikzlibrary{positioning}
+\usetikzlibrary{fit}
+\usetikzlibrary{calc}
 
-% %% Document
-% \begin{document}
+%% Document
+\begin{document}
 
@@
        (bonsai)
-       {\includegraphics[width=7em]{bonsai-lettering}\\
+       {\includegraphics[width=5em]{bonsai-lettering}\\
          Bonsai core (C\#)};
 
- \coordinate [below = 5em of bonsai.south] (A);
+  \coordinate [above right = 2.5ex and 13em of bonsai.east] (A);
+  \draw[<->, line width=1mm, gray] (bonsai.east) ++(-2em,0) -- +(2.5em,0);
 
@@
 
- \coordinate [below = 12em of A] (B);
+ \coordinate [below left = 11em and 5em of A] (B);
+ \draw[<->, line width=1mm, gray] (intr box.south) -- +(0,-2.5em);
 
@@
 
-  \draw[<->, line width=1mm, gray] (bonsai) -- (intr box);
-
-  \draw[<->, line width=1mm, gray] (intr box) -- (ml box);
-
   %% \node[below = 15em of bonsai, ML]
@@
 
-% \end{document}
+\end{document}
 

### 2021-11-17 application/narrativeCVSahani.tex
Updates from Goncalo

@@
 
-\subsection{Eligibility criteria}
+\subsection{Academic history and eligibility criteria}
 
-Prof.\ Sahani is a Professor of Theoretical Neuroscience and Machine
-Learning, and director of the Gatsby Computational Neuroscience Unit
-at UCL.
+I am currently Professor of Theoretical Neuroscience and Machine Learning, and
+director of the Gatsby Computational Neuroscience Unit at UCL.
 %
-As such, he is resident in the UK, and holds an open-ended academic
+As such, I am resident in the UK, and hold an open-ended academic
 employment contract with UCL.
+
+I
 
@@
 % From QQR:
+\subsection{Contributions to teams and development}
+
 As a PI, I have supervised a total of \textbf{X} students and \textbf{X} postdocs.  
@@
 % strategic leadership in influencing a research agenda.
+
+\subsection{Contributions to wider research community}
 

### 2021-11-17 application/resourceManagementandSustainability.tex
Updates from Overleaf

@@
 \section{Resource management}
-\textcolor{red}{I'm making this up -- Goncalo please comment/correct.}
 %
@@
 The source code for the core system and the majority of extension packages
-is hosted on GitHub at \url{github.com/bonsai-rx}.
+are hosted at \url{github.com/bonsai-rx}.
 %
 The documentation, including learning materials and all relevant community
-links are published in a publicly accessible website at \url{bonsai-rx.org}.
+links are on a publicly accessible website at \url{bonsai-rx.org}.
 %
-A community user forum is available as a public Google Group with an active
-user base of more than 700 users (Figure~\ref{fig:bonsai}f) and several
+A community user forum is a public Google Group with more than 700 active users (Figure~\ref{fig:bonsai}f) and several
 new questions and answers exchanged every day.
@@
 %
-This ensures that bugfixes and enhancements are made easily available but
+This ensures that bugfixes and enhancements are easily available but
 also importantly allows for reproducible deployment of user environments
@@
 Community technology previews are made available through the package manager,
-providing early adopters the possibility of opting-in to the latest features
-prior to making them generally available. This allows for the community
-to help identify bugs and report issues with new features on GitHub, thereby
+allowing early adopters to opt-in to the latest features
+before they are generally available. This allows the community
+to help report issues with new features on GitHub, thereby
 distributing bug reporting and triage and overall improving the quality of
@@
 
-\textcolor{red}{Testing}
-
-The core compiler and IDE is extensively tested, and also takes advantage of
+The core compiler and IDE is extensively tested, and takes advantage of
 a strong type system to greatly reduce bugs at compile-time. A set of best
@@
 
-Bonsai is a freely available open-source resource, already widely adopted within
+Bonsai is a freely available open-source resource, already widely adopted by
 the systems neuroscience and behavioural science communities (see Fig~\ref{fig:bonsai}).  
@@
 Current contributors include other large open projects (OpenEphys, DeepLabCut),
-as well as companies developing and supplying new hardware devices  (Neurophotometrics).
+and companies developing and supplying new hardware devices  (Neurophotometrics).
 %
-This level of community uptake, engagement, and commitment provides a foundation for
-long-term sustainability of the resource.
+This community uptake, engagement, and commitment provides a foundation for
+long-term sustainability.
 
@@
 Bonsai, by moderating the community user forums and collaborating with organisations
-who want to build new innovative research projects on top of Bonsai, as is the case with the
+who want to build new innovative research projects on top of Bonsai, as in the
 current proposal.
 
-The specific contributions developed as part of the current project will form new core
-components and packages within the Bonsai ecosystem. The new language integration packages are
-of particular relevance to the long-term sustainability of Bonsai, as they will dramatically
-increase the interoperability of Bonsai with existing software development communities which are
+The specific contributions developed as part of this project will form new core
+components and packages within the Bonsai ecosystem. In particular, the new language integration packages will support long-term sustainability, by dramatically
+increasing Bonsai's interoperability with existing software development communities which are
 already well established in basic science (MATLAB, Python, R).
 %
-The opportunity to develop these new packages with input from the core advisory panel of Bonsai,
+Developing these new packages with input from the core advisory panel of Bonsai,
 and early public distribution and user testing within the SWC and planned workshops, will ensure
@@
 %
-The close collaboration with the core maintainers of the Bonsai ecosystem at NeuroGEARS will
+Close collaboration with the core maintainers of the Bonsai ecosystem at NeuroGEARS will
 allow them to take over responsibility for maintenance and support of these packages beyond the

### 2021-11-17 application/trackRecord.tex

@@
-The team responsible for the proposed project comprises two academic
+The proposed project team comprises two academic
 institutes at University College London, the Gatsby Computational Neuroscience Unit (GCNU) and the Sainsbury
@@
 The GCNU is an authority in computational neuroscience and
-machine learning, while the Sainsbury Wellcome Centre is a world-leading experimental
+machine learning, while the SWC is a world-leading experimental
 neuroscience research institute.
-They share a common building, which facilitates regular interaction and collaboration across various research groups, including the two named PIs in this project, and also collaborate in many joint activities.
+They share a common building, which facilitates regular interaction and collaboration across research groups, including the two named PIs in this project, and also collaborate in many joint activities.
 %
-NeuroGEARS is a consulting company started by the creators of Bonsai, a unique ecosystem
-for experimental control and data acquisition. It is currently contributing to a major
+NeuroGEARS is a consulting company started by the creators of Bonsai. It is currently contributing to a major
 collaborative research project between the SWC and the GCNU. It continues to develop Bonsai,

### 2021-11-17 application/userEngagement.tex

@@
-In the past four years we have developed the bonsai-rx.org website with
-documentation on how to use the language, and as a central hub for learning
-resources, such as courses, video tutorials and examples. We have also engaged
-with leading training efforts in the field such as the CAJAL Neuroscience
+In the past four years we have developed bonsai-rx.org as a hub for documentation and learning
+resources (e.g. courses, video tutorials and examples) on how to use Bonsai. We have also engaged
+with leading, relevant training efforts such as the CAJAL Neuroscience
 Training Programme (Bonsai 0121) and the Transylvania Experimental Neuroscience
-Summer School (TENSS)  to increase awareness of the availability of Bonsai and
+Summer School, to increase awareness of Bonsai and
 its packages throughout the neuroscience community. Together with participation
 at smaller invited venues, there has been an average of 5 courses in Bonsai
-each year, with at least 20 students each, so about 100 neuroscience students
-every year get their introduction to experimental tools directly through a
+each year, with at least 20 students each; i.e.  around 100 neuroscience students annually are introduced to experimental tools directly through a
 course in the Bonsai programming language.
 
-The development of this new machine learning package will be deeply integrated
-into the Bonsai ecosystem which will complete its 10th anniversary in 2022, and
-will thus leverage planned venues, events, and infrastructure to disseminate
+Bonsai's 10th anniversary will take place in 2022. The proposed machine learning package will thus leverage planned venues, events, and infrastructure to disseminate
 awareness throughout the community, including presentations at conferences,
-workshops, and training sessions, as well as electronic newsletter, forums and
+workshops, and training sessions, as well as electronic newsletters, forums and
 social media articles.
 
-The new infrastructure that we will build on Bonsai to support the machine
-learning package will facilitate interfacing with other programming languages
-such as Python, R, and MATLAB, which will therefore attract a whole new
-community of users to Bonsai with expertise in software development. To take
-advantage of this new community and provide to them the full power of Bonsai,
+The proposed development will facilitate interfacing with other programming languages
+such as Python, R, and MATLAB. This will therefore attract a whole new
+community of Bonsai users with expertise in software development. To fully integrate this new community and provide them with the full power of Bonsai,
 we plan to organize hackathons to accelerate integration of their existing data
@@
  Given the modular and integrative nature of the machine learning package, we
- plan to advertise its development early in the project, so that interested
- partners have the opportunity to contribute to the requirements of key
+ plan to advertise its development early in the project, to allow interested
+ partners to contribute to key
  components of the platform, such as the Python and MATLAB integrations. We

### 2021-11-17 application/backgroundToTheResource.tex

@@
 
-To understand the neural basis of behaviour, experiments need to control and quantify animal behaviour whilst recording, and often manipulating, neural activity. 
+To understand the neural basis of behaviour, experiments need to control and quantify animal behaviour while recording, and often manipulating, neural activity. 
 %

### 2021-11-17 application/detailsOfResource.tex

@@
 \begin{wrapfigure}{R}{0pt}
-  \includegraphics[width=0.5\textwidth]{figures/extensions.pdf}
+  \includegraphics[width=0.5\textwidth]{figures/extensions.pdf}%
 %   \input{figures/extensions}
@@
 
-To avoid the need to re-implement a single data analysis algorithm to handle multiple different data types we will create a set of abstract machine-intelligence operators in Bonsai that can be configured to map between data substrates and machine-learning algorithms.
+To avoid the need to re-implement a single data analysis algorithm to handle multiple different data types we will create a set of abstract machine-intelligence operators in Bonsai that can be configured to map between data substrates and ML algorithms.
 %
@@
 
-\subsubsection{Video-based behavioral analysis}
+\subsubsection{Video-based behavioural analysis}
 \label{sec:videoBasedBehavioralAnalysis}
@@
         http://datta.hms.harvard.edu/research/behavioral-analysis/.}.
-        After trained MoSeq can be used to detect behavioral syllables online.
+        After trained MoSeq can be used to detect behavioural syllables online.
 
@@
 \texttt{LatentsCalculator}. Finally, a \texttt{Decoder} will extract the
-intended behavior of the animal from this low-dimensional representation.
+intended behaviour of the animal from this low-dimensional representation.
 
@@
 forum\footnote{https://groups.google.com/g/bonsai-users},
-and publications citing Bonsai that use th new machine larning functionality,
+and publications citing Bonsai that use the new ML functionality,
 to estimate the number of Bonsai users using this functionality.
 %
-Third, we will count the number of machine learning methods that are integrated
+Third, we will count the number of ML methods that are integrated
 into the Bonsai ecosystem by us and by other method developers.

### 2021-11-17 application/userEngagement.tex

@@
 
-Bonsai's 10th anniversary will take place in 2022. The proposed machine learning package will thus leverage planned venues, events, and infrastructure to disseminate
+Bonsai's 10th anniversary will take place in 2022. The proposed ML package will thus leverage planned venues, events, and infrastructure to disseminate
 awareness throughout the community, including presentations at conferences,
@@
 we plan to organize hackathons to accelerate integration of their existing data
-analysis and machine learning algorithms and methods into Bonsai using the new
+analysis and ML algorithms and methods into Bonsai using the new
 machine learning package.
 
- Given the modular and integrative nature of the machine learning package, we
+ Given the modular and integrative nature of the ML package, we
  plan to advertise its development early in the project, to allow interested

### 2021-11-17 application/caseForSupport.tex
Merging references conflict

--- 2021-11-17 application/caseForSupport.tex
@@
         testing of emerging theories of brain function and learning, and thus
-        contribute to the neuro-inspired machine learning algorithms that fuel
+        contribute to the neuro-inspired ML algorithms that fuel
         intelligent technologies, increasingly important in our society.  The
@@
         control or write software for state-of-the-art analysis of rich
-        neural and behavioural data, increasingly reliant on latest machine
-        learning algorithms. The proposed upgrading of the open-source and
+        neural and behavioural data, increasingly reliant on the latest ML algorithms. The proposed upgrading of the open-source and
         community-empowered Bonsai ecosystem with a toolbox of online and

### 2021-11-17 application/backgroundToTheResource.tex
Inserted references to Toms citations; fone

--- 2021-11-17 application/backgroundToTheResource.tex
@@
 standardised tools for experimental control and analysis, shared between
-laboratories \citep{baker16,ioannidis05}.
-Although such tools exist in genetics, astronomy, physics and medicine (Fish et al., 2016; Abdalla et al., 2018, CERN Education, Communications and Outreach Group, 2018, Dickinson et al., 2016, Bycroft et al., 2018), they are mostly lacking in neuroscience, with documented consequences for the reliability of ``discoveries'' (Baker, 2016; Botvinik-Nezer et al., 2020; Button et al., 2013).
+laboratories \citep{baker16,ioannidis05}. Although such tools exist in
+genetics, astronomy, physics and medicine 
+\citep{fishEtAl16,abdallaEtAl18,cern18,dickinsonEtAl16,bycroftEtAl18}, they are
+mostly lacking in neuroscience, with documented consequences for the
+reliability of ``discoveries'' \citep{baker16,botvinikNezerEtAl20,buttonEtAl13}.
 

### 2021-11-17 application/trackRecord.tex

@@
 
-In summary, the partners in this project have outstanding expertise on their focus areas (computational neuroscience and machine learning, experimental
+In summary, the partners in this project have outstanding expertise on their focus areas (computational neuroscience and ML, experimental
 neuroscience, implementation of experiments), as well as a rich collaborative

### 2021-11-17 application/backgroundToTheResource.tex
Updates from Overleaf

@@
 %
-Powerful experimental designs require behavioural feedback (e.g.\ reward, lights, sounds, actuated
+Powerful experimental designs require behavioural feedback (e.g.\ rewards, sounds, actuated
 devices) or neural perturbation (e.g.\ using light-sensitive polarising
@@
 %
-It has also become increasingly apparent that the neural signals that
+It has also become increasingly clear that the neural signals that
 drive complex behaviour are manifest in a broad population of neurons rather
-than in single cells. Current neural recording emphasises high-dimensional
+than in single cells. Neural recording now emphasises high-dimensional
 signals from 100s to 1000s of neurons monitored using multi-electrode arrays or cellular-resolution imaging, with a wide array of probe configurations and neural targets.
@@
 \textbf{This complexity of experiments limits the potential for database-driven progress in systems neuroscience}.
-The number of possible experiments is astronomical, and new questions are typically answered by designing and conducting new experiments rather than mining data from old ones.
+The number of possible experiments is astronomical, and new questions are typically answered by designing new experiments rather than mining data from old ones.
 
-At present, progress in the field is held back by the challenge of designing and implementing new experimental protocols, and the difficulty of integrating state-of-the-art data processing and neuroinformatics into custom experimental designs.
+Progress is thus held back by the challenges of designing and implementing new experimental protocols, and of integrating state-of-the-art data processing and neuroinformatics into custom experimental designs.
 %
-\textbf{Reproducibility}, in particular, depends on the availability of
-standardised tools for experimental control and analysis, shared between
-laboratories \citep{baker16,ioannidis05}. Although such tools exist in
-genetics, astronomy, physics and medicine 
-\citep{fishEtAl16,abdallaEtAl18,cern18,dickinsonEtAl16,bycroftEtAl18}, they are
-mostly lacking in neuroscience, with documented consequences for the
-reliability of ``discoveries'' \citep{baker16,botvinikNezerEtAl20,buttonEtAl13}.
+\textbf{Reproducibility}, in particular, requires standardised tools for experimental control and analysis, shared between
+laboratories \citep{baker16,ioannidis05}.
+Although such tools exist in genetics, astronomy, physics and medicine (Fish et al., 2016; Abdalla et al., 2018, CERN Education, Communications and Outreach Group, 2018, Dickinson et al., 2016, Bycroft et al., 2018), they are scarce in neuroscience, with documented consequences for the reliability of ``discoveries'' (Baker, 2016; Botvinik-Nezer et al., 2020; Button et al., 2013).
 
@@
 The proposed project will \textbf{enhance, reinforce and support} this key neuroinformatics
-resource to streamline \textbf{reproducible experimental design, data collection and analysis}, while making \textbf{standardised implementations of cutting-edge machine-learning methodologies} widely available to the community, potentially creating the most powerful experimental
+resource to streamline \textbf{reproducible experimental design, data collection and analysis}, while making \textbf{standardised implementations of cutting-edge machine-learning methodologies} widely available to the community.
+This will potentially create the most powerful experimental
 control and analysis tool available to neuroscientists, psychologists and

### 2021-11-17 application/communityDemand.tex
Updates from Overleaf

@@
 ecosystem. This integration will unlock new value
-and scientific directions from the rich behavioral and neural
+and scientific directions from the rich behavioural and neural
 experimental data that Bonsai already allows scientists to collect. This

### 2021-11-17 application/otherResources.tex
Updates from Overleaf

@@
 
-The space of technologies serving experimental control and behaviour monitoring
-is large, and is traditionally occupied either by domain-specific graphical
+The large field of technologies serving experimental control and behaviour monitoring is traditionally occupied either by domain-specific graphical
 user interfaces for control and recording of specific devices and experiment
@@
 Sanworks\footnote{https://sanworks.io/index.php}).
-These dedicated interfaces are typically very comfortable for experimenters
-operating within the specific domain that the tool is designed for, but tend to
-become unwieldy when there is a need to introduce a new device or variation of
-a task which is outside the scope of the framework. The alternative is to use a
-more general programming language like Python or MATLAB, with the disadvantage
+These dedicated interfaces are typically very comfortable for experimenters in the specific domain for which the tool is designed, but can become unwieldy with the introduction of a new device or task variation from outside their usual scope.
+Alternatively, one can use a more general programming language such as Python or MATLAB, with the disadvantage
 of the code being harder to understand, maintain, and change. Programming
-languages like LabVIEW straddle the middle ground and provide a high-level and
-flexible visual interface for composing data acquisition and control systems
-themselves. Unlike Bonsai, however, the graphical elements of LabVIEW are
-heterogeneous and very fine grained, creating the need for long and complicated
-logical structures to implement even a simple experimental control system. By
-providing an extremely simple, yet flexible, visual syntax, Bonsai opens the
-opportunity even for complete non-programmers to design and successfully
-customize relatively complex experiments from the ground up. It is mostly this
-capability which has made Bonsai so attractive as a standard tool in
+languages like LabVIEW straddle the middle ground and provide a high-level, flexible visual interface for composing data acquisition and control systems.
+Unlike Bonsai, however, the graphical elements of LabVIEW are
+heterogeneous and very fine grained, thus requiring long and complex
+logical structures to implement even a simple experimental control system. By providing an extremely simple, yet flexible visual syntax, Bonsai provides the
+opportunity for even complete non-programmers to design and successfully
+customise relatively complex experiments from the ground up. It is this
+capability in particular which has made Bonsai such an attractive standard tool in
 experimental neuroscience.

### 2021-11-17 application/userEngagement.tex
Updates from Overleaf

@@
 community of Bonsai users with expertise in software development. To fully integrate this new community and provide them with the full power of Bonsai,
-we plan to organize hackathons to accelerate integration of their existing data
+we plan to organise hackathons to accelerate integration of their existing data
 analysis and ML algorithms and methods into Bonsai using the new

### 2021-11-17 application/Je-S_sections.tex

@@
 
-The overall goal of this proposal is to extend, enhance, maintain and support Bonsai, a fully integrated software environment to enable cutting-edge reproducible laboratory systems neuroscience experiments in animal models.
+The overall goal of this proposal is to extend, enhance, maintain and support Bonsai, a fully integrated software environment enabling cutting-edge reproducible neuroscience and behavioural science experiments, with machine-intelligence-enabled, real-time neuroinformatics methods.
 
@@
 
-1.
-Develop, test and deploy a new software workflow to incorporate external packages into Bonsai's C# environment.
-\textcolor{red}{I believe this is already done.
-All existing bonsai community packages already depend on the existence of the extendable plugin architecture -- perhaps you mean "languages" rather than "packages"?
-i.e.
-MATLAB, Python, R, in that case I would rewrite as below}
-
-1. Develop, test and deploy new foreign language interfaces to allow extending the Bonsai environment with new packages developed in languages other than C#, namely MATLAB, Python and R.
+1. Develop, test and deploy new programming language interfaces to allow extending the Bonsai environment with new packages developed in languages other than C#, namely MATLAB, Python and R.
 

### 2021-11-17 application/backgroundToTheResource.tex

@@
 laboratories \citep{baker16,ioannidis05}.
-Although such tools exist in genetics, astronomy, physics and medicine (Fish et al., 2016; Abdalla et al., 2018, CERN Education, Communications and Outreach Group, 2018, Dickinson et al., 2016, Bycroft et al., 2018), they are scarce in neuroscience, with documented consequences for the reliability of ``discoveries'' (Baker, 2016; Botvinik-Nezer et al., 2020; Button et al., 2013).
+Although such tools exist in genetics, astronomy, physics and medicine \citep{fishEtAl16,abdallaEtAl18,cern18,dickinsonEtAl16,bycroftEtAl18}, they are scarce in neuroscience, with documented consequences for the reliability of ``discoveries'' \citep{baker16,botvinikNezerEtAl20,buttonEtAl13}.
 

### 2021-11-17 application/bonsai.tex

@@
 
-Bonsai has been in adopted in hundreds of laboratories worldwide and has the largest user base in the systems neuroscience community (Figure~\ref{fig:bonsai}D-F).
+Bonsai has been adopted in hundreds of laboratories worldwide and has the largest user base in the systems neuroscience community (Figure~\ref{fig:bonsai}D-F).
 In the last year alone, more than 1,000 new users incorporated Bonsai into their experimental protocols.
-The surprising rate of adoption of Bonsai in non-programmer experimental
+The rapid rate of adoption of Bonsai in non-programmer experimental
 labs highlights the need for accessible design tools that enable
@@
 outside their narrow domain of application. Bonsai has been successful because
-it has straddled this gap.
+it bridges this gap.
 
-The language has also helped to potentiate the growing wave of foundational
-open hardware initiatives, such as the OpenEphys \citep{siegleEtAl17} and UCLA
-Miniscope~\citep{caiEtAl16}, making it possible to quickly combine and
-integrate these tools into new experiments~\citep{buccinoEtAl18}.
+Bonsai also supports the growing wave of foundational
+open hardware initiatives, such as OpenEphys \citep{siegleEtAl17} and UCLA
+Miniscope~\citep{caiEtAl16}, allowing these tools to be quickly combined and integrated into new experiments~\citep{buccinoEtAl18}.
 %
@@
 Key to the wide adoption is the reproducibility that Bonsai ensures.  Data acquisition and experimental control protocols can be 
-replicated in any laboratory by just sharing a Bonsai configuration file. 
+replicated in any laboratory just by sharing a Bonsai configuration file. 
 %

### 2021-11-17 application/resourceManagementandSustainability.tex

@@
 management software included within the core Bonsai distribution and hosted
-in a shared centralized and curated repository.
+in a shared centralised and curated repository.
 %
@@
 Large-scale experimental projects such as the International Brain Laboratory
-have leveraged this capability of Bonsai to standardize data acquisition and
+have leveraged this capability of Bonsai to standardise data acquisition and
 control software across dozens of laboratories around the world.

### 2021-11-17 application/Je-S_sections.tex

@@
 
-5. Extend and support the Bonsai user community through code and forum engagement and development workshops.
+5. Extend and support the Bonsai user community through code, documentation, and forum engagement and development workshops.
 
@@
 
-Understanding the brain is a major scientific challenge of our era.
+Understanding the brain and the behaviour it generates is a major scientific challenge of our era.
 To succeed, scientists must be able to explain how animal behaviour relates to neural activity across different brain regions.
 This requires careful design and manipulation of behavioural experiments, where experimenters either record or manipulate neural activity while the animals (e.g.
-non-human primates, rodents, fish, insects) perform particular tasks.
+non-human primates, rodents, fish, insects) engage in specific behaviours which need to be carefully observed and quantified.
 

### 2021-11-17 application/detailsOfResource.tex

@@
 
-As it stands, Bonsai is designed to be extensible and is distributed with a package-management framework that controls installation of optional core packages as well as user-contributed extensions.
+Bonsai is designed to be extensible and is distributed with a package-management framework that controls installation of optional core packages as well as user-contributed extensions.
 %
-However extensions must be written in Bonsai's native C\# programming language.
+However, extensions must be written in Bonsai's native C\# programming language.
 %
-This specificity restricts the range of potential contributors to the ecosystem.
+This restricts the range of potential contributors to the ecosystem.
 %
-It also hinders easy incorporation of published analytic tools, many of which are implemented within other interpreted language systems such as Python, R, or MATLAB.  
+It also hinders incorporation of published analytic tools, many of which are implemented within other interpreted language systems such as Python, R, or MATLAB.  
 \begin{wrapfigure}{R}{0pt}
@@
 Precisely quantifying animal behaviour is an essential step toward understanding
-brain function.  Deeplabcut is a Python-based software for tracking animal body
-parts \citep{mathisEtAl18}, which it is currently well integrated with
+brain function.  DeepLabCut is a Python-based software for tracking animal body
+parts \citep{mathisEtAl18}, which is currently well integrated with
 Bonsai~\citep{kaneEtAl20}. Here we propose extensions to Bonsai to extract

### 2021-11-17 application/Je-S_sections.tex
Updates from Overleaf

@@
 
-Scientists studying brain and behaviour dedicate the majority of their time designing experiments and analysing the data, with least amount spent on data acquisition itself, which may impact the quality of data.
+Scientists studying brain and behaviour dedicate the majority of their time designing experiments and analysing the data, with least time spent on data acquisition itself, which may impact the quality of data.
 Moreover, hundreds of neuroscience research groups worldwide develop their own experimental and analytical tools, most using different programming languages, leading to  inefficiencies in data sharing and analysis, and impacting reproducibility (i.e.
-how easy it is for someone else to do the same experiment).
+how easy it is for someone else to repeat the same experiment).
 
-Here we propose to provide the neuroscience community with a software tool that will dramatically increase the efficiency of experimental control and data analysis.
+Here we propose to provide the scientific community with a software tool that will dramatically increase the efficiency of experimental control and data analysis.
 We will do so by developing a new set of functionalities to an existing software platform, Bonsai.
@@
 
-The main beneficiaries of this proposal will be experimental labs worldwide who currently lack the highly technical expertise required to apply Machine Intelligence to their experiments.
-While more and more automated Machine Learning platforms are becoming generally available across many data science fields, they either focus on making available completely general-purpose tools which require deep expertise and training to construct and train new algorithms and models from scratch, or they encapsulate some pre-trained and domain-specific model which is useful only for very specific classes of problems (e.g.
+The main beneficiaries of this proposal will be experimental brain and animal behaviour labs worldwide who currently lack the highly technical expertise required to apply Machine Intelligence to their experiments.
+While increasingly automated Machine Learning platforms are becoming available across many data science fields, they require deep expertise or training to construct and train new algorithms, or they are easy to deploy and use but are limited to very specific domains (e.g.
 face recognition, pose estimation).
 
-A middle ground where existing algorithms can be easily composed by non-experts in a flexible way to design entirely novel Machine Intelligence applications tailor-made for specific research questions is currently entirely absent.
+The scientific community will benefit from gains in efficiency and reproducibility of deploying Machine Intelligence pipelines and allowing non-experts to compose existing algorithms in a flexible way to design entirely novel Machine Intelligence applications tailor-made for specific research questions.
 In fields like systems neuroscience, where progress is made essentially by exploring the vast unknown space of animal behaviour and wide-scale brain physiology, the need for new bespoke algorithms and analysis procedures is constant and critical.
@@
 
-In addition to the researchers themselves, the standard analysis pipelines will allow everyone who is interested in asking questions about the behaviour and physiology of animals to analyse and understand their data.
-As an example, increasingly in research institutions worldwide Bonsai is being used to monitor animal welfare by in-house technicians who are concerned about a broader range of environmental variables which might impact animal housing quality but which are beyond the ones which are available in standard vivarium solutions.
+In addition to the researchers themselves, the standard analysis pipelines developed in this project will allow anyone who is interested in asking questions about the behaviour and physiology of animals to analyse and understand their data.
+As an example of these more general applications, increasingly in research institutions worldwide Bonsai is being used to monitor animal welfare by in-house technicians who are concerned about a broader range of environmental variables which might impact animal housing quality but which are beyond the ones which are available in standard vivarium solutions.
 They use Bonsai as it allows them to prototype and design entirely new monitoring systems on-site, without the need to engage an expensive software development process to answer simple questions.
-With access to the neuroinformatics packages developed for this proposal, staff at these institutions would be immediately empowered to address more sophisticated questions regarding interactions between housing conditions and animal physiology, which at the moment are entirely unthinkable, but would lead to instantly improving the welfare of hundreds of thousands of research animals in biological research facilities.
+With access to the neuroinformatics packages developed for this proposal, staff at these institutions would be immediately empowered to address more sophisticated questions regarding interactions between housing conditions and animal physiology, which at the moment may be unthinkable to them, but would lead to instantly improving the welfare of hundreds of thousands of research animals in biological research facilities.

### 2021-11-17 application/detailsOfResource.tex
Fixed dead link to lfpExtraction
Updates from Overleaf

@@
 
-    \item[Motion sequencing]\citep[MoSeq;][]{wiltschkoEtAl15}: extracts patterns
-        of behaviour that repeat over time (i.e., syllables of behaviour) from
+    \item[Motion sequencing]\citep[MoSeq;][]{wiltschkoEtAl15} extracts patterns
+        of behaviour that repeat over time (i.e. syllables of behaviour) from
         video data. For instance, it parses (in an unsupervised way) the
-        behaviour of a mouse in an arena into segments of time where a mouse
-        was running, rearing and grooming. It uses a hidden Markov model and it
+        behaviour of a mouse in an arena into segments of time where it        was running, rearing and grooming. It uses a hidden Markov model and
         is implemented in Python\footnote{Code for MoSeq can be requested from
@@
         Python\footnote{https://github.com/themattinthehatt/behavenet}.
-        After trained it can be used online to decode video data and detect
-        behavioral syllables.
+        After training it can be used online to decode video data and detect
+        behavioural syllables.
 
     \item[Interpretable latents for behavioral videos]\citep[Partitioned
-        Subspace Variational Autoencoder, PS-VAE;][]{whitewayEtAl21}: produces
-        interpretable low-dimensional representations of behavioral videos by
+        Subspace Variational Autoencoder, PS-VAE;][]{whitewayEtAl21} produces
+        interpretable low-dimensional representations of behavioural videos by
         combining the output of pose-estimation algorithms (e.g., DeepLabCut)
@@
         % Please refer to class \texttt{PSVAE} in texttt{behavenet.models.vaes.py}.
-        }. After trained it can be used
+        }. After training it can be used
         online to extract low-dimensional features and perform downstream
@@
 
-We propose to add functionality into Bonsai for the implementation of
+We propose to add functionality to Bonsai for the implementation of
 brain-computer-interface (BCI) applications illustrated in
 Figure~\ref{fig:proposedBonsaiExtensions}b. In
-this implementation voltage recordings from the brain will be converted into
+this implementation, voltage recordings from the brain will be converted into
 spikes fired by single neurons by a \texttt{SpikeDetector}
@@
 algorithms are designed to work offline (i.e., to use a complete recording
-session, after this session terminated). Online spike sorting (i.e., the task
-of assigning spikes to individual neurons as recordings are being aquired) is a
-still unsolved task, specially for recording devices with large number of
+session, after the session terminated). Online spike sorting (i.e., the task
+of assigning spikes to individual neurons as recordings are being acquired) is a
+still unsolved task, especially for recording devices with large number of
 electrodes.
@@
 by just detecting spikes, without assigning them to invidivual
-neurons~\citep{trautmannEtAl19,todorovaEtAl14}. Thus, for the BCI application
-proposed here we will only detect spikes with a simple zero-corssing method,
+neurons~\citep{trautmannEtAl19,todorovaEtAl14}. Thus, we will only detect spikes with a simple zero-corssing method,
 and not assign them to individual neurons.  We will, however, integrate into
-Bonsai two offline spike sorting methods, Kilosort~\citep[][written in Matlab]{pachitariuEtAl16} and MountainSort~\citep[][written in
-Python]{chungEtAl17}, to provide Bonsai users spike sorting functionality for
+Bonsai two offline spike sorting methods, Kilosort~\citep[][written in Matlab]{pachitariuEtAl16}, and MountainSort~\citep[][written in
+Python]{chungEtAl17}, to provide Bonsai users with spike sorting functionality for
 applications that benefit from it.
@@
 scientists to record neural data from a large number of devices. However, it
-lacks functionality to process these recordings. Here we describe software that
-we propose to integrate into Bonsai to extract interpretable summary statistics
+lacks functionality to process these recordings. Here we propose Bonsai software integrations to extract interpretable summary statistics
 (i.e., latent variables) from neural spiking activity.
@@
     \item[Gaussian Linear Dynamical
-        System]\citep[GLDS][]{andersonAndMoore12}: with sufficiently large
+        System]\citep[(GLDS)][]{andersonAndMoore12}: with sufficiently large
         bin sizes, spike counts can be modelled as Gaussian random processes.
@@
         parameters of linear dynamical system (LDS) models, as well as
-        inferences about their states. After models parameters have been
+        inferences about their states. After models' parameters have been
         learned, GLDS can be used online. A unique feature of GLDS is that the
@@
 %
-We propose to implement in Bonsai several decoding/classification algorithms:
+We propose implementing in Bonsai several decoding/classification algorithms:
 k-nearest neighbour, linear discriminative analysis, support vector machines,
@@
 
-We propose to add functionality to Bonsai to facilitate multi-method
+We propose adding functionality to Bonsai to facilitate multi-method
 comparisons in user-supplied datasets.
@@
 
-Multiple groups at the SWC currently use Bonsai for data acquisition and
+Multiple groups at the Sainsbury Wellcome Centre for Neural Circuits and Behaviour (SWC) currently use Bonsai for data acquisition and
 experimental control. Carefully testing the functionality of new software is
-essential to ensure its correct functionality. We propose to use the SWC large
-internal Bonsai user base to test the functionality we will add to Bonsai,
+essential to ensure correct functionality. We propose using the SWC's large Bonsai user base to test the added Bonsai functionality,
 before distributing it to the general public.
@@
 
-The current version of Bonsai facilitates reproducibility of data acquisition
+Bonsai currently facilitates reproducibility of data acquisition
 and experimental control across laboratories. Most data acquisition and
 experimental control aspects of an experiment can be reproduced across
-different laboratories by sharing a Bonsai configuration file.
+different laboratories by sharing a Bonsai configuration file. Adding data analysis functionality to Bonsai will also allow this to be reproduced by Bonsai users. 
 %
-By adding data analysis functionality to Bonsai, besides reproducing my data
-acquisition and experimental control, Bonsai users will also be able to reproduce
-data analysis functionality.
 
@@
 
-We propose to assess the progress of the proposed project using three different
+We propose assessing the progress of the proposed project using three 
 types of targets.

### 2021-11-17 application/communityDemand.tex

--- 2021-11-17 application/communityDemand.tex
@@
 package\footnote{https://groups.google.com/g/bonsai-users/c/BZ3zOOdv\_1c/m/x6OP75frBQAJ}. Even
-relatively modest contributions such as integrating support for specific models, such as DeepLabCut, or even the more specific zebrafish feature
-tracking package BonZeb, have attracted positive
+relatively modest contributions such as integrating support for specific models, such as DeepLabCut, or the more specific zebrafish feature-tracking package BonZeb, have attracted positive
 feedback from the community, \citep[e.g.,][]{kaneEtAl20,guilbeaultEtAl21}.
 
-The need for an easier interface into ML algorithms
+The need for an easier interface with ML algorithms
 is evidenced by the development of multiple graphical user interfaces for
@@
 user interfaces have no interaction with real-time experimental control, and
-their support libraries require expertise with the Python programming language
+their support libraries require expertise with the Python language
 to use flexibly. Bonsai addresses this gap, using its
@@
 technologies into Bonsai extremely timely. The proposed software infrastructure will facilitate the integration of novel ML methods into the Bonsai
-ecosystem. This integration will unlock new value
-and scientific directions from the rich behavioral and neural
-experimental data that Bonsai already allows scientists to collect. This
-integration will also provide experimentalists in the biological
-sciences who lack programming skills the opportunity to combine real-time ML with
-experimental control in neuroscience. To our knowledge
+ecosystem. This will unlock new value
+and scientific directions from the rich behavioural and neural data that Bonsai already allows scientists to collect. It will also give non-programming experimentalists in the biological
+sciences the opportunity to combine real-time ML with
+experimental control. To our knowledge
 this has not been done before on such a comprehensive scale, and will, we believe, make Bonsai into a fundamental and transformative technology for

### 2021-11-17 application/bonsai.tex
Fixed overhanging lines

@@
 International Brain
-Laboratory\footnote{https://www.internationalbrainlab.com/}
+Laboratory\\ \footnote{https://www.internationalbrainlab.com/}
 and the Allen Institute for Brain

### 2021-11-17 application/detailsOfResource.tex
Fixed overhanging lines

--- 2021-11-17 application/detailsOfResource.tex
@@
         neural analyses. It is based on autoencoders and is implemented in
-        Python\footnote{code for PS-VAE is embedded in the BehaveNet code
+        Python\footnote{code for PS-VAE is embedded in the BehaveNet code\\ 
         https://github.com/themattinthehatt/behavenet.%
@@
 \paragraph{LFP extraction}
-\label{sec:lfpExtractor}
+\label{sec:lfpExtraction}
 
@@
         which allows the use of external
-        inputs\footnote{https://github.com/joacorapela/kalmanFilter}.
+        inputs\\ \footnote{https://github.com/joacorapela/kalmanFilter}.
 
@@
         propose to interface Bonsai with a Matlab implementation of
-        PLDS\footnote{https://bitbucket.org/mackelab/pop\_spike\_dyn/src/master/}
+        PLDS\\ \footnote{https://bitbucket.org/mackelab/pop\_spike\_dyn/src/master/}
         that uses variational inference. PLDS does not provide online estimates
@@
         propose to use a Python implementation at
-        LFADS\footnote{https://github.com/tensorflow/models/tree/master/research/lfads.}.
+        LFADS\\ \footnote{https://github.com/tensorflow/models/tree/master/research/lfads.}.
         As GPFA, LFADS do not provide online estimates of states.

### 2021-11-17 application/otherResources.tex
Fixed overhanging lines

@@
 machine or similar formalisms (e.g. NIMH ML\footnote{https://monkeylogic.nimh.nih.gov/},
-pyControl\footnote{https://pycontrol.readthedocs.io/en/latest/},
+pyControl\\ \footnote{https://pycontrol.readthedocs.io/en/latest/},
 Autopilot\footnote{https://docs.auto-pi-lot.com/en/latest/},
-Sanworks\footnote{https://sanworks.io/index.php}).
+Sanworks\\ \footnote{https://sanworks.io/index.php}).
 These dedicated interfaces are typically very comfortable for experimenters in the specific domain for which the tool is designed, but can become unwieldy with the introduction of a new device or task variation from outside their usual scope.

### 2021-11-17 application/userEngagement.tex

@@
 Summer School, to increase awareness of Bonsai and
-its packages throughout the neuroscience community. Together with participation
-at smaller invited venues, there has been an average of 5 courses in Bonsai
+its packages throughout the neuroscience community. With participation
+at smaller invited venues, there have been on average 5 Bonsai courses
 each year, with at least 20 students each; i.e.  around 100 neuroscience students annually are introduced to experimental tools directly through a
@@
 
-Bonsai's 10th anniversary will take place in 2022. The proposed ML package will thus leverage planned venues, events, and infrastructure to disseminate
+Bonsai's 10th anniversary will take place in 2022. The proposed ML package will leverage planned events, and infrastructure to disseminate
 awareness throughout the community, including presentations at conferences,
 workshops, and training sessions, as well as electronic newsletters, forums and
-social media articles.
+social media.
 
 The proposed development will facilitate interfacing with other programming languages
-such as Python, R, and MATLAB. This will therefore attract a whole new
-community of Bonsai users with expertise in software development. To fully integrate this new community and provide them with the full power of Bonsai,
+such as Python, R, and MATLAB. This will attract a whole new
+community of Bonsai users with expertise in software development. To provide these new users with the full power of Bonsai,
 we plan to organise hackathons to accelerate integration of their existing data
 analysis and ML algorithms and methods into Bonsai using the new
-machine learning package.
+ML package.
 

### 2021-11-17 application/caseForSupport.tex
Updates from Overleaf

@@
         uncover how adaptive behaviour arises from computation in neural
-        circuits in different brain regions. Extending the Bonsai ecosystem
+        circuits across brain regions. Extending the Bonsai ecosystem
         will facilitate new types of experiment that causally connect neural
@@
         will support the development of new neural circuit biomarkers
-        and therapies. The Bonsai ecosystem will also enable  experimental
-        testing of emerging theories of brain function and learning, and thus
-        contribute to the neuro-inspired ML algorithms that fuel
+        and therapies. Bonsai will also enable experimental
+        testing of emerging theories of brain function and learning, thus
+        contributing to the neuro-inspired ML algorithms that fuel
         intelligent technologies, increasingly important in our society.  The
@@
     \item[Improve reproducibility in behavioural and brain sciences] Lack of
-        reproducible results slows down scientific progress and incurs personal
-        and economic costs to scientists and funders.  In neuroscience,  many
+        reproducible results slows scientific progress, increasing costs to scientists and funders.  Many
         behavioural and neural animal studies have been difficult to
-        reproduce, and experiments are rarely replicated across laboratories.
-        By creating a resource that standardises and enables reproducible
+        reproduce across laboratories.
+        By standardising and enabling reproducible
         experimental design/control, as well as analysis of neural and
-        behavioural data, the Bonsai ecosystem will allow multiple laboratories
+        behavioural data, Bonsai will allow multiple laboratories
         to implement and align experimental and analytical protocols as closely
@@
     \item[Improve efficiency of research] Most neuroscience laboratories
-        worldwide develop custom, unique solutions to their experimental
+        develop custom solutions to their experimental
         control and analysis needs. Some laboratories also lack the technical

### 2021-11-17 application/detailsOfResource.tex
Updates from Overleaf

--- 2021-11-17 application/detailsOfResource.tex
@@
 by just detecting spikes, without assigning them to invidivual
-neurons~\citep{trautmannEtAl19,todorovaEtAl14}. Thus, we will only detect spikes with a simple zero-corssing method,
+neurons~\citep{trautmannEtAl19,todorovaEtAl14}. Thus, we will only detect spikes with a simple zero-crossing method,
 and not assign them to individual neurons.  We will, however, integrate into
@@
         which allows the use of external
-        inputs\footnote{https://github.com/joacorapela/kalmanFilter}.
+        inputs\footnote{https://github.com/joacorapela/ kalmanFilter}.
 

### 2021-11-17 application/Je-S_sections.tex

@@
 The main beneficiaries of this proposal will be experimental brain and animal behaviour labs worldwide who currently lack the highly technical expertise required to apply Machine Intelligence to their experiments.
+In fields like systems neuroscience, where progress is made essentially by exploring the vast unknown space of animal behaviour and wide-scale brain physiology, the need for redesign of experimental protocols and rapid analysis of large and complex datasets is constant and critical.
 While increasingly automated Machine Learning platforms are becoming available across many data science fields, they require deep expertise or training to construct and train new algorithms, or they are easy to deploy and use but are limited to very specific domains (e.g.
 face recognition, pose estimation).
-
-The scientific community will benefit from gains in efficiency and reproducibility of deploying Machine Intelligence pipelines and allowing non-experts to compose existing algorithms in a flexible way to design entirely novel Machine Intelligence applications tailor-made for specific research questions.
-In fields like systems neuroscience, where progress is made essentially by exploring the vast unknown space of animal behaviour and wide-scale brain physiology, the need for new bespoke algorithms and analysis procedures is constant and critical.
 If successful, this proposal will entirely transform the landscape of Machine Intelligence applied to experimental lab research in biology, by empowering non-technical researchers to recombine and repurpose Machine Intelligence tools using a flexible visual programming environment and providing seamless integration with all the measurement, instrumentation and control packages already available in Bonsai.
 
-In addition to the researchers themselves, the standard analysis pipelines developed in this project will allow anyone who is interested in asking questions about the behaviour and physiology of animals to analyse and understand their data.
-As an example of these more general applications, increasingly in research institutions worldwide Bonsai is being used to monitor animal welfare by in-house technicians who are concerned about a broader range of environmental variables which might impact animal housing quality but which are beyond the ones which are available in standard vivarium solutions.
-They use Bonsai as it allows them to prototype and design entirely new monitoring systems on-site, without the need to engage an expensive software development process to answer simple questions.
-With access to the neuroinformatics packages developed for this proposal, staff at these institutions would be immediately empowered to address more sophisticated questions regarding interactions between housing conditions and animal physiology, which at the moment may be unthinkable to them, but would lead to instantly improving the welfare of hundreds of thousands of research animals in biological research facilities.
+Machine learning and computational neuroscience experts will also benefit from a standard platform to compare and benchmark different algorithms and stress test their tools against community datasets and experiments.
+The addition of new programming languages to develop and interface with Bonsai will also greatly expand the potential for application of their methods into new domains of research.
+
+The scientific community will collectively benefit from gains in efficiency and reproducibility of deploying Machine Intelligence pipelines and allowing non-experts to compose existing algorithms in a flexible way to design entirely novel Machine Intelligence applications tailor-made for specific research questions.
+The proposed tools will generate cost savings to individual laboratories as well as agencies who fund them by drastically reducing duplicated software development efforts across research groups.
+
+In addition to the research community, the standard analysis pipelines developed in this project may be applied to other domains of society where tracking, monitoring and understanding animal behaviour would yield benefits and improved outcomes.
+For example, Bonsai is being increasingly used in research institutions worldwide to monitor animal welfare by in-house technicians who are concerned about a broader range of environmental variables which might impact animal husbandry but which are beyond the ones which can be monitored in standard vivarium solutions.
+As a free and open-source software package, Bonsai would empower such users to prototype and design new animal monitoring systems on-site, without the need to engage in expensive software development to answer simple questions, and would lead to instantly improving the welfare of hundreds of thousands of research animals in biological research facilities.

### 2021-11-17 application/detailsOfResource.tex

--- 2021-11-17 application/detailsOfResource.tex
@@
         which allows the use of external
-        inputs\\ \footnote{https://github.com/joacorapela/kalmanFilter}.
+        inputs\footnote{https://github.com/joacorapela/ kalmanFilter}.
 

### 2021-11-17 application/resourceManagementandSustainability.tex

@@
 The documentation, including learning materials and all relevant community
-links are on a publicly accessible website at \url{bonsai-rx.org}.
+links are publicly accessible at \url{bonsai-rx.org}.
 %
-A community user forum is a public Google Group with more than 700 active users (Figure~\ref{fig:bonsai}f) and several
-new questions and answers exchanged every day.
+A community user forum is a public Google Group with more than 700 active users (Figure~\ref{fig:bonsai}f) exchanging several
+new questions and answers every day.
 %
@@
 %
-This ensures that bugfixes and enhancements are easily available but
-also importantly allows for reproducible deployment of user environments
-to ensure working experiments are not disrupted by changes in the system.
+This ensures that bugfixes and enhancements are easily available and, importantly, allows for reproducible deployment of user environments
+to ensure working experiments are not disrupted by system changes.
 Large-scale experimental projects such as the International Brain Laboratory
@@
 Community technology previews are made available through the package manager,
-allowing early adopters to opt-in to the latest features
+letting early adopters opt-in to the latest features
 before they are generally available. This allows the community
-to help report issues with new features on GitHub, thereby
-distributing bug reporting and triage and overall improving the quality of
+to report issues with new features on GitHub, thereby
+distributing bug reporting and triage, and improving the quality of
 stable releases.
 
-The core compiler and IDE is extensively tested, and takes advantage of
-a strong type system to greatly reduce bugs at compile-time. A set of best
-practices for leveraging this infrastructure for building new packages is
+The core compiler and IDE is extensively tested, and uses
+a strong type system to greatly reduce bugs at compile-time. Best
+practices for leveraging this infrastructure for building new packages are
 in place.
@@
 NeuroGEARS Ltd is a keystone of this community and a key collaborator in the current proposal.
-As a company founded by the original developer of Bonsai, it helps to support and expand the
+Founded by the original developer of Bonsai, it helps to support and expand the
 user base well beyond the life of the current proposal. NeuroGEARS currently provides
-quality control and curation for the core Bonsai code, as well as providing training and support
+quality control and curation for the core Bonsai code, and provides training and support
 services for research institutions around the world who want to start using Bonsai.
-Finally, NeuroGEARS also facilitates knowledge transfer across users and institutions adopting
+NeuroGEARS also facilitates knowledge transfer between users and institutions adopting
 Bonsai, by moderating the community user forums and collaborating with organisations
@@
 Developing these new packages with input from the core advisory panel of Bonsai,
-and early public distribution and user testing within the SWC and planned workshops, will ensure
+and early public distribution and user testing within the SWC and with planned workshops, will ensure
 solid and broad user engagement from the beginning of the development cycle and will therefore

### 2021-11-17 application/userEngagement.tex

@@
 such as Python, R, and MATLAB. This will attract a whole new
-community of Bonsai users with expertise in software development. To provide these new users with the full power of Bonsai,
+community of Bonsai users with expertise in software development. To provide these new users with Bonsai's full power,
 we plan to organise hackathons to accelerate integration of their existing data
@@
  components of the platform, such as the Python and MATLAB integrations. We
- also want to promote the co-development of algorithms by early adopters of the
- package, in true collaborative fashion. Given its modular nature and built-in
+ also want to promote co-development of algorithms by early adopters, in true collaborative fashion. Given its modular nature and built-in
  package manager capable of handling dependency resolution and curation of

### 2021-11-17 application/bonsai.tex
Merged overhaning lines conflict.

--- 2021-11-17 application/bonsai.tex
@@
 International Brain
-Laboratory\footnote{https://www.internationalbrainlab.com/}
+Laboratory\footnote{internationalbrainlab.com/}
 and the Allen Institute for Brain

### 2021-11-17 application/detailsOfResource.tex
Merged overhaning lines conflict.

--- 2021-11-17 application/detailsOfResource.tex
@@
         Python\footnote{code for PS-VAE is embedded in the BehaveNet code
-        https://github.com/themattinthehatt/behavenet.%
+        github.com/ themattinthehatt/behavenet.%
         % Please refer to class \texttt{PSVAE} in texttt{behavenet.models.vaes.py}.
@@
         propose to interface Bonsai with a Matlab implementation of
-        PLDS\footnote{https://bitbucket.org/mackelab/pop\_spike\_dyn/src/master/}
+        PLDS\footnote{https://bitbucket.org/mackelab/ pop\_spike\_dyn/src/master/}
         that uses variational inference. PLDS does not provide online estimates
@@
         propose to use a Python implementation at
-        LFADS\footnote{https://github.com/tensorflow/models/tree/master/research/lfads.}.
+        LFADS\footnote{https://github.com/tensorflow/models/tree/ master/research/lfads.}.
         As GPFA, LFADS do not provide online estimates of states.

### 2021-11-17 application/otherResources.tex
Merged overhaning lines conflict.

--- 2021-11-17 application/otherResources.tex
@@
 or by real-time control frameworks for specifying task logic using state
-machine or similar formalisms (e.g. NIMH ML\footnote{https://monkeylogic.nimh.nih.gov/},
-pyControl\footnote{https://pycontrol.readthedocs.io/en/latest/},
-Autopilot\footnote{https://docs.auto-pi-lot.com/en/latest/},
-Sanworks\footnote{https://sanworks.io/index.php}).
+machine or similar formalisms (e.g. NIMH ML\footnote{monkeylogic.nimh.nih.gov/},
+pyControl\footnote{pycontrol.readthedocs.io/ en/latest/},
+Autopilot\footnote{docs.auto-pi-lot.com/en/latest/},
+Sanworks\footnote{sanworks.io/index.php}).
 These dedicated interfaces are typically very comfortable for experimenters in the specific domain for which the tool is designed, but can become unwieldy with the introduction of a new device or task variation from outside their usual scope.

### 2021-11-17 application/Je-S_sections.tex

@@
 For example, Bonsai is being increasingly used in research institutions worldwide to monitor animal welfare by in-house technicians who are concerned about a broader range of environmental variables which might impact animal husbandry but which are beyond the ones which can be monitored in standard vivarium solutions.
-As a free and open-source software package, Bonsai would empower such users to prototype and design new animal monitoring systems on-site, without the need to engage in expensive software development to answer simple questions, and would lead to instantly improving the welfare of hundreds of thousands of research animals in biological research facilities.
+As a free and open-source software package, Bonsai would empower such users to prototype and design new animal monitoring systems on-site, without the need to engage in expensive software development to answer simple questions, and would lead to instantly improving the welfare of hundreds of thousands of research animals in biological research facilities, farming or .
+More broadly, Bonsai may also be applied to quantify dynamics of objects in wider settings such as videos obtained from drone or satellite imagery, to monitor the growth of crops, deforestation or climate dynamics of biological ecosystems.

### 2021-11-17 application/Je-S_sections.tex
merged

--- 2021-11-17 application/Je-S_sections.tex
@@
 
-The main beneficiaries of this proposal will be experimental brain and animal behaviour labs worldwide who currently lack the highly technical expertise required to apply Machine Intelligence to their experiments.
-In fields like systems neuroscience, where progress is made essentially by exploring the vast unknown space of animal behaviour and wide-scale brain physiology, the need for redesign of experimental protocols and rapid analysis of large and complex datasets is constant and critical.
-While increasingly automated Machine Learning platforms are becoming available across many data science fields, they require deep expertise or training to construct and train new algorithms, or they are easy to deploy and use but are limited to very specific domains (e.g.
+The main beneficiaries of the proposed resource will be experimental brain and animal behaviour labs worldwide who currently lack the highly technical expertise required to apply Machine Intelligence to their experiments.
+In fields like systems neuroscience, where progress is made by exploring the unknowns of animal behaviour and wide-scale brain physiology, the need for redesign of experimental protocols and rapid analysis of large and complex datasets is constant and critical.
+While increasingly-automated Machine Learning platforms are becoming available across many data science fields, they require deep expertise or training to construct and train new algorithms; or else they are easy to use but limited to very specific domains (e.g.
 face recognition, pose estimation).
-If successful, this proposal will entirely transform the landscape of Machine Intelligence applied to experimental lab research in biology, by empowering non-technical researchers to recombine and repurpose Machine Intelligence tools using a flexible visual programming environment and providing seamless integration with all the measurement, instrumentation and control packages already available in Bonsai.
+If successful, this proposal will empower an entirely new community of non-technical researchers to recombine and repurpose Machine Intelligence tools using a flexible visual programming environment and providing seamless integration with all the measurement, instrumentation and control packages already available in Bonsai.
 
-Machine learning and computational neuroscience experts will also benefit from a standard platform to compare and benchmark different algorithms and stress test their tools against community datasets and experiments.
+Machine Learning and computational neuroscience experts will also benefit from a standard platform to compare and benchmark different algorithms and stress-test their tools against community datasets and experiments.
 The addition of new programming languages to develop and interface with Bonsai will also greatly expand the potential for application of their methods into new domains of research.
 
-The scientific community will collectively benefit from gains in efficiency and reproducibility of deploying Machine Intelligence pipelines and allowing non-experts to compose existing algorithms in a flexible way to design entirely novel Machine Intelligence applications tailor-made for specific research questions.
-The proposed tools will generate cost savings to individual laboratories as well as agencies who fund them by drastically reducing duplicated software development efforts across research groups.
+The scientific community will collectively benefit from gains in efficiency and reproducibility of deploying Machine Intelligence pipelines, and from allowing non-experts to compose existing algorithms in a flexible way to design entirely novel Machine Intelligence applications tailor-made for specific research questions.
+The proposed tools will generate cost savings to individual laboratories as well as funding agencies, by drastically reducing duplicated software development efforts across research groups.
 
 In addition to the research community, the standard analysis pipelines developed in this project may be applied to other domains of society where tracking, monitoring and understanding animal behaviour would yield benefits and improved outcomes.
-For example, Bonsai is being increasingly used in research institutions worldwide to monitor animal welfare by in-house technicians who are concerned about a broader range of environmental variables which might impact animal husbandry but which are beyond the ones which can be monitored in standard vivarium solutions.
-As a free and open-source software package, Bonsai would empower such users to prototype and design new animal monitoring systems on-site, without the need to engage in expensive software development to answer simple questions, and would lead to instantly improving the welfare of hundreds of thousands of research animals in biological research facilities.
+For example, Bonsai is  increasingly used in research institutions worldwide to monitor animal welfare by in-house technicians who are concerned about a broader range of environmental variables which might impact animal husbandry but which cannot be monitored in standard vivarium solutions.
+As a free and open-source software package, Bonsai would empower such users to prototype and design new animal monitoring systems on-site, without the need to engage in expensive software development to answer simple questions.
+It could thus contribute to improving the welfare of hundreds of thousands of animals in research facilities, zoos or farms.
+More broadly, Bonsai may also be applied to quantify dynamics of objects in wider settings from drone or satellite imagery, to track and quantify the dynamics of crop growth, deforestation or other biological ecosystems.

### 2021-11-17 application/bonsai.tex
merged

@@
 International Brain
-Laboratory\footnote{internationalbrainlab.com/}
+Laboratory \footnote{internationalbrainlab.com/}
 and the Allen Institute for Brain

### 2021-11-17 application/communityDemand.tex
merged

@@
 explicit proposals to create a ML
-package\footnote{https://groups.google.com/g/bonsai-users/c/BZ3zOOdv\_1c/m/x6OP75frBQAJ}. Even
+package\footnote{groups.google.com/g/bonsai-users/c/BZ3zOOdv\_1c/m/x6OP75frBQAJ}. Even
 relatively modest contributions such as integrating support for specific models, such as DeepLabCut, or the more specific zebrafish feature-tracking package BonZeb, have attracted positive

### 2021-11-17 application/detailsOfResource.tex
merged

@@
 \begin{wrapfigure}{R}{0pt}
-  \includegraphics[width=0.5\textwidth]{figures/extensions.pdf}%
+  \includegraphics[width=0.6\textwidth]{figures/extensions.pdf}%
 %   \input{figures/extensions}
@@
 %
-Python.NET\footnote{https://github.com/pythonnet/pythonnet},
+Python.NET\footnote{github.com/pythonnet/pythonnet},
 %
-R.Net\break \footnote{https://github.com/rdotnet/rdotnet}
+R.Net\break \footnote{github.com/rdotnet/rdotnet}
 %
@@
 %
-If necessary, we will also incorporate direct inter-runtime communication using the messaging library ZeroMQ\footnote{https://zeromq.org/}.
+If necessary, we will also incorporate direct inter-runtime communication using the messaging library ZeroMQ\footnote{zeromq.org/}.\\
 
@@
         propose to interface Bonsai with a Matlab implementation of
-        PLDS\footnote{https://bitbucket.org/mackelab/ pop\_spike\_dyn/src/master/}
+        PLDS\footnote{bitbucket.org/mackelab/ pop\_spike\_dyn/ src/master/}
         that uses variational inference. PLDS does not provide online estimates
@@
 Second, we will monitor user activity in the Bonsai
-forum\footnote{https://groups.google.com/g/bonsai-users},
+forum\footnote{groups.google.com/g/bonsai-users},
 and publications citing Bonsai that use the new ML functionality,

### 2021-11-17 application/narrativeCVRapela.tex
merged

@@
 
-Therefore, Dr.~Rapela is excellently trained to implement all aspects
-(neuroscience, statistics, signal processing, and software development) of the
-proposed project.
+Therefore, Dr.~Rapela's expertise is perfectly suited to implement all aspects
+of the
+proposed project (neuroscience, statistics, signal processing, and software development).
 

### 2021-11-17 application/otherResources.tex
merged

@@
 types (e.g.\ Open Ephys
-GUI\footnote{https://open-ephys.org/gui/},
+GUI\footnote{open-ephys.org/gui/},
 Miniscope DAQ
-Software\footnote{https://github.com/Aharoni-Lab/Miniscope-DAQ-QT-Software})
+Software\footnote{github.com/ Aharoni-Lab/Miniscope-DAQ-QT-Software})
 or by real-time control frameworks for specifying task logic using state
 machine or similar formalisms (e.g. NIMH ML\footnote{monkeylogic.nimh.nih.gov/},
-pyControl\footnote{pycontrol.readthedocs.io/ en/latest/},
+pyControl\footnote{pycontrol. readthedocs.io/ en/latest/},
 Autopilot\footnote{docs.auto-pi-lot.com/en/latest/},
-Sanworks\footnote{sanworks.io/index.php}).
+Sanworks\footnote{sanworks.io/index.php}).\break
 These dedicated interfaces are typically very comfortable for experimenters in the specific domain for which the tool is designed, but can become unwieldy with the introduction of a new device or task variation from outside their usual scope.

### 2021-11-17 application/narrativeCVSahani.tex

@@
 
+{
+  \parindent 0pt
+  \parskip 1ex
+
 \section{Prof.\ Maneesh Sahani -- narrative CV}
@@
 
-I am currently Professor of Theoretical Neuroscience and Machine Learning, and
+I am currently \textbf{Professor of Theoretical Neuroscience and Machine Learning}, and
 director of the Gatsby Computational Neuroscience Unit at UCL.
 %
-As such, I am resident in the UK, and hold an open-ended academic
-employment contract with UCL.
+I am a UK resident and hold an open-ended academic employment contract
+with UCL.
 
-I
+I earned my PhD from the California Institute of Technology (Caltech)
+in 1999, working in the laboratories of Profs.\ Richard Andersen and
+John Hopfield.
+%
+After post-doctoral work at UCL/Gatsby (with Prof.\ Peter Dayan) and
+UCSF (Prof.\ Michael Merzenich) I joined the academic staff of the
+Gatsby Unit as a Lecturer in 2004, becoming Reader in 2009 and
+Professor in 2013.
+%
+I took over the Directorship of the Unit at the end of 2017.
 
@@
 
-\subsection{Contributions to ideas}
+\subsection{Contributions to ideas, hypotheses, tools and knowledge}
 
-Prof.\ Sahani has authored over 100 peer-reviewed scientific papers,
-with an h-index (computed by Google scholar) of 47.
+I have authored over 100 peer-reviewed scientific papers, with an
+h-index (computed by Google scholar) of 47.
 
-GPFA
+A substantial component of my research focuses on the development of
+advanced machine-learning based tools for neuroscience research.
+%
+This included my PhD research.  My thesis \textit{Latent variable
+  models for neural data analysis} introduced new techniques of
+probabilistic inference and applied these to the problem of ``spike
+sorting'' and clustering of neuronal activity profiles.  It has been cited
+over 180 times.
 
-PLDS
+Beginning around 2005, my group published a series of new
+neuroinformatics tools designed to characterise and understand
+population-scale activity using the large-scale multielectrode
+recording methods being developed.  These papers provided the backbone
+for a new analytic approach that is now being employed and extended by
+systems neuroscience laboratories worldwide.
+%
+A central component of the current proposal is to make this approach
+(and others) available within Bonsai, easing its adoption by a wider
+group of laboratories that lack in-house informatics expertise.
+%
+Key foundational papers include:
+\begin{itemize}[topsep=0pt,itemsep=0pt]
+\item  B. M. Yu, A. Afshar, G. Santhanam, S. I. Ryu, K. V. Shenoy, and M. Sahani.
+Extracting dynamical structure embedded in neural activity.
+In Y. Weiss, B. Schölkopf, and J. Platt, eds., Advances in Neural
+Information Processing Systems, vol. 18, pp. 1545–1552. MIT Press,
+Cambridge, Massachusetts, 2006.\\
+Cited 91 times.
+% \item M. M. Churchland, B. M. Yu, M. Sahani, and K. V. Shenoy.
+% Techniques for extracting single-trial activity patterns from large-scale neural recordings.
+% Current Opinion in Neurobiology, 17(5):609–618, 2007.\\
+% Review, cited 183 times.
+\item B. M. Yu, J. P. Cunningham, G. Santhanam, S. I. Ryu, K. V. Shenoy, and M. Sahani.
+Gaussian-process factor analysis for low-dimensional single-trial analysis of neural population activity.
+Journal of Neurophysiology, 102:614–635, 2009.\\
+Introduced the GPFA method; cited 561 times.
+% \item B. Petreska, B. M. Yu, J. P. Cunningham, G. Santhanam, S. I. Ryu, K. V. Shenoy, and M. Sahani.
+% Dynamical segmentation of single trials from population neural data.
+% In J. Shawe-Taylor, R. S. Zemel, P. Bartlett, F. C. N. Pereira, and
+% K. Q. Weinberger, eds., Advances in Neural Information Processing
+% Systems, vol. 24, pp. 756–764. Curran Associates, Inc., Red Hook, New
+% York, 2011.\\
+% Cited 64 times.
 
-Grant funding?
+\item J. H. Macke, L. Büsing, J. P. Cunningham, B. M. Yu, K. V. Shenoy, and M. Sahani.
+Empirical models of spiking in neural populations.
+In J. Shawe-Taylor, R. S. Zemel, P. Bartlett, F. C. N. Pereira, and
+K. Q. Weinberger, eds., Advances in Neural Information Processing
+Systems, vol. 24, pp. 1350–1358. Curran Associates, Inc., Red Hook,
+New York, 2011.\\
+Introduced effective methods to identify dynamical systems underlying
+population data.  Cited 210 times.
+\end{itemize}
+\noindent This remains an active thread of my research, with key recent
+papers including:
+\begin{itemize}[topsep=0pt,itemsep=0pt]
+% \item L. Buesing, J. H. Macke, and M. Sahani.
+% Spectral learning of linear dynamics from generalised-linear observations with application to neural population data.
+% In P. Bartlett, F. C. N. Pereira, L. Bottou, C. J. C. Burges, and K. Q. Weinberger, eds., Advances in Neural Information Processing Systems, vol. 25, 2012.
+% 
+% \item J. H. Macke, L. Buesing, and M. Sahani.
+% Estimating state and model parameters in state-space models of spike trains.
+% In Z. Chen, ed., Advanced State Space Methods for Neural and Clinical Data. Cambridge University Press, 2015.
+
+\item L. Duncker, G. Bohner, J. Boussard, and M. Sahani.
+Learning interpretable continuous-time models of latent stochastic dynamical systems.
+In K.
+Chaudhuri and R.
+Salakhutdinov, eds., Proceedings of the 36th International Conference on Machine Learning, vol.
+97 of Proceedings of Machine Learning Research, pp.
+1726–1734.
+PMLR, Long Beach, California, USA, 09–15 Jun 2019.
+\item V. Rutten, A. Bernacchia, M. Sahani, and G. Hennequin.
+Non-reversible gaussian processes for identifying latent dynamical structure in neural data.
+In H. Larochelle, M. Ranzato, R. Hadsell, M. F. Balcan, and H.-T. Lin,
+eds., Advances in Neural Information Processing Systems 33. Curran
+Associates, Inc., 2020.
+\item H. Soulat, S. Keshvarzi,  T. Margrie, and M. Sahani.
+  Probabilistic Tensor Decomposition of Neural Population Spiking
+  Activity.
+  Advances in Neural Information Processing Systems 34. Curran
+  Associates, Inc., 2021.
+\end{itemize}
+Many of the algorithms described in these papers form the basis of
+capabilities to be incorporated into Bonsai as part of the proposed
+project.
+
+I have also developed a number of other neuroinformatics tools
+relevant to the other capabiilties to be developed, and so my
+expertise here will be valuable for evaluation and guidance.
+%
+Key examples include  decoding methods:
+\begin{itemize}[topsep=0pt,itemsep=0pt]
+  \item B. M. Yu, G. Santhanam, M. Sahani, and K. V. Shenoy.
+Neural decoding for motor and communication prostheses.
+In K. G. Oweiss, ed., Statistical Signal Processing for Neuroscience,
+pp. 219–263. Elsevier, 2010.
+\item G. Santhanam, B. M. Yu, V. Gilja, S. I. Ryu, A. Afshar, M. Sahani, and K. V. Shenoy.
+Factor-analysis methods for higher-performance neural prostheses.
+Journal of Neurophysiology, 102:1315–1330, 2009. 
+\item B. M. Yu, C. Kemere, G. Santhanam, A. Afshar, S. I. Ryu, T. H. Meng, M. Sahani, and K. V. Shenoy.
+Mixture of trajectory models for neural decoding of goal-directed movements.
+Journal of Neurophysiology, 97(5):3763–3780, 2007. 
+\end{itemize}
+Cell segmentation methods for optical imaging:
+\begin{itemize}[topsep=0pt,itemsep=0pt]
+    \item M. Pachitariu, A. Packer, N. Pettit, H. Dalgleish, M. Hausser, M. Sahani.
+    Extracting regions of interest from biological images with convolutional sparse block coding
+    Advances in Neural Information Processing Systems 26, 1745-1753,
+    2013.
+  \item G. Bohner and M. Sahani.
+    Convolutional higher order matching pursuit.
+    In 2016 IEEE 26th International Workshop on Machine Learning for
+    Signal Processing (MLSP), 2016.
+\end{itemize}
 
@@
 
-As a PI, I have supervised a total of \textbf{X} students and \textbf{X} postdocs.  
-In the last five years, 4 of students have moved to postdoctoral positions (at
-Aalto, Cambridge, CMU and Stanford), 1 works at DeepMind, and 1 holds a faculty offer at Shanghai Tech.
-In all, 8 previous students have now secured faculty (or group leader) positions at Cambridge, CMU, Columbia, Copenhagen, Janelia (2), Pittsburgh and Shanghai.
-%
-Of the postdoctoral fellows I have supervised in the last five years, one is doing further postdoc work in Bonn; another holds an independent
-Research Fellowship at Radboud; and 2 have faculty positions: at Edinburgh and at UBC.
+To date, I have supervised a total of 14 PhDs and 14 postdocs.
 
-DARPA repair
+Of the students, 8 now hold faculty (or group leader) positions at
+Cambridge, CMU, Columbia, Copenhagen, Janelia (2), Pittsburgh and
+Shanghai; 2 work in corporate research groups (DeepMind and Advanced
+Bionics); and 4 recent graduates are currently in postdoctoral
+positions (Aalto, Cambridge, CMU, Stanford).
 
-Since 2017, I have been Director of the Gatsby Computational Neuroscience Unit, and have spearheaded its new strategic research direction.
-I created two new roles to support and develop the new strategy, and have recruited 2 new faculty members who have added new and highly relevant research strands to our research portfolio.
-As Director, I sit in the Executive Leadership Committee of the Faculty of Life Sciences at UCL.
+Of the postdoctoral fellows, 6 have faculty positions at British
+Columbia, Edinburgh, Oldenburg (2), Queensland and Tuebingen; 2 hold
+independent research positions (Livermore Labs, Radboud); and 4 work
+in industry.
+
+Since 2017, I have been Director of the Gatsby Computational
+Neuroscience Unit, leading strategy in research and teaching.  I have
+created two new roles to support and develop new strategy, and have
+recruited two new members of faculty who have expanded our research
+portfolio. As Director, I sit in the Executive Leadership Committee of
+the Faculty of Life Sciences at UCL.
 
@@
 
-Current collaborations/networks?
+I have been an active contributor to the academic community, shaping
+the development of programme content for many scientific meetings,
+critically assessing the intellectual content of research proposals
+and awards applications, and reviewing the research and strategic
+direction of various world-class research institutions.
 
-I have been an active member of the academic community by contributing the development of programme content for many scientific meetings, critically assessing the intellectual content of research proposals and awards applications, and reviewing the research and strategic direction of various world-class research institutions.
+I am a member of the Society for Neuroscience and IEEE, and in 2021
+became a Fellow of the European Laboratory for Learning and
+Intelligent Systems (ELLIS).
 
-I co-founded the Neural Coding, Computation and Dynamics meeting in 2007, and was in the organising committee in 2015 and 2017.
+I served on the Board of Directors of the Computational Neuroscience
+Organization between 2003-2006.
+%
+I was a member of the programme committees for 
+%
+the Neural Information Processing Systems (NeurIPS) meeting
+(2004,2006); and the Computational and Systems Neuroscience (COSYNE)
+conference (2007).
+%
+I was the Workshops Chair for NeurIPS in 2008 and for the Computational
+Neuroscience Meeting from 1999-2003.
+%
+In 2009 I was the Programme Chair for COSYNE, taking over as General
+Chair in  2010.
+%
+I co-founded the Neural Coding, Computation and Dynamics meeting in
+2007, and was in the organising committee in 2015 and 2017.
 
-I participated in the selection committee for the Swartz Prize for Theoretical and Computational Neuroscience in 2014-2016, and was in the programme committee for the Computational and Systems Neuroscience (COSYNE) conference in 2007, becoming a  Programme Chair in 2009 and General Chair in 2010.
+I have held editorial roles at various journals including, Current
+Opinion in Neurobiology (Special issue on Big Data and Neuroscience),
+Neural Computation, Neural Systems and Circuits, Network: Computation
+in Neural Systems, and Faculty of 1000, "Machine learning: life
+sciences" collection.
 
-I was also in the Programme Committee for NIPS (now NeurIPS) in 2004 and 2006, becoming a Workshops Chair in 2008.
-I was also a Workshops Chair for the Computational Neuroscience Meeting from 1999 - 2003.
+I have taken various advisory roles for international review and award
+panels. I reviewed the Donders Institute for Brain, Cognition and
+Behaviour (Netherlands), The Edmond and Lily Safra Center for Brain
+Sciences, Hebrew University (Israel), and the Barcelona Summer School
+on Advanced Modelling of Behaviour (Spain). I have advised on awards
+such as the Swartz Prize for Theoretical and Computational
+Neuroscience (SFN, US), the Bernstein Prize (BMBF, Germany), and
+COSYNE.
 
-I was a member of the Board of Directors of the Computational Neuroscience Organization between 2003 - 2006.
-I am currently a member of the Society for Neuroscience and IEEE, and in 2021 became a Fellow of the European Laboratory for Learning and Intelligent Systems (ELLIS).
+Recently, I have acted as an advisor to a new established
+cross-disciplinary effort at the Indian Institute of Science in
+Bangalore; visiting the site a number of times and hosting a joint
+workshop here at UCL.
 
-I currently have editorial roles in various journals including, Current Opinion in Neurobiology (Special issue on Big Data and Neuroscience), Neural Computation, Neural Systems and Circuits, Network:
-Computation in Neural Systems, and Faculty of
-1000, "Machine learning: life sciences" collection.
-
-I have taken various advisory roles for international review and award panels.
-I reviewed the Donders Institute for Brain, Cognition and Behaviour (Netherlands), The Edmond and Lily Safra Center for Brain Sciences, Hebrew
-University (Israel), and the Barcelona Summer School on Advanced Modelling of Behaviour (Spain).
-I have advised on awards such as the Swartz Prize for Theoretical and Computational Neuroscience (SFN, US), the Bernstein Prize (BMBF, Germany), and COSYNE.
-
-Recently, I have acted as an advisor to a new established cross-disciplinary effort at the Indian Institute of Science in Bangalore; visiting the site a
-number of times and hosting a joint workshop here at UCL.
-
-I have sat on various grant award panels for the BMBF and DFG (Germany), and the (US) NSF/NIH collaborative research
-in computational neuroscience (CRCNS) programme.
-I have also acted as a reviewer for proposals for the Bernstein prize, the Einstein Foundation, EPSRC, MRC, Novo Nordisk Foundation (Denmark) and NWO (Netherlands).
+I have sat on various grant award panels for the BMBF and DFG
+(Germany), and the (US) NSF/NIH collaborative research in
+computational neuroscience (CRCNS) programme. I have also acted as a
+reviewer for proposals for the Bernstein prize, the Einstein
+Foundation, EPSRC, MRC, Novo Nordisk Foundation (Denmark) and NWO
+(Netherlands).
 
@@
 % 
+}

### 2021-11-26 application/justificationOfResources.tex
Added

\documentclass[a4paper,11pt]{article}
\usepackage[margin=2cm]{geometry}

\title{Justification of Resources}
\author{}
\date{}

\begin{document}

\maketitle

\section{Directly Incurred - Staff}

\subsection*{Research software engineers (RSE)}

To implement the proposed project we will recruit two experienced RSEs, appointed
at grade 8. They will be responsible for developing and implementing all the
work packages throughout the project. We anticipate one of the RSEs to be Dr.
Joaquin Rapela, who will continue to work on integrating machine-learning
behavioural and neural analysis tools with Bonsai, in collaboration with
NeuroGEARS. He has the necessary machine learning and programming expertise to
deliver the proposed work. An additional RSE with a broad skillset (programming
in C\#, Python, R, Matlab, Windows) is required to deliver foreign programming language interface packages. They will work closely with the software engineering team at NeuroGEARS. Total cost to funder: £196,864 for each of the two RSEs.

\section{Directly Incurred - Travel \& Subsistence}

\subsection*{Conferences}

The success of the proposed project will depend on continued training and community engagement of employed RSEs, to stay abreast of recent developments in machine learning and neural and behavioural analysis. Conferences are key resources for such knowledge acquisition. We request funding to support the attendance of at least one
conference per year for each RSE. While many conferences now offer online formats, others are returning to in-person attendance, which is crucial for networking and broadening Bonsai user engagement. Possible conferences that could benefit the RSEs include SciPy, DotNet Conference, PyCon, Matlabexpo, UseR!, and SeptembRSE Conferences for Research Software Engineers.

Costs include travel and accommodation for 3 years for each RSE. Total cost to funder: £11,520 (£5,760 per RSE).

\subsection*{Networking Event}

We request funds for a user engagement event to mark the 10th anniversary of Bonsai. This small 3-day conference will have approximately 80-100 attendees and about 6 invited international scientists or developers who use or develop the Bonsai ecosystem. The conference will have 1 day of talks, followed by a separate 2-day hackathon (see below). The intention of this conference is to promote awareness of our software and bring together the community that uses our software to discuss possible future directions, in addition to bringing in new users. The conference would be a hybrid event with an online component for Bonsai users across the globe to also attend virtually. We request funds for travel and accommodation of invited speakers, in addition to catering for lunch and coffee breaks. We also request funds for an online platform for the conference. Travel and accommodation cost for speakers: £10,560. Catering and online platform costs: £1,920.

\subsection{Workshops}

We request funds for two Bonsai workshops to take place at the Sainsbury Wellcome Centre. The aim of the workshops is to increase dissemination and user engagement.

Bonsai for Software Developers Workshop. This will be a training course and workshop focusing explicitly on the extensibility features of the Bonsai programming languages, targeting software developers interested in making their own Bonsai packages and custom language integration. This workshop would take place at the end of year 2, towards the end of the development of the software infrastructure, as a way of presenting the development proof-of-concept to the community and possibly bringing new developers from the community to contribute to the resource from the outset.

Machine Learning in Bonsai Hackathon. This would be a collaborative workshop potentially combining scientists and engineers interested in developing applications of machine learning in Bonsai, either to address scientific questions, or to integrate new methods and algorithms into the resource (these could be developed as two alternate/complementary tracks running in parallel for participants). We expect to integrate this workshop with the user engagement event described above.

Each workshop will run over the course of two days. Registration for each workshop will be free of charge. We request funds for catering. Total cost for the two workshops: £1920.

\section{Other Directly Incurred}

\subsection*{Equipment}

We request funds for two high-end laptop computers to facilitate possible remote working for each RSE. Total cost to funder: £4,800 (£2400 per RSE).

\subsection*{Training}

To extend the knowledge base of RSEs working on this project, they may benefit from formal training courses in software design, programming, machine learning and/or computational neuroscience. Total cost of sign-up fees: £1,920.

\subsection*{Conference fees}

We request funds to cover the registration fees for the two RSEs to attend at least 3 conferences for the duration of the project.  
Total cost to funder: £2,112 (£1,056 per RSE).

\end{document}

### 2021-11-28 application/dataManagementPlan_long.tex
Added

\documentclass[a4paper,11pt]{article}
\usepackage[margin=2cm]{geometry}
%% Language and font encodings
\usepackage[english]{babel}
\usepackage[utf8x]{inputenc}
\usepackage[T1]{fontenc}
% \usepackage{microtype}      % fine pdf font control

\usepackage[scaled=0.92]{helvet} \renewcommand{\familydefault}{\sfdefault}
\usepackage{xcolor} % -- loaded by tikz

% \usepackage[colorlinks=true]{hyperref}
\usepackage[colorlinks = true,
            linkcolor = blue!50!black!50,
            urlcolor  = blue!50!black!50,
            citecolor = blue!50!black!50,
            anchorcolor = blue!50!black!50]{hyperref}
\usepackage[super]{natbib}
\renewcommand{\citet}[1]{[\citenum{#1}]}
% \usepackage{apalike}
\usepackage{graphicx}
\usepackage{verbatim}

\usepackage{tikz}
\usetikzlibrary{positioning}
\usetikzlibrary{fit}
\usetikzlibrary{calc}

\usepackage{float}
\usepackage{wrapfig}

\usepackage[font={footnotesize}]{caption}

\makeatletter
\renewcommand\paragraph{\@startsection{paragraph}{4}{\z@}%
            {-2.5ex\@plus -1ex \@minus -.25ex}%
            {1.25ex \@plus .25ex}%
            {\normalfont\normalsize\bfseries}}
\makeatother
\setcounter{secnumdepth}{4} % how many sectioning levels to assign numbers to
\setcounter{tocdepth}{4}    % how many sectioning levels to show in ToC
\usepackage[tiny]{titlesec} % format section titles
\def\headfmt{\color{blue!50!black!50}\bfseries}
\titleformat*{\section}{\headfmt}
\titlespacing*{\section}{0pt}{*1}{*0}[-4em]
\titleformat*{\subsection}{\headfmt}
\titlespacing*{\subsection}{0pt}{*0.5}{*0}
\titleformat*{\subsubsection}{\headfmt}
\titlespacing*{\subsubsection}{0pt}{*0}{*0}
\titleformat*{\paragraph}{\headfmt}
\titlespacing*{\paragraph}{0pt}{*0.5}{*0}

\usepackage{enumitem}
\setlist[description]{
    topsep=0pt,
    itemsep=0pt,
    partopsep=0pt,
    parsep=0pt,
    font=\headfmt,
    leftmargin=1em}

\renewcommand{\footnote}[1]{ [#1]}

\title{Data Management Plan}
\author{}
\date{}

\begin{document}

\maketitle

\setcounter{section}{-1}

\section{Proposal name}

\textcolor{orange}{\textit{Exactly as in the proposal that the DMP accompanies}}

Machine intelligence for (neuroscience) experimental control

\section{Description of the data}
\subsection{Type of study}

\textcolor{orange}{\textit{Up to three lines of text that summarise the type of study (or studies) for which the data are being collected}}

Enhancement of an existing software resource, the Bonsai software
ecosystem for data acquisition and experimental control.

\subsection{Types of data}

\textcolor{orange}{\textit{Types of research data to be managed: quantitative or
qualitative, sequencing data, images, models, software, scripts, protocols, and
procedures}}

We propose to incorporate into the Bonsai software ecosystem state-of-the-art
machine-learning models, as well as capabilities to enable the communication
between Bonsai, written in C\#, and machine-learning applications written in
Python, R and Matlab.

\subsection{Origin of the data}

\textcolor{orange}{\textit{Are you creating new primary data (i.e. collecting or
generating data)? Or are you re-using already available sources, including to
create new data (e.g. a new dataset created from transformation or integration
of existing data)? If data is being re-used, is it publicly accessible?}}

We will not create primary data for the proposed project. However, as outlined
in \emph{Section 2.5 Testing with neuroscience data} of the \emph{Case for
Support} document, the large community of Bonsai users at the Sainsbuy Wellcome
Centre (SWC) will test the new functionality added to Bonsai prior to its
distribution. For these tests, this community will use their own datasets, its
including high-density Neuropixels physiological recordings, with optogentic
manipulations and high-resolution video monitoring of animal behaviour. These
datasets will not be made publicly available for the purposes of the proposed
project. However, we are building infrastructure to make them available in the
near future.

The proposed enhancements to the Bonsai ecosystem will impact a wide range of
existing and to-be-created datasets collected by open-source initiatives
already integrated into the Bonsai ecosystem. For example, datasets of feeding
behaviour in flies (collected with FlyPAD
sensors\footnote{\url{https://flypad.rocks/}}), electrophysiology datasets
(collected with products supported by
OpenEphys\footnote{\url{https://open-ephys.org/}}), datasets of vivo imaging of
brain activity (collected with the UCLA
Miniscope\footnote{\url{http://miniscope.org/}} or with NeuroPhotometrics
products\footnote{\url{https://neurophotometrics.com/}}), datasets using rich
2D or 3D visual environments (created with
BonVision\footnote{\url{https://bonvision.github.io/}}), high-density
electrophysiology datasets (recorded with NeuroPixels
probes\footnote{\url{https://www.neuropixels.org/}}), and animal behavioral
datasets (with body parts tracked with
DeepLabCut\footnote{\url{http://www.mackenziemathislab.org/deeplabcut}}).

\subsection{Format and scale of the data}

\textcolor{orange}{\textit{File formats, software used, number of records, databases, sweeps, repetitions... (in terms that are meaningful in your field of research). An indication of the size of data to be stored and shared.}}

The source code for the core Bonsai ecosystem and the majority of extension
packages are hosted at \url{github.com/bonsai-rx}.

\section{Data management, documentation, and curation}

\textcolor{orange}{\textit{Keep this section concise and accessible to readers who are not data management experts.  Focus on principles, systems, and major standards. Focus on the main kind(s) of study data.  Give brief examples and avoid long lists.}}

% documentation
The documentation, including learning materials and all relevant community
links are publicly accessible at \url{bonsai-rx.org}. A forum where Bonsai
users daily exchange questions and answers can be accessed at
\url{https://groups.google.com/g/bonsai-users}. Currently more than 700 users
contribute to this forum (Figure~1f, \emph{Case for Support}).

% management
Package installation and updates are supported by standalone package management
software included within the core Bonsai distribution and hosted in a shared
centralised and curated repository. This ensures that bugfixes and enhancements
are easily available and, importantly, allows for reproducible deployment of
user environments to ensure working experiments are not disrupted by system
changes.
%
Large-scale experimental projects such as the International Brain Laboratory
have leveraged this capability of Bonsai to standardise data acquisition and
control software across dozens of laboratories around the world.
%
Community technology previews are made available through the package manager,
letting early adopters opt-in to the latest features before they are generally
available. This allows the community to report issues with new features on
GitHub, thereby distributing bug reporting and triage, and improving the
quality of stable releases.

The core compiler and IDE is extensively tested, and uses a strong type system
to greatly reduce bugs at compile-time. Best practices for leveraging this
infrastructure for building new packages are in place. The software engineers
recruited to this project will work with our collaborators at NeuroGEARS Ltd,
to strengthen and maintain these practices. Once fully familiar with the Bonsai
core architecture they will participate in review of pull requests and other
code contributions.

\subsection{Managing, storing and curating data}

\textcolor{orange}{\textit{Briefly describe how data will be stored, backed-up, managed and curated in the short to medium term. Specify any community agreed or other formal data standards used (with URL references).}}

Bonsai is a freely available open-source resource, already widely adopted by
the systems neuroscience and behavioural science communities (see Fig 1f).
Current contributors include other large open projects (OpenEphys, DeepLabCut),
and companies developing and supplying new hardware devices
(Neurophotometrics). This community uptake, engagement, and commitment provides
a foundation for long-term sustainability.

NeuroGEARS Ltd.\ is a keystone of this community and a key collaborator in the
current proposal. Founded by the original developer of Bonsai, it helps to
support and expand the user base well beyond the life of the current proposal.
NeuroGEARS currently provides quality control and curation for the core Bonsai
code, and provides training and support services for research institutions
around the world who want to start using Bonsai. NeuroGEARS also facilitates
knowledge transfer between users and institutions adopting Bonsai, by
moderating the community user forums and collaborating with organisations who
want to build new innovative research projects on top of Bonsai, as in the
current proposal.

The specific contributions developed as part of this project will form new core
components and packages within the Bonsai ecosystem. In particular, the new
language integration packages will support long-term sustainability, by
dramatically increasing Bonsai's interoperability with existing software
development communities which are already well established in basic science
(MATLAB, Python, R). Developing these new packages with input from the core
advisory panel of Bonsai, and early public distribution and user testing within
the SWC and with planned workshops, will ensure solid and broad user engagement
from the beginning of the development cycle and will therefore build momentum
for their use, support and dissemination. Close collaboration with the core
maintainers of the Bonsai ecosystem at NeuroGEARS will allow them to take over
responsibility for maintenance and support of these packages beyond the end of
the project, following the same principles of open-source and free sharing of
knowledge that characterises the entire Bonsai ecosystem.

\subsection{Metadata standards and data documentation}

\textcolor{orange}{\textit{What metadata is produced about the data generated from the research? For example descriptions of data that enable research data to be used by others outside of your own team.  This may include documenting the methods used to generate the data, analytical and procedural information, capturing instrument metadata alongside data, documenting provenance of data and their coding, detailed descriptions for variables, records, etc.}}

The fact that Bonsai is distributed as an open source project
(\url{https://bonsai-rx.org/}), that is has extensive online documentation
(\url{https://bonsai-rx.org/docs/}) and a vibrant user discussion forum
(\url{https://groups.google.com/g/bonsai-users}), and that is integrated with
many other open-source projects (Figure~1A and~1B, \emph{Case for Support})
enables the use of the Bonsai ecosystem by the general public.

\end{document}

### 2021-11-28 application/dataManagementPlan_short.tex
Added

--- 2021-11-28 application/dataManagementPlan_long.tex
@@
 
-Enhancement of an existing software resource, the Bonsai software
-ecosystem for data acquisition and experimental control.
+No primary data are being collected for the proposed software resource.
 
@@
 
-We propose to incorporate into the Bonsai software ecosystem state-of-the-art
-machine-learning models, as well as capabilities to enable the communication
-between Bonsai, written in C\#, and machine-learning applications written in
-Python, R and Matlab.
+We propose to manage software and models. We offer to incorporate into the
+Bonsai software ecosystem state-of-the-art machine-learning models, as well as
+capabilities to enable the communication between Bonsai, written in C\#, and
+machine-learning applications written in Python, R and Matlab.
 
@@
 (collected with products supported by
-OpenEphys\footnote{\url{https://open-ephys.org/}}), datasets of vivo imaging of
-brain activity (collected with the UCLA
+OpenEphys\footnote{\url{https://open-ephys.org/}} or with NeuroPixels
+probes\footnote{\url{https://www.neuropixels.org/}})), datasets of in-vivo
+imaging of brain activity (collected with the UCLA
 Miniscope\footnote{\url{http://miniscope.org/}} or with NeuroPhotometrics
@@
 2D or 3D visual environments (created with
-BonVision\footnote{\url{https://bonvision.github.io/}}), high-density
-electrophysiology datasets (recorded with NeuroPixels
-probes\footnote{\url{https://www.neuropixels.org/}}), and animal behavioral
+BonVision\footnote{\url{https://bonvision.github.io/}}), and animal behavioral
 datasets (with body parts tracked with
@@
 
-The source code for the core Bonsai ecosystem and the majority of extension
-packages are hosted at \url{github.com/bonsai-rx}.
+We believe this section does not applies to our proposed software resource that
+will not generate primary data.
 
@@
 
-% documentation
-The documentation, including learning materials and all relevant community
-links are publicly accessible at \url{bonsai-rx.org}. A forum where Bonsai
-users daily exchange questions and answers can be accessed at
-\url{https://groups.google.com/g/bonsai-users}. Currently more than 700 users
-contribute to this forum (Figure~1f, \emph{Case for Support}).
-
-% management
-Package installation and updates are supported by standalone package management
-software included within the core Bonsai distribution and hosted in a shared
-centralised and curated repository. This ensures that bugfixes and enhancements
-are easily available and, importantly, allows for reproducible deployment of
-user environments to ensure working experiments are not disrupted by system
-changes.
-%
-Large-scale experimental projects such as the International Brain Laboratory
-have leveraged this capability of Bonsai to standardise data acquisition and
-control software across dozens of laboratories around the world.
-%
-Community technology previews are made available through the package manager,
-letting early adopters opt-in to the latest features before they are generally
-available. This allows the community to report issues with new features on
-GitHub, thereby distributing bug reporting and triage, and improving the
-quality of stable releases.
-
-The core compiler and IDE is extensively tested, and uses a strong type system
-to greatly reduce bugs at compile-time. Best practices for leveraging this
-infrastructure for building new packages are in place. The software engineers
-recruited to this project will work with our collaborators at NeuroGEARS Ltd,
-to strengthen and maintain these practices. Once fully familiar with the Bonsai
-core architecture they will participate in review of pull requests and other
-code contributions.
+Please refer to \emph{Section 5 Resource management} of the \emph{Case for
+support} document for details about how we plan to manage the proposed resource
+development.
 
@@
 
-Bonsai is a freely available open-source resource, already widely adopted by
-the systems neuroscience and behavioural science communities (see Fig 1f).
-Current contributors include other large open projects (OpenEphys, DeepLabCut),
-and companies developing and supplying new hardware devices
-(Neurophotometrics). This community uptake, engagement, and commitment provides
-a foundation for long-term sustainability.
-
-NeuroGEARS Ltd.\ is a keystone of this community and a key collaborator in the
-current proposal. Founded by the original developer of Bonsai, it helps to
-support and expand the user base well beyond the life of the current proposal.
-NeuroGEARS currently provides quality control and curation for the core Bonsai
-code, and provides training and support services for research institutions
-around the world who want to start using Bonsai. NeuroGEARS also facilitates
-knowledge transfer between users and institutions adopting Bonsai, by
-moderating the community user forums and collaborating with organisations who
-want to build new innovative research projects on top of Bonsai, as in the
-current proposal.
-
-The specific contributions developed as part of this project will form new core
-components and packages within the Bonsai ecosystem. In particular, the new
-language integration packages will support long-term sustainability, by
-dramatically increasing Bonsai's interoperability with existing software
-development communities which are already well established in basic science
-(MATLAB, Python, R). Developing these new packages with input from the core
-advisory panel of Bonsai, and early public distribution and user testing within
-the SWC and with planned workshops, will ensure solid and broad user engagement
-from the beginning of the development cycle and will therefore build momentum
-for their use, support and dissemination. Close collaboration with the core
-maintainers of the Bonsai ecosystem at NeuroGEARS will allow them to take over
-responsibility for maintenance and support of these packages beyond the end of
-the project, following the same principles of open-source and free sharing of
-knowledge that characterises the entire Bonsai ecosystem.
+Please refer to \emph{Section 6 Long-term sustainability planning} of the
+\emph{Case for support} document for details about how we plan the long-term
+sustainability of the proposed resource development.
 
@@
 
-The fact that Bonsai is distributed as an open source project
-(\url{https://bonsai-rx.org/}), that is has extensive online documentation
-(\url{https://bonsai-rx.org/docs/}) and a vibrant user discussion forum
-(\url{https://groups.google.com/g/bonsai-users}), and that is integrated with
-many other open-source projects (Figure~1A and~1B, \emph{Case for Support})
-enables the use of the Bonsai ecosystem by the general public.
+We believe this section does not applies to our proposed software resource that
+will not generate primary data.
 

### 2021-11-29 application/dataManagementPlan_long.tex
Added reference to BonZeb and header of instructions

@@
 
+\textcolor{orange}{\textit{The following template is provided to applicants to assist in the development of a Data Management Plan (DMP) to accompany a research proposal.
+The notes (in italics) provide further context and guidance for its completion and should be deleted prior to submission.
+Applicants should consider submission of a DMP which is proportional to the quantity and nature of the data.
+Where substantial data is generated from the research, it is expected that the DMP will be longer and more detailed than studies generating small amounts of data, where DMPs may be significantly shorter.
+In either case, the DMP should be a maximum of 3 pages in length.}}
+
 \setcounter{section}{-1}
@@
 behaviour in flies (collected with FlyPAD
-sensors\footnote{\url{https://flypad.rocks/}}), electrophysiology datasets
-(collected with products supported by
-OpenEphys\footnote{\url{https://open-ephys.org/}}), datasets of vivo imaging of
-brain activity (collected with the UCLA
+sensors\footnote{\url{https://flypad.rocks/}}), high-resolution zebrafish
+tracking datasets (collected with
+BonZeb\footnote{\url{https://github.com/ncguilbeault/BonZeb/}}),
+electrophysiology datasets (collected with products supported by
+OpenEphys\footnote{\url{https://open-ephys.org/}} or with NeuroPixels
+probes\footnote{\url{https://www.neuropixels.org/}})), datasets of in-vivo
+imaging of brain activity (collected with the UCLA
 Miniscope\footnote{\url{http://miniscope.org/}} or with NeuroPhotometrics

### 2021-11-29 application/dataManagementPlan_short.tex
Added reference to BonZeb and header of instructions

@@
             anchorcolor = blue!50!black!50]{hyperref}
-\usepackage[super]{natbib}
-\renewcommand{\citet}[1]{[\citenum{#1}]}
-% \usepackage{apalike}
+% \usepackage[super]{natbib}
+% \renewcommand{\citet}[1]{[\citenum{#1}]}
+\usepackage{natbib}
+\usepackage{apalike}
 \usepackage{graphicx}
@@
 behaviour in flies (collected with FlyPAD
-sensors\footnote{\url{https://flypad.rocks/}}), electrophysiology datasets
-(collected with products supported by
+sensors\footnote{\url{https://flypad.rocks/}}), high-resolution zebrafish
+tracking datasets (collected with
+BonZeb\footnote{\url{https://github.com/ncguilbeault/BonZeb/}}),
+electrophysiology datasets (collected with products supported by
 OpenEphys\footnote{\url{https://open-ephys.org/}} or with NeuroPixels
@@
 
+% \bibliographystyle{unsrtveryabbrv}
+\bibliographystyle{apalike}
+\bibliography{bonsai}
+
 \end{document}

### 2021-11-29 application/dataManagementPlan_long.tex
Updates from Overleaf

--- 2021-11-28 application/dataManagementPlan_long.tex
@@
 
+% \textcolor{orange}{\textit{The following template is provided to applicants to assist in the development of a Data Management Plan (DMP) to accompany a research proposal.
+The notes (in italics) provide further context and guidance for its completion and should be deleted prior to submission.
+Applicants should consider submission of a DMP which is proportional to the quantity and nature of the data.
+Where substantial data is generated from the research, it is expected that the DMP will be longer and more detailed than studies generating small amounts of data, where DMPs may be significantly shorter.
+In either case, the DMP should be a maximum of 3 pages in length.}}
+
 \setcounter{section}{-1}
@@
 
-\textcolor{orange}{\textit{Exactly as in the proposal that the DMP accompanies}}
+% \textcolor{orange}{\textit{Exactly as in the proposal that the DMP accompanies}}
 
@@
 
-\textcolor{orange}{\textit{Up to three lines of text that summarise the type of study (or studies) for which the data are being collected}}
+% \textcolor{orange}{\textit{Up to three lines of text that summarise the type of study (or studies) for which the data are being collected}}
 
-Enhancement of an existing software resource, the Bonsai software
-ecosystem for data acquisition and experimental control.
+The project will generate software modules appropriate for use in behavioural neuroscience experiments in animal models, and other types of study that involve a combination of reactive experimental (particularly behavioural) control and streamed data acquisition.
+
+% Enhancement of an existing software resource, the Bonsai software
+% ecosystem for data acquisition and experimental control.
 
@@
 
-\textcolor{orange}{\textit{Types of research data to be managed: quantitative or
-qualitative, sequencing data, images, models, software, scripts, protocols, and
-procedures}}
+% \textcolor{orange}{\textit{Types of research data to be managed: quantitative or
+% qualitative, sequencing data, images, models, software, scripts, protocols, and
+% procedures}}
 
-We propose to incorporate into the Bonsai software ecosystem state-of-the-art
-machine-learning models, as well as capabilities to enable the communication
-between Bonsai, written in C\#, and machine-learning applications written in
-Python, R and Matlab.
+This is a software project.
+We will produce and manage new modules to be integrated into the Bonsai software ecosystem.
+These will include communication modules to be written in C\#, and machine-learning applications written in Python, R and Matlab.
+
+% We propose to incorporate into the Bonsai software ecosystem state-of-the-art
+% machine-learning models, as well as capabilities to enable the communication
+% between Bonsai, written in C\#, and machine-learning applications written in
+% Python, R and Matlab.
 
@@
 
-\textcolor{orange}{\textit{Are you creating new primary data (i.e. collecting or
-generating data)? Or are you re-using already available sources, including to
-create new data (e.g. a new dataset created from transformation or integration
-of existing data)? If data is being re-used, is it publicly accessible?}}
+% \textcolor{orange}{\textit{Are you creating new primary data (i.e. collecting or
+% generating data)? Or are you re-using already available sources, including to
+% create new data (e.g. a new dataset created from transformation or integration
+% of existing data)? If data is being re-used, is it publicly accessible?}}
 
-We will not create primary data for the proposed project. However, as outlined
-in \emph{Section 2.5 Testing with neuroscience data} of the \emph{Case for
-Support} document, the large community of Bonsai users at the Sainsbuy Wellcome
-Centre (SWC) will test the new functionality added to Bonsai prior to its
-distribution. For these tests, this community will use their own datasets, its
-including high-density Neuropixels physiological recordings, with optogentic
-manipulations and high-resolution video monitoring of animal behaviour. These
-datasets will not be made publicly available for the purposes of the proposed
-project. However, we are building infrastructure to make them available in the
-near future.
+This is a software project.
+New modules will be created and tested by software research engineers at UCL in consultation with staff at NeuroGEARS Ltd (a project collaborator).
+These will be integrated into the existing Bonsai software ecosystem.
 
-The proposed enhancements to the Bonsai ecosystem will impact a wide range of
-existing and to-be-created datasets collected by open-source initiatives
-already integrated into the Bonsai ecosystem. For example, datasets of feeding
-behaviour in flies (collected with FlyPAD
-sensors\footnote{\url{https://flypad.rocks/}}), electrophysiology datasets
-(collected with products supported by
-OpenEphys\footnote{\url{https://open-ephys.org/}}), datasets of vivo imaging of
-brain activity (collected with the UCLA
-Miniscope\footnote{\url{http://miniscope.org/}} or with NeuroPhotometrics
-products\footnote{\url{https://neurophotometrics.com/}}), datasets using rich
-2D or 3D visual environments (created with
-BonVision\footnote{\url{https://bonvision.github.io/}}), high-density
-electrophysiology datasets (recorded with NeuroPixels
-probes\footnote{\url{https://www.neuropixels.org/}}), and animal behavioral
-datasets (with body parts tracked with
-DeepLabCut\footnote{\url{http://www.mackenziemathislab.org/deeplabcut}}).
+We will access experimental data during the testing phase, but will not seek to manage these data within the project.
+Such data will have been collected in the course of other projects and will be managed in accordance with the original protocols.
+
+% We will not create primary data for the proposed project. However, as outlined
+% in \emph{Section 2.5 Testing with neuroscience data} of the \emph{Case for
+% Support} document, the large community of Bonsai users at the Sainsbuy Wellcome
+% Centre (SWC) will test the new functionality added to Bonsai prior to its
+% distribution. For these tests, this community will use their own datasets, its
+% including high-density Neuropixels physiological recordings, with optogentic
+% manipulations and high-resolution video monitoring of animal behaviour. These
+% datasets will not be made publicly available for the purposes of the proposed
+% project. However, we are building infrastructure to make them available in the
+% near future.
+
+% The proposed enhancements to the Bonsai ecosystem will impact a wide range of
+% existing and to-be-created datasets collected by open-source initiatives
+% already integrated into the Bonsai ecosystem. For example, datasets of feeding
+% behaviour in flies (collected with FlyPAD
+% sensors\footnote{\url{https://flypad.rocks/}}), high-resolution zebrafish
+% tracking datasets (collected with
+% BonZeb\footnote{\url{https://github.com/ncguilbeault/BonZeb/}}),
+% electrophysiology datasets (collected with products supported by
+% OpenEphys\footnote{\url{https://open-ephys.org/}} or with NeuroPixels
+% probes\footnote{\url{https://www.neuropixels.org/}})), datasets of in-vivo
+% imaging of brain activity (collected with the UCLA
+% Miniscope\footnote{\url{http://miniscope.org/}} or with NeuroPhotometrics
+% products\footnote{\url{https://neurophotometrics.com/}}), datasets using rich
+% 2D or 3D visual environments (created with
+% BonVision\footnote{\url{https://bonvision.github.io/}}), and animal behavioral
+% datasets (with body parts tracked with
+% DeepLabCut\footnote{\url{http://www.mackenziemathislab.org/deeplabcut}}).
 
@@
 
-\textcolor{orange}{\textit{File formats, software used, number of records, databases, sweeps, repetitions...
+% \textcolor{orange}{\textit{File formats, software used, number of records, databases, sweeps, repetitions...
 (in terms that are meaningful in your field of research).
@@
 
-The source code for the core Bonsai ecosystem and the majority of extension
-packages are hosted at \url{github.com/bonsai-rx}.
+Software modules will comprise source files in a number of programming languages: C\#, Python, MATLAB and R.  These will be accompanied by extensive documentation (discussed below).
+
+NEED: Approximate lines of code??? (order of magnitude is fine)
+
+NEED: testing files, makefiles?, other stuff,
+
+% The source code for the core Bonsai ecosystem and the majority of extension
+% packages are hosted at \url{github.com/bonsai-rx}.
 
@@
 Give brief examples and avoid long lists.}}
+
+\subsection{Managing, storing and curating data}
 
@@
 code contributions.
-
-\subsection{Managing, storing and curating data}
 
@@
 
+\subsection{Data preservation strategy and standards}
+
+Plans and place for long-term storage, preservation, and planned retention period for the research data.
+Formal preservation standards, if any.
+Indicate which data may not be retained (if any).
+Also, briefly describe what will be stored long-term, raw data, processed data?
+
+\section{Data sharing and access}
+\subsection{Where will data be shared?}
+% Identify any data repository (-ies) that are, or will be, entrusted with storing, curating and/or sharing data from your study, where they exist for particular disciplinary domains or data types.
+Software modules and documentation will be available on github, with links from the Bonsai project site (https://bonsai-rx.org) and available through the Bonsai system's package management interface.
+
+NEED: github org and status?
+
+\subsection{When will data be available?}
+% It is expected that timely release would generally be no later than the release through publication of the main findings and should be in line with established best practice in the field.
+Where best practices do not exist release within three years of generation of the dataset is suggested as a guide.
+
+Release management: repositories will be public throughout, but marked as developmental.
+Alpha-level code will be available but not publicised until after extensive testing.
+Will be integrated into package management and advertised once it achieves release quality.
+
+\subsection{How will data be made findable and accessible?}
+
+Package management
+Bonsai website
+Community fora
+Workshops
+Paper?
+
+% Indicate how potential new users (outside of your organisation) can find out about your data and identify whether it could be suitable for their research purposes, also consider how it will be licenced.
+The approach should encompass data in all its forms as well as software.
+This section could also outline access procedures, how your data will be able to be cited and tracked.
+
+\subsection{How will data be made reusable?}
+
+This is an open source software project.  Code will be distributed under ??? license.
+
+% What steps will you take to ensure that your data could be re-analysed by other researchers? (e.g. availability of meta data, worked examples, vignettes). How will metadata be made available?
+
+\subsection{Restrictions or delays to sharing, with planned actions to limit such restrictions} 
+Restriction to data sharing may be due to subject confidentiality or to ensure you can gain IP protection.
+Applicants should ensure they have obtained necessary clearances from relevant collaborators with regards to the content of the proposal including the data sharing plan in line with the BBSRC Research Grants Guide.
+Where possible please indicate when data will be made available following a period of restriction.
+
+All distributed software will be developed as part of the project, based on published and unrestricted algorithms.  Thus it is unencumbered by IP restrictions.  Will be available as above.
+
+4. Data security (where relevant)
+4.1	Formal information/data security standards
+% Identify formal information standards with which your study is or will be compliant. An example is ISO 27001. If your organisation is ISO compliant, please state the registration number.
+
+Not applicable.
+
+4.2	Main risks to data security
+% All personal data has an element of risk.
+Summarise the main risks to the confidentiality and security of information related to human participants or users, the level of risk and how these risks will be managed.
+Cover the main processes or facilities for storage and processing of personal data, data access, with controls put in place and any auditing of user compliance with consent and security conditions.
+It is not sufficient to write ‘not applicable’ under this heading.
+
+This is a freely-available open source software package, which does not contain participants or personal data.
+User access to the repository is anonymous and user personal information will not be collected.
+
 \end{document}

### 2021-11-29 application/dataManagementPlan_long_joaco.tex
Several modifications wrt dataManagementPlan.tex

--- 2021-11-29 application/dataManagementPlan_long.tex
@@
 
-This is a software project.
-We will produce and manage new modules to be integrated into the Bonsai software ecosystem.
-These will include communication modules to be written in C\#, and machine-learning applications written in Python, R and Matlab.
+This is a software project.  We will produce and manage new modules to be
+integrated into the Bonsai software ecosystem.  These will include extensions
+to incorporate non-native code packages into Bonsai (Section 2.1, Case for
+Support), a general abstraction framework for data processing (Section 2.2,
+Case for Support), machine learning functionality added to Bonsai (Section 2.3,
+Case for Support) and documentation.
 
@@
 \subsection{Managing, storing and curating data}
+% \textcolor{orange}{\textit{Briefly describe how data will be stored, backed-up, managed and curated in the short to medium term.
+Specify any community agreed or other formal data standards used (with URL references).}}
 
-% documentation
-The documentation, including learning materials and all relevant community
-links are publicly accessible at \url{bonsai-rx.org}. A forum where Bonsai
-users daily exchange questions and answers can be accessed at
-\url{https://groups.google.com/g/bonsai-users}. Currently more than 700 users
-contribute to this forum (Figure~1f, \emph{Case for Support}).
+% code
+The source code for the core Bonsai ecosystem and the majority of extension
+packages are hosted at \url{github.com/bonsai-rx} (MIT license with no
+copyright and licencing restrictions).
 
@@
 
-\textcolor{orange}{\textit{Briefly describe how data will be stored, backed-up, managed and curated in the short to medium term.
-Specify any community agreed or other formal data standards used (with URL references).}}
+\subsubsection*{Team responsibilities}
 
-Bonsai is a freely available open-source resource, already widely adopted by
-the systems neuroscience and behavioural science communities (see Fig 1f).
-Current contributors include other large open projects (OpenEphys, DeepLabCut),
-and companies developing and supplying new hardware devices
-(Neurophotometrics). This community uptake, engagement, and commitment provides
-a foundation for long-term sustainability.
+\begin{description}
 
-NeuroGEARS Ltd.\ is a keystone of this community and a key collaborator in the
-current proposal. Founded by the original developer of Bonsai, it helps to
-support and expand the user base well beyond the life of the current proposal.
-NeuroGEARS currently provides quality control and curation for the core Bonsai
-code, and provides training and support services for research institutions
-around the world who want to start using Bonsai. NeuroGEARS also facilitates
-knowledge transfer between users and institutions adopting Bonsai, by
-moderating the community user forums and collaborating with organisations who
-want to build new innovative research projects on top of Bonsai, as in the
-current proposal.
+    \item[Goncalo Lopes] Bonsai related software development,
+        software releases, organization of Bonsai workshops
+        and co-supervision of research software engineer (RSE).
 
-The specific contributions developed as part of this project will form new core
-components and packages within the Bonsai ecosystem. In particular, the new
-language integration packages will support long-term sustainability, by
-dramatically increasing Bonsai's interoperability with existing software
-development communities which are already well established in basic science
-(MATLAB, Python, R). Developing these new packages with input from the core
-advisory panel of Bonsai, and early public distribution and user testing within
-the SWC and with planned workshops, will ensure solid and broad user engagement
-from the beginning of the development cycle and will therefore build momentum
-for their use, support and dissemination. Close collaboration with the core
-maintainers of the Bonsai ecosystem at NeuroGEARS will allow them to take over
-responsibility for maintenance and support of these packages beyond the end of
-the project, following the same principles of open-source and free sharing of
-knowledge that characterises the entire Bonsai ecosystem.
+    \item[Joaquin Rapela] machine learning (ML) related software development,
+        interaction with ML functionality methods developers, validating of ML
+        functionality with behavioral and neural recordings and co-supervision
+        of RSE.
+
+    \item[Maneesh Sahani] advise on ML.
+
+    \item[Tom Mrsic-Flogel] advise on neuroscience topics and communication
+        with the steering committee.
+
+\end{description}
+
+\subsection*{Resourcing and responsibility}
+
+\begin{description}
+
+    \item[software development model] we will use the Kanban software
+        development model, which we both Dr.~Lopes and Dr.~Rapela are already
+        using in a collaborative project at the SWC/GCNU.
+
+    \item[weekly meetings] between Goncalo Lopes, Joaquin Rapela and the RSE
+        will be an essential component of the software development process.
+        These will ensure that every member of the team is informed about both
+        the Bonsai integration and the machine learning developments and will
+        make the software development team more reliable upon the absence of
+        any of its members.
+
+    \item[code review and pair-programming] sessions will be held on a monthly basis among Goncalo Lopes, Joaquin Rapela and the recruited software engineer to deliver readable code sharing a standard programming style.
+
+\end{description}
+
+\subsubsection*{Major risks}
+
+\begin{description}
+
+    \item[language integration] we proposed to use existing software libraries
+        to enable the communication between Bonsai, written in C\#, and other
+        machine learning algorithms, written in Python (Python.NET), R (R.Net)
+        and Matlab (native Matlab .Net interface). We have not yet tested any
+        of these existing libraries, and we may found that they don't function
+        reliably enough for our purposes. If this were the case, we proposed to
+        build our own interface protocol using the ZeroMQ messaging library.
+        This risk will only delay the progress of the proposed project, but
+        will not hinder the deployment of its proposed outputs.
+
+    \item[disfunctional machine learning functionality] in Section 2.3 of the
+        Case for Support document (Figure 2, green box) we proposed to
+        integrate existing machine learning algorithms into the Bonsai
+        ecsosystem. It could be the case that some of these algorithms do not
+        work as reliably as needed for our resource. I such a case we will use
+        another algorithm with similar functionality.  This risk should not be
+        a major problem, since most of the machine learning functionality we
+        proposed to incorporate into Bonsai has been developed by us, or we are
+        experienced using it.
+
+    \item[lack of interest from Bonsai users] We will minimize the probability
+        of this problem by advertising broadly the new functionality upon
+        release.  However, we don't think this would be a real problem, since
+        Bonsai users have reapeatly asked the addition of machine learning
+        functionality to Bonsai (Section 3 Community demand, Case for Support),
+        and because the machine learning functionality we proposed to add to
+        Bonsai are popular among experimental-control neurocientists.
+
+    \item[lack of interest from machine-learning methods developers] To
+        minimize the probability that this risk turns into a large problem we
+        will take three actions. First, we will provide machine-learning
+        methods developers and easy-to-use API to interface their methods into
+        the Bonsai ecosystem. Second, we will show machine-learning methods
+        developers the large user exposure that Bonsai provides to integrated
+        ML functionality. Third, we will organize Bonsai workshops for methods
+        developers showcasing machine learning functionality integrated into
+        the Bonsai ecosystem and instructing them how to integrate new methods
+        into Bonsai.
+
+    \end{description}
+
+% Bonsai is a freely available open-source resource, already widely adopted by the systems neuroscience and behavioural science communities (see Fig 1f).
+Current contributors include other large open projects (OpenEphys, DeepLabCut), and companies developing and supplying new hardware devices (Neurophotometrics).
+This community uptake, engagement, and commitment provides a foundation for long-term sustainability.
+
+% NeuroGEARS Ltd.\ is a keystone of this community and a key collaborator in the current proposal.
+Founded by the original developer of Bonsai, it helps to support and expand the user base well beyond the life of the current proposal.
+NeuroGEARS currently provides quality control and curation for the core Bonsai code, and provides training and support services for research institutions around the world who want to start using Bonsai.
+NeuroGEARS also facilitates knowledge transfer between users and institutions adopting Bonsai, by moderating the community user forums and collaborating with organisations who want to build new innovative research projects on top of Bonsai, as in the current proposal.
+
+% The specific contributions developed as part of this project will form new core components and packages within the Bonsai ecosystem.
+In particular, the new language integration packages will support long-term sustainability, by dramatically increasing Bonsai's interoperability with existing software development communities which are already well established in basic science (MATLAB, Python, R).
+Developing these new packages with input from the core advisory panel of Bonsai, and early public distribution and user testing within the SWC and with planned workshops, will ensure solid and broad user engagement from the beginning of the development cycle and will therefore build momentum for their use, support and dissemination.
+Close collaboration with the core maintainers of the Bonsai ecosystem at NeuroGEARS will allow them to take over responsibility for maintenance and support of these packages beyond the end of the project, following the same principles of open-source and free sharing of knowledge that characterises the entire Bonsai ecosystem.
 
@@
 
-\textcolor{orange}{\textit{What metadata is produced about the data generated from the research?
+% \textcolor{orange}{\textit{What metadata is produced about the data generated from the research?
 For example descriptions of data that enable research data to be used by others outside of your own team.
@@
 
-The fact that Bonsai is distributed as an open source project
-(\url{https://bonsai-rx.org/}), that is has extensive online documentation
-(\url{https://bonsai-rx.org/docs/}) and a vibrant user discussion forum
-(\url{https://groups.google.com/g/bonsai-users}), and that is integrated with
-many other open-source projects (Figure~1A and~1B, \emph{Case for Support})
-enables the use of the Bonsai ecosystem by the general public.
+% The fact that Bonsai is distributed as an open source project
+% (\url{https://bonsai-rx.org/}), that is has extensive online documentation
+% (\url{https://bonsai-rx.org/docs/}) and a vibrant user discussion forum
+% (\url{https://groups.google.com/g/bonsai-users}), and that is integrated with
+% many other open-source projects (Figure~1A and~1B, \emph{Case for Support})
+% enables the use of the Bonsai ecosystem by the general public.
+
+% documentation
+The documentation, including learning materials and all relevant community
+links are publicly accessible at \url{bonsai-rx.org}. A forum where Bonsai
+users daily exchange questions and answers can be accessed at the public
+repository \url{https://groups.google.com/g/bonsai-users}. Currently more than
+700 users contribute to this forum (Figure~1f, \emph{Case for Support}).
+
+Classes, methods and functions will be documented in the code, and automatic
+documentation tools (e.g., Sphinx for Python code, or Sphinx for .Net for C\#)
+will be used to automatically generate API documentation. API documentation
+will be complemented with class and interaction diagrams, documenting the
+structure and dynamics of the application objects.
 
@@
 
-Plans and place for long-term storage, preservation, and planned retention period for the research data.
-Formal preservation standards, if any.
-Indicate which data may not be retained (if any).
-Also, briefly describe what will be stored long-term, raw data, processed data?
+\textcolor{orange}{\textit{Plans and place for long-term storage, preservation,
+and planned retention period for the research data. Formal preservation
+standards, if any. Indicate which data may not be retained (if any). Also,
+briefly describe what will be stored long-term, raw data, processed data?}}
+
+For persistent and long term preservation we propose to use
+Zenodo\footnote{\url{https://zenodo.org/}}.
 
@@
 % Identify any data repository (-ies) that are, or will be, entrusted with storing, curating and/or sharing data from your study, where they exist for particular disciplinary domains or data types.
-Software modules and documentation will be available on github, with links from the Bonsai project site (https://bonsai-rx.org) and available through the Bonsai system's package management interface.
+
+Software modules and documentation will be available on Github, with links from
+the Bonsai project site (https://bonsai-rx.org) and available through the
+Bonsai system's package management interface.
 
@@
 
-Release management: repositories will be public throughout, but marked as developmental.
+% Release management: repositories will be public throughout, but marked as developmental.
 Alpha-level code will be available but not publicised until after extensive testing.
 Will be integrated into package management and advertised once it achieves release quality.
+
+Method to accept a project output: software will be automatically tested with a
+comprehensive set of unit test and integration tests, implemented as  GitHub
+Actions. In addition, prior to distribution to the general public, Bonsai users
+inside the SWC will evaluate the functionality of the software with their own
+state-of-the-art datasets. Project outputs will be accepted when all tests run
+successfully and no major functional problems are reported by SWC Bonsai users.
+
+Release management: we will create a new release after each of the machine
+learning functionalities (described in Section~2.3 of the Case for Support) is
+integrated into the Bonsai ecosystem following the above method to accept a
+project output. Credentials for release will be granted initially by Dr.
+Gonçalo Lopes, Head Engineer at Neurogears and the current curator of the
+Bonsai official package feed, but might be expanded to other team members after
+training.  Acceptance criteria for release will be a group decision between
+Dr.~Lopes, Dr.~Rapela, Prof.~Sahani and Prof.~Mrsic-Flogel.
 

### 2021-11-29 application/dataManagementPlan_long.tex
Updates from Overleaf

@@
 
-This is a software project.
 New modules will be created and tested by software research engineers at UCL in consultation with staff at NeuroGEARS Ltd (a project collaborator).
@@
 We will access experimental data during the testing phase, but will not seek to manage these data within the project.
-Such data will have been collected in the course of other projects and will be managed in accordance with the original protocols.
+Such data will have been collected in the course of other projects and will continue to be managed by the relevant laboratories in accordance with the protocols they have defined.
 
@@
 
-Software modules will comprise source files in a number of programming languages: C\#, Python, MATLAB and R.  These will be accompanied by extensive documentation (discussed below).
+Software modules will comprise source files in various programming languages: C\#, Python, MATLAB and R.  These will be accompanied by extensive documentation (discussed below).
 
-NEED: Approximate lines of code??? (order of magnitude is fine)
+**NEED**: Approximate lines of code??? (order of magnitude is fine)
 
-NEED: testing files, makefiles?, other stuff,
+**NEED**: testing files, makefiles?, other stuff,
 
@@
 
-\textcolor{orange}{\textit{Keep this section concise and accessible to readers who are not data management experts.
+% \textcolor{orange}{\textit{Keep this section concise and accessible to readers who are not data management experts.
 Focus on principles, systems, and major standards.
@@
 
-% documentation
-The documentation, including learning materials and all relevant community
-links are publicly accessible at \url{bonsai-rx.org}. A forum where Bonsai
-users daily exchange questions and answers can be accessed at
-\url{https://groups.google.com/g/bonsai-users}. Currently more than 700 users
-contribute to this forum (Figure~1f, \emph{Case for Support}).
+% Briefly describe how data will be stored, backed-up, managed and curated in the short to medium term. Specify any community agreed or other formal data standards used (with URL references).
+
+Software will be developed within the git distributed version control system, with public repositories made available within a specially created organisation at github.com.
+
+Copies of all repositories will be maintained within the Sainsbury Wellcome Centre data storage system at UCL.  **NEED**: short/medium term backup protocols from John Pelan's DMP
 
@@
 
-\textcolor{orange}{\textit{Briefly describe how data will be stored, backed-up, managed and curated in the short to medium term.
-Specify any community agreed or other formal data standards used (with URL references).}}
+% Bonsai is a freely available open-source resource, already widely adopted by
+% the systems neuroscience and behavioural science communities (see Fig 1f).
+% Current contributors include other large open projects (OpenEphys, DeepLabCut),
+% and companies developing and supplying new hardware devices
+% (Neurophotometrics). This community uptake, engagement, and commitment provides
+% a foundation for long-term sustainability.
 
-Bonsai is a freely available open-source resource, already widely adopted by
-the systems neuroscience and behavioural science communities (see Fig 1f).
-Current contributors include other large open projects (OpenEphys, DeepLabCut),
-and companies developing and supplying new hardware devices
-(Neurophotometrics). This community uptake, engagement, and commitment provides
-a foundation for long-term sustainability.
+% NeuroGEARS Ltd.\ is a keystone of this community and a key collaborator in the
+% current proposal. Founded by the original developer of Bonsai, it helps to
+% support and expand the user base well beyond the life of the current proposal.
+% NeuroGEARS currently provides quality control and curation for the core Bonsai
+% code, and provides training and support services for research institutions
+% around the world who want to start using Bonsai. NeuroGEARS also facilitates
+% knowledge transfer between users and institutions adopting Bonsai, by
+% moderating the community user forums and collaborating with organisations who
+% want to build new innovative research projects on top of Bonsai, as in the
+% current proposal.
 
-NeuroGEARS Ltd.\ is a keystone of this community and a key collaborator in the
-current proposal. Founded by the original developer of Bonsai, it helps to
-support and expand the user base well beyond the life of the current proposal.
-NeuroGEARS currently provides quality control and curation for the core Bonsai
-code, and provides training and support services for research institutions
-around the world who want to start using Bonsai. NeuroGEARS also facilitates
-knowledge transfer between users and institutions adopting Bonsai, by
-moderating the community user forums and collaborating with organisations who
-want to build new innovative research projects on top of Bonsai, as in the
-current proposal.
-
-The specific contributions developed as part of this project will form new core
-components and packages within the Bonsai ecosystem. In particular, the new
-language integration packages will support long-term sustainability, by
-dramatically increasing Bonsai's interoperability with existing software
-development communities which are already well established in basic science
-(MATLAB, Python, R). Developing these new packages with input from the core
-advisory panel of Bonsai, and early public distribution and user testing within
-the SWC and with planned workshops, will ensure solid and broad user engagement
-from the beginning of the development cycle and will therefore build momentum
-for their use, support and dissemination. Close collaboration with the core
-maintainers of the Bonsai ecosystem at NeuroGEARS will allow them to take over
-responsibility for maintenance and support of these packages beyond the end of
-the project, following the same principles of open-source and free sharing of
-knowledge that characterises the entire Bonsai ecosystem.
+% The specific contributions developed as part of this project will form new core
+% components and packages within the Bonsai ecosystem. In particular, the new
+% language integration packages will support long-term sustainability, by
+% dramatically increasing Bonsai's interoperability with existing software
+% development communities which are already well established in basic science
+% (MATLAB, Python, R). Developing these new packages with input from the core
+% advisory panel of Bonsai, and early public distribution and user testing within
+% the SWC and with planned workshops, will ensure solid and broad user engagement
+% from the beginning of the development cycle and will therefore build momentum
+% for their use, support and dissemination. Close collaboration with the core
+% maintainers of the Bonsai ecosystem at NeuroGEARS will allow them to take over
+% responsibility for maintenance and support of these packages beyond the end of
+% the project, following the same principles of open-source and free sharing of
+% knowledge that characterises the entire Bonsai ecosystem.
 
@@
 
-The fact that Bonsai is distributed as an open source project
-(\url{https://bonsai-rx.org/}), that is has extensive online documentation
-(\url{https://bonsai-rx.org/docs/}) and a vibrant user discussion forum
-(\url{https://groups.google.com/g/bonsai-users}), and that is integrated with
-many other open-source projects (Figure~1A and~1B, \emph{Case for Support})
-enables the use of the Bonsai ecosystem by the general public.
+% documentation
+Current Bonsai documentation, including learning materials and all relevant community
+links are publicly accessible at \url{bonsai-rx.org}.
+
+Documentation will be created for the new modules produced in this project that meets the existing standards, and will be integrated into the Bonsai documentation framework.
+
+% The fact that Bonsai is distributed as an open source project
+% (\url{https://bonsai-rx.org/}), that is has extensive online documentation
+% (\url{https://bonsai-rx.org/docs/}) and a vibrant user discussion forum
+% (\url{https://groups.google.com/g/bonsai-users}), and that is integrated with
+% many other open-source projects (Figure~1A and~1B, \emph{Case for Support})
+% enables the use of the Bonsai ecosystem by the general public.
 
@@
 
-Plans and place for long-term storage, preservation, and planned retention period for the research data.
+% Plans and place for long-term storage, preservation, and planned retention period for the research data.
 Formal preservation standards, if any.
@@
 % Identify any data repository (-ies) that are, or will be, entrusted with storing, curating and/or sharing data from your study, where they exist for particular disciplinary domains or data types.
-Software modules and documentation will be available on github, with links from the Bonsai project site (https://bonsai-rx.org) and available through the Bonsai system's package management interface.
+Software modules and documentation will be available through the github.com service, with links from the Bonsai project site (https://bonsai-rx.org) and available through the Bonsai system's package management interface.
 
-NEED: github org and status?
+**NEED**: github org and status?
+
+Repositories will be mirrored within the Sainsbury Wellcome Centre scientific computing system.  **NEED**: long-term backup from John's DMP
 
@@
 
+% Indicate how potential new users (outside of your organisation) can find out about your data and identify whether it could be suitable for their research purposes, also consider how it will be licenced.
+The approach should encompass data in all its forms as well as software.
+This section could also outline access procedures, how your data will be able to be cited and tracked.
+
 Package management
@@
 Paper?
-
-% Indicate how potential new users (outside of your organisation) can find out about your data and identify whether it could be suitable for their research purposes, also consider how it will be licenced.
-The approach should encompass data in all its forms as well as software.
-This section could also outline access procedures, how your data will be able to be cited and tracked.
+ A forum where Bonsai
+users daily exchange questions and answers can be accessed at
+\url{https://groups.google.com/g/bonsai-users}. Currently more than 700 users
+contribute to this forum (Figure~1f, \emph{Case for Support}).
 
@@
 
-This is an open source software project.  Code will be distributed under ??? license.
-
 % What steps will you take to ensure that your data could be re-analysed by other researchers? (e.g. availability of meta data, worked examples, vignettes). How will metadata be made available?
 
+This is an open source software project.  Code will be distributed under **NEED** licence, and thus will be freely reusable and modifiable.
+
 \subsection{Restrictions or delays to sharing, with planned actions to limit such restrictions} 
-Restriction to data sharing may be due to subject confidentiality or to ensure you can gain IP protection.
+% Restriction to data sharing may be due to subject confidentiality or to ensure you can gain IP protection.
 Applicants should ensure they have obtained necessary clearances from relevant collaborators with regards to the content of the proposal including the data sharing plan in line with the BBSRC Research Grants Guide.
@@
 
-All distributed software will be developed as part of the project, based on published and unrestricted algorithms.  Thus it is unencumbered by IP restrictions.  Will be available as above.
+All distributed software will be developed as part of the project, based on published and unrestricted algorithms.
+Thus it is unencumbered by IP restrictions and will be available throughout development as described above.
 
-4. Data security (where relevant)
-4.1	Formal information/data security standards
-% Identify formal information standards with which your study is or will be compliant. An example is ISO 27001. If your organisation is ISO compliant, please state the registration number.
+\section{Data security (where relevant)}
 
-Not applicable.
-
-4.2	Main risks to data security
-% All personal data has an element of risk.
-Summarise the main risks to the confidentiality and security of information related to human participants or users, the level of risk and how these risks will be managed.
-Cover the main processes or facilities for storage and processing of personal data, data access, with controls put in place and any auditing of user compliance with consent and security conditions.
-It is not sufficient to write ‘not applicable’ under this heading.
-
-This is a freely-available open source software package, which does not contain participants or personal data.
-User access to the repository is anonymous and user personal information will not be collected.
+The software produced will be freely available and open sourced.
+It will contain no personal data.
+User access to the repository will be anonymous and user personal information will not be collected.
+Thus security concerns do not apply.
 

### 2021-11-30 application/dataManagementPlan_long_joaco.tex
Multiple modificiations

@@
 
-Software modules will comprise source files in a number of programming languages: C\#, Python, MATLAB and R.  These will be accompanied by extensive documentation (discussed below).
+Software modules will comprise source files in various programming languages:
+C\#, Python, MATLAB and R.  These will be accompanied by documentation
+(Section~\ref{sec:documentation}).
 
-NEED: Approximate lines of code??? (order of magnitude is fine)
+% **NEED**: Approximate lines of code??? (order of magnitude is fine)
 
-NEED: testing files, makefiles?, other stuff,
+% **NEED**: testing files, makefiles?, other stuff,
 
@@
 The source code for the core Bonsai ecosystem and the majority of extension
-packages are hosted at \url{github.com/bonsai-rx} (MIT license with no
-copyright and licencing restrictions).
+packages are hosted, with MIT license (i.e., no copyright and licencing
+restrictions), at \url{github.com/bonsai-rx}
 
@@
 
-\subsubsection*{Team responsibilities}
+Team responsibilities: Dr.~Lopes (Bonsai related software development, software
+releases, organization of Bonsai workshops and co-supervision of research
+software engineer), Dr.~Rapela (machine learning software development, testing
+maching learning functionality with behavioral and neural recordings and
+co-supervision of research software engineer), Prof.~Sahani (supervision on
+maching learning and neural data analysis), Prof.~Mrsic-Flogel (superivision on
+neuroscience topics and communication with the steering committee).
+
+We will use the Kanban software development model, which both Dr.~Lopes and
+Dr.~Rapela are already using in a collaborative project at the SWC/GCNU.
+
+Weekly meetings between Dr.~Lopes, Dr~Rapela and the to-be-hired research
+software engineer will be an integral component of the software development
+team. For instance, these meetings will discuss software engineering
+topics, perform code reviews, or practice paired programming. They will help
+every member of the software development team be informed about all aspect of
+the software development project, which will ensure that knowledge is not lost
+if a team member leaves the project. In addition, code review and
+pair-programming sessions will guarantee that the generated code shares a
+uniform style, which will help new contributors to the project (e.g., machine
+learning methods developers) easily understand the code functionality.
+
+We speculate the following risks of the project:
 
@@
 
-    \item[Goncalo Lopes] Bonsai related software development,
-        software releases, organization of Bonsai workshops
-        and co-supervision of research software engineer (RSE).
-
-    \item[Joaquin Rapela] machine learning (ML) related software development,
-        interaction with ML functionality methods developers, validating of ML
-        functionality with behavioral and neural recordings and co-supervision
-        of RSE.
-
-    \item[Maneesh Sahani] advise on ML.
-
-    \item[Tom Mrsic-Flogel] advise on neuroscience topics and communication
-        with the steering committee.
-
-\end{description}
-
-\subsection*{Resourcing and responsibility}
-
-\begin{description}
-
-    \item[software development model] we will use the Kanban software
-        development model, which we both Dr.~Lopes and Dr.~Rapela are already
-        using in a collaborative project at the SWC/GCNU.
-
-    \item[weekly meetings] between Goncalo Lopes, Joaquin Rapela and the RSE
-        will be an essential component of the software development process.
-        These will ensure that every member of the team is informed about both
-        the Bonsai integration and the machine learning developments and will
-        make the software development team more reliable upon the absence of
-        any of its members.
-
-    \item[code review and pair-programming] sessions will be held on a monthly basis among Goncalo Lopes, Joaquin Rapela and the recruited software engineer to deliver readable code sharing a standard programming style.
-
-\end{description}
-
-\subsubsection*{Major risks}
-
-\begin{description}
-
-    \item[language integration] we proposed to use existing software libraries
-        to enable the communication between Bonsai, written in C\#, and other
-        machine learning algorithms, written in Python (Python.NET), R (R.Net)
-        and Matlab (native Matlab .Net interface). We have not yet tested any
-        of these existing libraries, and we may found that they don't function
-        reliably enough for our purposes. If this were the case, we proposed to
-        build our own interface protocol using the ZeroMQ messaging library.
-        This risk will only delay the progress of the proposed project, but
-        will not hinder the deployment of its proposed outputs.
+    \item[problem in language integration] we proposed to use existing software
+        libraries to enable the communication between Bonsai, written in C\#,
+        and other machine learning algorithms, written in Python (Python.NET),
+        R (R.Net) and Matlab (native Matlab .Net interface). We have not yet
+        tested any of these existing libraries, and we may found that they
+        don't function reliably enough for our purposes. If this were the case,
+        we proposed to build our own interface protocol using the ZeroMQ
+        messaging library.  This risk will only delay the progress of the
+        proposed project, but will not hinder the deployment of its proposed
+        outputs.
 
@@
         Case for Support document (Figure 2, green box) we proposed to
-        integrate existing machine learning algorithms into the Bonsai
-        ecsosystem. It could be the case that some of these algorithms do not
-        work as reliably as needed for our resource. I such a case we will use
-        another algorithm with similar functionality.  This risk should not be
-        a major problem, since most of the machine learning functionality we
-        proposed to incorporate into Bonsai has been developed by us, or we are
-        experienced using it.
+        integrate existing machine learning methods into the Bonsai ecsosystem.
+        It could be that some of these methods do not work as reliably as
+        needed for our resource. In such a case we will use another existing
+        methods with similar functionality.  This risk should not be a major
+        problem for the proposed project, since most of the machine learning
+        methods we proposed to incorporate into Bonsai have been
+        developed by us, or we are substantial experienced using them.
 
-    \item[lack of interest from Bonsai users] We will minimize the probability
-        of this problem by advertising broadly the new functionality upon
-        release.  However, we don't think this would be a real problem, since
-        Bonsai users have reapeatly asked the addition of machine learning
-        functionality to Bonsai (Section 3 Community demand, Case for Support),
-        and because the machine learning functionality we proposed to add to
-        Bonsai are popular among experimental-control neurocientists.
+    \item[lack of interest from Bonsai users] we will minimize the probability
+        of this problem by demonstrating broadly among Bonsai users the unique
+        capabilities of the new machine learning functionality.  We don't think
+        this would be a serious problem, since Bonsai users have reapeatly
+        asked the addition of machine learning functionality to Bonsai (Section
+        3 Community demand, Case for Support), and because the functionality we
+        proposed to add to Bonsai (Section~2.3, Case for Support) is of large
+        interest among experimental-control neurocientists.
 
-    \item[lack of interest from machine-learning methods developers] To
-        minimize the probability that this risk turns into a large problem we
-        will take three actions. First, we will provide machine-learning
-        methods developers and easy-to-use API to interface their methods into
-        the Bonsai ecosystem. Second, we will show machine-learning methods
-        developers the large user exposure that Bonsai provides to integrated
-        ML functionality. Third, we will organize Bonsai workshops for methods
-        developers showcasing machine learning functionality integrated into
-        the Bonsai ecosystem and instructing them how to integrate new methods
-        into Bonsai.
+    \item[lack of interest from machine-learning methods developers] to reduce
+        the probability of this problem we will take three actions. First, we
+        will provide machine-learning methods developers and easy-to-use API to
+        interface their methods into the Bonsai ecosystem. Second, we will
+        demonstrate them the large user exposure that Bonsai offers to
+        integrated maching learning functionality. Third, we will organize
+        Bonsai workshops for methods developers showcasing machine learning
+        methods already integrated into the Bonsai ecosystem and instructing
+        them on how to integrate new methods into the Bonsai ecosystem.
 
@@
 \subsection{Metadata standards and data documentation}
-
+\label{sec:documentation}
 % \textcolor{orange}{\textit{What metadata is produced about the data generated from the research?
@@
 700 users contribute to this forum (Figure~1f, \emph{Case for Support}).
+Documentation will be created for the new modules produced in this project that
+meets the existing standards, and will be integrated into the Bonsai
+documentation framework.
 
@@
 documentation tools (e.g., Sphinx for Python code, or Sphinx for .Net for C\#)
-will be used to automatically generate API documentation. API documentation
+will be used to automatically generate API documentation. This documentation
 will be complemented with class and interaction diagrams, documenting the
@@
 
-\textcolor{orange}{\textit{Plans and place for long-term storage, preservation,
-and planned retention period for the research data. Formal preservation
-standards, if any. Indicate which data may not be retained (if any). Also,
-briefly describe what will be stored long-term, raw data, processed data?}}
+% \textcolor{orange}{\textit{Plans and place for long-term storage, preservation, and planned retention period for the research data.
+Formal preservation standards, if any.
+Indicate which data may not be retained (if any).
+Also, briefly describe what will be stored long-term, raw data, processed data?}}
 
 For persistent and long term preservation we propose to use
-Zenodo\footnote{\url{https://zenodo.org/}}.
+Zenodo\footnote{\url{https://zenodo.org/}}, which is well integrated with
+GitHub.
 
@@
 
-Software modules and documentation will be available on Github, with links from
-the Bonsai project site (https://bonsai-rx.org) and available through the
-Bonsai system's package management interface.
-
-NEED: github org and status?
+Software modules will be available on Github and on Zenodo for long-term
+preservation, with links from the Bonsai project site
+(\url{https://bonsai-rx.org}) and available through the Bonsai system's package
+management interface. Tutorial on how to use the added machine learning
+functionality will be available at the Bonsai project site
+(\url{https://bonsai-rx.org}) and automatically generated API documentation for
+this functionality will be available through
+ReadTheDocs\footnote{\url{https://readthedocs.org/}}.
 
@@
 
-Package management
-Bonsai website
-Community fora
-Workshops
-Paper?
-
 % Indicate how potential new users (outside of your organisation) can find out about your data and identify whether it could be suitable for their research purposes, also consider how it will be licenced.
@@
 
+% Package management
+% Bonsai website
+% Community fora
+% Workshops
+% Paper?
+ % A forum where Bonsai
+% users daily exchange questions and answers can be accessed at
+% \url{https://groups.google.com/g/bonsai-users}. Currently more than 700 users
+% contribute to this forum (Figure~1f, \emph{Case for Support}).
+
+Currently Bonsai is a well-known software. A strategy that its author has
+used to gain this visiability is to build interoperability with a large number
+of open-source source projects for data acquistion. With the addition of the
+proposed machine learning functionality, Bonsai will interact with popular
+machine learning algorithms, which will confer Bonsai further visibility.
+
 \subsection{How will data be made reusable?}
-
-This is an open source software project.  Code will be distributed under ??? license.
 
@@
 
+Several factors will contribute to the reusability of new functionality
+incorporated into Bonsai. Code implementing this functionality will be
+available in Github and Zenodo, under MIT license. Emphasis on standard and
+uniform code style, achieved with the help of code review and paired
+programming, will generate readable code, easier to read and extend by third
+parties. Automated build methods, will simplify code building and encourage
+modification by new developers. Test cases with large coverage will motivate
+software developers to modify Bonsai, without the risk of introducing
+undetected bugs into the software. Carefully designed software, with simple and
+clear object interfaces, will simplify the resuse of the proposed software.
+
 \subsection{Restrictions or delays to sharing, with planned actions to limit such restrictions} 
-Restriction to data sharing may be due to subject confidentiality or to ensure you can gain IP protection.
+% Restriction to data sharing may be due to subject confidentiality or to ensure you can gain IP protection.
 Applicants should ensure they have obtained necessary clearances from relevant collaborators with regards to the content of the proposal including the data sharing plan in line with the BBSRC Research Grants Guide.

### 2021-11-30 application/dataManagementPlan_longJR.tex
Fixed typos

--- 2021-11-30 application/dataManagementPlan_long_joaco.tex
@@
 Team responsibilities: Dr.~Lopes (Bonsai related software development, software
-releases, organization of Bonsai workshops and co-supervision of research
+releases, organisation of Bonsai workshops and co-supervision of research
 software engineer), Dr.~Rapela (machine learning software development, testing
-maching learning functionality with behavioral and neural recordings and
+machine learning functionality with behavioral and neural recordings and
 co-supervision of research software engineer), Prof.~Sahani (supervision on
-maching learning and neural data analysis), Prof.~Mrsic-Flogel (superivision on
+machine learning and neural data analysis), Prof.~Mrsic-Flogel (supervision on
 neuroscience topics and communication with the steering committee).
@@
 
-    \item[disfunctional machine learning functionality] in Section 2.3 of the
+    \item[dysfunctional machine learning functionality] in Section 2.3 of the
         Case for Support document (Figure 2, green box) we proposed to
-        integrate existing machine learning methods into the Bonsai ecsosystem.
+        integrate existing machine learning methods into the Bonsai ecosystem.
         It could be that some of these methods do not work as reliably as
@@
 
-    \item[lack of interest from Bonsai users] we will minimize the probability
+    \item[lack of interest from Bonsai users] we will minimise the probability
         of this problem by demonstrating broadly among Bonsai users the unique
         capabilities of the new machine learning functionality.  We don't think
-        this would be a serious problem, since Bonsai users have reapeatly
+        this would be a serious problem, since Bonsai users have repeatedly
         asked the addition of machine learning functionality to Bonsai (Section
@@
         demonstrate them the large user exposure that Bonsai offers to
-        integrated maching learning functionality. Third, we will organize
+        integrated machine learning functionality. Third, we will organise
         Bonsai workshops for methods developers showcasing machine learning
@@
 Currently Bonsai is a well-known software. A strategy that its author has
-used to gain this visiability is to build interoperability with a large number
-of open-source source projects for data acquistion. With the addition of the
+used to gain this visibility is to build interoperability with a large number
+of open-source source projects for data acquisition. With the addition of the
 proposed machine learning functionality, Bonsai will interact with popular
@@
 undetected bugs into the software. Carefully designed software, with simple and
-clear object interfaces, will simplify the resuse of the proposed software.
+clear object interfaces, will simplify the reuse of the proposed software.
 

### 2021-11-30 application/dataManagementPlan_long.tex
Updates from Overleaf

@@
 
-The project will generate software modules appropriate for use in behavioural neuroscience experiments in animal models, and other types of study that involve a combination of reactive experimental (particularly behavioural) control and streamed data acquisition.
+The project will generate software modules appropriate for use in behavioural neuroscience experiments using animal models, and other types of study that involve a combination of reactive experimental (particularly behavioural) control and streamed data acquisition.
 
@@
 
-Software modules will comprise source files in various programming languages: C\#, Python, MATLAB and R.  These will be accompanied by extensive documentation (discussed below).
-
-**NEED**: Approximate lines of code??? (order of magnitude is fine)
-
-**NEED**: testing files, makefiles?, other stuff,
+Software modules will comprise source files in various programming languages: C\#, Python, MATLAB and R, along with supporting files for version control, compilation, testing, packaging and documentation.
 
@@
 
-Software will be developed within the git distributed version control system, with public repositories made available within a specially created organisation at github.com.
+Software will be developed within the git distributed version control
+system, with public repositories made available within a specially
+created organisation at github.com and linked from \url{github.com/bonsai-rx}.
 
-Copies of all repositories will be maintained within the Sainsbury Wellcome Centre data storage system at UCL.  **NEED**: short/medium term backup protocols from John Pelan's DMP
+Mirrors of all repositories will be maintained within the Sainsbury Wellcome Centre data storage system at UCL.
 
@@
 are easily available and, importantly, allows for reproducible deployment of
-user environments to ensure working experiments are not disrupted by system
+user environments to guarantee existing experiments are not disrupted by system
 changes.
@@
 
-The core compiler and IDE is extensively tested, and uses a strong type system
-to greatly reduce bugs at compile-time. Best practices for leveraging this
+The core compiler and the Integrated Development Environment (IDE) is extensively tested, and uses a strongly typed language
+that greatly reduce bugs at compile-time. Best practices for leveraging this
 infrastructure for building new packages are in place. The software engineers
@@
 % documentation
-Current Bonsai documentation, including learning materials and all relevant community
+Bonsai documentation, including learning materials and all relevant community
 links are publicly accessible at \url{bonsai-rx.org}.
+% A forum where Bonsai users daily exchange questions and answers can be accessed at the public
+% repository \url{https://groups.google.com/g/bonsai-users}. Currently more than
+% 700 users contribute to this forum (Figure~1f, \emph{Case for Support}).
+Documentation will be created for the new modules produced in this project that
+meets the existing standards, and will be integrated into the Bonsai
+documentation framework.
 
-Documentation will be created for the new modules produced in this project that meets the existing standards, and will be integrated into the Bonsai documentation framework.
-
-% The fact that Bonsai is distributed as an open source project
-% (\url{https://bonsai-rx.org/}), that is has extensive online documentation
-% (\url{https://bonsai-rx.org/docs/}) and a vibrant user discussion forum
-% (\url{https://groups.google.com/g/bonsai-users}), and that is integrated with
-% many other open-source projects (Figure~1A and~1B, \emph{Case for Support})
-% enables the use of the Bonsai ecosystem by the general public.
+Classes, methods and functions will be documented in the code, and automatic
+documentation tools (e.g., Sphinx for Python code, or Sphinx for .Net for C\#)
+will be used to automatically generate API documentation. This documentation
+will be complemented with class and interaction diagrams, documenting the
+structure and dynamics of the application objects.
 
@@
 
+Persistent storage and long term preservation will be achieved using
+Zenodo [\url{https://zenodo.org/}], which is well integrated with
+GitHub.
+
 \section{Data sharing and access}
@@
 % Identify any data repository (-ies) that are, or will be, entrusted with storing, curating and/or sharing data from your study, where they exist for particular disciplinary domains or data types.
-Software modules and documentation will be available through the github.com service, with links from the Bonsai project site (https://bonsai-rx.org) and available through the Bonsai system's package management interface.
-
-**NEED**: github org and status?
-
-Repositories will be mirrored within the Sainsbury Wellcome Centre scientific computing system.  **NEED**: long-term backup from John's DMP
+Software modules will be available on Github and on Zenodo for long-term
+preservation, with links from the Bonsai project site
+(\url{https://bonsai-rx.org}) and available through the Bonsai system's package
+management interface. Tutorials on how to use the added machine learning
+functionality will be available at the Bonsai project site
+(\url{https://bonsai-rx.org}) and automatically generated API documentation for
+this functionality will be available through ReadTheDocs [\url{https://readthedocs.org/}].
 
@@
 
-Release management: repositories will be public throughout, but marked as developmental.
+% Release management: repositories will be public throughout, but marked as developmental.
 Alpha-level code will be available but not publicised until after extensive testing.
 Will be integrated into package management and advertised once it achieves release quality.
+
+Modules will be made available once they have passed rigorous internal quality checks.
+Software will be automatically tested during development with a comprehensive set of unit test and integration tests, implemented as  GitHub Actions.  Once complete, each 
+module will be tested by a group of beta adopters (largely drawn from within the SWC to enable rapid feedback and debugging).
+Project outputs will be accepted when all tests run successfully and no major functional problems are reported in beta tests.
+
+Release management: we will create a new release after each of the machine
+learning functionalities (described in Section~2.3 of the Case for Support) is
+integrated into the Bonsai ecosystem following the above method to accept a
+project output. Credentials for release will be granted initially by Dr.
+Gonçalo Lopes, Head Engineer at Neurogears and the current curator of the
+Bonsai official package feed, but might be expanded to other team members after
+training.  Acceptance criteria for release will be a group decision between
+Dr.~Lopes, Dr.~Rapela, Prof.~Sahani and Prof.~Mrsic-Flogel.
 
@@
 
-Package management
-Bonsai website
-Community fora
-Workshops
-Paper?
- A forum where Bonsai
-users daily exchange questions and answers can be accessed at
-\url{https://groups.google.com/g/bonsai-users}. Currently more than 700 users
-contribute to this forum (Figure~1f, \emph{Case for Support}).
+Bonsai is already a widely adopted software ecosystem.
+Community users will engage with the development of new capabilities in this project through planned workshops and the Bonsai users forum [\url{https://groups.google.com/g/bonsai-users}].
+
+On release, modules will be announced on the Bonsai website and available for installation through the package management system.
+
+We anticipate that incorporation of these new capabilities, associated workshops, announcements, and possible future publications will drive further adoption of Bonsai in behavioural systems neuroscience laboratories worldwide.
 
@@
 
-This is an open source software project.  Code will be distributed under **NEED** licence, and thus will be freely reusable and modifiable.
+Several factors will contribute to the reusability of new functionality
+incorporated into Bonsai. Code implementing this functionality will be
+available in Github and Zenodo, under the permissive MIT license. Emphasis on standard and
+uniform code style, achieved with the help of code review and paired
+programming, will generate readable code, easier to read and extend by third
+parties. Automated build methods will simplify code building and encourage
+modification by new developers. Test cases with large coverage will motivate
+software developers to modify Bonsai, without the risk of introducing
+undetected bugs into the software. Carefully designed software, with simple and
+clear object interfaces, will simplify the reuse of the proposed software.
 
@@
 
+\section{Capabilities}
+
+The project collaborator, NeuroGEARS Ltd, and its lead, Dr.~Gon\c{c}alo Lopes, have extended experience of software management and distribution, and will support our efforts.
+In addition we will benefit from advice from UCL's recently established Centre for Advanced Research Computing.
+
+\section{Responsibilities}
+
+Drs.~Rapela (team member) and Lopes (collaborator) will jointly take responsibility for technical quality of version control, testing, documentation, and distribution.
+Oversight will be provided by Profs.~Sahani (co-I) and Mrsic-Flogel (PI) along with a planned scientific advisory board committee that will include international experts.
+
+\section{Relevant institutional, departmental or study policies on data sharing and data security}
+
+Not applicable.
+
+\section{Author}
+
+The plan was prepared jointly by the team.  Enquiries should be directed to the PI and Co-I.
+
 \end{document}

### 2021-11-30 application/dataManagementPlan_long.tex
Added UCL Research Data Management group
Added input from John Pelan

@@
 
-The project collaborator, NeuroGEARS Ltd, and its lead, Dr.~Gon\c{c}alo Lopes, have extended experience of software management and distribution, and will support our efforts.
-In addition we will benefit from advice from UCL's recently established Centre for Advanced Research Computing.
+The project collaborator, NeuroGEARS Ltd, and its lead, Dr.~Gon\c{c}alo Lopes,
+have extended experience of software management and distribution, and will
+support our efforts.  In addition we will benefit from advice from UCL's
+recently established Centre for Advanced Research Computing and the UCL
+Research Data Management group. Local storage platforms at the SWC/GCNU, with
+more than six peta bytes of space, will be used to mirrow Github repositories.
 

### 2021-11-30 application/dataManagementPlan_long.tex
Ammended Sec 5 and fixed typos

--- 2021-11-30 application/dataManagementPlan_long.tex
@@
 % Identify any data repository (-ies) that are, or will be, entrusted with storing, curating and/or sharing data from your study, where they exist for particular disciplinary domains or data types.
-Software modules will be available on Github and on Zenodo for long-term
+Software modules will be available on GitHub and on Zenodo for long-term
 preservation, with links from the Bonsai project site
@@
 incorporated into Bonsai. Code implementing this functionality will be
-available in Github and Zenodo, under the permissive MIT license. Emphasis on standard and
+available in GitHub and Zenodo, under the permissive MIT license. Emphasis on standard and
 uniform code style, achieved with the help of code review and paired
@@
 
-The project collaborator, NeuroGEARS Ltd, and its lead, Dr.~Gon\c{c}alo Lopes, have extended experience of software management and distribution, and will support our efforts.
-In addition we will benefit from advice from UCL's recently established Centre for Advanced Research Computing.
+The project collaborator, NeuroGEARS Ltd, and its lead, Dr.~Gon\c{c}alo Lopes,
+have extended experience of software management and distribution, and will
+support our efforts. In addition we will benefit from advice from the UCL
+Research Data Management group and from the recently established UCL's Centre
+for Advanced Research Computing. Local storage platforms at the SWC/GCNU,
+totaling more than six petabytes in storage space, will be used to mirror                                                                            
+GitHub repositories.
 

### 2022-03-03 reviews/responseReviewer_155296694.tex
Added

\documentclass[12pt]{letter}

\usepackage[margin=1.25in]{geometry}

\begin{document}

\begin{letter}{
    BBSRC-BBR Evaluation Committee\\
    Polaris House\\
    North Star Avenue\\
    Swindon, Wiltshire\\
    United Kingdon SN2 1UH
}
\opening{Dear Committee Members,}

We thank the reviewer for his/her comments, that we address below, indexed by
the corresponding sections.

\noindent\textbf{Excellence}

    \texttt{The problem is that building high-level ML objects may hide caveats
    of such algorithms and may make people who use them not understand what
    they are actually doing.}

We share these concerns with the reviewer, that we will address as follows.

We will invest substantial resources to educate Bonsai users about the
functionality of the machine learning methods that we will add to Bonsai. We
will provide extensive documentation, and devote ample time in workshops to
explain this functionality.
%
The Gatsby Computational Neuroscience Unit is a centre of excellence in
statistical neuroscience and machine learning. Its contributions will be
essential to build simple and high-quality statistical documentation related to
the interfaced methods.

In addition, for every machine-learning method that we will integrate into
    Bonsai, we will provide recommendations on how to report analysis results.

Critical to help Bonsai users understand the function of the integrated machine
learning methods, and interpret their estimated models' parameters, is to
provide extensive visualisations of these parameters. We will provide such
visualisations.

The Bonsai interfaces that we will develop to machine learning methods will
allow Bonsai users to set any parameter of these methods, and the proposed
documentation will describe in detail the function of all parameters. Although,
we will provide good default values, users will be able to set any parameter
value. In this sense, Bonsai will not be a high-level interface to these
methods.

A central goal of Bonsai is to allow people without programming experience to
perform sophisticated experiments and, with the addition of the proposed
machine-learning functionality, to analyse their results. However, this does
not mean that non-programmers should be ignorant about how the statistical
methods that they are using work. We will help Bonsai users understand the
statistical methods that they use. Programming experience, or deep statistical
knowledge, should not required for this.

Our vision is that Bonsai will not only allow its users to run advanced
machine-learning methods, but it will also serve as an educational tool, that
will provide its users deep understanding of what these methods do.

\noindent\textbf{Uniqueness and Availability}

    \texttt{The added high-level ML tools are already available as free
    packages, but this proposal puts them all together and within the Bonsai
    framework, so the uniqueness is somewhat lower.}

It is true that the machine learning functionality that we propose to
incorporate into Bonsai is already available as free packages. However, people
without much programming experience will most commonly not be able to use this
functionality to control or analyse their experiments. Thus, a first unique
contribution from our proposal is that we will provide machine learning
functionality to people without programming experience, which will most
probably accelerate discovery.

Experimental control and neural data analysis are currently highly
disassociated. Most neuroscientists collect their data and only afterwards
analyse it. This split has had negative consequences for the progress of
neuroscience, as the vast majority of neural data analysis methods are offline
and do not provide key functionality for online experimental control. Hence, a
second unique contribution from our proposal is that it will generate the need
of online data analysis methods, which will generate advances in the field.

\end{letter}

\end{document}

### 2022-03-03 reviews/responseReviewer_155296694.tex
Fixed typo in Kingdom
More updates

@@
     Swindon, Wiltshire\\
-    United Kingdon SN2 1UH
+    United Kingdom SN2 1UH
 }
@@
 
-    \texttt{The problem is that building high-level ML objects may hide caveats
-    of such algorithms and may make people who use them not understand what
-    they are actually doing.}
+\texttt{The problem is that building high-level ML objects may hide caveats of
+such algorithms and may make people who use them not understand what they are
+actually doing.}
 
-We share these concerns with the reviewer, that we will address as follows.
+We share these concerns with the reviewer. Some data-analysis open-source
+software packages allow users to fit complex statistical models by just
+clicking a few buttons in a graphical user interface. And sometimes, this ease
+of use leds users to apply these statistical models without sufficient
+understanding. We want to avoid this problem in Bonsai.
 
 We will invest substantial resources to educate Bonsai users about the
-functionality of the machine learning methods that we will add to Bonsai. We
-will provide extensive documentation, and devote ample time in workshops to
-explain this functionality.
+functionality of the machine learning methods added to Bonsai.  We will provide
+extensive written and video documentation explaining these methods.  In
+addition, in collaboration with methods authors, we will deliver talks in
+workshops explaining the function of added methods.
 %
@@
 Bonsai, we will provide recommendations on how to report analysis results.
+This documentation will be targeted to methods' users, as well as journal and
+grant reviewers. It will inform these reviewers what they could expect from
+statistical analysis perfromed with methods integrated into the Bonsai
+ecosysstem.
 
@@
 machine-learning methods, but it will also serve as an educational tool, that
-will provide its users deep understanding of what these methods do.
+will provide users deep understanding of what these methods do.
 
@@
 
-    \texttt{The added high-level ML tools are already available as free
-    packages, but this proposal puts them all together and within the Bonsai
-    framework, so the uniqueness is somewhat lower.}
+\texttt{The added high-level ML tools are already available as free packages,
+but this proposal puts them all together and within the Bonsai framework, so
+the uniqueness is somewhat lower.}
 
@@
 second unique contribution from our proposal is that it will generate the need
-of online data analysis methods, which will generate advances in the field.
+of online data analysis methods, which may generate advances in the field.
 

### 2022-03-04 reviews/responseReviewer_155296694.tex
Reviewer reply comments from Goncalo.

@@
 clicking a few buttons in a graphical user interface. And sometimes, this ease
-of use leds users to apply these statistical models without sufficient
+of use leads users to apply these statistical models without sufficient
 understanding. We want to avoid this problem in Bonsai.
@@
 We will invest substantial resources to educate Bonsai users about the
-functionality of the machine learning methods added to Bonsai.  We will provide
-extensive written and video documentation explaining these methods.  In
-addition, in collaboration with methods authors, we will deliver talks in
+functionality of the machine learning methods added to Bonsai. Furthermore, we will provide
+extensive written and video documentation explaining the integrated methods.
+We are currently developing a technical reference  website for Bonsai which is open-source, integrated into the language editor, and open for community contributions.
+We aim to make extensive use of this website to raise awareness of the details of each method very close to the users.
+In
+addition, and in collaboration with method authors, we will deliver talks in
 workshops explaining the function of added methods.
@@
 statistical neuroscience and machine learning. Its contributions will be
-essential to build simple and high-quality statistical documentation related to
+essential to building simple and high-quality statistical documentation related to
 the interfaced methods.
@@
 grant reviewers. It will inform these reviewers what they could expect from
-statistical analysis perfromed with methods integrated into the Bonsai
-ecosysstem.
+statistical analysis performed with methods integrated into the Bonsai
+ecosystem.
 
-Critical to help Bonsai users understand the function of the integrated machine
+Critical to helping Bonsai users understand the function of the integrated machine
 learning methods, and interpret their estimated models' parameters, is to
 provide extensive visualisations of these parameters. We will provide such
-visualisations.
+visualisations by taking advantage of Bonsai's built-in data visualisation functionality.
 
-The Bonsai interfaces that we will develop to machine learning methods will
+The Bonsai interfaces that we will develop with machine learning methods will
 allow Bonsai users to set any parameter of these methods, and the proposed
@@
 we will provide good default values, users will be able to set any parameter
-value. In this sense, Bonsai will not be a high-level interface to these
+value. In this sense, Bonsai will not by itself be a high-level interface to these
 methods.
@@
 statistical methods that they use. Programming experience, or deep statistical
-knowledge, should not required for this.
+knowledge, should not be required for this.
 
@@
 probably accelerate discovery.
+Bonsai has a good track record of providing advanced functionality to a broad community of users, for example in the existing integration with DeepLabCut (https://elifesciences.org/articles/61909) or the support for dynamic presentation of visual stimuli in BonVision (https://elifesciences.org/articles/65541).
 
@@
 of online data analysis methods, which may generate advances in the field.
+For example, we have discussed the possibility of enabling and standardising systems for online closed-loop control of behavioural tasks where the results of statistical inference directly feedback to modulate the environment to which animals are exposed.
+Such novel tasks remain very hard to implement at the moment, and their reproducibility is complicated by the fact that there is no standardised way to easily share such interactive systems.
+Bonsai is well positioned to address this problem and we expect that the impact of such practices will be very high in the experimentalist community.
 

### 2022-03-04 reviews/responseReviewer_155296694.tex
Added a few changes to Goncalo\'s latest edition

@@
 
+\usepackage[hypertexnames=false,colorlinks=true,breaklinks]{hyperref}
 \usepackage[margin=1.25in]{geometry}
@@
 
-We share these concerns with the reviewer. Some data-analysis open-source
+We share these concerns with the reviewer. Some current data-analysis open-source
 software packages allow users to fit complex statistical models by just
@@
 statistical neuroscience and machine learning. Its contributions will be
-essential to building simple and high-quality statistical documentation related to
+essential to building simple and accurate statistical documentation related to
 the interfaced methods.
@@
 machine-learning methods, but it will also serve as an educational tool, that
-will provide users deep understanding of what these methods do.
+will provide users deep understanding of machine learning functionality for
+experimental control and neural data analysis.
 
@@
 
-It is true that the machine learning functionality that we propose to
-incorporate into Bonsai is already available as free packages. However, people
-without much programming experience will most commonly not be able to use this
-functionality to control or analyse their experiments. Thus, a first unique
-contribution from our proposal is that we will provide machine learning
-functionality to people without programming experience, which will most
-probably accelerate discovery.
-Bonsai has a good track record of providing advanced functionality to a broad community of users, for example in the existing integration with DeepLabCut (https://elifesciences.org/articles/61909) or the support for dynamic presentation of visual stimuli in BonVision (https://elifesciences.org/articles/65541).
+It is not the aim of the proposed resource to develop new machine learning
+functionality. We intend to build a resource that can easily integrate into the
+experimental control loop existing reliable machine learning functionality
+developed by the machine learing community. In this way, the
+proposed resource should continuously provide to experimental neuroscientists
+up-to-date machine learning functionality. In addition, the resource should
+provide methods developers rich datasets on which to assess the performance of
+their contributed methods. We believe that these contributions are unique.
+
+Bonsai has a good track record of providing advanced functionality to a broad
+community of users, for example in the existing integration with
+DeepLabCut\footnote{\url{https://elifesciences.org/articles/61909}} or the
+support for dynamic presentation of visual stimuli in
+BonVision\footnote{\url{https://elifesciences.org/articles/65541}}. The
+proposed machine learning extensions should significantly enhance these
+contributions.
 
@@
 second unique contribution from our proposal is that it will generate the need
-of online data analysis methods, which may generate advances in the field.
-For example, we have discussed the possibility of enabling and standardising systems for online closed-loop control of behavioural tasks where the results of statistical inference directly feedback to modulate the environment to which animals are exposed.
-Such novel tasks remain very hard to implement at the moment, and their reproducibility is complicated by the fact that there is no standardised way to easily share such interactive systems.
-Bonsai is well positioned to address this problem and we expect that the impact of such practices will be very high in the experimentalist community.
+of online data analysis methods, which may generate advances in the field. For
+example, we have discussed the possibility of enabling and standardising
+systems for online closed-loop control of behavioural tasks where the results
+of statistical inference directly feedback to modulate the environment to which
+animals are exposed. Such novel tasks remain very hard to implement at the
+moment, and their reproducibility is complicated by the fact that there is no
+standardised way to easily share such interactive systems. Bonsai is well
+positioned to address this problem and we expect that the impact of such
+practices will be very high in the community of experimental neuroscientists.
 

### 2022-03-04 reviews/responseReviewers.tex
Added

--- 2022-03-04 reviews/responseReviewer_155296694.tex
@@
-\documentclass[12pt]{letter}
+\documentclass[11pt]{letter}
 
 \usepackage[hypertexnames=false,colorlinks=true,breaklinks]{hyperref}
-\usepackage[margin=1.25in]{geometry}
+\usepackage[a4paper, margin=2cm]{geometry}
 
@@
 \begin{letter}{
-    BBSRC-BBR Evaluation Committee\\
-    Polaris House\\
-    North Star Avenue\\
-    Swindon, Wiltshire\\
-    United Kingdom SN2 1UH
 }
-\opening{Dear Committee Members,}
 
-We thank the reviewer for his/her comments, that we address below, indexed by
-the corresponding sections.
+Thank you for the reviewers' feedback and the opportunity to respond to
+comments.
 
-\noindent\textbf{Excellence}
+Response to reviewer 175134370:
 
-\texttt{The problem is that building high-level ML objects may hide caveats of
-such algorithms and may make people who use them not understand what they are
-actually doing.}
+Regarding the specific comment on reaching new user communities: the proposed
+expansion of the Bonsai ecosystem, and the hackathons, workshops etc. mentioned
+in the proposal, will be disseminated via numerous methods and fora. Many of
+these methods have been successfully used to promote Bonsai thus far, including
+professional networks, social media, and engagement with open-source hardware
+projects such as Open-Ephys and Neurophotometrics, both of which have actively
+promoted Bonsai for use with their systems. As part of the proposed extension,
+we will demonstrate Bonsai and its new machine learning functionality at both
+experimental and computational neuroscience venues (e.g., BNA's Festival of
+Neuroscience, SFN's annual meeting, FENS meetings, Neuromatch Academy,
+NeuroIPS, CoSyNe), to engage experimental neuroscience users, as well as new
+neural data analysis method developers.
 
-We share these concerns with the reviewer. Some current data-analysis open-source
-software packages allow users to fit complex statistical models by just
-clicking a few buttons in a graphical user interface. And sometimes, this ease
-of use leads users to apply these statistical models without sufficient
-understanding. We want to avoid this problem in Bonsai.
+From SWC and GCNU, we will put together a comprehensive communications plan to
+engage new communities with Bonsai. We will use a range of communications
+channels including owned (SWC and GCNU websites, newsletters), shared (organic
+social media) and paid (targeted/boosted social media and paid search)
+channels. This could include publishing and publicising longer-form content on
+how Bonsai is currently used by the global neuroscience community (e.g. at the
+Allen Institute, Janelia Research Campus, etc.) and how the proposed expansion
+will enhance this.
+
+We will also empower current users to act as advocates for Bonsai, thereby
+maximising reach through existing networks. Group Leaders will promote Bonsai
+at conferences during research talks, and we will make targeted approaches to
+Group Leaders not currently using the tools who may both benefit from and be
+valuable promoters of Bonsai. This will include laboratories both within
+neuroscience and from other fields— potentially leveraging user contacts in the
+biological and drug screening groups already utilising Bonsai.
+
+\pagebreak
+
+Response to reviewer 155296694:
+
+We share the concerns of the reviewer regarding people using machine learning
+methods without understanding them. Some current data-analysis software
+packages allow users to fit complex statistical models by just clicking a few
+buttons in a graphical user interface. And sometimes, this ease of use leads
+users to apply these statistical models without sufficient understanding. We
+want to avoid this problem in Bonsai.
 
@@
 
-\noindent\textbf{Uniqueness and Availability}
-
-\texttt{The added high-level ML tools are already available as free packages,
-but this proposal puts them all together and within the Bonsai framework, so
-the uniqueness is somewhat lower.}
+Regarding the comment about the lower uniquess of our proposal by the fact that
+the machine learning functionality that we proposed to add to Bonsai already
+exists as free software package, we would like the make the following
+clarifications.
 
@@
 
+\pagebreak
+
+A response to reviewer 206073742 is not required.
+
 \end{letter}

### 2022-03-04 reviews/responseReviewers.tex
Removed extra fat

@@
 We will invest substantial resources to educate Bonsai users about the
-functionality of the machine learning methods added to Bonsai. Furthermore, we will provide
-extensive written and video documentation explaining the integrated methods.
-We are currently developing a technical reference  website for Bonsai which is open-source, integrated into the language editor, and open for community contributions.
-We aim to make extensive use of this website to raise awareness of the details of each method very close to the users.
-In
-addition, and in collaboration with method authors, we will deliver talks in
-workshops explaining the function of added methods.
+functionality of the machine learning methods added to Bonsai. In addition, we
+will provide extensive written and video documentation explaining the
+integrated methods. We are currently developing a technical reference  website
+for Bonsai which is open-source, integrated into the language editor, and open
+for community contributions. We aim to make extensive use of this website to
+explain relevant aspects of machine learning methods to the Bonsai community.
+Furthermore, in collaboration with method authors, we will deliver talks at
+workshops explaining the added methods.
 %
@@
 
-Critical to helping Bonsai users understand the function of the integrated machine
-learning methods, and interpret their estimated models' parameters, is to
-provide extensive visualisations of these parameters. We will provide such
-visualisations by taking advantage of Bonsai's built-in data visualisation functionality.
-
-The Bonsai interfaces that we will develop with machine learning methods will
-allow Bonsai users to set any parameter of these methods, and the proposed
-documentation will describe in detail the function of all parameters. Although,
-we will provide good default values, users will be able to set any parameter
-value. In this sense, Bonsai will not by itself be a high-level interface to these
-methods.
-
-A central goal of Bonsai is to allow people without programming experience to
-perform sophisticated experiments and, with the addition of the proposed
-machine-learning functionality, to analyse their results. However, this does
-not mean that non-programmers should be ignorant about how the statistical
-methods that they are using work. We will help Bonsai users understand the
-statistical methods that they use. Programming experience, or deep statistical
-knowledge, should not be required for this.
+Critical to helping Bonsai users understand the function of the integrated
+machine learning methods, and interpret their estimated models' parameters, is
+to provide extensive visualisations of these parameters. We will provide such
+visualisations by taking advantage of Bonsai's built-in data visualisation
+functionality.
 
@@
 machine-learning methods, but it will also serve as an educational tool, that
-will provide users deep understanding of machine learning functionality for
+will provide users deep understanding of machinelearning functionality for
 experimental control and neural data analysis.
@@
 the machine learning functionality that we proposed to add to Bonsai already
-exists as free software package, we would like the make the following
-clarifications.
-
-It is not the aim of the proposed resource to develop new machine learning
-functionality. We intend to build a resource that can easily integrate into the
-experimental control loop existing reliable machine learning functionality
-developed by the machine learing community. In this way, the
-proposed resource should continuously provide to experimental neuroscientists
-up-to-date machine learning functionality. In addition, the resource should
-provide methods developers rich datasets on which to assess the performance of
-their contributed methods. We believe that these contributions are unique.
-
-Bonsai has a good track record of providing advanced functionality to a broad
-community of users, for example in the existing integration with
-DeepLabCut\footnote{\url{https://elifesciences.org/articles/61909}} or the
-support for dynamic presentation of visual stimuli in
-BonVision\footnote{\url{https://elifesciences.org/articles/65541}}. The
-proposed machine learning extensions should significantly enhance these
-contributions.
-
-Experimental control and neural data analysis are currently highly
-disassociated. Most neuroscientists collect their data and only afterwards
-analyse it. This split has had negative consequences for the progress of
-neuroscience, as the vast majority of neural data analysis methods are offline
-and do not provide key functionality for online experimental control. Hence, a
-second unique contribution from our proposal is that it will generate the need
-of online data analysis methods, which may generate advances in the field. For
-example, we have discussed the possibility of enabling and standardising
-systems for online closed-loop control of behavioural tasks where the results
-of statistical inference directly feedback to modulate the environment to which
-animals are exposed. Such novel tasks remain very hard to implement at the
-moment, and their reproducibility is complicated by the fact that there is no
-standardised way to easily share such interactive systems. Bonsai is well
-positioned to address this problem and we expect that the impact of such
-practices will be very high in the community of experimental neuroscientists.
+exists as free software package, we would like to clarify that it is not the
+aim of the proposed resource to develop new machine learning functionality. We
+intend to build a resource that can easily integrate into the experimental
+control loop existing reliable machine learning functionality developed by the
+machine learing community. In this way, the proposed resource should provide to
+experimental neuroscientists up-to-date machine learning functionality, even
+after the finalization of the proposed development, in a sustainable manner. In
+addition, the resource should provide methods developers rich datasets on which
+to assess the performance of their contributed methods.  We believe that these
+contributions are quite unique.
 

### 2022-03-08 reviews/responseReviewers.tex
Added first comment from Maneesh

@@
 
+Besides describing clearly in the documentation, and in the graphical user
+interface, if suitable, the assumptions of every statistical method, in some
+cases we may be able to add automatic tests checking if assumptions of a
+statistical method (e.g., gaussianity, or homoscedasticity) are satisfied by the
+data on which the method is applied.
+
 In addition, for every machine-learning method that we will integrate into
@@
 
-Our vision is that Bonsai will not only allow its users to run advanced
-machine-learning methods, but it will also serve as an educational tool, that
-will provide users deep understanding of machinelearning functionality for
-experimental control and neural data analysis.
+Our vision is that Bonsai will not only allow to run advanced machine-learning
+methods, but it will also serve as an educational tool, that will provide users
+deep understanding of machine learning functionality for experimental control
+and neural data analysis.
 
-Regarding the comment about the lower uniquess of our proposal by the fact that
+Regarding the comment about the lower uniqueness of our proposal by the fact that
 the machine learning functionality that we proposed to add to Bonsai already
@@
 control loop existing reliable machine learning functionality developed by the
-machine learing community. In this way, the proposed resource should provide to
+machine learning community. In this way, the proposed resource should provide to
 experimental neuroscientists up-to-date machine learning functionality, even
-after the finalization of the proposed development, in a sustainable manner. In
+after the finalisation of the proposed development, in a sustainable manner. In
 addition, the resource should provide methods developers rich datasets on which

### 2022-03-08 reviews/responseReviewers.tex
Added one more suggestion from Maneesh

@@
 We will invest substantial resources to educate Bonsai users about the
-functionality of the machine learning methods added to Bonsai. In addition, we
-will provide extensive written and video documentation explaining the
-integrated methods. We are currently developing a technical reference  website
-for Bonsai which is open-source, integrated into the language editor, and open
-for community contributions. We aim to make extensive use of this website to
+functionality of the added machine learning methods. In addition, we will
+provide extensive written and video documentation explaining the integrated
+methods. We are currently developing a technical reference  website for Bonsai
+which is open-source, integrated into the language editor, and open for
+community contributions. We aim to make extensive use of this website to
 explain relevant aspects of machine learning methods to the Bonsai community.
@@
 
-Our vision is that Bonsai will not only allow to run advanced machine-learning
+Our aim is that Bonsai will not only allow to run advanced machine-learning
 methods, but it will also serve as an educational tool, that will provide users
@@
 
-Regarding the comment about the lower uniqueness of our proposal by the fact that
-the machine learning functionality that we proposed to add to Bonsai already
-exists as free software package, we would like to clarify that it is not the
-aim of the proposed resource to develop new machine learning functionality. We
-intend to build a resource that can easily integrate into the experimental
-control loop existing reliable machine learning functionality developed by the
-machine learning community. In this way, the proposed resource should provide to
-experimental neuroscientists up-to-date machine learning functionality, even
-after the finalisation of the proposed development, in a sustainable manner. In
-addition, the resource should provide methods developers rich datasets on which
-to assess the performance of their contributed methods.  We believe that these
-contributions are quite unique.
+One of the largest dangers of allowing users to run complex statistical models
+without proper understanding relates to testing statistical significance, which
+is a subtle problem, as noted by the reviewer.
+%
+However, most statistical models we propose to interface with Bonsai will
+provide predictive or exploratory functionality, avoiding problems with
+significance testing.
+%
+Some of these models (e.g., latent variable models) may require hypothesis
+testing (e.g., comparing which of several models best explains a given
+dataset). But most estimations will use Bayesian methods, which are better
+suited for hypothesis testing, as noted elsewhere (e.g., Price et al., 2020).
+
+Regarding the comment about the lower uniqueness of our proposal by the fact
+that the machine learning functionality that we proposed to add to Bonsai
+already exists as free software package, we would like to clarify that it is
+not the aim of the proposed resource to develop new machine learning
+functionality. We intend to build a resource that can easily integrate into the
+experimental control loop existing reliable machine learning functionality
+developed by the machine learning community. In this way, the proposed resource
+should provide to experimental neuroscientists up-to-date machine learning
+functionality, even after the finalisation of the proposed development, in a
+sustainable manner. In addition, the resource should provide methods developers
+rich datasets on which to assess the performance of their contributed methods.
+We believe that both contributions are quite unique.
 
